### 2019-10-19 repro/caam641/fd/paper.tex

\title{Notes on Convergence of FD Methods}
\maketitle

\section{Energy Estimates}
Abstract setup: 
\begin{itemize} 
\item Hilbert space $H=H_1 \oplus H_2$ (``continuum fields'')
\item Family of Hilbert spaces $G(h) =U(h) \oplus V(h), h>0$ (``grids'')
\item bounded self-adjoint positive definite operators 
\[
A(h): U(h) \rightarrow U(h), B(h): V(h) \rightarrow V(h)
\]
upper and lower bounds uniform in $h$ (``coefficients''):
\[
A_*\|u\|^2 \le \langle Au,u \rangle \le A^*\|u\|^2, \, B_*\|v\|^2 \le
\langle B v, v\rangle \le B^*\|v\|.
\]
\item bounded operators $D(h,r), D_1(h):U(h) \rightarrow V(h)$,
\[
\|D_1(h)\| \le D^* \mbox{ for all }h>0, D(h,r) = rD_1(h)
\]
(``stress divergence'').
\item Bounded operator $S(h):H \rightarrow G(h)$ (``sampling'')
\end{itemize}

Staggered grid dynamics: $(u^n(h),v^{n+1/2}(h)) \in G(h)$ satisfy 
\begin{eqnarray}
\label{eqn:sg}
A(u^{n+1} - u^n) & = & -D^T v^{n+1/2} \nonumber\\
B(v^{n+1/2}-v^{n-1/2}) & = & D u^n
\end{eqnarray}

Energy form:
\[
E = \frac{1}{2}(\langle Au,u\rangle + \langle Bv,v\rangle + \langle D^Tv,u\rangle ) 
\]
\[
= \frac{1}{2} \left(
\begin{array}{c}
u\\
v 
\end{array}
\right)^T
\left(
\begin{array}{cc}
A & \frac{1}{2}D^T\\
\frac{1}{2}D & B
\end{array}
\right)
\left(
\begin{array}{c}
u\\
v 
\end{array}
\right)
\]

\noindent Claim: for a solution $\{u^n,v^{n-1/2}$ of the system \ref{eqn:sg},
\[
E^n = E(u^n,v^{n-1/2})
\]
is independent of $n$.

Proof: 
\[
E^{n+1} - E^n =\langle A(u^{n+1}-u^n), (u^{n+1}+u^n)\rangle 
+ \langle B(v^{n+1/2}-v^{n-1/2}), (v^{n+1/2}+v^{n-1/2})\rangle +
\]
\[
\langle D^Tv^{n+1/2},u^{n+1}\rangle ) - \langle D^Tv^{n-1/2},u^n\rangle )
\]
\[
= \langle -D^Tv^{n+1/2},(u^{n+1}+u^n) \rangle 
+ \langle Du^n. (v^{n+1/2}+v^{n-1/2}) \rangle
+\langle D^Tv^{n+1/2},u^{n+1}\rangle ) - \langle D^Tv^{n-1/2},u^n\rangle )
\]
\[ 
=0.
\]
Q.E.D.

\noindent Claim: for sufficiently small $r$, there exist $0 < C_* \le C^*$ such that 
\begin{equation}
\label{eqn:pd}
C_*\|u,v\|^2 \le E(u,v) \le C^*\|u,v\|^2
\end{equation}

Proof: upper bound is clear, from $h$-uniform bounds on $A$, $B$,
$D$. 
\[
E(u,v) \ge A_*\|u\|^2 + B_*\|v\|^2 - rD^*\|u\|\|v\|
\]
which is a positive definite form in $\|u\|,\|v\|$ when $rD^* <
\sqrt{A^*B^*}$.
Q.E.D.

\section{$L^2$ Convergence of Smooth Solutions}
Inhomogeneous system:

\begin{eqnarray}
\label{eqn:in}
A(u^{n+1} - u^n) & = & -D^T v^{n+1/2} + hf^{n+1/2}\nonumber \\
B(v^{n+1/2}-v^{n-1/2}) & = & D u^n + h g^n
\end{eqnarray}

\noindent Claim: for $r$ sufficiently small that $E$ is positive definite, there
exists $K \ge 0$, $\lambda > 1$ so that so that
\[
E^n = E(u^n,v^{n-1/2}) \le \lambda^nE(u^0,v^{-1/2}) + Kh\sum_{m=0}^{n-1}\lambda^{n-m+1}(\|f^{m-1/2}\|^2 + \|g^m\|^2)
\]
Proof: The same arithmetic as used in the homogeneous case leads to
\[
E^{n+1}-E^n = h\langle f^{n+1/2}, u^{n+1}+u^n\rangle + h\langle g^n,
v^{n+1/2}+v^{n-1/2}\rangle
\]
\[
\le \frac{h}{2\alpha^2} (\|f^{n+1/2}\|^2 + \|g^n\|^2) +
\frac{\alpha^2h}{C_*}(E^{n+1}+E^n)
\]
for any $\alpha \in (0,1)$. Choose $\alpha$ so that $\alpha^2 h < C_*$
(that is, can make fixed choice of $\alpha$ for $h$ small enough), and
set 
\[
\lambda = \left(1+\frac{\alpha^2h}{C_*}\right) \left(1-\frac{\alpha^2h}{C_*}\right)^{-1}
\]
and 
\[
K = \frac{\frac{1}{2\alpha^2}}{ \left(1+\frac{\alpha^2h}{C_*}\right) }
\]
Then
\[
\lambda^{-1}E^{n+1} - E^n = Kh (\|f^{n+1/2}\|^2 + \|g^n\|^2)  
\] 
so 
\[
\lambda^{-(n+1)}E^{n+1} - \lambda^{-n}E^n = \lambda^{-n}Kh (\|f^{n+1/2}\|^2 + \|g^n\|^2) 
\]
whence
\[
\lambda^{-n}E^n = E^0 +
Kh\sum_{m=0}^{n-1}\lambda^{-m+1}(\|f^{m-1/2}\|^2 + \|g^m\|^2)
\]
Q.E.D.

Suppose that $(U^n,V^{n-1/2}) \in H$. Set
\begin{eqnarray}
\label{eqn:tr}
f^{n+1/2}(h)& =& h^{-1}(A(h)(S(h)U^{n+1} - S(h)U^n) + D(h,r)^T S(h)V^{n+1/2}))\\
g^{n}(h)& =& h^{-1}(B(h)(S(h)V^{n+1/2}-S(h)V^{n-1/2}) - D(h,r) S(h)U^n)
\end{eqnarray}

Suppose $(u^n,v^{n-1/2})$ solves \ref{eqn:sg} with 
\[
(u^0, v^{-1/2})=S(h)(U^0,V^{-1/2})
\]

\noindent Claim: Define $K, \lambda$ as above, then
\[
E((u^n,v^{n-1/2})-S(h)(U^n,V^{n-1/2})) \le
Kh\sum_{m=0}^{n-1}\lambda^{n-m+1}(\|f^{m-1/2}\|^2 + \|g^m\|^2)
\]
Proof: follows immediately from previous claim.

\noindent Corollary: Suppose $T>0$ given independent of $h$. If
$\|f^{m-1/2}\|^2 + \|g^m\|^2 \le L^2 h^{2p}$ for $0\le m \le N$, $Nh=T$, then 
\begin{equation}
\label{eqn:ee}
E((u^n,v^{n-1/2})-S(h)(U^n,V^{n-1/2})) \le M^2L^2h^{2p}
\end{equation}
Proof: from the preceding estimate,
\[
E((u^n,v^{n-1/2})-S(h)(U^n,V^{n-1/2})) \le KL^2
h^{2p+1}\frac{\lambda^N-1}{\lambda-1}
\]
Note that if $h$ is sufficiently small, then $\lambda < 1 +
2\frac{\alpha^2h}{C_*}$ so $\lambda^N \le
exp(2\frac{\alpha^2}{C_*}T)$, where $T=Nh$. If $h$ is perhaps smaller
yet, then $\lambda > 1+\frac{\alpha^2h}{2C_*}$. Putting this all
together, there is $M^2$ depending on $T$ and all of the other constants
in the setup, so that if $nh <T$ so that \ref{eqn:ee} holds.
Q. E. D.

We apply these results to 3D acoustics and elasticity. In either case,
$H_1 = L^2(\bR^3,\bR^N)$,  $H_2 = L^2(\bR^3,\bR^3)$. Members of $H_1$
are stresses ($N=6$) or pressures ($N=1$), members of $H_2$ are velocities.

$U(h) = $ square integrable $\bR^N$-valued functions on rectangular
grid of mesh $h$, $V(h) = $ square integrable $\bR^3$-valued functions
on half-cell shifted grids ($V_i$ shifted along $i$th axis) (discrete velocities)

$A$ = smooth symmetric $N \times N$ matrix-valued function on $\bR^3$,
uniformly bounded with univormly bounded inverse

$B$ = smooth symmetric $3 \times 3$ matrix-valued function on $\bR^3$,
uniformly bounded with univormly bounded inverse

View $(U,V)$ as $\bR^{N+3}$ valued, then 
\[
S(h)(U,V)_{i,j,k} = \left(
\begin{array}{c}
U(ih,jh,kh)\\
V_1((i+1/2)h,jh,kh)\\
V_2((ih,(j+1/2)h,kh)\\
V_3((ih,jh,(k+1/2)h)
\end{array}\right)
\]

Corresponding discretizations of $A$ and $B$: in the acoustic case, $A$ is scalar,
$(A(h) u)_{i,j,k} = \frac{1}{\kappa(ih,jh,kh)}u_{i,j,k}$. For
elasticity, it is a more complex evaluation of the Hooke tensor at
various grid points, which I will not give explicitly. In all cases,
$B(h)$ is a diagonal sampling of the density: 
\begin{eqnarray*}
(B(h)_{i,j,k})_{1,1} & = & \rho((i+1/2)h,jh,kh),\\
(B(h)_{i,j,k})_{2,2} & = & \rho(ih,(j+1/2)h,kh),\\
(B(h)_{i,j,k})_{3,3} & = & \rho(ih,jh,(k+1/2)h),\\
(B(h)_{i,j,k})_{l,m} & = & 0,\,l \ne m
\end{eqnarray*}

$D$ is a finite difference approximation to the strain (or gradient,
for acoustics) of order $2p$. For example, for $p=1$, one possibility is 
\[
D u =
\{(u_{i+1,j,k}-u_{i,j,k},u_{i,j+1,k}-u_{i,j,k},u_{i,j,k+1}-u_{i,j,k}\}
\]
This discrete gradient operator is a second-order approximation to the
partial derivatives of $u$ at half-cell positions along each
coordinate axis, corresponding to velocity sample points as described
above.

$D^T$ is discrete divergence, $r = \Delta t/h$.

Standard truncation error calculation: if $(U(t), V(t))$ is smooth solution of
acoustic (pressure-velocity) system with smooth right-had sides $F(t),G(t)$, 
\begin{eqnarray}
\label{vdawe.fd}
\frac{1}{\kappa}\frac{\partial U}{\partial t} + \nabla \cdot V & = &
                                                                     F,\\
\rho \frac{\partial V}{\partial t} + \nabla U & = & G
\end{eqnarray}
each time slice of compact
support, $(U^n(h),V^{n-1/2}(h)) = S(h)(U(n\Delta t),V((n-1/2)\Delta t))$, then
\begin{eqnarray}
\label{eqn:tr2}
f^{n+1/2}(h)& =& h^{-1}(A(h)(U^{n+1}(h) - S(h)U^n(h)) + D(h,r)^T
                 V^{n+1/2}(h))) = O(h^{2p})\\
g^{n}(h)& =& h^{-1}(B(h)(V^{n+1/2}(h)-V^{n-1/2}(h)) - D(h,r)
             U^n(h)) = O(h^{2p}+\Delta t^2)
\end{eqnarray}
pointwise, uniformly in $i,j,k$. Since the support is uniformly bounded if
$0\le n\Delta t = nh/r \le T$, the $O$ statements apply  to the $L^2$
norms as well.

Let $u$ denote the solution of \ref{eqn:sg} with
$(u^0,v^{-1/2})=S(h)(U^0,V^{-1/2})$. Then $(u,v)-S(h)(U,V)$ solves
\ref{eqn:in} with $f,g$ as just given.  Therefore the error estimate
\ref{eqn:ee} implies that 
\[
\|(u,v)-S(h)(U,V)\| = O(h^{2p} + \Delta t^2),
\]
that is, $(2,2p)$-th order convergence in the $L^2$ sense.

\section{Weak Convergence of Singular Solutions}