### 2019-10-19 repro/papers/asg/paper1_madplot.tex

\inputdir{project}
\plot{frame13}{width=15cm}{Point source field, homogeneous medium with
$v_p=1.5$ km/s, at 1.2 s}
\plot{frame40-1}{width=15cm}{Point source field at 4.0 s, after interaction with reflecting (zero-pressure) boundaries}
\plot{frame40-2}{width=15cm}{Point source field at 4.0 s, after interaction
with 250 m PML boundary zones on bottom and sides ($\eta_0=1.0$) -
same grey scale as Figure \ref{fig:frame40-1}. Longest wavelength
carrying significant energy is roughly 500 m.}
\plot{frame40-3}{width=15cm}{Point source field at 4.0 s, after interaction
with 100 m PML boundary zones on bottom and sides ($\eta_0=1.0$) -
same grey scale as Figure \ref{fig:frame40-1}.}
\plot{bm1}{width=15cm}{Dome bulk modulus}
\plot{by1}{width=15cm}{Dome buoyancy}
\plot{data1}{width=15cm}{2D shot record, (2,4) staggered grid scheme, $\Delta x = \Delta z =$ 5 m,
  appropriate $\Delta t$, 301 traces: shot x = 3300 m, shot z = 40 m, receiver x =
 100 - 6100 m, receiver z = 20 m, number of time samples = 1501, time
sample interval = 2 ms. Source pulse = zero phase trapezoidal [0.0,
2.4, 15.0, 20.0] Hz bandpass filter.}

\plot{trace}{width=15cm}{Trace 100 (receiver x = 2100 m) for $\Delta x = \Delta z = $
 20 m (black), 10 m (blue), 5 m (green), and 2.5 m (red). Note
 arrival time discrepancy after 1 s: this is the interface error
 discussed in \cite[]{SymesVdovina:09}. Except for the 20 m result,
 grid dispersion error is minimal.}

\plot{wtrace}{width=15cm}{Trace 100 detail, 1.8-2.5 s, showing more clearly the
 first-order interface error: the time shift between computed events
 and the truth (the 2.5 m result, more or less) is proportional to
 $\Delta t$, or equivalently to $\Delta z$.}

\plot{data8k1}{width=15cm}{2D shot record, (2,8) scheme, other
  parameters as in Figure \ref{fig:data1}.}

\plot{trace8k}{width=15cm}{Trace 100 computed with the (2,8) scheme,
  other parameters as described in the captions of Figures
  \ref{fig:data1} and \ref{fig:trace}.}

\plot{wtrace8k}{width=15cm}{Trace 100 detail, 1.8-2.5 s, (2,8) scheme..
Comparing to Figure \ref{fig:wtrace}, notice that the dispersion error for
the 20 m grid is considerably smaller, but the results for finer grids
are nearly identical to those produced by the (2,4) grids - almost all
of the remaining error is due to the presence of discontinuities in
the model.}