### 2019-10-19 repro/papers/iva/old.tex

\title{Inversion Velocity Analysis}
\date{}
\author{William W. Symes, The Rice Inversion Project}

\maketitle
\parskip 12pt
\begin{abstract}
Minimization of semblance subject to regularized linearized inversion with fixed penalty parameter(s) results in computable gradient approximation with controllable error. This is an intermediate result for the constrained formulation.
\end{abstract}

\section{Introduction}

\section{Mapping properties of separable extended modeling operators}
This is a Hilbert space story. For any Hilbert space $H$, the inner
product will be denoted $\langle \cdot, \cdot \rangle_H$, the norm
denoted $\| \cdot \|_H$, with the subscript omitted whenever the
choice of space is clear from context.

The natural objects in this study are {\em Hilbert space scales}, that
is sequences of Hilbert spaces $H = \{ H^s, s \in \bZ\}$, decreasing in
the sense that
$H^{s+1} \subset H^s, s \in \bZ$. Define the sets
\[
H^{\infty} = \cap_{s \in \bZ} H^s, \,H^{-\infty} = \cup_{s \in
  \bZ}H^s.
\]
An {\em operator} $L$  {\em of order $k \in \bZ$} from
a scale $H_1$ to a scale $H_2$ is linear function from $H_1^{-\infty}$ to
$H_2^{-\infty}$ for which $L|_{H_1^s} \in \sB(H_1^s,H_2^{s-k})$ for
all $ s \in
\bZ$. Denote by $OP^k(H_1,H_2)$ the set of all operators of order $k$.

The motivating example of this structure are scales of $L^2$ Sobolev
spaces, and (pseudo)differential operators on them.

The domain of a separable extended model consists of two
components. The first is a (background) model Hilbert space $M_b$ and
an open set of admissible models
$U \subset M_b$. In all examples, $M_b$ is a space of (possibly)
vector-valued) smooth ($C^{\infty}$, ``low frequency'') functions on a suitable physical
domain representing the physical parameters of a wave dynamics model, and members of $U$ obey additional constraints (such as
bounds) required to make the dynamical laws they parametrize
well-posed.

The other component of the model domain is a Hilbert space scale of extended
models $\oM =\{\oM_k, k\in \bZ\}$. 
A distinguished subscale $M = \{M_k, k\in \bZ\}, M_k \subset \oM_k$,
represents the non-extended or physical models. 
The  {\em annihilator} $A \in OP^0(\oM,N)$ has $M$ as its kernel:
\begin{equation}
\label{eqn:ann}
\om \in \oM,\,\,A\om=0 \Leftrightarrow \om \in M.
\end{equation}.
The {\em extension map} $E \in OP^0(M,\oM)$ is injective.

The model range or data space is another Hilbert scale $D$.

The extended modeling operator, or forward map, is a function
\[
\oF: U \rightarrow OP^0(\oM,D)
\]
It is {\em smooth with loss of regularity}: $m_b \mapsto \oF[m_b]|_{\oM^k}$ is a 
of class $C^p$ as a map from $U \rightarrow \sB(\oM^k,D^{k-p})$.

The physical modeling operator is a similar map
\[
F:U \rightarrow OP^0(M,D)
\]
with the same smoothness with loss of regularity.

The extended and physical modeling operators are related through the
extension map: for each $m_b \in U$,
\[
F[m_b] = \oF[m_b] \circ E.
\]

\section{Idealized extended inversion}
The main alleged advantage of extended modeling over standard modeling
is that {\em data can always be fit}, thus avoiding any possibility of
cycle-skipping. For the moment, then, idealize the extended modeling
operator $\oF[m_b]$ as bijective, for each admissible
background model $m_b \in U$: that is, assume that for any admissible
background model, it is always possible to perfectly fit data (certainly untrue
of standard modeling) with proper choice of extended model, and in
precisely one way. In some simple cases, this property actually holds,
while in others it holds in an approximate sense; in the latter case,
regularization is required, as will be explained in the next section.

In terms of the framework introduced in the last section, the inverse
problem is: given data $d \in D^0$, find a model $m_b \in U, m \in M^0$ so that
$F[m_b]m \approx d$. 
Call $d \in D$ {\em model-consistent} if there
exists $m_b^* \in U, m^* \in M$ so that $F[m_b^*]m^*=d$. 
In view of the extension property, and the assumed
bijectivity of $\oF[m_b]$, for any $m_b \in U$ there exists a unique
$\om[m_b,d] \in \oM$ so that
\be
\label{eqn:perfect}
\oF[m_b]\om[m_b,d] = d
\ee

{\em Inversion velocity analysis} (IVA) is the minimization of the objective
\be
\label{eqn:iva}
J[m_b] = \frac{1}{2}\|A\om[m_b,d]\|_0^2
\ee
over $m_b \in U$. If the data $d$ is model-consistent, then the global
minimum value is zero, since $M$ is the kernel of $A$. Therefore in
this case any global minimizer of $J$ is the first component of a
solution of the inverse problem, or equivalently a global minimizer of
the mean-square error $\|F[m_b]m-d\|_0$.

\ref{eqn:perfect}), is necessarily the background component of a
solution to the inverse problem. Note that this solution is not necessarily the
same as $m_b^*,m^*$, since $\oF$ (as opposed to $\oF[m_b], m_b \in U$)
has not been assumed injective.
If the data is not model-consistent, 
then $J[m_b]$ measures the distance to the feasible set ($M$) amongst
all extended models that fit the data, so a projection (left inverse
to the injection $M \subset \oM$) from $m_b,\om[m_b,d]$ onto $M$ will produce an
approximate solution.

Evidently (\ref{eqn:iva}) could be viewed as a constrained problem,
with constraint (\ref{eqn:perfect}). On the other hand, FWI could be
posed as minimzation of the extended residual (error in
(\ref{eqn:perfect})) subject to the constraint $A\om=0$. Thus IVA
may be viewed as FWI with the roles of objective and
constraint switched. While the two problems have the same solutions
for model-consistent data, the solutions are different for
model-inconsistent data. Also, all nontrivial examples of output least
squares inversion tend to generate a 
multiplicity of uninformative stationary points arbitrarily close to a
consistent-data global minimizer (the ``local min''
problem). In contrast, it can sometimes be shown that IVA
has only global minimizers, or at least that the IVA objective
is convex in a ball about the global solution.

Local minimization via Newton-like optimization requires computation
of the gradient. According to the smoothness assumptions made in the
last section,  $m_b \mapsto \om[m_b,d]$ is differentiable; by implicit
differentiation of (\ref{eqn:perfect}),
\begin{equation}
\label{eqn:dperfect}
(D\oF[m_b]\delta m_b)\om[m_b,d] + \oF[m_b]D\om[m_b,d]\delta m_b = 0,
\end{equation}
so
\[
DJ[m_b]\delta m_b = \langle D\om[m_b]\delta m_b, A^*A\om[m_b]\rangle
\]
\begin{equation}
\label{eqn:naive_jderiv}
= -\langle \oF[m_b]^{-1}(D\oF[m_b]\delta m_b)\om[m_b],A^*A\om[m_b]\rangle.
\end{equation}
For $m_b \in M_b$ and $\om \in \oM$, denote by $D\oF[m_b]\om$ the
linear map: $M_b \rightarrow D$ defined by
\[
\delta m_b \mapsto D(\oF[m_b]\om)\delta m_b
\]
and by $(D\oF[m_b]\om)^T$ its adjoint (note the distinction with the
adjoint $(D\oF[m_b]\delta m_b)^*$). Then 
whence the formal gradient of $J$ is
\be
\label{eqn:naive0_g}
\nabla J[m_b] = - (D\oF[m_b]\om[m_b])^T\bq[m_b]
\ee
in which $\bq \in D$ is the solution of
\be
\label{eqn:naive0_q}
\oF[m_b]^*\bq = A^*A\om[m_b].
\ee

Thus computation of the gradient appears to involve solution of two
linear systems ((\ref{eqn:perfect}) and (\ref{eqn:naive0_q})) and
application of the  {\em tomographic operator}
$(D\oF[m_b]\om[m_b,d])^T$. As explained elsewhere
\cite[]{KerSy:94,BiondiSava:04}. the tomographic operator is quite
computable via a variant of the {\em adjoint state method}, so modulo
the solution of the linear systems, (\ref{eqn:naive0_g}) appears to be
a computational {\em entr\'{e}} into IVA via quasi-Newton or conjugate
gradient methods. However several obstacles arise to practical use of
these ideas:
\begin{itemize}
\item for all but the simplest examples of extended seismic modeling, the linearized extended forward map $\oF$ is not actually surjective, therefore the equality constraint (\ref{eqn:perfect}) must be replaced by a best-fit criterion, such as least squares.
\item for all but the simplest examples of extended seismic modeling, the linearized extended forward map is not coercive, that is, the ratio $\|\oF[m_b]\om\|/\|\om\|$ attains arbitrarily small values (this is the case even if $\oF[m_b]$ is injective, which hardly matters). Therefore the least squares problem replacing (\ref{eqn:perfect}) must be regularized. On the other hand, for a natural choice of norm, the cone defined by $\|\oF[m_b]\om\| \ge \alpha \|\om\|$ for suitable $\alpha > 0$ contains a subspace whose dimension grows without bound as the frequency content of the modeled data grows without bound - that is an expression of the high information content of carried by propagating waves.
\item for all but the simplest examples of extended seismic modeling,
  the optimization problem replacing (\ref{eqn:perfect}) must be
  solved by an iterative process, simply because of computational size - therefore only an estimate of $\om[m_b,d]$ is available, and its error propagates into error in (regularized versions of) $J$ and $\nabla J$.
\item if the model parameters include wave velocities (that is to say,
almost always), the tomographic operator $(D\oF[m_b]\om[m_b,d])^T$ is not
continuous in the choice of norms for which the $\oF[m_b]$ is bounded
and has the ``partly coercive'' property mentioned in the second
bullet above. Otherwise put, in a choice of norms for which
$(D\oF[m_b]\om[m_b,d])^T$ is continuous, $\oF[m_b]$ is compact. If on the other hand
the iterative estimate oof $\om[m_b,d]$ relies on a choice of norms
for which $\oF[m_b]$ has a ``large'' well-conditioned subspace, then
no bound is possible for the effect of $(D\oF[m_b]\om[m_b,d])^T$ on
these error, hence no error control of the gradient is possible. In
practice, after discretization, this means that even a minimally accurate gradient requires
an excessively precise, hence computationally costly, estimation of $\om[m_b,d]$.
\end{itemize}

\section{Regularized extended inversion}
Overcoming the obstacles outlined at the end of the last section
requires introduction of additional abstract structure mirroring the
properties of important exaamples.

First, the extended domain space $\oM=\oM_0$ is member of a scale of Hilbert
spaces $\{\oM_s: s \in \bf{Z}\}$, each compactly embedded in the next: $\oM_s \subset \subset \oM_{s+1}$. Denote by $\langle \cdot ,\cdot \rangle_{s}$ the inner product
on $\oM_s$. Typically, $\{\oM_s\}$ is some version of the Sobolev
scale. Motivated by that example, assume that the inner product on
$\oM_0$ extends to a pairing: $\oM_s \times \oM_{-s} \rightarrow
\bf{R}$, realizing an isomorphism of $\oM_{-s}$ with the dual of
$\oM_s$, all $s \in \bf{Z}$.

We will assume that the data space $D = D_0$ is also a member of a
Hibert scale $\{D_s: s \in \bf{Z}\}$.

A linear map $L: \cup_{s \in \bf{Z}} X_s \rightarrow \cup_{s \in
  \bf{Z}}Y_s$ between scales $X, Y$ is {\em of order} $k$ iff $L: \in
{\cal L}(X_s,Y_{s-k})$ for every $s \in \bf{Z}$.

\noindent {\bf Assumption 1:} Assume that the operators $\oF[m_b]$ and $A$ extend to
the  {\em order 0} operators on the domain and range scales: $\oF[m_b]
\in {\cal L}(\oM_s,D_s)$ for every $s \in \bf{Z}$, similarly for $A$. The
notation $*$ is reserved for the adjoint of
$\oF[m_b]: \oM_0 \rightarrow D_0$.

\noindent {\bf Assumption 2:}  The {\em normal operator} $ m_b
\mapsto \oF[m_b]^*\oF[m_b]$ defines a smooth
${\cal L}(\oM_s,\oM_s)$-valued function on $U$ for every $s\in
\bf{Z}$, mapping $U$ to a bounded set. Similarly, the (``abnormal''?) operator $ m_b
\mapsto \oF[m_b]\oF[m_b]^*$ defines a smooth
${\cal L}(D_s,D_s)$-valued function on $U$ with bounded image, for every  $s \in \bf{Z}$.

\noindent {\bf Remark 1:} $\oF$ itself is {\em not} assumed to define a differentiable
map $\oM_s \rightarrow D_s$: that is because it does not, when wave
velocities are amongst the parameters in $m_b$. The (Gateaux) derivative does
exist, but only as an operator of order 1 - more on this below.

\noindent {\bf Remark 2:} Except in trivial instances, extended
modeling is {\em necessary} in order that $\oF\oF^*$ has the
smoothness property indicated.

Denote by $\bre[m_b,d]$ the solution of the $L^2$ least-squares problem,
\be
\label{eqn:reg}
\mbox{min}_{\br \in \oM_0}\, \frac{1}{2}(\|\oF[m_b]\br -\de\|_0^2 + \lambda \|\br\|_0^2).
\ee
This condition is of course a regularized version of (\ref{eqn:perfect}).
The first order condition for this problem is
\be
\label{l2normal}
\Ne[m_b]\br = \oF[m_b]^*\de
\ee
in which 
\[
\Ne[m_b] = \oF[m_b]^*\oF[m_b] + \lambda I
\]
Note that $\Ne$ is positive definite and self-adjoint on $M_0$, so the
Lax-Milgram Theorem asserts the existence of a unique $\bre[m_b,d] \in
L^2(X)$ satisfying (\ref{l2normal}).

In general the best possible estimate for the solution of (\label{l2normal})
is
\be
\label{eqn:genlest}
\|\om[m_b,d]\|_0  = O(\lambda^{-1}\|d\|_0)
\ee
However (\ref{eqn:genlest}) does not reflect the information content
of wave data, as remarked at the end of the last section.  To capture
this additional control possible in certain instances, posit the
existance of self-adjoint approximate projectors
\be
\label{eqn:dataproj}
\Pi_D,\bar{\Pi}_D \in \cap_{s \in \bf{Z}}{\cal L}(D_s, D_s), \,\Pi_D^*
= \Pi_D, \,\bar{\Pi}_D^* = \bar{\Pi}_D,
\ee
so that $\bar{\Pi}_D$ acts as an approximate identity on the range of
  $\Pi_D$, in the sense that 
\be
\label{eqn:projsmoothing}
(I-\bar{\Pi}_D)\Pi_D \in {\cal L}(D_s, D_{s+k}) \mbox{ for every } s
\in \bf{Z}, k \in \bf{Z}^+.
\ee

\noindent {\bf Assumption 3:} There exist $\gamma >0$, and for every $s \in \bf{Z}$, $k \in
\bf{Z}^+$ $C_{s,k} \ge 0$, so that
\be
\label{eqn:microell}
\|\oF[m_b]\oF[m_b]^*\bar{\Pi}_Dd\|_s \ge \gamma \|\bar{\Pi}_Dd\|_s -
\|d\|_{s-k}
\ee

\noindent {\bf Remark 3:} It follows from (\ref{eqn:projsmoothing})
that a similar inequality holds for $\Pi_D$.

\noindent {\bf Remark 4:} The inequality (\ref{eqn:microell})
abstracts a key feature of several important examples, a microlocal version of G\aa rding's inequality
\cite[]{Tay:81}; it is established by appeal to microlocal ellipticity
of the extended modeling operator.

The final two assumptions concern properties of the data, expressed in
terms of homogeneous inequalities defining cones in $D_s$. The results
to follow are {\em aymptotic}, so concern, not a single datum but a family
$\{\de\}_{\lambda \in (0,\lambda_{\rm max}}$ of data. The index
parameter $\lambda$ will turn out to signify (roughly) wavelength.

The first property expresses concentration of energy in ``imageable''
parts of data space.

\noindent {\bf Assumption 4:} For every $s \in \bf{Z}$, there exists
$L_s \ge 0$ so that for all $k \in \bf{Z}^+$, $\lambda \in
(0, \lambda_{\rm max}]$,
\be
\label{micro}
\|(I- \Pi_D)\de\|_s \le L_s \|\de\|_{s-k}.
\ee

The second property makes the connection with wavelength, and expresses
the assumption that the data is concentrated at high frequency:

\noindent {\bf Assumption 5 -  High Frequency Cone Condition:} 
For each $s \in \bf{Z}$ there exists $K_s
\ge 0$ so that for  $\lambda \in
(0, \lambda_{\rm max}]$,
\be
\label{hfcd}
\|\de\|_{s} \le K_s \lambda\|\de\|_{s+1}
\ee

\noindent{\bf Theorem 1:} 
Under assumptions 1 - 5 above, there exist $C,\lambda_0>0$ so that $\lambda_0 \le \lambda_{\rm max}$ and
\[
\|\bre[m_b,d]\|_0 \le C\|\de\|_0
\]
uniformly in $\lambda \in (0,\lambda_0]$, $m \in U$.

\noindent{\bf Proof:} See following section. % Take $M=\oF[m_b]^*\oF[m_b]$, $P=I$,
%$q_{\lambda}=\oF[m_b]^*\de$, $s=0$ and $k=1$ in the statement of Corollary A6. Lemma 2 verifies the required cond%itions on $M$, $P$, and $q_{\lambda}$. Q. E. D.

The semblance optimization functional [REF TO DEF IN INTRO] may be written as
\be
\label{tgt}
\Je[m_b] = \frac{1}{2}\|A\bre[m_b,d]\|^2
\ee
\[
=\frac{1}{2}\langle \de, \oF[m_b]\Ne[m_b]^{-1} A^*A \Ne[m_b]^{-1}\oF[m_b]^*\de \rangle
\]
%From Egorov's Theorem and Assumption 1 follows that operator in this expression is $\Psi$DO with symbol depending smoothly on $m$. Therefore $\Je$ is smooth in $m$.

\noindent {\bf Proposition 1:} $\Ne \in C^1(U,P_0)$, and for $\br \in \oM_1$, $\delta m_b \in M$,
\[
(D\Ne[m,\delta m]\br = (D\oF[m,\delta m]\oF[m_b]+\oF[m_b]D\oF[m,\delta m])\br
\]
\noindent {\bf Proof:} ...

For $\delta m \in M$,
\[
D\Je[m_b]\delta m = \langle D\bre[m_b,d]\delta m, A^*A\bre[m_b,d]\rangle.
\]
Thanks to Proposition 1,
\[
D\bre[m_b,d]\delta m = 
D\Ne[m_b]^{-1}(\delta m,\oF^*[m_b]\de) + \Ne[m_b]^{-1}D\oF[m_b]^*(\delta m, \de))
\]
\[
= -\Ne[m_b]^{-1}D\Ne[m_b](\delta m,\Ne[m_b]^{-1}\oF^*[m_b]\de) + \Ne[m_b]^{-1}D\oF[m_b]^*(\delta m, \de)
\]
\[
=-\Ne[m_b]^{-1}D\oF[m_b]^*(\delta m,\oF[m_b]\bre[m_b,d]) - \Ne[m_b]^{-1}\oF[m_b]^*D\oF[m_b](\delta m,\bre[m_b,d])
\]
\be
\label{prerderiv}
+ \Ne[m_b]^{-1}D\oF[m_b]^*(\delta m, \de)
\ee
\be
\label{rderiv}
=\Ne^{-1}(D\oF[m_b]^*(\delta m, \de - \oF[m_b]\bre[m_b,d]) - \oF[m_b]^*D\oF[m_b](\delta m, \bre[m_b,d])).
\ee
Define
\be
\label{qdef}
\bqe[m_b]=\Ne[m_b]^{-1}A^*A\bre[m_b,d].
\ee
Then
\be
\label{Jderiv}
D\Je[m_b]\delta m =\langle D\oF[m_b]^*(\delta m, \de - \oF[m_b]\bre[m_b,d]) - \oF[m_b]^*D\oF[m_b](\delta m, \bre[m_b,d]), \bqe[m_b]\rangle 
\ee
from which it follows that the formal ($L^2$) gradient of $\Je$ is 
\be
\label{grad}
\nabla \Je[m_b] = D\oF[m_b]^*(\bqe[m_b],\de - \oF[m_b]\bre[m_b,d])- D\oF[m_b]^*(\bre[m_b,d],\oF[m_b]\bqe[m_b]),
\ee

\noindent {\bf Remark:} That is, two linear least squares solves are required to compute the gradient of $\Je[m_b]$: (\ref{l2normal}) for $\bre[m_b,d]$ and (\ref{qdef}) for $\bqe[m_b]$.

\noindent {\bf Factorization Lemma:} $D\oF[m_b](\delta m,\br) = \oF[m_b]Q[m,\delta m]\br$, in which $Q[m,\delta m]$ is a smooth $OPS^1$-valued function of $(m,\delta m) \in U \times M$, linear in $\delta m$, whose values are essentially skew-adjoint:
\[
Q[m,\delta m] + Q[m,\delta m]^* = R[m,\delta m] \in OPS^0.
\]
For $u \in H^1$, 
\be
\label{Qnorm}
\|Q[m,\delta m]u\|_0 \le C\|\delta m\|_M \|u\|_1,
\ee
\be
\label{Rnorm}
\|R[m,\delta m]u\|_0 \le C\|\delta m\|_M \|u\|_0,
\ee
with $C$ uniform for $m\in U$.

It will be convenient to view $Q[m,\cdot]\br, m\in U, \br \in \oM$, as a continuous linear map $:M \rightarrow \oM$, and denote its adjoint by $Q^*[m,\cdot]\br$. That is,
\[
\langle Q^*[m,\br_2]\br_1,\delta m\rangle_M = \langle Q[m,\delta m]\br_1,\br_2\rangle.
\]
In the statement (\ref{Qnorm}), I have used $Q[m,\delta m]^*$ to denote the ``other'' adjoint, defined by 
\[
\langle Q[m,\delta m]^*\br_1,\br_2\rangle = \langle \br_1,Q[m,\delta m]\br_2\rangle
\]

\noindent {\bf Theorem 5:} 
\be
\label{qrderiv}
D\bre[m_b,d]\delta m = -Q[m,\delta m]\bre[m_b,d] + \lambda \Ne^{-1}R[m,\delta m]\bre[m_b,d].
\ee
\be
\label{eqn:gradq}
\nabla \Je[m_b] = (-Q^*[m,A^*A\bre[m_b,d]] + \lambda R^*[m,\bqe[m_b]])\bre[m_b,d].
\ee
For $C \ge 0$ uniform in small $\lambda$, $m \in U$,
\be
\label{eqn:firstgradest}
\|\nabla \Je[m_b] - -Q^*[m,A^*A\bre[m_b,d]]\bre[m_b,d] \|_M \le C\lambda \|\de\|_0.
\ee
Also, the $M$-valued quadratic form on $H^1$, defined by
\be
\label{qquad}
\br \mapsto -Q^*[m,A^*A\br]\br
\ee
extends continuously to $H^0$.

\noindent {\bf Remark:} The importance of the last statement stems from the effect of approximating $\bre[m_b,d]$, of course, as will be described below.

\noindent {\bf Proof:} From (\ref{prerderiv}) above and the definition of $Q$,
\[
D\bre[m_b,d]\delta m = 
-\Ne[m_b]^{-1}(Q[m,\delta m]^*\oF[m_b]^*\oF[m_b]\bre[m_b,d] + \oF[m_b]^*\oF[m_b]Q[m,\delta m])\bre[m_b,d]
\]
\[
 + \Ne[m_b]^{-1}Q[m,\delta m]^*\oF[m_b]^*\de)
\]
\[
= 
-\Ne[m_b]^{-1}([Q[m,\delta m]^*,\oF[m_b]^*\oF[m_b]] + \oF[m_b]^*\oF[m_b] (Q^*[m,\delta m]+Q[m,\delta m]))\bre[m_b,d]
\]
\[
 + \Ne[m_b]^{-1}Q[m,\delta m]^*\oF[m_b]^*\de)
\]
Using (\ref{l2normal}) and the definitions of $Q,R$, this is
\[
(-\Ne[m_b]^{-1}[Q[m,\delta m]^*,\oF[m_b]^*\oF[m_b]]-(I-\lambda \Ne[m_b]^{-1})R[m,\delta m])\bre[m_b,d]
\]
\[
+ \Ne[m_b]^{-1}([Q[m,\delta m]^*,\oF[m_b]^*\oF[m_b]] + \Ne[m_b]Q[m,\delta m]^*)\bre[m_b,d]
\]
\[
=(Q[m,\delta m]^*-(I-\lambda \Ne[m_b]^{-1})R[m,\delta m])\bre[m_b,d]
\]
\[
= -Q[m,\delta m]\bre[m_b,d] + \lambda \Ne^{-1}R[m,\delta m]\bre.
\]
as claimed.

The stated expression for $\nabla \Je[m_b]$ follows. The estimate amounts to a bound on the second term on the RHS. To see this, note that according to Theorem 4, $\bre[m_b,d]$ satisfies the hypotheses of Lemma A-4 with $s=0, k=1, M \leftarrow \oF[m_b]^*\oF[m_b], q \leftarrow \bre[m_b,d]$, and $\Pi \leftarrow \Pi[m_b]$. Accordingly from (\ref{a4:res1}) follows
\[
\|\bqe[m_b]\|_0 \le C\|\bre[m_b,d]\|_0.
\]
Since $R[m,\delta m]$ is continuous $OPS^0$-valued, the estimate follows from Theorem 1.

Finally, note that
\[
|\langle Q[m,\delta m]\br,A^*A \br\rangle| = |\langle ([A,Q[m,\delta m]]+Q[m,\delta m]A)\br,A\br\rangle|
\]
\[
=|\langle[A,Q[m,\delta m]]\br+\frac{1}{2}R[m,\delta m]A\br,A\br\rangle| \le C\|\br\|_0^2,
\]
since the operators appearing in the last expression are $\in OPS^0$.
Q. E. D.

The estimate (\ref{eqn:firstgradest}) suggests the use of the first
term in (\ref{eqn:gradq}) as an approximation to the gradient: in
essence, the use of the Factorization Lemma has isolated the result
$\bqe$ of the otherwise apparently necessary second system solve in a
term which is uniformly $O(\lambda)$. As $\lambda$ will in
applications be proportional to a wavelength, the achievable
resolution of the inversion, it seems reasonable to neglect this
term. Thus $\bqe$ need not be computed.

However the key player in the first term in (\ref{eqn:gradq}), namely
the operator $Q$, is not directly computable. Fortunatly, key examples
add yet another property to the extended modeling operator, formalized
in

\noindent {\bf Assumption 6:} For $m_b \in U$, $\oF[m_b]^{\dagger}$
denotes a {\em partial parametrix} for $\oF[m_ b]$, in the sense that
for every $s \in \bf{Z}$, $k \in \bf{Z}^*$,
\be
\label{eqn:paradef}
\bar{\Pi}_D (I-\oF[m_b]\oF[m_b]^{\dagger}) \in {\cal L}(D_s,D_{s-k}).
\ee

\noindent {\bf Remark 7}: In fact, the inequality (\ref{eqn:paradef}) will be required only for 
a finite range of $k$, actually for $k=2$, and can even be concocted
from an operator satisfying (\ref{eqn:paradef}) for $k=1$. This fact
is important, as common algorithms for constructing such partial
parametrices conveniently produce operators satisfying
(\ref{eqn:paradef}) only for $k=1$. %In fact write $\bar{\Pi}_D(I-

\section{Linearized Extended Inversion}
The first order of business is to describe the key model
properties. An example possessing these properties will be described
later.

\begin{itemize}
\item The space of space of p-parameter reference models $M \subset
  (H_{\rm loc}^m(\bR^3))^p,\,m>>3/2$. Such models must have a number of continuous
  derivatives, so $m$ must be quite large.
\item The feasible reference model set $U \subset M$ is defined by
  pointwise bounds.
\item Perturbational extended model spaces $\oM_s \subset (H^s_{\rm
    loc}(\bR^3 \times \bR^2))^p$, $s=0, \pm 1, \pm 2,...$, Hilbert norms $\|\cdot\|_{s} = \|\Lambda^s \cdot\|_0$, where $\Lambda = \sqrt{I-l^2\nabla^2}$ is the usual Sobolev weight operator, extended to act diagonally on vector-valued functions, with nondimensionalizing weight $l$ (units of length). 
\item Corresponding data spaces $D_s \subset \subset (H^s_{\rm
    loc}(\bR^3 \times \bR^2))^d$, $s=0, \pm 1, \pm 2,...$, 
\item Linearized extended forward map = $\oF_{\rm raw} \cap_{s \in
    {\rm Z}} C^1(U , {\cal L}(\oM_s,D_s)$ 
\item Semblance operator (``annihilator'') = $A \in \cap_{s \in
    {\rm Z}} {\cal L}(\oM_s,\oM_s)$;
\item Precompact model and data supports $\Omega_M, \Omega_D$. Pre-
  and post-cutoff versions of $\oF, A$ so that $\oM_s$ can be
  replaced by $(H^s_0(\Omega_M))^p$, $D_s$ by
  $(H^s_0(\Omega_D))^d$. $\Omega_M$ is precompact in $\bR^5$,
  $\Omega_D$ is a precompact smooth manifold of dimension 5 in $R^6$.
\end{itemize}

These abstract the structure of a typical wave inverse problem setting for seismology, with a number of critical assumptions, detailed in the appendix, along with an important example. The examples have a number of crucial additional features:

\noindent {\bf Assumption 1:} For $m \in U$, both $\oF[m_b]^*\oF[m_b]$ and $\oF[m_b]\oF[m_b]^*$ are uniformly bounded on each $\oM_s$ respectively $D_s$ for every $s \in {\bf Z}$, uniformly in $m \in U$.

\noindent {\bf Assumption 2:} $\oF[m_b]^*\oF[m_b] \in OPS^0(\Omega_M)$,
$\oF[m_b]\oF[m_b]^* \in OPS^0(\Omega_D)$. Here $OPS^k$ denotes $p \times
p$, resp. $d \times d$, matrix pseudodifferential operators of order
k. Officially, reference models must be $\in C^{\infty}$ for this to
literally true, however the exponent $m$ in the first bullet above
should be chosen large enough that it makes no
difference. Alternative, restrict $M$ to a finite-dim'l subspace of
smooth models.

\noindent {\bf Remark:} Extended modeling is {\em necessary} in order that $\oF\oF^*$ is $\Psi$DO;

\noindent {\bf Assumption 3 - Uniform Microlocal Ellipticity:}  $\Gamma_D \subset T^*(Y)$ is conic so that $\oF[m_b]\oF[m_b]^*$ is elliptic in $\Gamma_D$ uniformly for $m \in U$.

\noindent {\bf Remark:} Denote by ${\cal C}[m_b]$ the canonical relation of $\oF[m_b]$. According to Egorov's Theorem, $\oF[m_b] P \oF[m_b]^*$ is elliptic in $\Gamma_D$ if $P \in OPS^0$ is elliptic in ${\cal C}[m_b]^{-1}\Gamma_D \equiv \Gamma[m_b]$. Also $\oF[m_b]^*\oF[m_b]$ is elliptic in $\Gamma[m_b]$.

The final two assumptions concern properties of the data, expressed in
terms of homogeneous inequalities defining cones in $D_s$. The results
to follow are {\em aymptotic}, so concern, not a single datum but a family
$\{\de\}_{\lambda \in (0,\lambda_{\rm max}}$ of data. The index
parameter $\lambda$ will turn out to signify (roughly) wavelength.

The first property expresses microlocal concentration of energy.
Pseudodifferential ``projection'' operators $\Pi_D, \bar{\Pi}_D \in
OPS^0$ will appear throughout the following results: these are chosen
to satisfy 
\[
\mbox{ess supp } \bar{\Pi}_D \subset \Gamma_D,
\]
\[
\sigma_0(\bar{\Pi}_D) = 1 \mbox{ on ess supp } \Pi_D
\]

In the following two Assumptions, $k_0>0$ is 
independent of $m \in U$ and $\lambda \in (0,\lambda_{\rm max}]$.

\noindent {\bf Assumption 4 - Data Microlocalization:} There exists $L
\ge 0$ so that for $0 \le s,  k \le k_0, \lambda \in
(0, \lambda_{\rm max}]$,
\be
\label{micro}
\|(I- \Pi_D)\de\|_s \le L_s \|\de\|_{s-k}.
\ee

The last property makes the connection with wavelength, and expresses
the assumption that the data is concentrated at high frequency:

\noindent {\bf Assumption 5 -  High Frequency Cone Condition:} 
For each $s \ge -k_0$ there exists $K_s
\ge 0$ so that for  $\lambda \in
(0, \lambda_{\rm max}]$,
\be
\label{hfcd}
\|\de\|_{s} \le K_s \lambda\|\de\|_{s+1}
\ee

The estimates (\ref{micro}) and (\ref{hfcd}) involve constants independent of the parameter $\lambda \in (0,\lambda_0]$, so it is not immediately obvious that these assumptions define non-empty cones. To build a family $\{\de: \lambda \in (0,\lambda_0]\}$ satisfying estimates of the forms (\ref{micro}) and (\ref{hfcd}). choose $u \in C_0^{\infty}(\bR^3)$ and define $\Phi_{\lambda}(\bx)=(x_1/\lambda,x_2,x_3)$, $u_{\lambda} = \Phi_{\lambda}^*u, \de= \Lambda^{k_0}u_{\lambda}$. Recall the simple anisotropic Poincar\'{e} inequality:
\[
\int u^2 = \int_0^1dx_2 \int_0^1dx_3 \int_0^1dx_1\left(\int_0^{x_1} D_1u\right) ^2
\]
\[
\le \int_0^1dx_2 \int_0^1dx_3 \int_0^1dx_1\left(x_1\int_0^{x_1} (D_1u)^2\right)
\le \int (D_1u)^2
\]
whence
\[ 
\int u_\lambda^2 = \int (\Phi_{\lambda}^{-1})^*u_{\lambda} |det D(\Phi_{\lambda}^{-1})|
 = \lambda \int u^2 \le \lambda \int (D_1u)^2 = \lambda^3 \int (\Phi_{\lambda}^{-1})^*(D_1 u_{\lambda})^2 
\]
\[
= \lambda^2 \int (D_1 u_{\lambda})^2 \le \lambda^2 (u^2 + l^2|\nabla u|^2), 
\]
so
\[
\|u_{\lambda}\|_0 \le \lambda \|D_1 u_{\lambda}\| \le \lambda \|u_{\lambda}\|_1.
\]
Since $D^{\alpha}u_{\lambda}$ is of the same type as $u_{\lambda}$, the same inequality holds, hence for $s \ge 0$  
\[
\|u_{\lambda}\|_s \le C_s \lambda \|u_{\lambda}\|_{s+1}.
\]
Then
\[
\|\de\|_s = \|u_{\lambda}\|_{s+k_0} \le C_{s+k_0}\lambda\|u_{\lambda}\|_{s+k_0+1} = C_{s+k_0}\lambda\|\de|_{s+1}
\]
for $s \ge -k_0$, that is, $\{\de:\lambda > 0\}$ satistfies (\ref{hfcd}) with $K_s = C_{s+k_0}$.

Choose a pseudodifferential partition of unity $\{\Pi_D,I-\Pi_D\}$ with
the scalar matrix symbol of $\Pi$ being
\begin{itemize}
\item independent of $\bx$,
\item homogeneous of order 0 outside a neighborhood of the zero, 
  section, with  values between 0 and 1;
\item $\equiv 1$ in a conic neighborhood of $\{(\bx,{\bf
    \xi}):\xi_2=\xi_3=0\}$, say  $\{(\bx,{\bf \xi}):|\xi_1|\le \beta
  |\xi'|\}$ ($\xi'=(\xi_2,\xi_3)$); 
\item $\equiv 0$ in a conic neighborhood of  $\{(\bx,{\bf
    \xi}):\xi_1=0\}$. 
\end{itemize}
Then for $0\le k \le k_0, s\ge-k_0+k$, 
\[
\|(I-\Pi)\de\|_s^2  = \int d\xi |\hat{u_\lambda}(\xi)|^2\Pi(\xi)^2\Lambda(\xi)^{2(s+k_0)} 
\]
\[
\le \int_{|\xi_1| \le \beta|\xi'|} d\xi |\hat{u_\lambda}(\xi)|^2\Lambda(\xi)^{2(s+k_0)} 
\]
\[
= \lambda^2 \int_{|\xi_1| \le \beta|\xi'|} d\xi
\left|\hat{u}\left(\lambda \xi_1,\xi'\right)\right|^2\Lambda(\xi)^{2(s+k_0)} .
\]
Since $\Lambda(\xi) \le (1+\beta)\Lambda(\xi')$ provided that $|\xi_1|
\le \beta |\xi'|$, this is
\[
\le (1+\beta)^{2k} \lambda^2 \int_{|\xi_1| \le \beta|\xi'|} d\xi
\left|\hat{u}\left(\lambda
    \xi_1,\xi'\right)\right|^2\Lambda(\xi')^{2k}\Lambda(\xi)^{2(s+k_0-k)} 
\]
\[
=(1+\beta)^{2k} \lambda^2 \int_{|\xi_1| \le \beta|\xi'|} d\xi
\left|\widehat{\Lambda(D')^ku}\left(\lambda
    \xi_1,\xi'\right)\right|^2\Lambda(\xi)^{2(s+k_0-k)}  \le (1+\beta^{2k})\|\Lambda(D')^ku_{\lambda}\|_{s+k_0-k}
\]
Since $s+k_0-k \ge 0$, the alternate expression for the norm is
available:
\[
\le C(1+\beta)^2 \sum_{ |\alpha| \le s+k_0-k}
\|D^{\alpha}\Lambda(D')^ku_{\lambda}\|_0^2 
\]
\[
\le \le C(1+\beta)^2 \sum_{
  0 \le \alpha_1 \le s+k_0-k, |\alpha'| \le s+k_0-k-\alpha_1}
\lambda^{1-2\alpha_1}\|D_1^{\alpha_1}(D')^{\alpha'}\Lambda(D')^ku\|_0^2
\]
\[
 \le C_1C(1+\beta)^2 \sum_{
  0 \le \alpha_1 \le s+k_0-k, |\alpha'| \le s+k_0-k-\alpha_1}
\lambda^{1-2\alpha_1}\|D_1^{\alpha_1}(D')^{\alpha'}u\|_0^2
\]
\[
 \le C_1C(1+\beta)^2 \sum_{
  0 \le \alpha_1 \le s+k_0-k, |\alpha'| \le s+k_0-k-\alpha_1}
\|D_1^{\alpha_1}(D')^{\alpha'}u_{\lambda}\|_0^2
\]
\[
\le C_1C(1+\beta)^2 \sum_{ |\alpha| \le s+k_0-k}
\|D^{\alpha}u_{\lambda}\|_0^2 \le C_2 C_1 C (1+\beta)^2
\|u_\lambda\|_{s+k_0-k} \le L \|\de\|_{s-k}
\]
in which 
\[
C_1 = \max_{|\alpha| \le
  s+k_0-k}\frac{\|D^{\alpha}\Lambda(D')^ku\|_0^2}
{\|D^{\alpha}u\|_0^2}
\]

\noindent{\bf Lemma 1:} There exist $C_*,\lambda_0>0$, $\lambda_0 \le \lambda_{\rm max}$ depending on the same quantities as $K,L$ so that for $\lambda \in (0,\lambda_0]$, $m \in U$, $s=-k_0,...,t$, 
\be
\label{lemma1:res}
\|\de\|_s \le C_*\|\oF[m_b]^*\de\|_s.
\ee

\noindent{\bf Proof:} 
The principal symbol of $\oF[m_b]\oF[m_b]^*$ is nonnegative, and positive in $\Gamma_D$:
\[
\sigma_0(\oF[m_b]\oF[m_b]^*) \ge \gamma^2 \mbox{ in } \Gamma_D,
\]
so by a standard construction there exists a microlocal square root $B[m_b] \in OPS^0$ with real nonnegative principal symbol, for which 
\[
\sigma_0(B[m_b]) \ge \gamma \mbox{ in } \Gamma_D,
\]
\[
(\oF[m_b]\oF[m_b]^*-B[m_b]^*B[m_b]) \Pi_D \in OPS^{-2},
\]
whence if $\lambda \in (0,\lambda_{\rm max}]$,
\[
\|\oF[m_b]^*\de\|^2_s = \langle \de, \oF[m_b]\oF[m_b]^* \de \rangle_s =
\|B[m_b]\de\|_s+ \langle \de,(\oF[m_b]\oF[m_b]^*-B[m_b]^*B[m_b]) \Pi_D\de
\rangle_s
\]
\[
+ \langle \de,(\oF[m_b]\oF[m_b]^*-B[m_b]^*B[m_b]) (I-\Pi_D)\de \rangle_s \ge
\|B[m_b]\Pi_D\de\|^2_s  -\|B[m_b](I-\Pi_D)\de\|^2_s  - C\|\de\|^2_{s-1}
\]
\[
\ge \gamma \|\de\|^2_s - C\|\de\|^2_{s-1} \ge (\gamma - C_1\lambda^2)\|\de\|^2_s,
\]
using Lemma A1 (microlocal G\aa rding inequality), $L^2$ estimates, and Assumptions 4 and 5 (with $k=1$). Take $\lambda_0 = \mbox{min }(\lambda_{\rm max}, (\gamma/2C_1))^{1/2})$ (for instance): then for $0 < \lambda \le \lambda_0$,
\be
\label{lemma1:eq5}
\|\oF[m_b]^*\de\|_s \ge \frac{\gamma}{2} \|\de\|_s
\ee

Q. E. D.

From her on replace $\lambda_{\rm max}$ by $\lambda_0$.

\noindent{\bf Lemma 2:} There exist $K_0, L_0 \ge 0$ so that for $0
\le s \le t, 1 \le k \le k_0, \lambda \in (0, \lambda_{\rm max}]$,
\be
\label{lemma2:res1}
\|(I-\Pi[m_b])\oF[m_b]^*\de\|_s \le L_0 \|\oF[m_b]^*\de\|_{s-k}
\ee
\be
\label{lemma2:res2}
\|\oF[m_b]^*\de\|_{s-1} \le K_0\lambda\|\oF[m_b]^*\de\|_{s}
\ee

\noindent{\bf Proof:} 
Egorov's Theorem (REF) implies that $\oF[m_b]^*\Pi_D = \Pi[m_b]\oF[m_b]^*+R[m_b]$, in which $\Pi[m_b] \in OPS^0$ has principal symbol $=1$ in $\Gamma[m_b]$, and $R[m_b] \in OPS^{-\infty}$.  
\[
\|(\Pi[m_b]-I)\oF[m_b]^*\de\|_s = \|\oF[m_b]^*(\Pi_D-I)\de\|_s + \|R[m_b]\|_{s-k,s}\|\de\|_{s-k}  
\]
\[
\le (\|\oF[m_b]^*\|_{s}L_k+\|R[m_b]\|_{s-k,s})\|\de\|_{s-k} \le L_{0,k}\|\oF[m_b]^*\de\|_{s-k},
\]
from (\ref{lemma1:res}), with $L_{0,k} = (\|\oF[m_b]^*\|_{s}L_k+\|R[m_b]\|_{s-k,s})C_*$.

Similarly,
\[
\|\oF[m_b]^*\de\|_{s-1} \le \lambda\|\oF[m_b]^*\|_{s-1}\|\de\|_{s} \le \lambda K_0\|\oF[m_b]^*\de\|_s,
\]
with $K_0 = \|\oF[m_b]^*\|_{s-1}C_*$.

Q. E. D.

\noindent {\bf Lemma A1: Standard Microlocal G\aa rding Estimate:} Suppose that $\Gamma \subset T^*\bR^n$ is conic open, and let $\Pi$, $\bar{\Pi} \in OPS^0$ be such that 
\begin{itemize}
\item[1. ]$\mbox{ess supp} \,\bar{\Pi} \subset \Gamma$;
\item[2. ]$(\bar{\Pi}-I)\Pi \in OPS^{-\infty}$
\end{itemize}
Suppose that $M \in OPS^0$ is microlocally elliptic in $\Gamma$: for certain $\gamma, R>0$,  $\sigma_0(M)(\bx,\xi)  \ge \gamma > 0$ for $(\bx,\xi) \in \Gamma$, $|\xi|>R$.   Then for $u \in H^s(\bR^n)$, any $k \ge 0$, there exists $C_{s,k} \ge 0$ depending on $\Gamma$ and on the symbol semi-norms of  $\bar{\Pi}, \Pi$ and $M$ so that
\be
\label{garding}
\|M\Pi u\|_s \ge \gamma \|\Pi u\|_s - C_{s,k} \|u\|_{s-k}.
\ee

\noindent Proof: Define
\[
\tilde{M} = M \bar{\Pi} + \gamma (I-\bar{\Pi}).
\]
Then $\sigma_0(\tilde{M})(\bx,\xi) \ge \gamma$ so by standard G\aa rding, 
\[
\|\tilde{M}(\Pi u)\|_s \ge \gamma \|\Pi u\|_s - C \|u\|_{s-k},
\]
with $C \ge 0$ as described. Since $(I-\bar{\Pi})\Pi \in OPS^{-\infty}$, the conclusion follows with a different $C$ of same type. Q. E. D.

\noindent {\bf Lemma A2: Standard Microlocal Parametrix Construction:} With the same hypotheses, there exist $M^{\dagger} \in OPS^0$ so that for $s \in {\bf Z}, k \in {\bf Z}_+$, there are $K_s, C_{s,k} \ge 0$ so that for $u \in H^s(\bR^n)$,
\[
\|\Pi (MM^{\dagger}u - u)\|_s \le C_{s,k} \|u\|_{s-k},
\]
\[
\|\Pi (M^{\dagger}Mu - u)\|_s \le C_{s,k} \|u\|_{s-k},
\]
\[
\|M^{\dagger}u\|_s \le \frac{1}{\gamma}\|u\|_s + K_s \|u\|_{s-1}.
\]
The constants may be taken uniform over any bounded subset of $M \in OPS^0$ that are uniformly elliptic in $\Gamma$.

\noindent {\bf Lemma A3:} Same hypotheses as Lemma A1, and $P \in OPS^0$. Suppose $t \ge 0$ and $q \in H^t(\bR^n)$ satisfies
\be
\label{microlocal}
\|\Pi q - q \|_s \le L\|q\|_{s-k}
\ee
for $0 \le s \le t, 1 \le k \le k_0 > 0$. Then there exists $p\in H^t(\bR^n)$ so that
\[
\|Mp - Pq\|_s \le C_{s,k}\|q\|_{s-k} 
\]
for $0 \le s \le t, 1 \le k \le k_0 > 0$, with $C$ depending on the constant in the $\Gamma$-ellipticity estimate of $M$,  semi-norms of $\Pi, \bar{\Pi}, M$, and $P$, and on $L$.

\noindent Proof: Let $M^{\dagger}$ be a $\Gamma$-microlocal parametrix as in Lemma A2, set
\[
p = M^{\dagger}Pq.
\]
Then for $0 \le s \le t$, 
\[
\|Mp-Pq\|_s \le \|\bar{\Pi}(MM^{\dagger}-I)Pq\|_s + \|(I-\bar{\Pi})(MM^{\dagger}-I)P\Pi q\|_s 
\]
\[
+ \|(I-\bar{\Pi})(MM^{\dagger}-I)P(I-\Pi) q\|_s
\]
For $k=1,...,k_0$, the first term is bounded by a multiple of $\|q\|_{s-k}$ as a consequence of the parametrix bound. The operator in the second term is $\in OPS^{-\infty}$, and the third term is also bounded by a multiple of $\|q\|_{s-k}$ due to the microlocal bound on $q$. All of the constants are uniform as indicated. Q. E. D.

\noindent {\bf Lemma A4:} Suppose that $M \in OPS^0$ is positive semidefinite on $H^0$. Then $M_{\lambda} \equiv M + \lambda I$ is invertible on $H^s$ for $s \in Z^+$, and for $C_{s} \ge 0$ independent of $\lambda \in (0,\infty)$,
\be
\label{a4:res0}
\|M_{\lambda}^{-1}q \|_s \le C_s\sum_{j=0}^s \lambda^{-(j+1)} \|q\|_{s-j},\,\,q \in H^s.
\ee
\noindent {\bf Proof:} $M_{\lambda}$ is invertible on $H^0$, as a consequence of the bounds
\[
C \|u\|^2  \ge \langle u, M_{\lambda}u\rangle_0 \ge \lambda \|u\|_0^2
\]
and the Lax-Milgram Theorem. In particular $\|M_{\lambda}^{-1}q\|_0\le \lambda^{-1}\|q\|_0$, that is, the case $s=0$ is established.

For $q \in H^s$, $M_{\lambda}^{-1}q \in H^s$ as a consequence of the existence of a parametrix $\in  OPS^0$. Note that for $u \in H^s$, 
\[
\frac{\lambda}{4}\|u\|_s^2 + \frac{1}{\lambda}\|M_{\lambda}u\|_s^2 \ge \|u\|_s\|M_{\lambda}u\|_s \ge \langle \Lambda^s u, \Lambda^s M_{\lambda} u\rangle_0  
\]
\[
= \langle \Lambda^s u, M_{\lambda} \Lambda^s u\rangle_0 + \langle \Lambda^s u, [\Lambda^s, M_{\lambda}]u\rangle_0 
\ge \lambda \|u\|_s^2 - C\|u\|_s \|u\|_{s-1} \ge \frac{\lambda}{2}\|u\|_s^2-\frac{C^2}{2\lambda}\|u\|_{s-1}^2,
\]
whence
\[
\|M_{\lambda}u\|_s + \frac{C^2}{2}\|u\|_{s-1} \ge \sqrt{2} (\|M_{\lambda}u\|_s^2 + \frac{C^2}{2}\|u\|_{s-1}^2)^{\frac{1}{2}} \ge \frac{\lambda}{\sqrt{2}}\|u\|_s
\]
Take $u=M_{\lambda}^{-1}q$, and suppose that the conclusion holds for $s-1$. Then
\[
\frac{\lambda}{\sqrt{2}}\|M_{\lambda}^{-1}q\|_s \le \|q\|_s + \frac{C^2}{2}\|M_{\lambda}^{-1}q\|_{s-1} 
\]
\[
\le \|q\|_s + \frac{C^2}{2}C_{s-1}\sum_{j=0}^{s-1}\lambda^{-(j+1)}\|q\|_{s-1-j}.
\]
Rearranging and renaming constants completes the induction step. Q. E. D.

\noindent {\bf Lemma A5:} Suppose that $M \in OPS^0$ is microlocally elliptic in $\Gamma$, self-adjoint and positive semidefinite on $H^0$. Suppose $t \ge 0$ and $q \in H^t(\bR^n)$ satisfies (\ref{microlocal}) for $0 \le s \le t, k=1,...,k_0 > 0$. 
Then there exist $C_{s,k} \ge 0$ and $\lambda_0>0$ so that for $\lambda \in (0,\lambda_0]$,
 \be
 \label{a4:res1}
\| M_{\lambda}^{-1} P q \|_s \le C_{s,k} (\|Pq\|_s + \sum_{j=0}^s \lambda^{-(j+1)} \|q\|_{s-j-k}).
\ee
Denote by $M_{\lambda}^{\dagger}$ the $\Gamma$-microlocal parametrix for $M_{\lambda}$ constructed in Lemma A-2, and note that $M_{\lambda}^{\dagger}$ is bounded on each $H^s$ with bound independent of $\lambda$. Then
\be
\label{a4:res2}
\|M_{\lambda}^{-1} P q - M_{\lambda}^{\dagger} Pq\|_s \le C_{s,k} \sum_{j=0}^s \lambda^{-(j+1)} \|q\|_{s-j-k}.
\ee
Finally,
\be
\label{a4:res3}
\|(I-\bar{\Pi})M_{\lambda}^{-1}Pq\|_s \le C_{s,k} \sum_{j=0}^s \lambda^{-(j+1)} \|q\|_{s-j-k}.
\ee
for $0 \le s \le t, k=1,...,k_0 > 0$. 
The constants $C_{s,k}$ depend on $s,k$ and the symbol semi-norms of $M, P, \Pi, \bar{\Pi}$ and on $L$, but are independent of $\lambda \in (0,\lambda_0]$.

\noindent {\bf Proof:} Apply Lemma A3 to $M_{\lambda}$ to construct $p \in H^t(\bR^n)$ so that for $0 \le s \le t$, $0 \le k \le k_0$,
\[
\|M_{\lambda}p-Pq\|_s \le C\|q\|_{s-k}.
\]
Note that $\{M_{\lambda}: \lambda \in (0,\lambda_0]\}$ is bounded in $OPS^0$ and uniformly $\Gamma$-elliptic, hence the constants $C$ from Lemma A3 are also independent of $\lambda$. 
Then
\[
\|M_{\lambda}^{-1}Pq\|_s \le \|p\|_s + \|M_{\lambda}^{-1}(Pq-M_{\lambda}p)\|_s 
\]
\[
\le \|p\|_s + C_s\sum_{j=0}^s \lambda^{-(j+1)}\|Pq-M_{\lambda}p\|_{s-j} 
\]
\[
\le \|M_{\lambda}^{\dagger}Pq\|_s + C_s\sum_{j=0}^s \lambda^{-(j+1)} C_{s-j,k}\|q\|_{s-j-k}
\]
\[
\le \|M^{\dagger}\|\|Pq\|_s + C_{s,k} \sum_{j=0}^s \lambda^{-(j+1)} \|q\|_{s-j-k}
\]
after suitable redefinition of constants.
Also
\[
\|M_{\lambda}(M_{\lambda}^{-1} P q - M_{\lambda}^{\dagger} Pq)\|_s = \|Pq-M_{\lambda}p\|_s
\]
\[
\le C\|q\|_{s-k},
\]
from Lemma A3, whence (\ref{a4:res2}) follows from (\ref{a4:res0}). Finallly, (\ref{a4:res2}) implies
\[
\|(I-\bar{\Pi})M_{\lambda}^{-1}Pq\|_s \le \|M_{\lambda}^{-1} P q - M_{\lambda}^{\dagger} Pq\|_s + \|(I-\bar{\Pi})M_{\lambda}^{\dagger}Pq\|_s
\]
\[
\le C_{s,k}\sum_{j=0}^{s}\lambda^{-(j+1)}\|q\|_{s-j-k} + \|(I-\bar{\Pi})M_{\lambda}^{\dagger}P \Pi q\|_s
\]
\[
+\|(I-\bar{\Pi})M_{\lambda}^{\dagger}P(I-\Pi)q\|_s
\]
The operator inside the second term on the RHS is smoothing, and the third term is $O(\|q\|_{s-k})$ for any $k=1,...,k_0$ by hypothesis, hence bounded by a sum of the given form, in fact by its $j=0$ term. Once again redefining the $C_{s,k}$ yields (\ref{a4:res3}).

Q. E. D.

\noindent {\bf Corollary A6:} Suppose that in addition to the assumptions of Lemma A6, $\{q_{\lambda}: 0<\lambda<\lambda_{\rm max}\} \subset H^t$ satisfies 
\be
\label{hfc}
\|q_{\lambda}\|_{k} \le K \lambda \|q_{\lambda}\|_{k+1}
\ee
for $-k_0 \le k < t$, $K$ independent of $\lambda$, $k_0 \ge 2$. Then there exists
$C >0$ so that for $0 \le s
\le t$, 
\be
\label{main_est}
\|M_{\lambda}^{-1}Pq_{\lambda}\|_s \le C \|q_{\lambda}\|_s.
\ee
and for $1 \le k \le k_0$,
\be
\label{microlocal_est}
\|(I-\bar{\Pi})M_{\lambda}^{-1}Pq_{\lambda}\|_s \le C \lambda\|q\|_{s-k+2}.
\ee
If additionally $P=I$ and $k_0 \ge 2$, then $L_1, K_1>0$ exist so that for $0 \le s \le t, 1 \le k \le \min(k_0,s+1)$,
\be
\label{hfc_sol}
\|M_{\lambda}^{-1}q_{\lambda}\|_{s-1} \le K_1\lambda\|M_{\lambda}^{-1}q\|_{s}
\ee
and
\be
\label{microlocal_sol}
\|(I-\bar{\Pi})M_{\lambda}^{-1}q_{\lambda}\|_s \le L_1\lambda\|M_{\lambda}^{-1}q_\lambda\|_{s}.
\ee
for $s=0,...t$.

\noindent{\bf Proof:} 
The inequalities (\ref{hfc}) imply that 
\[
\|q_{\lambda}\|_{s-j-k} \le K \lambda \|q_{\lambda}\|_{s-(j-1)-k} \le ...\le (K\lambda)^{j} \|q_{\lambda}\|_{s-k}.
\]
So all terms in the sum on the RHS of (\ref{a4:res0}) are bounded by multiples of $\lambda^{-1}\|q_{\lambda}\|_{s-k}$, and add up to
\[
\|M_{\lambda}^{-1}Pq_{\lambda}\|_s \le C (\|Pq_{\lambda}\|_s + \lambda^{-1} \|q_{\lambda}\|_{s-k} )
\]
Since $k \ge 1$, (\ref{hfc}) implies (\ref{main_est}), with the constant $C$ inheriting its properties from standard bounds on $P$, $K$, and the $C_{s,k}$.

Apply the same trick to (\ref{a4:res3}) to obtain 
\[
\|(I-\bar{\Pi})M_{\lambda}^{-1}Pq_{\lambda}\|_s \le C
\lambda^{-1}\|q\|_{s-k}.
\]
Since $s \ge 0$, $s-k \ge -k_0$, so you can apply (\ref{hfc}) twice to obtain (\ref{microlocal_est}).

To show (\ref{hfc_sol}), write
\[
\|M_{\lambda}^{-1} q \|_{s-1} \le \|M_{\lambda}^{-1} q_{\lambda} - M_{\lambda}^{\dagger} q_{\lambda}\|_s + \|M_{\lambda}^{\dagger} q_{\lambda}\|_{s-1}
\]
For $s =0, ..., t$, (\ref{a4:res2}) together with the trick applied twice already using (\ref{hfc}) estimates the first term as
\[
\le  C \lambda^{-1}\|q_{\lambda}\|_{s-k} \le C \lambda \|q_{\lambda}\|_{s}
\]
taking $k=2$, as we may thanks to $k_0 \ge 2$. Lemma A-2 provides a uniform (in $\lambda$) estimate for $\|M_{\lambda}^{\dagger}\|_{s-1,s-1}$ for any $s \in {\bf Z}$: 
\[
\|M_{\lambda}^{\dagger} q\|_{s-1} \le C \|q\|_{s-1} \le C\lambda \|q\|_{s}
\]
Combining these estimates, obtain
\[
\|M_{\lambda}^{-1} q \|_{s-1} \le C \lambda \|q\|_{s} \le C\lambda \|M_{\lambda}^{-1} q \|_{s}.
\]

Finally, for $P=I$ and $k=2$, (\ref{microlocal_est}) takes the form
\[
\|(I-\bar{\Pi})M_{\lambda}^{-1}q_{\lambda}\|_s \le C \lambda\|q\|_{s}
\]
\[
\le C\lambda\|M_{\lambda}\|_{s,s}\|M_{\lambda}^{-1}q_{\lambda}\|_{s}
\]
which gives (\ref{microlocal_sol}) with proper choice of $L_1$.

Q. E. D.

\bibliographystyle{seg}
\bibliography{../bib/masterref}