### 2019-10-19 repro/papers/iva/paper.tex

\title{Gradients of Reduced Objective Functions for Separable Wave
  Inverse Problems}
\date{}
\author{William W. Symes, The Rice Inversion Project}
\lefthead{Symes}
\righthead{Accurate Reduced Gradients}
\maketitle
\parskip 12pt
\begin{abstract}
The derivative of the forward (modeling) map for wave inverse problems
is unbounded in the natural metric of the model and data spaces. This
unboundedness complicates the accurate computation of the reduced
gradient in separable wave inverse problems: iterative estimation of
the linear variables may not lead to convergent computation of the
gradient. If the linear part of the modeling operator possesses a computable
inverse up to a smoothing error, however, and if the data is
high-frequency in a precise sense, then the gradient may be decomposed
into a computable non-iterative part and a convergent remainder. 
The resulting form of the
gradient error control is well-suited to integration with inexact trust-region
optimization methods.

\end{abstract}

\section{Introduction}
Separable (partly linear) modeling operators for wave inverse problems
arise either via Born approximation
\cite[]{KerSy:94,ChaurisNoble:01,MuldertenKroode:01,
  ShenSymes:08,WeibullArntsen:13,BiondiAlmomin:14} or by including
source (right-hand side) parameters as model degrees of freedom
\cite[]{SongSymes:94b,Plessix:00,PrattSymes:02,LuoSava:11,Warner:14,LeeuwenHerrmannWRI:13}.
Extended model spaces add yet more degrees of freedom to the linear
parameters of a separable modeling operator, to permit data fit for a
large set of nonlinear parameter (``velocity model'') values. The
non-extended (``physical'') subspace is the kernel of an
``annihilator''. Minimizing the norm of the annihilator output over
all data-fitting models picks out an approximation to the solution of
the physical least squares inverse problem. A natural approach to this
minimization is via elimination of the linear variables to create a
reduced function of the nonlinear variables
\cite[]{KerSy:94,vanLeeuwenMulder:09,LiRickettAbubakar:13}. If the
nonlinear parameters include velocities, indices of refraction, or
equivalent, then the modeling operator is differentiable with loss of
regularity: a convergent sequence of approximate solutions of the
data-fit problem does not (necessarily) lead to a convergent sequence
of derivative values. Consequently, the computable approximate
solutions of the data-fit problem do not lead to convergent gradient
approximations for the reduced objective, which may reduce the
effectiveness of iterative optimization strategies
\cite[]{HuangSymes:SEG15}.

In the paragraphs to follow, I show that the gradient computation may
be modified to become convergent if three conditions are satisfied:
\begin{itemize}
\item the linear part of the forward modeling operator possesses a
  computable parametrix (inverse modulo compacts, or ``asymptotic
  inverse'') in the form of its adjoint with respect to suitable
  weighted norms, and
\item the inverse problem is embedded in a family of problems indexed
  by a data frequency parameter.
\item the (inevitable) Tihonov regularization parameter is coupled to
  data frequency.
\end{itemize}
Then the reduced gradient 
computation may be modified to become convergent as well.

Reliance on parametrices might seem surprising. However, parametrices
for linearized inverse wave problems have a long history, mostly as a
subtopic of computational geometric optics
\cite[]{CohBlei:77,Beyl:85,BeylBurr:90,VirJinMadLam:92,ArayaSymes:96,Operto:00}. Recently
several authors have described direct (``wave equation'') parametrix
constructions without any ray-geometric computations
\cite[]{ZhangBleistein:05,ZhangSunGray:07,XuZhangTang:11,tenKroode:12,HouSymes:15}. For
inversion of source parameters, even the usual ``exact'' inversion
by Fourier division is actually a parametrix, due to the inevitable
presence of time cut-off.

The existence of the parametrix permits the decomposition of the
gradient into two summands: the first does not involve the solution of
the inner (linear) problem at all, while the second does but is
manifestly continuous in the linear variables. Hence approximation of
the linear part of the solution by a convergent iteration results
in a convergent gradient approximation.

While convergence is assured by the three conditions above, properly
interpreted, control over the actual error in the gradient is
indirect, depending on quantities that are difficult or impossible ot
estimate accurately. However I also show that gradient error control
of the type established here is compatible with so-called inexact
trust-region optimization algorithms \cite[]{HeinkenVicente:01b}, with
(in principle) assured convergence to a stationary point. That is,
control over gradient error must be linked to convergence metrics of
the optimization algorithm.

Some uses of model extension for inverse wave problems have been
criticized for so-called ``gradient artifacts'', that is, computed
gradients that appear to give rise to slow convergence or undesired
model features \cite[]{FeiWilliamson:10,VyasTang:10}. However it now
appears that nothing was wrong with the gradients - instead, the
reduced objective functions of separable least squares problems were
defined with insufficiently precise solution of the linear (inner)
inverse problem \cite[]{KerSy:94,LiuSymesLi:14,ChaurisGP:14}. The present paper
could be viewed as an elaboration upon the latter work.

\section{Operators}
This is a Hilbert space story. 
In fact, the natural objects in this study are {\em Hilbert space scales}, that
is sequences of Hilbert spaces $H = \{ H^s, s \in \bZ\}$, decreasing in
the sense that
$H^{s+1} \subset H^s, s \in \bZ$. Define the sets
\[
H^{\infty} = \cap_{s \in \bZ} H^s, \,H^{-\infty} = \cup_{s \in
  \bZ}H^s.
\]
The inner 
product in $H^s$ will be denoted $\langle \cdot, \cdot \rangle_s$, and
the corresponding norm is
$\| \cdot \|_s$. Usually several scales will be involved in most of
our assertions; the scale to which the inner product or norm belongs
should be clear from context.

An {\em operator} $L$  {\em of order $k \in \bZ$} from
a scale $H_1$ to a scale $H_2$ is linear function from $H_1^{-\infty}$ to
$H_2^{-\infty}$ for which $L|_{H_1^s} \in \sB(H_1^s,H_2^{s-k})$ for
all $ s \in
\bZ$. Denote by $Op^k(H_1,H_2)$ a set of operators of order $k$,
satisfying the additional order properties: for $L_i \in
Op^{k_i}_1(H_1,H_2), i=1,2$,
\begin{equation}
\label{eqn:comm}
L_1 L_2 \in Op^{k_1+k_2}(H_1,H_2),\,[L_1,L_2] \in
Op^{k_1+k_2-1}(H_1,H_2).
\end{equation}
$Inv^k(H_1,H_2) \subset Op^k(H_1,H_2)$ is the subset of invertible
operators of order $k$.

The motivating example of this structure are scales of $L^2$ Sobolev
spaces, and scalar (pseudo-)\-differential operators on them.

The domain of a separable extended model consists of two
components. The first is a (background) model Hilbert space $M_b$ and
an open set of admissible models
$U \subset M_b$. In all examples, $M_b$ is a space of (possibly)
vector-valued) smooth ($C^{\infty}$, ``low frequency'') functions on a suitable physical
domain representing the physical parameters of a wave dynamics model, and members of $U$ obey additional constraints (such as
bounds) required to make the dynamical laws they parametrize
well-posed.

The other component of the model domain is a Hilbert space scale of extended
models $\oM =\{\oM_k, k\in \bZ\}$. 
A distinguished subscale $M = \{M_k, k\in \bZ\}, M_k \subset \oM_k$,
represents the non-extended or physical models. 
The  {\em annihilator} $A \in Op^0(\oM,N)$ has $M$ as its kernel:
\begin{equation}
\label{eqn:ann}
\om \in \oM,\,\,A\om=0 \Leftrightarrow \om \in M.
\end{equation}.
The {\em extension map} $E \in Op^0(M,\oM)$ is injective.

The model range or data space is another Hilbert scale $D$.

The extended modeling operator, or forward map, is a continuous function
\[
\oF: U \rightarrow Op^0(\oM,D)
\]
It is {\em smooth with loss of regularity}: 
\begin{equation}
\label{eqn:nonsmooth1} 
m_b \mapsto \oF[m_b]|_{\oM^k} \in C^p(U,\sB(\oM^k,D^{k-p})),
\end{equation}
but
\begin{equation}
\label{eqn:nonsmooth2} 
m_b \mapsto \oF[m_b]|_{\oM^k} \notin C^p(U,\sB(\oM^k,D^{k-p+1})),
\end{equation}
and in particular
\begin{equation}
\label{eqn:nonsmooth3} 
m_b \mapsto \oF[m_b]|_{\oM^0} \notin C^p(U,\sB(\oM^0,D^0)). 
\end{equation}
However, the normal operator is smooth 
\begin{equation}
\label{eqn:smoothnormal}
F^TF \in C^{\infty}(U,Op^0(\oM,\oM),
\end{equation}
as follows from the
{\em factorization lemma}:
\begin{equation}
\label{eqn:factor}
D\oF[m_b] \delta m_b = F[m_b](Q_1[m_b]\delta m_b)
\end{equation}
in which $Q_1 \in C^{\infty}(U\times M_b,Op^1(\oM,\oM))$ is {\em essentially
  skew-symmetric}:
\begin{equation}
\label{eqn:skew}
Q_1 + Q_1^T = Q_0 \in C^{\infty}(U\times M_b,Op^0(\oM,\oM)).
\end{equation}
Finally, assume that $\oF$ has the {\em Egorov property}: if $L_1 \in
C^{\infty}(U,Op^k(\oM,\oM))$, then there exists $L_2 \in
C^{\infty}(U,Op^k(D,D))$ so that for every $m_b \in U$,
\begin{equation}
\label{eqn:egorov}
\oF[m_b] L_1[m_b] = L_2[m_b] \oF[m_b].
\end{equation}
and vis-versa.

\noindent {\bf Remark:} In the examples, $\oF[m_b]$ is a Fourier
Integral Operator whose canonical relation is a local canonical graph,
and whose symbol and canonical relation depends smoothly on $m_b$.
Both the regularity with loss of a derivative
\cite[]{BlazekStolkSymes:13} and normal regularity of $\oF^T\oF$
follow from these facts: in this case, $\oF^T\oF$ is a
pseudodifferential operator. For examples of the factorization lemma,
see \cite{tenKroode:IPTA14,Symes:IPTA14}. The Egorov
property follows from the usual Egorov's theorem (see for example \cite{Tay:81}).

The physical modeling operator is a similar map
\[
F:U \rightarrow Op^0(M,D)
\]
with the same smoothness with loss of regularity.

The extended and physical modeling operators are related through the
extension map: for each $m_b \in U$,
\[
F[m_b] = \oF[m_b] \circ E.
\]

\section{Objective}
In the notation just introduced, the separable inverse problem of this
paper may be crudely stated as:
\begin{equation}
\label{eqn:basicip}
\mbox{Given }d\in D^0,\mbox{ find }m\in M_b,m_l \in M^0 \mbox{ so 
  that }\\
F[m]m_l \approx d 
\end{equation}
In view of the extension structure, this problem is equivalent to
\begin{equation}
\label{eqn:ip}
\mbox{Given }d\in D^0,\mbox{ find }m\in M_b,\om \in \oM^0 \mbox{ so 
  that }\\
\oF[m]\om \approx d, \,\,A\om \approx 0
\end{equation}
For consistent data (that is, for which ``$\approx$'' can be replaced
with ``$=$'' in either of the above statements, a solution is also a
global minimizer of both
\begin{equation}
\label{eqn:residuals}
\|\oF[m]\om - d\|^2 \mbox{ and } \|A\om\|^2
\end{equation}
The fundamental difficulty of this type of problem lies in the
mapping properties of $m \mapsto F[m] \mbox{ or } \oF[m]$: because of
these maps are differentiable only with loss of regularity (statements
\ref{eqn:nonsmooth1}, \ref{eqn:nonsmooth2}, and \ref{eqn:nonsmooth3}),
the first of the two functions in display \ref{eqn:residuals} (the
``data residual'') is not smooth as a map $M_b \times \oM \rightarrow
\bR$. This lack of smoothness is the ultimate reason that solution of the basic
``full waveform inversion'' problem \ref{eqn:basicip} is still a topic
of active research, rather than a routine tool in exploration and
academic seismology, more than thirty years after it was introduced.

Two objectives such as \ref{eqn:residuals} can be combined in various
ways to produce constrained and unconstrained optimization
problems. Both are quadratic in $\om$, so minimization of either in
this variable seems a tractable problem. Minimizing the second
(``semblance'') function so that the data residual becomes a function
of $m$ only (a so called reduced objective) does not improve matters:
the resulting unconstrained optimization is equivalent to
\ref{eqn:basicip} and has a non-smooth objective.

Remarkably, eliminating $\om$ by minimizing the data residual,  yields a smooth reduced semblance
objective: this conclusion will be justified below.
Generally it cannot
be assumed that $\oF$ is coercive, even if it is
injective. Therefore some form of regularization must be added to the
data residual to guarantee existence of a stable minimizer. It is also
turns out to be essential to
permit more freedom in the metric structure of
domain and range spaces also.

Introduce a data weighting operator $W_d\in C^{\infty}(U,Inv^0(D,D))$.
When restricted to $D^0$, $W_d$ is self-adjoint and positive definite,
hence induces a norm on the order $0$ data space:
\begin{equation}
\label{eqn:wdata}
\langle d_1,d_2 \rangle_d = \langle d_1, W_d d_2\rangle_0 \mbox{ for }
d_1,d_2 \in D^0 
\end{equation}
Similarly, introduce a model weight operator
$W_m\in C^{\infty}(U, Inv^0(\oM,\oM))$ on the model space, self-adjoint and positive definite when restricted to
$\oM^0$, with corresponding norm and inner product
\begin{equation}
\label{eqn:wmodel}
\langle \om_1,\om_2 \rangle_m = \langle \om_1, W_d \om_2\rangle_0 \mbox{ for }
\om_1,\om_2 \in \oM^0 
\end{equation}
Reformulate data residual minimization for the linear 
variables $\om$ as
\begin{equation}
\label{eqn:wredn}
\mbox{min}_{\om}\|\oF[m]\om-d\|^2_d + \lambda^2\|R\om\|_m^2 
\end{equation}
in which $R$ is a bounded and coercive regularization operator, whose
range is some other Hilbert space. The normal equation is
\begin{equation}
\label{eqn:normal}
N_{\lambda}[m]\om = \odF[m]d,\,\,N_{\lambda}[m]
=\odF[m]\oF[m]+\lambda^2 \dR R,
\end{equation}
in which $\odF=W_m^{-1}\oF[m]^TW_d[m]$ is the adjoint of $\oF$ with
respect to the weighted norms and $\oF^T$ is the adjoint with respect
to the scale 0-norms. Similarly $\dR$ is the adjoint of $R$ with respect to
the weighted model-space norm.

Denote by $\bre \in C^{\infty}(U \times D,\oM^0)$ the solution of
\ref{eqn:normal}, and define the reduced objective function 
\begin{equation}
\label{eqn:aredn}
J_{\lambda}[m,d]= \frac{1}{2}\|A\bre[m,d]\|_0^2
\end{equation}
If $J_{\lambda}[m,d]$ attains its obvious lower bound, then $\bre[m,d] \in \oM^0$ and
$(m,\bre[m,d])$ solves the extended inverse problem \ref{eqn:ip}.

\section{Gradient}

{\bf Calculation}
 $W = W_m$, $W_d=I$
\[
\langle u,v\rangle_m = \langle u, W[m]^{-1} v \rangle_0
\]
\[
W[m]S[m]^T S[m] = I + R[m], \, R \in Op^{-1}
\]
also 
\[ 
SWS^T \approx I
\]
\[
S[m]^{\dagger} = W[m]S[m]^T
\]
\[
J[m] = \frac{1}{2}\|AS[m]^{\dagger}d\|_m^2 =
\frac{1}{2}\|AW[m]S[m]^Td\|_m^2
\]
\[
= \frac{1}{2}\langle AW[m]S[m]^Td, W[m]^{-1}AW[m]S^T[m]d\langle_0
\]
\[
DJ[m] = \langle A D(W[m]S[m]^T)d, W[m]^{-1}AW[m]S^T[m]d \rangle_0
- \frac{1}{2}\langle AW[m]S[m]^Td,
W[m]^{-1}DW[m]W[m]^{-1}AW[m]S[m]^Td\rangle_0
\]
\[
=\langle A (DWW^{-1})(WS^T)d, W^{-1}AWS^Td
\rangle_0
+\langle A WDS^Td, W^{-1}AWS^Td \rangle_0
\]
\[
- \frac{1}{2}\langle AWS^Td,W^{-1}DWW^{-1}AWS^Td\rangle_0
\]
Definition of $Q$:
\[
Q \approx WS^TDS, DS = SQ, DS^T = Q^TS^T
\]
so 
\[
DJ[m] =
\langle A (DWW^{-1})(WS^T)d, W^{-1}AWS^Td
\rangle_0
+\langle A WQ^TS^Td, W^{-1}AWS^Td \rangle_0
\]
\[
- \frac{1}{2}\langle AW[m]S[m]^Td,
W[m]^{-1}DW[m]W[m]^{-1}AW[m]S[m]^Td\rangle_0 
\]
\[
=
\langle A (DWW^{-1})(WS^T)d, W^{-1}AWS^Td
\rangle_0
+\langle A WQ^TW^{-1}WS^Td, W^{-1}AWS^Td \rangle_0
\]
\[
- \frac{1}{2}\langle AWS[m]^Td,
W[m]^{-1}DW[m]W[m]^{-1}AW[m]S[m]^Td\rangle_0 
\]
Symmetry of $Q$:
\[
Q + WQ^TW^{-1} = WS^TDS + W(DS^TSW)W^{-1} = W(S^TDS + DS^TS)
\]
\[
= WD(S^TS) = D(WS^TS) - DW W^{-1}WS^TS \approx -DW W^{-1}
\]
\[
QW + WQ^T \approx -DW
\]
So
\[
\langle A WQ^TW^{-1}WS^Td, W^{-1}AWS^Td \rangle_0 =
\langle A WQ^TS^Td, W^{-1}AWS^Td \rangle_0
\]
\[
=-\langle A QWS^Td, W^{-1}AWS^Td \rangle_0 -\langle ADWW^{-1}(WS^Td),
W^{-1}AWS^Td\rangle_0 
\]
\[ 
=-\langle W^{-1}AWA QWS^Td, W^{-1}(WS^Td) \rangle_0 -\langle ADWW^{-1}(WS^Td),W^{-1}AWS^Td\rangle_0
\]
On the other hand,
\[
\langle A WQ^TW^{-1}WS^Td, W^{-1}AWS^Td \rangle_0 =
\langle Q^TS^Td, WA^TW^{-1}AWS^Td \rangle_0
\]
\[
=\langle WS^Td, W^{-1}Q WA^TW^{-1}AWS^Td \rangle_0
\]
Adding left and right hand sides of last two equations and dividing by
2, get
\[
\langle A WQ^TW^{-1}WS^Td, W^{-1}AWS^Td \rangle_0 
\]
\[
= \frac{1}{2}\left[\langle WS^Td, W^{-1}[Q,WA^TW^{-1}A]WS^Td\rangle_0
 - \langle ADWW^{-1}(WS^Td),W^{-1}AWS^Td\rangle_0\right]
\]
Combine with previous calc to get
\[
DJ =
\langle A (DWW^{-1})(WS^Td), W^{-1}AWS^Td \rangle_0
+\frac{1}{2}\left[\langle WS^Td, W^{-1}[Q,WA^TW^{-1}A]WS^Td\rangle_0
 - \langle A(DWW^{-1})(WS^Td),W^{-1}AWS^Td\rangle_0\right]
\]
\[
- \frac{1}{2}\langle W^{-1} AWS^Td,(DWW^{-1})A(WS^Td)\rangle_0 
\]
\[
=\frac{1}{2}\langle WS^Td, W^{-1}[Q,WA^TW^{-1}A]WS^Td\rangle_0
- \frac{1}{2}\langle W^{-1} AWS^Td,[(DWW^{-1}),A](WS^Td)\rangle_0 
\]
Since $[(DWW^{-1}),A] \in Op^{-1}$, the last term is negligible, and
we get
\[
DJ \approx \frac{1}{2}\langle WS^Td,
  W^{-1}[Q,WA^TW^{-1}A]WS^Td\rangle_0
\]
\[
= \frac{1}{2}\langle S^{\dagger}d, [Q,A^{\dagger}A]S^{\dagger}
d\rangle_m
\]

Minimization of $J_{\lambda}$ 
by any variant of Newton's method requires computation of the gradient  
of $J_{\lambda}$. Formally,
\[
DJ_{\lambda}[m,d]\delta m = \langle D\bre[m,d]\delta m,A^TA\bre[m,d]\rangle_0
\]
\[
= \langle -\Ne[m]^{-1}(D\Ne[m]\delta m) \bre[m,d]
+ \Ne[m]^{-1}(D\odF[m]\delta m)d ,A^TA\bre[m,d]\rangle_0
\]
Inserting $W_mW_m^{-1}$, referring to the definition of $\langle
\cdot,\cdot \rangle_m$, and noting that $W_m^{-1}A^T = A^{\dagger}$
and that $\Ne$ is self-adjoint in the model norm, obtain
\begin{equation}
\label{eqn:deriv}
= \langle -(D\Ne[m]\delta m)\bre[m,d] +(D\odF[m]\delta m)d, \bqe[m,d] \rangle_m
\end{equation}
where $\bqe[m,d]$ is the solution of
\begin{equation}
\label{eqn:q}
\Ne[m] \bqe[m,d] = A^{\dagger}A\bre[m].
\end{equation}

The first term in right-hand side of equation \ref{eqn:deriv}  defines
a continuous function of $\bre[m,d]$. 
Since $\Ne = W_m^{-1}\oF^TW_d\oF + \lambda^2 \dR R$, the rule \ref{eqn:egorov} implies
the existence of $P \in C^{\infty}(U,Inv^0(\oM,\oM))$ so that $\Ne =
W_m^{-1}\oF^T\oF P + \lambda^2 \dR R$. This shows in turn that $\Ne \in
C^{\infty}(U,Op^0(\oM,\oM))$, thanks to the assumptions on $W_d, W_m$,
\ref{eqn:egorov} once again, and the \ref{eqn:smoothnormal}. In
particular,
$D\Ne[m]\delta m \in Op^0(\oM,\oM)$, so the bilinear form 
\begin{equation}
\label{eqn:norcont}
\bq, \br \mapsto \langle (D\Ne[m]\delta m)\br,\bq \rangle_m
\end{equation}
is continuous in $\oM^0 \times \oM^0$. Also,
\[
\langle (D\Ne[m]\delta m)\br,\bq\rangle_m=
\langle (D(\odF[m])\delta m)\oF[m]\br,\bq \rangle_m
\]
\[
+\langle(\odF[m] (D\oF[m]\delta m)\br,\bq \rangle _m
= -\langle W_m[m]^{-1}(DW_m[m]\delta m)\odF[m]\oF[m]\br,\bq\rangle_m
\]
\[
+ \langle \odF[m] W_d[m]^{-1}DW_d[m]\oF[m]\br,\bq\rangle_m
+ \langle (D\oF[m]\delta m)^{\dagger}\oF[m]\br,\bq \rangle_m
\]
\[
+ \langle \odF[m](D\oF[m]\delta m)\br,\bq \rangle_m
\]
\[
= -\langle (DW_m[m]\delta m)\odF[m]\oF[m]\br,\bq\rangle_0
+ \langle (DW_d[m]\delta m)\oF[m]\br,\oF[m]q \rangle_0
\]
\begin{equation}
\label{eqn:firstterm}
+ \langle \oF[m]\br,(D\oF[m]\delta m) \bq \rangle_d
+ \langle (D\oF[m]\delta m)\br,\oF[m]\bq \rangle_d
\end{equation}

Concerning the second term in \ref{eqn:deriv}, note that
\[
(D\odF[m]\delta m) = D[(W_m[m])^{-1}\oF[m]^TW_d[m]d](\delta m)
\]
\begin{equation}
\label{eqn:1}
=-W_m[m]^{-1}(DW_m[m]\delta m)\odF[m]d + \odF[m]W_d[m]^{-1}DW_d[m]d 
+ (D\oF[m]\delta m)^{\dagger}d
\end{equation}
The first two terms in equation \ref{eqn:1} involve only bounded
operators on $\oM^0$. The contribution to the second
term in the right hand side of equation \ref{eqn:deriv} is
\[
\langle (D\odF[m]\delta m)d,\bq\rangle_m =
\]
\[
- \langle W_m[m]^{-1}(DW_m[m]\delta m)\odF[m]d, \bq\rangle_m + 
\langle \odF[m]W_d[m]^{-1}DW_d[m]d,\bq\rangle_m
\]
\[
+ \langle (D\oF[m]\delta m)^{\dagger}d,\bq\rangle_m
\]
\[
= -\langle (DW_m[m]\delta m)\delta m)\odF[m]d,\bq\rangle_0 +
+ \langle (DW_d[m]\delta m)d,\oF[m]\bq\rangle_0
\]
\begin{equation}
\label{eqn:secondterm} 
+ \langle d, (D\oF[m]\delta m)\bq \rangle_d
\end{equation}
Now the difficulty is plain to see: if $\bq=\bqe[m,d]$ in the last
term on the right hand side of equation \ref{eqn:secondterm} is replaced with an
approximation in the sense of the $0$ norm, as would be the result of
an iterative process for solving \ref{eqn:q} (or equation
\ref{eqn:normal}), no bound on the resulting error in the inner
product above, hence in the resulting approximation to $DJ_{\lambda}[m,d]\delta
m$ or the gradient of $J_{\lambda}$, can be asserted, since $D\oF[m]\delta m$ is not continuous in the sense of $\oM^0$.

Ignoring for the moment the apparent instability, combine equations
\ref{eqn:firstterm} and \ref{eqn:secondterm} to obtain an expression for the
derivative of $J_{\lambda}[m,d]$:
\[
DJ_{\lambda}[m,d]\delta m = \langle (DW_m[m]\delta m)
\odF[m](\oF[m]\bre[m,d]-d),\bqe[m,d]\rangle_0 
\]
\[
+ \langle (DW_d[m]\delta
m)(\oF[m]\bre[m,d]-d),\oF[m]\bqe[m,d]\rangle_0
\]
\begin{equation}
\label{deriv1}
- \langle F[m]\bre[m,d]-d,(D\oF[m]\delta m)\bqe[m,d]\rangle_d
- \langle (D\oF[m]\delta m)\bre[m,d],F[m]\bqe[m,d]\rangle_d
\end{equation}
There follows an expression for the gradient, using the partial
duals $D\oF[m]^*$, a continuous quadratic form $: D^0 \times \oM^1
\rightarrow M_b$ defined by
\[
\langle \delta m, D\oF[m]^*(d,\br)\rangle_{M_b} = \langle d,
(D\oF[m]\delta m)\br\rangle_d = \langle (D\oF[m]\delta
m)^{\dagger}d,\br\rangle_m
\]
\begin{equation}
\label{eqn:dual}
= \langle W_d[m]^{-1}d,(D\oF[m]\delta m)\br\rangle_0
\end{equation}
and $DW_m[m]^t$, a continuous bilinear map $\oM^0 \times \oM^0
\rightarrow M_b$ defined by 
\[
\langle \delta m, DW_m[m]^t(\br,\bq) \rangle_{M_b} = 
\langle (DW_m[m]\delta m)\br,\bq\rangle_0
\]
$DW_d[m]^t$ is defined similarly.
Then the $m$-gradient of $J_{\lambda}$ is given by
\[
\nabla_m J_{\lambda}[m,d] = DW_m[m]^t(\odF[m](\oF[m]\bre[m,d]-d),\bqe[m,d])
\]
\[
+ DW_d[m]^t(\oF[m]\bre[m,d]-d),\oF[m]\bqe[m,d]
\]
\[
- D\oF[m]^*(\oF[m]\bre[m,d],\bqe[m,d]) -
D\oF[m]^*(\bre[m,d],\oF[m]\bqe[m,d])
\]
\begin{equation}
\label{eqn:gradient}
+D\oF[m]^*(d,\bqe[m,d])
\end{equation}
Every part of this expression is manifestly {\em stable}, that is,  substitutions of
approximations $\brea,\bqea$ in their computation will entail 
$O(\|\bre - \brea\|_0, \|\bqe-\bqea\|_0)$ error, {\em except} for the
last term

\section{Stablility}
In this section, I will show that the last term in \ref{eqn:gradient} can 
expressed as the value at $\br=\bre[m,d]$ and $\bq=\bqe[m,d]$ of 
of continuous function of $d, \br,\bq$, thus giving a stable
computation of the gradient in the sense explained at the end of the
last section.

The construction depends on the availability of a {\em parametrix} or
approximate inverse modulo lower order error, of a particular
form. The utility of the construction depends on the computability of
the parametrix, that is, on its being a straightforward modification of the
transpose. \cite[]{HouSymes:15} demonstrated the existence of such special
parametrices for a particular separable inverse problem (extended
linearized constant density acoustic inversion with horizontal
subsurface offset extension) and parametrices with similar properties
exist for other extended modeling operators as well.

In fact, most parametrices of the type I've mentioned are only
microlocal, and moreover can only be computed approximately. These
limitations will be addressed in coming sections.

Assume that the data and model weight operators $W_d, W_m$ can be
chosen so that $\oF$ is approximately unitary with respect to the
norms $\|\cdot\|_d, \|\cdot\|_m$:
satisfies 
\begin{equation}
\label{eqn:parametrix}
\odF\oF - I \equiv S \in C^{\infty}( U, Op^{-1}(\oM,\oM)). 
\end{equation}
Thus $\Ne = S +
I +\lambda^2\dR R)$. Thus equations \ref{eqn:normal}, \ref{eqn:q} are equivalent to 
\begin{eqnarray}
\label{eqn:magic}
\bre &=& (I+\lambda^2\dR R)^{-1}(\odF d - S\bre)\\
\bqe &=&(I+\lambda^2\dR R)^{-1}(A^{\dagger}A\bre - S\bqe)\\
   & = &(I+\lambda^2\dR R)^{-1}(A^{\dagger}A (I+\lambda^2 \dR R)^{-1}(\odF d -
             S \bre)-(I+\lambda^2\dR R)^{-1}S\bqe
\end{eqnarray}
Replacing $\bqe$ in \ref{eqn:gradient} with the right hand side in the last equality of
\ref{eqn:magic}, obtain
\[
D\oF[m]^*(d,\bqe[m,d])=
D\oF[m]^*(d,(I+\lambda^2\dR R)^{-1}A^{\dagger}A (I+\lambda^2 \dR R)^{-1}\odF[m]d 
\]
\[
- (I+\lambda^2\dR R)^{-1}A^{\dagger}A (I+\lambda^2 \dR R)^{-1}
S[m]\bre[m,d] 
\]
\begin{equation}
\label{eqn:protogradient}
- (I+\lambda^2\dR R)^{-1}S[m]\bqe[m,d]).
\end{equation}
The operators appearing in the last two terms of
\ref{eqn:protogradient} are of
order $-1$, mapping $\oM^0$ continuously to $\oM^1$. So the continuity
property of $D\oF[m]^*$ noted above implies that the last two terms
above are stable, that is, substitution of approximations $\brea$ and
$\bqea$ for $\bre$ and $\bqe$ in the right-hand side of equation
\ref{eqn:protogradient} will result in an
$O(\|\brea-\bre\|_0,\|\bqea-\bqe\|_0)$ error.

It remains to be seen that the first term on the right hand side of
equation \ref{eqn:protogradient} is stable: {\em a priori}, it only
makes sense for $d\in D^1$ (hence $\bre \in \oM^1$). A continuous
extension to $d \in D^0$ follows however from the factorization
property \ref{eqn:factor}. For convenience, define 
\[
B = (1+\lambda^2\dR R)^{-1}A^{\dagger}A (1+\lambda^2 \dR R)^{-1}.
\]
Then the first term on the right-hand side of \ref{eqn:protogradient} is
\[
\langle \delta m, D\oF[m]^*(d,B\odF[m]d) \rangle_{M_b}
\]
\[
= \langle d, (D\oF[m]\delta m)B\odF[m]d \rangle_d 
\]
\[
= \langle d, \oF[m](Q_1[m]\delta m)B\odF[m]d\rangle_d
\]
\begin{equation}
\label{eqn:2}
=\langle \odF[m]d, (Q_1[m]\delta m)B\odF[m]d\rangle_m 
\end{equation}
The essentially-skew property of $Q$ (equation \ref{eqn:skew}) holds
also for the weighted inner product $\langle \cdot, \cdot \rangle_m$,
since the weight operator $W_m$ is invertible and of order 0:
specifically,
\begin{equation}
\label{eqn:skew-m}
(Q_1[m]\delta m)^{\dagger} = -(Q_1[m]\delta m) + (Q_{0,m}[m]\delta m),
\end{equation}
with $Q_{0,m} = Q_0 + W_m^{-1}[Q_1,W_m] \in C^{\infty}(U, Op^0(\oM,\oM))$.
 So the
right-hand side of equation \ref{eqn:2} is
\[
=\langle \odF[m]d, ([B,(Q_1[m]\delta m)] + B(Q_1[m]\delta m))
\odF[m]d\rangle_m 
\]
\begin{equation}
\label{eqn:3}
+\langle \odF[m]d, ([B,(Q_1[m]\delta m)] + ((Q_{0,m}[m]- Q_1[m])\delta m)B)\odF[m]d\rangle_m.
\end{equation}
Add the right hand sides of equations \ref{eqn:2} and \ref{eqn:3} and
divide by 2 to obtain
\[
\langle \delta m, D\oF[m]^*(d,B\odF[m]d) \rangle_{M_b}
\]
\begin{equation}
\label{eqn:4}
=\frac{1}{2}\langle \odF[m]d,([B,(Q_1[m]\delta m)] +
(Q_{0,m}[m]\delta m))\odF[m]d\rangle_m.
\end{equation}
The right-hand side of equation \ref{eqn:4} is a quadratic form in
$\odF[m]d$, defined by a self-adjoint operator (the sum of the various
commutators and products in the above expression) of order
zero. Therefore 
\[
D\oF[m]^*(d,B\odF[m]d)
\]
is a $\|\cdot\|_0$-continuous $M_b$-valued quadratic form in $d$,
hence extends continuously to $d \in D^0$, whence $\nabla_m J_{\lambda}[m,d]$ is
$\|\cdot\|_0$-continuous in $d$.

For Hilbert spaces $H_1, H_2$, denote by ${\cal Q}(H_1,H_2)$ the
Banach space of continuous $H_2$-valued quadratic forms on $H_1$.
Then the result of the foregoing calculations is summarized in

\begin{thm} Define $G_{\lambda} \in C^{\infty}(U, {\cal
  Q}(D^0 \times \oM^0 \times \oM^0, M_b)$ by
\begin{equation}
\label{eqn:grad_approx}
\begin{split}
G_{\lambda}[m](d,\br,\bq) = 
& D\oF[m]^*(d,(1+\lambda^2\dR R)^{-1}A^{\dagger}A (1+\lambda^2 \dR R)^{-1}\odF[m]d \\
& - (1+\lambda^2\dR R)^{-1}A^{\dagger}A (1+\lambda^2 \dR R)^{-1} S[m]\br\\
& - (1+\lambda^2\dR R)^{-1}S[m]\bq]) \\
& + DW_m[m]^t(\odF[m](\oF[m]\br-d),\bq)\\
&+ DW_d[m]^t(\oF[m]\br-d,\oF[m]\bq)\\
&- D\oF[m]^*(\oF[m]\br,\bq) \\
& -D\oF[m]^*(\br,\oF[m]\bq)
\end{split}
\end{equation}
Then
\begin{equation}
\label{eqn:thm1}
\nabla_m J_{\lambda}[m,d] = G_{\lambda}[m](d,\bre[m,d],\bqe[m,d]),
\end{equation}
in which $\bre[m,d]$ and $\bqe[m,d]$ are solutions of equations \ref{eqn:normal} and
\ref{eqn:q} respectively.
\end{thm}

\begin{proof} The content of the definition \ref{eqn:grad_approx} of
  $G_{\lambda}$ is that it is a continuous quadratic-form-valued
  function. This fact has been established for each of the summands:
  for the first term by equation \ref{eqn:4}, for the second and third
  by equation \ref{eqn:protogradient} and following discussion, the
  fourth and fifth by equation \ref{eqn:1}, and the last two by
  equations \ref{eqn:norcont} and \ref{eqn:firstterm} and surrounding
  discussion. These calculations also established equation
  \ref{eqn:thm1}.
\end{proof}

\begin{rem} For computational purposes, it is more convenient to work
directly with the defining Hilbert space structure of the background, domain and
range spaces, rather than with the background, model and data norms, and to
formulate the background model space norm also as a weighted norm. For example,
\[
\langle \delta m, D\oF[m]^*(d,A^{\dagger}A\odF[m]d) \rangle_{M_b}
\]
\[
= \langle d, (D\oF[m]\delta m)A^{\dagger}A\odF[m]d \rangle_d 
\]
\[
= \langle W_d d, (D\oF[m]\delta m)W_m^{-1}A^TA\odF[m]d \rangle_0.
\]
So
\begin{equation}
\label{eqn:thm1b}
\nabla_m J_{\lambda}[m,d] = W_b^{-1}D\oF[m]^t(W_d[m]d,W_m[m]^{-1}A^TA\odF[m]d) +
K[m](\bre,\bqe),
\end{equation}
where $W_b$ is the background model space weight operator defining the
norm $\|\cdot\|_{M_b}$, and $D\oF[m]^t$ is the transpose with respect
to the 0-norms (usually, $L^2$, or in the discrete case Euclidean
length) rather than the $M_b$ and $\|\cdot\|_d$ norms:
\begin{equation}
\label{eqn:dual0}
\langle \delta m, D\oF[m]^t(d,\br)\rangle_0 = \langle d,
(D\oF[m]\delta m)\br\rangle_0.
\end{equation}
Assuming that $\oF$ is implemented by a time-stepping finite
difference or finite element method, this notion of transpose can be
computed by a variant of the adjoint state method. For the case that
$\oF$ is a linearization (that is, Born approximation), so
that $D\oF$ is actually a second derivative, this
application of the adjoint state method was introduced by
\cite{SymSant:88}, and employed by \cite{KerSy:94} in computations
similar to those presented here.
\end{rem}

\section{Controllability}
The stability result of the last two sections is not in itself
sufficient foundation for a convergent optimization
algorithm, for two reasons: 
\begin{enumerate}
\item The solution errors $\brea-\bre$ and $\bqea-\bqe$ are not
directly observable, since neither $\bre$ nor $\bqe$ are known in
practice;
\item typical iterative solution algorithms for
equations \ref{eqn:normal} and \ref{eqn:q} do not necessarily reduce
their solution errors.
\end{enumerate}
Observable quantities actually reduced by iterative algorithms include the
data residual $\|\oF[m]\brea - d\|_d$, the residual for each of the
equations \ref{eqn:normal} and \ref{eqn:q}, and the norms of the
approximate solutions $\brea$ and $\bqea$.

It might be objected that the solution errors and residuals for
equations \ref{eqn:normal} and \ref{eqn:q} are related, the former
being at most $\lambda^{-2}$ times the latter ($\lambda^2$ being a
lower bound for the normal operator $\Ne$). Of course such bounds are
not uniform in $\lambda$, hence are not useful in formulation of an
effective algorithm.

As will be explained in the next section, the necessary form for the
gradient error estimate is
\begin{equation}
\label{eqn:graderr}
\|g - \nabla J_{\lambda}[m,d]\|_{M_b} \le K \epsilon
\end{equation}
in which $K \ge 0$ and $\epsilon$ is a parameter of the estimation process
that produces $\brea,\bqea$, in effect that $\brea,\bqea$
are functions of $m \in U, d\in D^0, e >0$. Only minimal assumptions
will be made about the estimation process - as will be seen below,
only requirements on the residuals produced in equations
\ref{eqn:normal} and \ref{eqn:q}.

However, such estimates appear to
require explicit use of scale separation, abstracted in part by introducing a
{\em family} $d=\{\de: \lambda > 0\} \subset D^0$ of data, and
corresponding separable least squares problems. The {\em high
  frequency cone condition}
\begin{equation}
\label{eqn:hfc}
\|\de\|_k \le C_k\lambda^2 \|\de\|_{k+1},\, k \ge k_0
\end{equation}
gives a quantitative expression of scale separation. The lower index
bound $k_0$ is presumed to be at most $-1$.

\begin{rem} For the Sobolev scale, this condition identifies the
  regularization parameter $\lambda^2$ with wavelength.
\end{rem}

From
here on, $\bre[m,d]$ and $\bqe[m,d]$ denote the solutions of the
parametrized systems
\begin{equation}
\label{eqn:normalbis}
N_{\lambda}[m]\bre[m,d] = \odF[m]\de  
\end{equation} 
and
\begin{equation}
\label{eqn:qbis}
N_{\lambda}[m]\bqe[m,d] = A^{\dagger}A\bre[m,d]. 
\end{equation}

Linking the size of the residuals
\begin{eqnarray}
\label{eqn:mres}
e_{m,\lambda}[m,d] & = & \Ne[m]\brea[m,d]-\odF[m]\de,\\
\label{eqn:qres}
e_{q,\lambda}[m,d] & = & \Ne[m]\bqea[m,d]-A^{\dagger}A\brea[m,d,e] 
\end{eqnarray}
to the solution errors $\brea-\bre$ and $\bqea-\bqe$ requires some
assumption about the approximate solutions $\brea$ and $\bqea$. Note
that any Krylov method for solution of the systems
$\ref{eqn:normalbis}$ and $\ref{eqn:qbis}$, preconditioned by $(I+\lambda^2\dR
R)^{-1}$, yields first iterates
\begin{eqnarray}
\label{eqn:krylov1}
\breone[m,d]& =& (I+\lambda^2\dR R)^{-1}\odF[m]\de\\
\bqeone[m,d]&=&(I+\lambda^2 \dR R)^{-1}A^{\dagger}A\breone[m,d]
\end{eqnarray}
for which the residuals are
\begin{eqnarray}
\label{eqn:mres1}
e_{m,\lambda,1} & = & S[m](I+\lambda^2\dR R)^{-1}\odF[m]\de,\\
e_{q,\lambda,1} & = & S[m](I+\lambda^2\dR R)^{-1}A^{\dagger}A\breone[m,d],
\end{eqnarray}

\begin{thm} Suppose that for $\epsilon > 0$, $\brea[m,d,\epsilon]$ and
  $\bqea[m,d,\epsilon]$ are approximate solutions of \ref{eqn:normalbis} and
  \ref{eqn:qbis} for which the residuals $e_{m,\lambda,\epsilon}$ and
  $e_{q,\lambda,\epsilon}$ satisfy
\begin{eqnarray}
\label{eqn:mresn}
\|e_{m,\lambda,\epsilon}\|_m & \le & \epsilon \|e_{m,\lambda,1}\|_m\\
\|e_{q,\lambda,\epsilon}\|_m & \le & \epsilon \|e_{q,\lambda,1}\|_m 
\end{eqnarray}
Then there exists $K>0$ and $\lambda_0 \ge 0$ so that for $\lambda \le
\lambda_0$,
\begin{eqnarray}
\label{eqn:postbd1}
\|\brea[m,d,\epsilon] -\bre[m,d]\|_m & \le & K \epsilon\|\de\|_d\\
\label{eqn:postbd2}
\|\bqea[m,d,\epsilon] -\bqe[m,d]\|_m & \le & K \epsilon \|\de\|_d
\end{eqnarray}
uniformly in $m \in U$, for fixed choice of $\{C_k\}$ in
\ref{eqn:hfc}.
\end{thm}

\begin{proof}  Since 
\[
\Ne[m] (\brea[m,d,\epsilon] -\bre[m,d]) = e_{m,\lambda,\epsilon},
\]
\[
\|\brea[m,d,\epsilon] -\bre[m,d]\|_0 \le \frac{K}{\lambda^2}\| e_{m,\lambda,\epsilon}\|_0
\le \frac{K\epsilon}{\lambda^2}\| e_{m,\lambda,1}\|_0
\]
\[
\le \frac{K\epsilon}{\lambda^2}\|\de\|_{-1} \le K \epsilon \|\de\|_d
\]
The next-to-the last inequality follows from the definition of
$e_{m,\lambda,1}$ and the uniform $Op^{-1}$ bound on $S[m]$, the last
from \ref{eqn:hfc} for $k=-1$. Thus \ref{eqn:postbd1} is established. Since $e_{q,\lambda,1}$
takes the same form, a similar estimate holds for it.
\end{proof}
\begin{cor}
Under the assumptions of Theorem 2, there exists $K \ge 0$ for which
\begin{eqnarray}
\label{eqn:graderr1}
\left|\frac{1}{2}\|A\brea[m,d,\epsilon]\|^2 - J_{\lambda}[m,d]\right |
  \le K \epsilon\\
\|G_{\lambda}[m](d,\brea[m,d,\epsilon],\bqea[m,d,\epsilon]) - \nabla J_{\lambda}[m,d]\|_{M_b} \le K \epsilon
\end{eqnarray}
 for $m \in U$ and sufficiently small $\lambda$.
\end{cor}

\begin{proof}
Follows from continuity of $G_{\lambda}$ as described in Theorem 1,
and bounds on the solution errors established in Theorem 2. 
\end{proof}

\section{Optimization}

In view of the error inherent in the 
calculations explained in the previous sections, standard variants of
Newton's method for local minimization cannot be guaranteed to
converge to a local minimizer of $J_{\lambda}$. Minimization of the
reduced objective requires use of iterations that converge in the
presence of inexact gradient and objective function evaluations. Such
an algorithm would necessarily need to couple error and step size control.

Since error-contaminated function and gradient computations are hardly
rare, it is unsurprising that this topic has a fairly large
literature
\cite[]{Dembo:82,Carter:91,Deuflhard:91,Carter:93,EisenstatWalker:94}. All
of these works and many more recent ones share a major drawback: they
mandate {\em absolute} control of function and/or gradient error,
implying in particular that the computed gradient be a descent
direction. While of course this condition must eventually hold, it is
nearly impossible to check in practice: many sources of error are like
those considered in the preceding sections, in that they give no
direct measure of gradient or function error, but only allow indirect
control.

One exception to this pattern is the work of \cite{HeinkenVicente:01b,DPKouri_MHeinkenschloss_DRidzal_BGvanBloemenWaanders_2013a,DPKouri_MHeinkenschloss_DRidzal_BGvanBloemenWaanders_2014a}
on error control in the context of {\em trust region} methods for
sequential qudratic programming, an approach to constrained
optimization. Trust region globalization is a general concept that
applies also to unconstrained formulations
\cite[]{ConnGouldToint:00,NocedalWright}. The condition that
\cite{HeinkenVicente:01b} place on gradient error takes the form:
error $\le K \epsilon$, where $\epsilon$ is a control parameter but
$K$ is unknown or at least poorly controlled. Thus nothing can be said
{\em a priori} about the size of the gradient error, or even whether
the computed gradient is a descent direction, except in an asymptotic
sense. Reference to Corollary 1 shows that the theory developed above
provides exactly this sort of condition.

The trust region approach to minimization of a $C^2$ function $f$
defined on a Hilbert space $H$ bases its $k$th step on a quadratic model
approximating $f$ near $x_k$.
\[
m_k(s) = f_k + \langle s,g_k \rangle + \frac{1}{2}\langle s,H_k s
\rangle.
\]
Here $f_k$ is the computed value and $g_k$ the computed gradient at
$x_k$. $H_k$ is an approximation, possibly quite crude, to the Hessian
of $f$ at $x_k$. The basic trust region algorithm seeks the step $s=x_{k+1}-x_k$
as the optimum of the constrained problem
\begin{equation}
\label{eqn:trustregion}
\mbox{minimize }m_k(s) \mbox{ subject to } \|s\| \le \Delta_k,
\end{equation}
The {\em trust radius} $\Delta_k$ is also subject to update as the iteration
proceeds, so as to satisfy a {\em sufficient decrease} criterion
leading to assured global convergence to a local minimizer, under
various conditions (smoothness, Hessian definiteness, etc.) some of
which will be mentioned below.

I will describe a simple variant of the trust radius step, depending
on two computed quantities: actual reduction,
\[
actred = f(x_k) - f(x_k +s)
\]
and predicted reduction, 
\[
predred = m_k(0) - m_k(s) = -(\langle s,g_k \rangle +
\frac{1}{2}\langle s, H_k s \rangle)
\],
and on four magic numbers, $0 < \eta_1 <\eta_2 < 1$ and 
$0 < \gamma_1 <1 < \gamma_2 $.

Choose a {\em search direction} $p$ by
minimizing the model as an unconstrained problem: that is,
\[
p = -H_k^{-1}g_k.
\]
Note that $H_k$ is only an approximation to the Hessian of $f$, and
that in a rather loose sense: it may be computed by building up a
Krylov or quasi-Newton approximation to the Hessian or its
Gauss-Newton simplification, for example. Thus $p$ may be only a crude
approximation to a Newton or Gauss-Newton step: this algorithm
description encompasses so-called truncated Newton-Krylov methods, for
example, and the step selection may interact with the computation of
$H_k$.

Next enter the step update loop:
\begin{enumerate}
\item set $s=\Delta_k p$
\item compute $actred, predred$.
\begin{enumerate}

\item if $actred \le \eta_1 predred$, $\Delta_k \leftarrow \gamma_1 
  \Delta_k$, go to 1.
\item if $actred \ge \eta_2 predred$, $\Delta_k \leftarrow \min(1,\gamma_2 
  \Delta_k)$, 
\end{enumerate}
\item $x_{k+1}=x_k + s$, $\Delta_{k+1} = \Delta_k$, exit.
\end{enumerate}

To see how this algorithm accommodates function and gradient error,
first presume that there is none: that is, $f_k=f(x_k), g_k = \nabla
f(x_k)$, and $H_k \ge \beta > 0$ for all $k$.  The theory presumes that the
function $f$ is twice continuously differentiable, bounded below, and
has uniform bounds on the gradient and condition number of the Hessian
over the sub-level sets of the objective. It is easy to see that the
condition for a successful step implies that
\begin{equation}
\label{eqn:trstep}
f_k - f_{k+1} \ge C \|g_k\|\Delta_k
\end{equation}
references, these assumptions imply a natural floor
under the trust radius $\Delta_k$. The trust radius decreases only when the actual
function 
reduction is not at least the lower proportion $\eta_1$ of the
predicted reduction. However as the trust radius, hence the step, gets
smaller, the actual reduction and the predicted reduction become
close, since the model becomes close to the function to first
order. Once the trust radius is small enough, the conditions for
further decrease are never met, and the trust radius never decreases
beyond a fixed positive thresshold. Since the function values are
decreasing from step to step and bounded below, they must
converge. Therefore the actual reduction converges to zero, hence the
gradient must converge to zero thanks to inequality
\ref{eqn:trstep}. Note that this reasoning does not establish that
$\{x_k\}$ converges - that requires a bit more reasoning. However, if
the sequence of iterates does converge, it must converge to a
stationary point.

In fact the expected behaviour of this algorithm is that it eventually
settles down at $\Delta_k = 1$, takes the full step, and converges at
whatever rate the underlying approximation to Newton's method yields.

Convergence with inexact function and gradient evaluations follows
from an extension of the same reasoning. Assume for the moment that
the function values are exact. Then the primary condition to be imposed on the
gradient, according to \cite{HeinkenVicente:01b}, is of the form
\begin{equation}
\label{eqn:hv}
\|g_k - \nabla f(x_k)\| \le \xi \min(\|g_k\|,\Delta_k).
\end{equation}
in which it is merely required that the constant $\xi \ge 0$ be
uniform over the sublevel set of $f$ in which the iterates lie. Then
once again reduction of $\Delta_k$ during a step eventually stops
short of a uniform positive lower bound, since the inequality
\ref{eqn:hv} forces $g_k$, hence the step $s$, towards the steepest
descent step. For short enough steepest descent steps the
actual reduction exceeds $\eta_1$ times the predicted reduction, which
leads both to the step being taken and in the computed gradient $g_k$
being reduced in length.

I embed this mechanism in an algorithm for minimization of
$J_{\lambda}$ following the pattern of \cite{HeinkenVicente:01b},
section 5. Add $\xi >0$ to the list of magic numbers required by the
trust region algorithm; in theory at least $\xi=1$ is adequate.

Define $(f,g,H) = JET(m,\epsilon)$ via the algorithm

  \begin{enumerate}
    \item compute initial approximate solutions $\br_1,\bq_1$
      \begin{eqnarray}
      \label{eqn:krylov1alg}
      \br_1& =& (I+\lambda^2\dR R)^{-1}\odF[m]\de\\
      \bq_1&=&(I+\lambda^2 \dR R)^{-1}A^{\dagger}A\br_1
      \end{eqnarray}
       with residuals $e_{m,1}, e_{q,1}$;
    \item compute approximate solutions $\br,\bq$ of the system
      \ref{eqn:normalbis},\ref{eqn:qbis} satisfying with residuals
      $e=(e_{m}, e_{q})$ satisfying
      \begin{eqnarray}
      \label{eqn:mresnalg}
      \|e_{m}\|_m & \le & \epsilon \|e_{m,1}\|_m\\
      \|e_{q}\|_m & \le & \epsilon \|e_{q,1}\|_m 
      \end{eqnarray}
    \item compute approximate function value, gradient
      \begin{eqnarray}
      \label{eqn:fgalg}
      f& =& \frac{1}{2}\|A\br\|^2 \\
      g& = & G_{\lambda}[m](d,\br,\bq)
      \end{eqnarray}
      and approximate Hessian $H$
  \end{enumerate}

Formulation of inexact Trust Region algorithm:

\begin{enumerate}
  \item $(f_k,g_k,H_k) =JET(m_k,\epsilon_k)$
  \item while $\epsilon_k > \xi \min(\|g_k\|_m,\Delta_k)$ 
  \begin{enumerate}
    \item $\epsilon_k \leftarrow \xi \min(\|g_k\|_m,\Delta_k)$
    \item $(f_k,g_k,H_k) =JET(m_k,\epsilon_k)$
  \end{enumerate}
  \item $s = -\Delta_k H_k^{-1}g_k, m_{k+1}=m_k+s, \epsilon_{k+1} = \epsilon_k,
    (f_{k+1},g_{k+1},H_{k+1}) = JET(m_{k+1},\epsilon_{k+1})$
    \begin{eqnarray}
    \label{eqn:actpred}
    actred &=&f_k-f_{k+1}\\
    predred &=& \left(1-\frac{\Delta_k}{2}\right)
                \Delta_k\langle g_k,H_k^{-1}g_k\rangle_m
    \end{eqnarray}
  \item if $actred \le \eta_1 predred$, $\Delta_k \leftarrow \gamma_1 
      \Delta_k$, go to 3.
  \item if $actred \ge \eta_2 predred$, $\Delta_k \leftarrow \min(1,\gamma_2 
      \Delta_k)$, 
  \item next k
\end{enumerate}

\begin{cor} The sequence $\{m_k\}$ produced by the
  inexact Trust Region algorithm just described converges to a
  stationary point of $J_{\lambda}$.
\end{cor}

\section{Microlocalization}

With a few exceptions, no examples of forward maps $\oF$ defined by
separable inverse wave problems have actual parametrices, that is,
operators $\odF$ satisfying the definition
\ref{eqn:parametrix}. Instead, they have {\em microlocal}
parametrices. Abstractly, the microlocal property is captured in an
approximate projector $\Pi \in Op^0(\oM,\oM)$, self-adjoint with
respect to $\|\cdot\|_m$ and close to idempotent in
the sense that 
\begin{equation}
\label{eqn:approxidem}
\|\Pi - \Pi^2\|_m \le \frac{1}{4},\,\,0 \le \Pi \le 1.
\end{equation}
A microlocal parametrix $\odF$ satisfies
\begin{equation}
\label{eqn:micropar}
\odF[m]\oF[m] - \Pi = S[m] \in Op^{-1}(\oM,\oM).
\end{equation}
In this section, the approximate projector $\Pi$ is locally constant,
that is, the relation \ref{eqn:micropar} holds in a neighborhood of
$m_0 \in M$.

Augment the linear least squares problem \ref{eqn:wredn} by adding
$\langle \om, (I-\Pi)\om\rangle_m$ to objective. The corresponding
effect on the normal operator is:
\[
\Ne[m] = \odF[m]\oF[m] + (I-\Pi) + \lambda^2 \dR R = I + S[m] +
\lambda^2\dR R.
\]
Define as before $\bre[m,d]$ to be the solution of $\Ne[m]
\bre[m,d] = \odF[m]d$, and the reduced objective $J_{\lambda}$ by
equation \ref{eqn:aredn}.

Computation of $J_{\lambda}$ and $\nabla J_{\lambda}$ goes exactly as
before, and Theorems 1 and 2 and Corollary 1 hold verbatim. The only
change comes in the relation between the minimizer of $J_{\lambda}$
and the solution of the inverse problem $m,\om$: the data misfit is
small only if one assumes that $d$ can be fit with $\om$ nearly
annihilated by $I-\Pi$. That is an {\em a priori} assumption about the
solution, which in examples entails a similar assumption about data. 
% explore meaning in an example

\bibliographystyle{seg}
\bibliography{../../bib/masterref}