### 2019-10-19 repro/papers/lsmabs18/paper.tex

\title{Accelerated Acoustic Least Squares Migration}
\renewcommand{\thefootnote}{\fnsymbol{footnote}} \author{Raanan
  Dafni, William W. Symes\footnotemark[1], The Rice Inversion Project,
  Rice University}
\captionsetup{font=scriptsize,labelfont=scriptsize}
\righthead{Accelerated LSM}
\lefthead{Dafni and Symes}
\maketitle

\begin{abstract}
Least squares migration cost depends on the number of iterations
required to produce a satisfactory image volume and/or fit to data. The
cost can be reduced considerably through effective
preconditioning. One such preconditioner derives from true-amplitude
or asymptotic approximate inversion. We describe implementation details of a
true-amplitude preconditioned least squares migration algorithm, and illustrate
its performance using a synthetic 2D example.
\end{abstract}

\section{Introduction}
Least squares migration (LSM) based on two-way wave equation modeling
produces optimized image volumes - in fact inversions - that correct
many of the basic defects of reverse time migration images
\cite[]{Dutta:SEG14,DaiBS:2010,DaiSchuster2013,Nemeth:99,ChaventPlessix:98}. LSM
accommodates any wave physics that can be modeled in the linearized
(``Born'') sense \cite[]{duan2017elastic,sun:SEG15,DuttaSchuster:14},
and can produce information about the earth's structure well beyond a
mere image. However LSM is necessarily an iterative process, as the
linear least squares problem it defines is far to large to admit
matrix solution methods at any interesting scale. Each iteration
requires a combination of modeling (demigration) and migration. The
computational expense can therefore be considerable.

\cite{HouSymes:EAGE16} have described the use of asymptotic approximate
  inverse formulae (``true amplitude migration'') of an appropriate form, derived by
  \cite{HouSymes:15}, to accelerate the
  convergence of conjugate gradient LSM. Here we base the construction
  of a preconditioner instead on an alternative construction of approximate inversion 
  \cite[]{HouSymes:17,tenKroode:12}, leading to a simpler and equally
  effective algorithm. We explain several crucial 
  implementation details which have not been described elsewhere but
  which are essential for successful convergence
  acceleration, and illustrates the speedup obtainable in the acoustic
  case via a synthetic example.

\section{Method}
In this section we explain our approach to acoustic modeling and RTM, the
structure of the approximate inverse and its key role in successful
preconditioning of the LSM iteration, and several computational
constraints and short-cuts that lead to a practical and efficient
algorithm.

\subsection{Modeling and Migration}
Born acoustic modeling is the solution of the system
\ref{eqn:awe}-\ref{eqn:icwe} of partial differential equations for the
excess pressure and particle velocity fields $p, \bv$ and their
perturbations $\delta p, \delta \bv$ in response to perturbations
$\delta \kappa, \delta \beta$ in the bulk modulus and
buoyancy (reciprocal density) fields $\kappa$, $\beta$:
\begin{equation}
\label{eqn:awe}
\frac{\partial p}{\partial t} = - \kappa \nabla \cdot \bv + f;
\,\frac{\partial \bv}{\partial t} = - \beta \nabla p
\end{equation} 
\begin{equation}
\label{eqn:bwe}
\frac{\partial \delta p}{\partial t} = -\delta  \kappa \nabla \cdot
\bv - \kappa \nabla \cdot \delta \bv ;
\,\frac{\partial \delta \bv}{\partial t} = - \delta \beta \nabla p -
\beta \nabla \delta p
\end{equation} 
\begin{equation}
\label{eqn:icwe}
p, \delta p, \bv, \delta \bv, f =0,\, t<<0
\end{equation}
In addition to the initial conditions \ref{eqn:icwe}, boundary
conditions are also required to make the solution unique. In all of
the examples presented here, we solve the system
\ref{eqn:awe}-\ref{eqn:icwe} in a rectangular subdomain of the 2D
half-space $z \ge 0$, use PML absorbing boundary conditions \cite[]{Habashy:07} on the
lateral sides and bottom, and impose either the pressure-release
condition $p=0$ or a PML condition (both are needed, as will be seen)
on the top surface $z=0$

The source fields $f$ represent isotropic point sources
$f(\bx,t)=w(t)\delta(\bx-\bx_s)$ for a line of source positions
$\bx_s$. The linearized, or Born, traces $\delta d(\bx_s,\bx_r,t)$ are
the samples of $\delta p$ at source-receiver pairs
$\bx_s,\bx_r$. Since these depend on both the background parameters
$\kappa$, $\beta$ and the perturbations $\delta \kappa$, $\delta
\beta$, and are linear in the latter, this relation defines the
absorbing and free surface Born modeling operators $F$ and $F_0$ respectively:
\begin{eqnarray}
\label{eqn:afdef}
\{\delta p(\bx_r,t;\bx_s)\} &=& F[\kappa,\beta][\delta \kappa, \delta \beta]\mbox{
  (abs. surface)}\\
\label{eqn:ffdef}
\{\delta p(\bx_r,t;\bx_s)\} &=& F_0[\kappa,\beta][\delta \kappa, \delta \beta]\mbox{ (free. surface)}
\end{eqnarray}
Since in LSM the background fields $\kappa, \beta$ are not updated, we
drop them from the notation. Also, we will set $\delta \beta = 0$
throughout, for simplicity, thus writing $F \delta \kappa$ for the
output of the Born modeling operator.

 We use the well-known staggered grid
finite difference method, of order 2 in time and 2$k$ in space
($k=$ 4 in the example below) to approximate solutions of the system
\ref{eqn:awe}-\ref{eqn:icwe}.

Reverse-time migration is taken to mean application of the {\em
  transpose} $F^T$ or $F^T_0$ of the Born modeling operator. We use the
well-known adjoint state method to compute the action of these operators
\cite[]{Plessix:06}. We
do not specify the adjoint system of differential
equations  here, as
we do not directly discretize it, but rather create the transpose
mapping of the discrete Born modeling operator as produced by the
finite difference method. This approach guarantees that the computed
$F^T$ (resp. $F_0^T$) is actually the transpose of the discrete $F$
(resp. $F_0$), that is, passes the ``dot
product test'' to machine precision. This discrete adjointness
relation is very important in ensuring that the conjugate gradient
method exhibits its theoretical convergence properties.

Our method for computing $F^T$, $F_0^T$ with these properties relies on the
automatic differentiationn software {\em TAPENADE}
\cite[]{TapenadeRef13} to create finite
difference kernels for the Born fields (equation \ref{eqn:bwe}) and
their corresponding adjoint kernels, and iterating these in an
abstract loop framework adaptable to any wave physics \cite[]{GeoPros:11}. The adjoint state method requires backwards-in-time
reconstruction of the source wavefields $p,\bv$: we use optimal
checkpointing \cite[]{Griewank:00,Symes:06a-pub} to avoid introducing any error
beyond round-off in this step.

We also note that in the theoretical derivation of the adjointness
relation, and the asymptotics calculations in the next section,
$F^T$ and $F_0^T$ involve an {\em integral} over source and
receiver positions. Thus the sum over source and receiver positions
found in naive descriptions of RTM must be replaced by a
quadrature. If the source and receiver spacings are uniform, then a
simple cell-size factor is sufficient - we have taken advantage of
that observation in the synthetic example below.

\subsection{Asymptotic Inversion}
Beginning with \cite{CohBlei:77}, high frequency asymptotics has been
used to construct approximate inverses to Born modeling. All such
constructions require normalization of the energy source. In terms of
the system \ref{eqn:awe}-\ref{eqn:icwe}, the relevant normalization is
the step/Heaviside point source, $f(\bx,t) =
H(t)\delta(\bx-\bx_s)$. In practice this idealized source must be
modified by a bandpass filter. For data supplied with a source pulse
estimate, a shaping filter should be constructed outputting a
bandlimited Heaviside pulse, and applied to the data.

We focus first on the absorbing surface operator
$F$. The results derived in \cite{HouSymes:17} imply that with the source normalized as just
described,
\begin{equation}
\label{eqn:alt1}
F^{\dagger} = V_m^{-1}F^TV_d,\, V_m^{-1}= 32\beta \kappa^3 \partial_z,\, 
V_d = I_t (I_t\partial_{z_s})(I_t\partial_{z_r}) 
\end{equation}
defines an asymptotic (in frequency) inverse $F^{\dagger}$ for $F$,
that is,
\begin{equation}
\label{eqn:appinv}
F^{\dagger}F \approx I.
\end{equation}
In equation
\ref{eqn:alt1}, $I_t$ represents the indefinite $t$ integral, and
$\partial$ signifies the partial derivative with respect to the
variable in the subscript.

Actually, the result of \cite{HouSymes:17} pertains to the {\em
  subsurface offset} extension of Born modeling, not to the operators
discussed here which do not require extended modeling calculations. In
fact equation \ref{eqn:alt1}, with the meaning of the various symbols
as defined here, expresses the {\em zero (subsurface) offset} section
of the inversion defined by \cite{HouSymes:17}, possible because the
operators $V_m^{-1}$ and $V_d$ operate trace-by-trace so can be
restricted to the zero subsurface offset section. In the extended
modeling context, the analogs of $F^{\dagger},F$ are approximately
inverse to each other, as asserted in equation \ref{eqn:appinv}. The
zero-offset section of the extended approximate inverse, namely
$F^{\dagger}$ as defined here, cannot be exactly inverse to $F$, as
the energy in the extended inverse is necessarily spread over at least
a wavelength in the offset direction. Recovery of the bulk modulus
perturbation from the extended image volume requires a {\em stack},
approximating the integral over subsurface offset, discretely
the sum multiplied by the offset step (which must be the same as the
horizontal space step $\Delta x$). Thus the zero offset section gives
an image smaller by roughly the number of gridpoints per wavelength,
multiplied by $\Delta x$, than would be the case for an approximate
inverse. We introduce a factor of $\Delta x$ in the approximate
inverse output, so that the remaining scale difference is roughly the
number of gridpoints per wavelength - a less well-defined quantity. As
shall be shown below this residual scale error is absorbed in the LSM
process.

Two amendments are necessary to render $F^{\dagger}$ as defined above
a useful accelerator for LSM.
$F^{\dagger}$ would be the adjoint of $F$ with respect to weighted inner
products with Gram operators $V_m$ and $V_d$, if it were the case that these have
the necessary properties of Gram operators, namely symmetry and
positivity. Neither $V_m$ (or $V_m^{-1}$) nor $V_d$, as defined above,
have either property. We can however alter the definitions to recover
these properties. Note that $\partial_t = H |\partial_t|$, in which
$H$ is the Hilbert transform and $|\partial_t|$ is represented in the
Fourier domain by multiplication with
the absolute value of frequency. A well-known result from analysis
(Egorov's Theorem, \cite{Tay:81}) allows us to pass $H$ from one
side of $F^T$ to the other in equation \ref{eqn:alt1}, where it is
``absorbed'' by its inverse $-H$ coming from the first $I_t$. The
upshot is an alternate expression 
\begin{eqnarray}
\label{eqn:alt2}
F^{\dagger} & = & W_m^{-1}F^TW_d, \nonumber \\
 W_m^{-1}&=&32\beta^{1/2} \kappa^{3/2} K_z \beta^{1/2}
             \kappa^{3/2},\nonumber \\
W_d &=& -K_t^{-3} \partial_{z_s}\partial_{z_r}.
\end{eqnarray}
Here $K_z$ is defined in the frequency domain by multiplication with
$|k_z|$, supplemented with an appropriate bandpass filter, and $K_t$
is defined similarly. $W_m^{-1}$ is manifestly symmetric and positive.

Concerning $W_d$, \cite{HouSymes:15} demonstrated
that $-\partial_{z_s}\partial_{z_r}$ is symmetric and
positive if the source wavefield is downgoing near the surface, and
the receiver wavefield is upcoming. In general guaranteeing this
property requires decomposition of the fields into up- and down-going
components. However, when sources and receivers are embedded in a
reflector-free layer (simulating the ocean) then these properties
necessarilly hold. Therefore it is only necessary in that setting to
apply a taper operator on the input side of $F$ (and the output side
of $F^T$) to make sure that no artificial reflectors
appear in the water column. Note that if this step is left
out, reflectors inevitably develop in the water column, $W_d$ loses
its positivity, and the algorithm defined in the next section fails to
converge.

We have idenfitied several methods to compute the $z_s, z_r$ derivatives in the definition of $W_d$. If the $z$ component of motion is measured at the receiver (eg. OBN or
Geostreamer$^{TM}$ data), then the second equation of motion
\ref{eqn:bwe} supplies $\partial_{z}\delta p|_{\bx=\bx_r}
= \partial_{z_r}F$. In general, if the source and receiver wavefields
are separated into up- and down-going components, then each obeys a
one-way wave equation, which specifies the $z$ derivative in terms of
the $t$ derivative and other spatial derivatives.

In the example reported below, we have used yet another method,
mentioned briefly by \cite{HouSymes:15}. We
presume that the source and receiver depths for the absorbing surface
fields are both $=0$. The corresponding free surface ($z=0$) acquisition geometry satisfies the {\em shallow tow depth}
assumption if
\begin{equation}
\label{eqn:std}
\partial_{z_s}\partial_{z_r} \delta p_{z_s = z_r=0} \approx
\frac{1}{4z_sz_r}Z\delta p_0
\end{equation}
In equation \ref{eqn:std}, $\delta p_0$ is the free surface Born field at prescribed $z_s, z_r$, and $Z$ sets the source and receiver depths in
each free surface trace to zero and leaves all other header words and
data samples unchanged. The level of error implicit in the $\approx$
symbol should be on the order of the discretization error in the finite
difference scheme, which is the case if $z_s$ and $z_r$ are comparable
to the grid step in $z$. If the condition \ref{eqn:std} holds, then 
\begin{equation}
\label{eqn:stdop}
\partial_{z_s}\partial_{z_r} F \approx \frac{1}{4z_sz_r}Z F_0.
\end{equation}
Therefore the definition \ref{eqn:alt2} translates to
\begin{equation}
\label{eqn:appinvab}
F^{\dagger} \approx - W_m^{-1} \frac{1}{4z_sz_r}F_0^TZ^TK_t^{-3}.
\end{equation}
That is, we have used the ghosted traces to emulate the $z_s,z_r$
derivatives. Such emulation is only valid up to the first ghost notch,
however that is precisely the frequency range for which the criterion
mentioned above assures accuracy of the finite difference simulation.

The examples below employ the free surface condition. Similar logic
leads to the approximate inverse formula
\begin{equation}
\label{eqn:appinvfr}
F_0^{\dagger} \approx - W_m^{-1} \frac{1}{4z_sz_r}F^TK_t^{-3}Z.
\end{equation}

Note that the absorbing surface RTM operator appears in the expression
for the free surface true amplitude migration, and vis-versa.

\subsection{Preconditioned Conjugate Gradient Iteration}
Use of the inner products with weight operators $W_m$ and $W_d$ in
model and data spaces and the transpose $F^{\dagger}$ with respect to
these inner products in conjugate gradient iteration produces a rapidly convergent algorithm, since
with respect to the weighted inner products $F$ is approximately
unitary (up to a scale factor). The computation of the weighted inner products is
tedious. Fortunately, the preconditioned conjugate gradient method
\cite[]{Golub:2012} is
mathematically equivalent, and uses only the Euclidean inner
products. Define $F_0^*$ to be the transpose of $F_0$ with respect to the
Euclidean inner product in model space, and the $W_d$-weighted inner
product in data space:
\begin{equation}
\label{eqn:uwadj}
F_0^* = F_0^T W_d = - \frac{1}{4z_sz_r}F^TK_t^{-3}Z
\end{equation}
Use $F_0$, $F_0^*$ to define the least squares problem and and
$W_m^{-1}$ as the preconditioner. The iteration converges to the minimizer of the
data-weighted mean square error
\begin{equation}
\label{eqn:obj}
J[\delta \kappa] = \frac{1}{2}(F_0\delta \kappa -\delta d)^T W_d
(F_0\delta \kappa -\delta d).
\end{equation}

\section{Numerical Example}
We present a 2D synthetic numerical illustration based on the Marmousi
blind test model \cite[]{BoLaVe:91}. We take $\beta = 1, \delta \beta = 0$
throughout. The smoothed background bulk modulus $\kappa$ appears in
Figure \ref{fig:bulk}, and the target bulk modulus perturbation in
Figure \ref{fig:dbulk}.  We discretized the model on a 12 m $\times$
12 m grid, and used a [2.5, 5, 20, 25] Hz zero-phase bandpass filter, integrated
once to yield a bandlimited Heaviside function, as the source in both
modeling and migration. 180 source positions at $z_s=12 m$ and 360
receiver positions at $z_r=12 m$ were spread across the top of the
model (free surface, full aperture, fixed spread). 
%A typical shot gather appears as
%Figure \ref{fig:shot151}.

LSM via 20 iterations of (unpreconditioned) conjugate gradient
iteration produces the estimated bulk modulus perturbation in Figure
\ref{fig:lsm}, plotted on the same color scale as Figure
\ref{fig:dbulk}. While relative amplitudes above 2500 m depth are
reasonably well recovered, the deeper part of the model is not, and
the data residual is reduced only to 40\% of its initial value. 
Application of the approximate inverse yields the estimate in Figure
\ref{fig:appinv}, with relative amplitudes that are actually more
accurate (for the price of 1 RTM) than was achieved with 20 LSM
iterations. The preconditioned LSM achieves the result displayed in
Figure \ref{fig:plsm} after 20 iterations. The recovered approximation
of the bulk modulus perturbation is roughly as accurate as the data
bandwidth allows, and the data residual is under 10\%. The convergence
curves in Figure \ref{fig:conv} show clearly the cost reduction
achieved through true amplitude preconditioning.

\section{Discussion}

%\cite{HouSymes:SEG16b} show that the same technique used here to
%accelerate LSM may be used to accelerate Full Waveform
%Inversion (FWI). Acceleration does not mean avoidance of cycle-skipping: if
%basic FWI does not converge to a useful result, the accelerated
%algorithm simply achieves the useless result faster. However the
%improvement in convergence is as dramatic as shown here for LSM, and
%could conceivably be combined with other techniques that render FWI
%more robust.

The preconditioning approach presented here is not the
only one suggested for acceleration of LSM iteration
\cite[]{Aoki:09,Berkhout:09d,HuangNammourSymes:16,RochaSavaGuitton:17,Guitton:17}. A
very limited comparison suggests that true-amplitude preconditioning
may be more effective than some of the other techniques discussed in
these references, when it applies. However, as described here, it is
limited to recovery of a single parameter (we have used bulk
modulus). True amplitude recovery of multiple parameters is possible,
but in principle requires the use of extended modeling and migration
\cite[]{YuZhang:14}. On the other hand, analogous results should be
possible for various forms of
pseudo-acoustic anisotropy.

\section{Conclusion}
We have presented details of a true-amplitude acceleration of LSM. The
accelerated iteration has essentially the same cost per step as
ordinary LSM, but converges substantially more rapidly. We have
presented the 2D version here, corresponding to our example, but
similar acceleration formulae are available for 3D, at the same low cost.

\section{ACKNOWLEDGMENTS}
We are grateful to the sponsors of The Rice Inversion Project (TRIP)
for their long-term support, to Shell International Exploration
and Production Inc for its support of Raanan Dafni's postdoctoral work
and  Jie Hou's PhD studies in TRIP, and to Fons ten Kroode for inspiring our
work on this topic. We acknowledge the Texas Advanced Computing
Center (TACC) and the Rice University Research Computing Support Group
(RCSG) for providing essential HPC resources.

\plot{bulk}{width=0.45\textwidth}{Marmousi bulk modulus sampled to 12
  m $\times$ 12 m grid, smoothed with 4 iterations of a 3 $\times$ 3
  rectangular moving average.}

\plot{dbulk}{width=0.45\textwidth}{Marmousi bulk modulus perturbation
  = difference between original bulk modulus and slightly milder
  smoothing than that shown in Figure \ref{fig:bulk}.}

\plot{lsm}{width=0.45\textwidth}{LSM from data described in text, via
  20 conjugate gradient iterations. Note that color scale
  is same as in Figure \ref{fig:dbulk}.}

\plot{appinv}{width=0.45\textwidth}{True amplitude migration of
  Marmousi data described in text. Note that relative amplitudes are
  quite similar to target bulk modulus perturbation (Figure
  \ref{fig:dbulk}), however overall amplitude is smaller by a nearly
  position independent factor.}

\plot{plsm}{width=0.45\textwidth}{LSM from data described in text, via
  20 preconditioned conjugate gradient iterations. Note that color
  scale is same as that in Figure \ref{fig:dbulk}.}

\plot{conv}{width=0.45\textwidth}{Mean square data error as function
  of iteration, with (blue) and without (red) true amplitude preconditioning.}
\bibliographystyle{seg}
\bibliography{../../bib/masterref}