### 2019-10-19 repro/papers/ms/f.tex

\title{Figures}
\author{William W. Symes}

\begin{abstract}
abstract goes here
\end{abstract}

\inputdir{project}

\plot{res0}{width=\textwidth}{res0}
\plot{res1}{width=\textwidth}{res1}
\plot{res2}{width=\textwidth}{res2}
\plot{res3}{width=\textwidth}{res3}
\plot{res4}{width=\textwidth}{res4}
\plot{res5}{width=\textwidth}{res5}
\plot{res6}{width=\textwidth}{res6}
\plot{res7}{width=\textwidth}{res7}
\plot{res8}{width=\textwidth}{res8}
\plot{res9}{width=\textwidth}{res9}
\plot{res10}{width=\textwidth}{res10}

\plot{shot6kmbigdecon0tscale}{width=\textwidth}{shot6kmbigdecon0tscale}
\plot{shot6kmbigdecon1tscale}{width=\textwidth}{shot6kmbigdecon1tscale}
\plot{shot6kmbigdecon2tscale}{width=\textwidth}{shot6kmbigdecon2tscale}
\plot{shot6kmbigdecon3tscale}{width=\textwidth}{shot6kmbigdecon3tscale}
\plot{shot6kmbigdecon4tscale}{width=\textwidth}{shot6kmbigdecon4tscale}
\plot{shot6kmbigdecon5tscale}{width=\textwidth}{shot6kmbigdecon5tscale}
\plot{shot6kmbigdecon6tscale}{width=\textwidth}{shot6kmbigdecon6tscale}
\plot{shot6kmbigdecon7tscale}{width=\textwidth}{shot6kmbigdecon7tscale}
\plot{shot6kmbigdecon8tscale}{width=\textwidth}{shot6kmbigdecon8tscale}
\plot{shot6kmbigdecon9tscale}{width=\textwidth}{shot6kmbigdecon9tscale}
\plot{shot6kmbigdecon10tscale}{width=\textwidth}{shot6kmbigdecon10tscale}

\plot{shot6kmbig}{width=\textwidth}{shot6kmbig}
\plot{shot6kmhom}{width=\textwidth}{shot6kmhom}
\plot{fwiscanlo}{width=\textwidth}{fwiscanlo}
\plot{ivascan}{width=\textwidth}{ivascan}
\plot{waveletbase}{width=\textwidth}{waveletbase}