### 2019-10-19 repro/papers/ms/q.tex

\title{Computing Q}
\author{William W. Symes}

\begin{abstract}
abstract goes here
\end{abstract}

Basic physics: acoustics
\be
\label{awe}
\frac{\partial^2 u}{\partial t^2} - m \nabla^2 u = S
\ee

Forward map: $F(m) = Pu$, $P$ = trace operator on receiver surface $\Sigma$ (sampling).

Source extension: $\oF(m,\oS) = Pu$ with $S$ in (\ref{awe}) replaced by $\oS$.

Derivative: $D\oF(m)\delta m = P \delta u$ where
\be
\label{dawe}
\frac{\partial^2 \delta u}{\partial t^2} - m \nabla^2 \delta u = \delta m \nabla^2 u
\ee

Duhamel's principle:
\[
\delta u(x,z,t) = \int_0^t ds v(x,z,t,s)
\]
where
 \begin{eqnarray}
\label{vawe}
\frac{\partial^2 v}{\partial t^2} - m \nabla^2 v &=& 0,\,\,t \ge s; \nonumber\\
v(x,z,s,s)&=&0; \nonumber\\
\frac{\partial v}{\partial t}(x,z,s,s)&=& (\delta m \nabla^2 u)(x,z,s). 
\ee

\noindent {\bf Assumption 1:} 
\be
\label{supp}
\mbox{supp }\delta m \cap \mbox{supp }\oS = \emptyset
\ee

Then can rewrite 3rd condition in (\ref{vawe}) as 
\be
\label{vawe3}
\frac{\partial v}{\partial t}(x,z,s,s)&=& \left(\frac{\delta m}{m} \frac{\partial^2 u}{\partial t^2}\right)(x,z,s) 
\ee

Set 
\[
w(x,z,t,s) = \frac{1}{2}\left( \frac{\partial u}{\partial t}(x,z,t) - \frac{\partial u}{\partial t}(x,z,2s-t)\right)
\]

Then in complement of $\mbox{supp }\oS$, $w$ solves (\ref{vawe}), provided $\delta m/m \equiv 1$.

\noindent {\bf Assumption 2:} $u$ is supported near $t \mapsto (X(t),Z(t))$, and there exists exactly one $t_r$ so that $(X(t_r),Z(t_r)) \in \Sigma$.

\noindent {\bf Assumption 3:} for each $t$, the support of $u$ is small enough that $\delta m$ may be approximated by a constant on $\mbox{supp }u(\cdot,\cdot,t)$.

Claim: 
\[
v(x,z,t,s) \approx  \frac{\delta m}{m}(X(s),Z(s)) w(x,z,t,s)
\]
so for $t near t_r$,
\[
\delta u(x,z,t) = \frac{1}{2}\frac{\partial u}{\partial t}(x,z,t)\int_0^t ds  \frac{\delta m}{m}(X(s),Z(s)) 
\]
\be
\label{a1}
+ \frac{1}{2} \int_0^t ds  \frac{\delta m}{m}(X(s),Z(s))\frac{\partial u}{\partial t}(x,z,2s-t)
\ee
The second term integrates $u$ {\em backwards} along the trajectory $t \mapsto (X(t),Z(t))$, so if the trajectory intersects $\Sigma$ at $t=t_r$, this involves only times $<< t_r$, hence does not contribute to the trace. Upshot:
under assumptions 1 - 3,
\be
\label{a2}
\oF(m,\oS) \approx \frac{1}{2}\frac{\partial }{\partial t}\oF(m,\oS)(x,z,t)\int_0^t ds  \frac{\delta m}{m}(X(s),Z(s)) 
\ee

Application: transmission configuration, construct $\oS$ as integral over localized oscillatory data generating solutions remaining close to rays. Read off symbol of Q.

Note that this will be correct even when caustics are present, by Huyghens construction.