### 2019-10-19 repro/papers/ms/r.tex

\title{Volume/Surface Source Equivalence}
\author{William W. Symes}

\begin{abstract}
abstract goes here
\end{abstract}

\section{Transmission Problem}

Model problem: $u: \bR^{d+1} \rightarrow \bR$
\bea
\label{awe}
\frac{\partial^2 u}{\partial t^2} - m \nabla^2 u &=& f(t)
\delta(\bx-\bx_s)\\
u&=&0, t<0
\eea

$\Omega \subset \bR^d$, $\bx_s \in \partial \Omega$,
\[
\partial \Omega = \Gamma_R \cup \Gamma_D
\]
Forward map: $F(m)f = Pu$, $P$ = trace operator on receiver
surface $\Gamma_R \times [0,T]$

Volume source extension:
\[
\mbox{supp }\bar{f} \subset \bar\Omega \times [0,\infty)
\]
\bea
\label{extawe}
\frac{\partial^2 \bar{u}}{\partial t^2} - m \nabla^2 \bar{u} & = &
\bar{f}(\bx_s,\bx,t)\\
\bar{u}&=&0, t<0
\eea
Extended forward map: $\bar{F}(m)\bar{f} = P\bar{u}$, $P$ = trace operator on receiver
surface $\Gamma_R \times [0,T]$

Write $\bar{u} = u_0 + u_1,$, with $u_0 = 0$ in $(\bR^d \setminus
\bar{\Omega})\times \bR$, 
\bea
\label{awe0}
\frac{\partial^2 u_0}{\partial t^2} - m \nabla^2 u_0 &=& \bar{f}
\mbox{ in }\Omega
\times \bR,\\
u_0 & = & 0 \mbox{ on } \partial \Omega\\
u_0 & = & 0, t < 0
\eea
and 
\bea
\label{awe1}
\frac{\partial^2 u_1}{\partial t^2} - m \nabla^2 u_1 &=& 0 \mbox{ in }\Omega
\times \bR,\\
u_1 & = & \bar{u} \mbox{ on } \partial \Omega\\
u_1 & = & 0, t < 0
\eea
Note that $\bar{u}=u_1$ in $(\bR^d \setminus
\bar{\Omega})\times \bR$ . Standard uniqueness for WE $\Rightarrow \bar{u}=u_0+u_1 $ in $\bR^d \times \bR$.

SInce $u_0$ satisfies homogeneous Dirichlet condition on all of
$\partial \Omega$, $\bar{F}(m)\bar{f} = Pu_1. $

From the definition of  that in a sufficiently small neighborhood of any point of
$\partial \Omega \times \bR$, 
\bea
\label{awe2}
 \frac{\partial^2 u_1}{\partial t^2} - m \nabla^2 u_1 &=&[\partial_n
 u_1] \delta_{\partial \Omega \times \bR}
\eea
in which $[\cdot]$ is the trace of the jump on $\partial \Omega$ and
$\delta_{\partial \Omega \times \bR}$ is the delta distribution on the
space-time boundary. In view
of (\ref{awe1}) and $\bar{u}=u_1$ outside of $\Omega \times \bR$, in
fact $u_1$ solves (\ref{awe2}) globally, and $u_1=0$ for $t<0$, which
determines $u_1$ uniquely in terms of $G[m,\bar{f}] = [\partial_n
u_1]|_{\partial \Omega \times \bR}$.

For $g \in H^{-1/2}(\partial \Omega \times \bR), g=0$ for $t<0$,
define the boundary trace extension $\tilde{F}(m)g = Pw$ for the
solution $w$ of 
\bea
\label{awe3}
 \frac{\partial^2 w}{\partial t^2} - m \nabla^2 w &=& g
 \delta_{\partial \Omega \times \bR},\\
w&=&0,t<0
\eea

The foregoing shows that
\[
\tilde{F}[m]G[m,\bar{f}] = \bar{F}[m]\bar{f}.
\]
That is, the boundary trace extension is equivalent to the volume
source extension, in the sense that the two have the same
range. Since $\bx_s \in \partial \Omega$, a suitable annihilator for
the boundary trace extended IP is still $A =$ multiply by $|\bx-\bx_s|$.