### 2019-10-20 repro/papers/varden/paper.tex
added repro/papers/varden

\title{Asymptotic Inversion of the Variable Density Acoustic
  Model\\Part 2: Implementation and Examples}

\author{
Raanan Dafni\thanks{Department of Computational and Applied Mathematics, Rice University, 
Houston, TX, 77005, USA, 
{\tt symes@rice.edu}},
William Symes\thanks{Department of Computational and Applied Mathematics, Rice University, 
Houston, TX, 77005, USA, 
{\tt symes@rice.edu}}
}

\lefthead{Dafni Symes}

\righthead{Asymptotic Acoustic Inversion}

\maketitle
\parskip 12pt

\begin{abstract}
Techniques developed for asymptotic wave equation based linearized inversion of
subsurface offset extended constant density acoustics also apply to variable density
acoustics. The data may be matched by proper choice of an extended
relative bulk modulus perturbation, with zero density
perturbation. This bulk modulus perturbation may be treated as an
offset-dependent reflectivity volume, and AVA-like techniques applied
to extract asymptotically correct estimates of velocity and density
perturbations, assuming that the background velocity is kinematically
consistent with the data.

The first part of this paper presents a 
theoretical analysis of the linearized inversion, leading to the
asymptotic wave equation algorithm. This second part
describes implementation details, explains how the asymptotic inverse
may be used to accelerate iterative inversion algorithms (``least
squares migration''), and  presents numerical examples of both
asymptotic and iterative inversion.
\end{abstract}
\setlength{\parindent}{0cm}

\section{Preliminaries}

Position vectors are $\bx = (x,z)$; $x$ is the horizontal position
vector, 1D or 2D for 2D or 3D modeling respectively.

The techniques explained here apply to acoustic data, or data that can
be satisfactorily treated as acoustic, recorded in a water layer at
the top of the half-space $z>0$. It's conceivable that similar
techniques may be developed in the future for other wave physics and
other acquisition geometries, but so far as I know such extensions
have not yet been worked out.

The pair $\kappa,\beta$ of bulk modulus and buoyancy (reciprocal
density) respectively is a convenient parametrization of the acoustic
model for present purposes. The acoustic fields $p,\bv$ solve the
acoustic wave equations
\begin{eqnarray}
\label{eqn:awe}
\frac{\partial p}{\partial t}& =& -\kappa (\nabla \cdot \bv +
w(t)\delta(\bx-\bx_s)),\nonumber \\
\frac{\partial \bv}{\partial t} & = & -\beta \nabla p. 
\end{eqnarray}  
The perturbational acoustic fields $\delta p, \delta \bv$ due to
perturbations $\delta \kappa, \delta \beta$ solve the
related equations 
\begin{eqnarray}
\label{eqn:pawe}
\frac{\partial \delta p}{\partial t}& =& -\kappa \nabla \cdot \delta \bv -
\delta \kappa (\nabla \cdot \bv + w(t)\delta(\bx-\bx_S),\nonumber \\
\frac{\partial \delta \bv}{\partial t} & = & -\beta \nabla \delta p - \delta
                                      \beta \nabla p. 
\end{eqnarray}  
in which $p,\bv$ are solutions of equation \ref{eqn:awe}.
The source characteristics (wavelet $w(t)$, locations $\bx_s$) are
assumed known.

$\dF$ denotes the acoustic Born modeling operator in the half-space $z>0$
adjoined to an infinite homogeneous half-space in $z<0$, that is, an
infinitely deep ocean. An implementable approximation of this
configuration uses an
absorbing boundary condition (PML) at the $z=0$ surface of a
rectangular domain, and also use absorbing boundary conditions at the
other boundary surfaces of the rectangle. Assume that the recorded
data are simply the samples of pressure at receiver
locations, {\em for both sources and receivers located on the
  bounndary $z=0$}. Explicitly,
\[
\dF[\kappa,\beta][\delta \kappa,\delta \beta](x_r,t;x_s) = \{\delta p((x_r,0),t;(x_s,0))\}
\]
with $x_r, x_s$ ranging over survey source-receiver horizontal
locations.

$\dF_0$ denotes the free surface (pressure-release) Born modeling operator in
the same half-space. The free surface acoustic field $p_0,\bv_0$ solves
the acoustic system \ref{eqn:awe} with the additional
Dirichlet boundary condition $p_0=0$ on $z=0$. The perturbational
field $\delta p_0,\delta \bv_0$ is defined similarly. Sources and receivers
must be positioned at positive depths $z_r>0$, $z_s>0$ (else the
output is boring) 
and the same horizontal locations as the hypothetical
absorbing-surface survey.

Define ``shallow tow depth'' to mean that for the data
spectrum of interest, the finite difference approximation 
\begin{equation}
\label{eqn:dipole}
\left(\frac{\partial}{\partial z_r}\frac{\partial}{\partial
  z_s}\delta p\right)((x_r,0),t; (x_s,0)) \approx
\frac{1}{4z_rz_s}\delta p_0((x_r,z_r),t;(x_s,z_s))
\end{equation}
is acceptably accurate.  
Assuming that $\delta \kappa, \delta \rho$ are non-zero only in $z>0$,
the absorbing surface perturbational field $\delta p,\delta \bv$ is
upcoming at $z=0$, that is, its high frequency energy is carried along
rays for which $z$ decreases as $t$ increases. Consequently, the depth
derivatives in equation \ref{eqn:dipole} can be computed
asymptotically to any order in frequency from the surface data $\dF$
alone, since asymptotically the fields satisfy one-way wave
equations. Let the operators on the surface data corresponding to the
partial derivatives in $z_s$, $z_r$ be denoted $D_{z_s}$ and $D_{z_r}$
respectively. \cite{HouSymes:15} show that $-D_{z_s}D_{z_r}$ is a
positive semi-definite operator of order 2. The definition of
``shallow tow depth'' \ref{eqn:dipole}
can be re-written
\begin{equation}
\label{eqn:opdipole}
D_{z_s}D_{z_r}\dF[\kappa,\rho][\delta \kappa,\delta \beta](x_r,t;x_s) \approx 
\frac{1}{4z_rz_s} Z \dF_0[\kappa,\beta][\delta \kappa,\delta \beta](x_r,t;x_s) 
\end{equation}
The operator $Z$ assigns traces with source and receiver depths $z_s,
z_r$ to traces with source and receiver depths $=0$, with all other
header and sample data the same.

The subsurface offset flavor of extended modeling makes the mechanical
parameters $\kappa,\beta$ into operators. In the Born case, only the
perturbational parameters $\delta \kappa,\delta \beta$ are regarded as
operators. In fact, it turns out that there is really no point in
making $\delta \beta$ into an operator, as it plays no necessary role at
all in subsequent developments so set
\begin{equation}
\label{eqn:zerodbeta}
\delta \beta = 0
\end{equation}
and drop $\delta \beta$ from the argument list of the linearized operators.

The extended bulk modulus
perturbation (operator kernel) $\delta \bar{\kappa}$ has a convenient
parametrization in terms of sunken midpoint $(m,z)$ and (horizontal)
half-offset $h$. The extended perturbation fields
$\delta \bar{p}, \delta \bar{\bv}$ are solutions of the extended
version of the system \ref{eqn:pawe}:
\begin{eqnarray}
\label{eqn:pawe}
\frac{\partial \delta \bar{p}}{\partial t}& =& -\kappa \nabla \cdot \delta \bar{\bv} -
\int\,dh\,\delta \bar{\kappa}(x-h,z,h) \nabla \cdot \bv(x-2h,z,t),\nonumber \\
\frac{\partial \delta \bar{\bv}}{\partial t} & = & -\beta \nabla \delta \bar{p} 
\end{eqnarray}  
Here is how to interpret the somewhat peculiar form of the
$h$-integral: $x$ is the horizontal sunken receiver position, $m=x-h$ the
sunken midpoint, and $x-2h$ the sunken source.

With this
interpretation, the absorbing surface extended modeling operator defined by 
\begin{equation}
\label{eqn:extop}
\doF[\kappa,\beta][\delta \bar{\kappa}] =
\{\bar{p}((x_r,0),t;(x_s,0))\}
\end{equation}
is the adjoint of Claerbout's survey sinking migration operator,
interpreted as a reverse-time construction.

The free surface extended modeling operator $\doF_0$ has a similar
definition. The exended absorbing surface fields $\delta
\bar{p},\delta \bar{\bv}$ are also upcoming at $z=0$, and ``shallow
tow depth'' can also be taken to mean
\begin{equation}
\label{eqn:opdipole}
D_{z_s}D_{z_r}\doF[\kappa,\rho][\delta \bar{\kappa}]\approx 
\frac{1}{4z_rz_s} Z \doF_0[\kappa,\beta][\delta \bar{\kappa}]
\end{equation}
to acceptable accuracy.

\section{Asymptotic Unitarity}
The first part of this paper analyzes unitarity relations for a
normalized version of $\doF$, acting on {\em extended reflectivities} $\bar{r}_{\kappa}$
and $\bar{r}_{\rho}$ as 
\begin{eqnarray}
\label{eqn:relfs}
\bar{r}_{\kappa}(\bx,\bh)& = & \frac{\delta \bar{\kappa}(\bx,\bh)}{\sqrt{\kappa(\bx+\bh)\kappa(\bx-\bh)}} \nonumber \\
\bar{r}_{\rho}(\bx,\bh)& = & \frac{\delta \bar{\rho}(\bx,\bh)}{\sqrt{\rho(\bx+\bh)\rho(\bx-\bh)}}
\end{eqnarray}
The relative linearized forward operator $\bar{F}[\kappa,\rho]$ is
\begin{equation}
\label{eqn:nondim}
\bar{F}[\kappa,\rho][\bar{r}_{\kappa},\bar{r}_{\rho}] =
D\bar{{\cal F}}[\kappa,\rho][\delta 
\bar{\kappa},\delta\bar{\rho}]
\end{equation}
where $\bar{r}_{\kappa},\bar{r}_{\rho}$ are related to $\delta 
\bar{\kappa},\delta\bar{\rho}$ by equation \ref{eqn:relfs}. That is, 
\begin{equation}
\label{eqn:nondimbis}
\bar{F}[\kappa,\rho][\frac{\delta \bar{\kappa}}{\sqrt{\kappa_+\kappa_-}}, \frac{\delta \bar{\rho}}{\sqrt{\rho_+\rho_-}}] =
D\bar{{\cal F}}[\kappa,\rho][\delta 
\bar{\kappa},\delta\bar{\rho}]
\end{equation}
using the notation $\kappa_{\pm}(\bx,\bh) = \kappa(\bx \pm \bh)$ and
so on, introduced in part 1.

Recall from the last section that $\delta \bar{\rho} = 0$ in the
sequel, and that $D\bar{\cal F}$ is regarded here as acting only on
$\delta \bar{\kappa}$. Therefore the same is assumed for the
normalized operator $\bar{F}$. Also, since $\kappa_{\pm}$ differs from
$\kappa$ by a (multiplication) $\Psi$DO whose symbol is $=1$ at $h=0$,
substituting $\kappa$ for $\kappa_{\pm}$ and $\rho$ for $\rho_{\pm}$
commits an error of the type we will ignore in the approximations to
follow, so we make this substitution without comment. Combining these
two observations, we write
\begin{equation}
\label{eqn:nondimterce}
\bar{F}[\kappa,\rho]\left[\frac{\delta \bar{\kappa}}{\kappa}\right] \approx  
D\bar{\cal F}[\kappa,\rho]\delta \bar{\kappa} 
\end{equation} 
or more concisely 
\begin{equation}
\label{eqn:nondimquatro}
\bar{F} \approx\doF \kappa
\end{equation}
in which $\kappa$ is shorthand for the operator of multiplication by $\kappa$.

We will use only the second of the unitarity results for $\bar{F}$,
Theorem 2, from the first part of the paper. We repeat this result for
convenience. First, define asymptotic approximations of the Hilbert transforms of the $z$ and $t$
derivatives:
\begin{eqnarray}
\label{eqn:kzkt1} 
K_zu(x,z,h) &=& \int dk_z e^{i(k_xx+k_z z+k_hh)}
                (1-\phi(k_z))|k_z|\hat{u}(k_x,k_z,k_h)\\
\label{eqn:kzkt3} 
K_tu(x_r,t;x_s) &=& \int d\omega e^{i(k_{x_r} x_r+k_{x_s} x_s+\omega
                    t)}
                    (1-\phi(\omega))|\omega|\hat{u}(k_{x_r},\omega;k_{x_s}) \\
\end{eqnarray}
in which $\phi \in C^{\infty}_0(\bR)$ is $=1$ in a neigborhood of
$0$. 
Define the model and data space weight operators by
\begin{eqnarray}
\label{eqn:defbarWm}
\bar{W}_m^{-1} &=& 32 v^2 K_z \\
\label{eqn:defbarWd}
\bar{W}_d & = & K_t I_t^2 I_t
D_{z_s}I_t
D_{z_r}I_t^2
\end{eqnarray}
Then
\begin{equation}
\label{eqn:defdd}
\bar{F}^{\ddagger} = \bar{W}_m^{-1}(\bar{F}^T\bar{W}_d)_{\kappa}.
\end{equation}
satisfies
\begin{equation}
\label{eqn:rk6sigmaterce}
\bar{F}[\kappa,\rho]^{\ddagger}\bar{F}[\kappa,\rho][\bar{r}_{\kappa}]
\approx  
\frac{1}{8 \pi^3}\int dk_x dk_z dk_h e^{i(xk_{x}+zk_{z}+hk_{h})}B P
 (\hat{\bar{r}}_{\kappa}) .
\end{equation}

Even though no ray-based computations are required to evaluate $\bar{F}^{\ddagger}$,
the derivation of its properties is
ray-theoretic, and depends on the assumptions
that the background fields $\kappa, \beta$ are smooth (slowly varying
on wavelenghth scale) and satisfy the ``DSR hypothesis'': that all
rays from source or receiver to scattering point maintain a nonzero
vertical velocity throughout their extent. A consequence of the DSR
hypothesis is that the energetically significant part of the source
wavefield should be downgoing, and the receiver wavefield should be
upcoming, at least near the sources and receivers.

The DSR assumption can be
stated: turning rays carry negligible energy. This hypothesis is
sometime true, sometimes not, and constitutes a real limitation on the
approximate inverse construction.

As discussed in part 1, for physical $\bar{r}_{\kappa} =
r_{\kappa}\delta(h)$, the operators $B$ and $P$ act as the identity, 
whence equation \ref{eqn:rk6sigmaterce} exhibits approximate unitarity 
of $\doF$ with respect to the weighted norms 
\begin{equation}
\label{eqn:wnorms}
\langle m_1,m_2\rangle_m = \langle m_1,W_m m_2 \rangle,\,\, 
\langle d_1,d_2\rangle_d = \langle d_1,W_d d_2 \rangle. 
\end{equation}

A unitarity result for $\doF$ follows immediately from that for
$\bar{F}$. Indeed, from \ref{eqn:nondimquatro} and
\ref{eqn:rk6sigmaterce} obtain
\[
PB \approx \bar{F}^{\ddagger} \bar{F} = \bar{W}_m^{-1} \bar{F}^T
\bar{W}_d  \bar{F} 
\]
\[
\approx \bar{W}_m^{-1} \kappa \doF^T \bar{W}_d \doF \kappa
\]
\begin{equation}
\label{eqn:rk6sigmaquatro}
\approx W_m^{-1}\doF^T \bar{W}_d \doF 
\end{equation}
in which
\begin{eqnarray}
\label{eqn:wmdefk}
W_m^{-1} = \kappa^2 \bar{W_m} &=& 32 \kappa^2 v^2 K_z = 32 \kappa^3 \beta
                             K_z
\end{eqnarray}

A rearrangement of $\bar{W}_d$ implicitly expresses the Green's
function perturbation, and corresponds to the implementation discussed
below. From the first sections of part 1, the pressure perturbation
corresponding to a reference field with a space-time delta source on
the RHS of the pressure equation is actually the time derivative of
the Green's function perturbation. Its trace is $\doF \delta \bar{\kappa}$, so
$I_t \doF$ is the trace of the Green's function perturbation. This
trace is already implicit in the data space weight operator:
\[
\doF^T \bar{W}_d \doF = \doF^T K_t I_t^2 I_t 
D_{z_s}I_t
D_{z_r}I_t^2 \doF
\]
\[
= -(I_t \doF)^T K_t I_t^2 
D_{z_s}
D_{z_r}I_t^2 (I_t \doF)
\]
\begin{equation}
\label{eqn:wdrearr}
=-(I_t \doF)^T K_t^{-3} 
D_{z_s}
D_{z_r} (I_t \doF)
\end{equation}
since $I_t^2 \approx -K_t^{-2}$. Define 
\begin{equation}
\label{eqn:wdredef}
W_d = -K_t^{-3}D_{z_s}D_{z_r}
\end{equation}
Then the approximation \ref{eqn:rk6sigmaquatro} is equivalent to the
unitarity relation for $I_t\doF$:
\begin{equation}
\label{eqn:rk6sigmaquinze}
PB \approx W_m^{-1} (I_t\doF)^T W_d (I_t \doF), 
\end{equation}
so that 
\begin{equation}
\label{eqn:dddef}
\iddoF = W_m^{-1} (I_t\doF)^T W_d
\end{equation}
is both the adjoint of $(I_t \doF)$ with respect to the weighted norms
defined by $W_m$ and $W_d$, and its approximate inverse for
near-physical extended $\delta \bar{\kappa}$.
[Note that we are re-purposing the symbols $W_m^{-1}$ and $W_d$ used 
in the statement of Theorem 1, part 1.]

Computation of $I_t \doF$ and its transpose simply requires use of the
$\delta(\bx-\bx_s)H(t)$, or a bandlimited approximation, in the RHS of
the reference pressure equation, that is, a bandpass filtered version of $w=H$ in \ref{eqn:awe}.

\section{Asymptotic Unitarity for Free Surface modeling}
An approximate inversion of
the free-surface modeling operator $\doF_0$ can be constructed on the
same principles. Begin with the dipole relation \ref{eqn:opdipole},
which we abbreviate as 
\begin{equation}
\label{eqn:opdipolebis}
D_{z_s}D_{z_r}\doF \approx \frac{1}{4z_rz_s} Z \doF_0
\end{equation}
The ``ghosting'' operator $D_{z_s}D_{z_r}$ (without time integrations)
is microlocally negative
definite symmetric as a consequence of the DSR hypothesis
\cite[]{HouSymes:15}, and in particular microlocally
invertible. Abusing notation a bit, write
\begin{equation}
\label{eqn:opdipoleterce}
\doF = (4z_rz_s D_{z_s}D_{z_r})^{-1} Z \doF_0
\end{equation}
and substitute \ref{eqn:opdipoleterce} in \ref{eqn:rk6sigmaquinze} to
obtain
\[
PB \approx W_m^{-1} (ZI_t\doF_0)^T (4z_rz_s D_{z_s}D_{z_r})^{-1} W_d
(4z_rz_s D_{z_s}D_{z_r})^{-1} (ZI_t \doF_0) 
\]
\begin{equation}
\label{eqn:rk6sigmads}
= W_m^{-1} (ZI_t\doF_0)^T (4z_rz_s)^{-2} K_t^{-3} D_{z_s}D_{z_r})^{-1} (ZI_t \doF_0).
\end{equation}
The relation \ref{eqn:rk6sigmads} may be re-written
\begin{equation}
\label{eqn:rk6sigmadsbis}
PB = W_m^{-1} (I_t\doF_0)^T W_d^0 (I_t\doF_0) 
\end{equation}
with 
\begin{equation}
\label{eqn:wddeffs}
W_d^0 = (4z_sz_r)^{-2} Z^T K_t^{-3} (D_{z_s}D_{z_r})^{-1} Z
\end{equation}
thus exhibiting the weighted adjoint
\begin{equation}
\label{eqn:appinvfs}
\iddoFf= W_m^{-1} (I_t \doF_0)^T W_d^0
\end{equation}
as an approximate inverse for near-physical arguments.

Construction of $\iddoFf$ according to \ref{eqn:appinvfs} appears to
require application of the ``de-ghosting'' operator
$(D_{z_s}D_{z_r})^{-1}$, per the definition \ref{eqn:wddeffs} of
$W_d^0$. As we will point out below, the shallow tow-depth hypothesis
proides an approximation to $\iddoFf$ that does not require
de-ghosting amplitude and phase adjustment.

\section{Preconditioned Conjugate Gradient Iteration}
The availability of an approximate inverse for $I_t \doF$ (and $I_t
\doF_0$) suggests that the least squares migration problem should be
reformulated as
\begin{equation}
\label{eqn:elsm}
\mbox{min}_{\delta \bar{\kappa}} \|I_t\doF[\kappa,\beta][\delta 
\bar{\kappa}] -I_t\delta d\|^2_d. 
\end{equation}
In this section, we will discuss the Conjugate Gradient (CG) algorithm
for generic problems of this form. Because the algorithms have general
applicability, and for brevity, we will use the notation $A$ for a
linear operator such as $I_t \doF$, $W_m$ and $W_d$ for positive
definite symmetric weight operators defining weighted inner products
in the domain and range of $A$ respectively, $b$ for a data vector in
the range ($I_t \delta d$ in our application), and $x$ for a model
vector in the domain ($=\delta \bar{\kappa}$ here). 
The CG algorithm is defined in terms of the weigted inner products
domain space $\langle x_1, x_2 \rangle_m = x_1^TW_m x_2$ and similarly
in the range. Denote by $A^{\dagger}$ the adjoint of $A$ with respect to the
weighted inner products.
Near-unitarity of $A$, that is, $A^{\dagger}A \approx I$, implies
rapid convergence of the CG iteration. 
\begin{algorithm}[H]
\caption{Conjugate Gradient Algorithm}
\begin{algorithmic}[1]
  \State Choose $x_0$ somehow
  \State $g_0 \gets A^{\dagger}(b-Ax_0)$
  \State $p_0 \gets g_0$ 
  \State $w_0 \gets A^{\dagger}A p_0$
  \State $k \gets 0$
  \Repeat
  \State $\alpha_k \gets \frac{\langle g_k,g_k \rangle_m}{\langle p_k,w_k\rangle_m}$
  \State $x_{k+1} \gets x_k + \alpha_k p_k$
\State $g_{k+1} \gets g_k - \alpha_kw_k$
  \State $\beta_{k+1} \gets \frac{\langle g_{k+1},g_{k+1}\rangle_m}{\langle g_k,g_k\rangle_m}$
  \State $p_{k+1}\gets g_{k+1}+\beta_{k+1}p_k$
  \State $w_{k+1} \gets A^{\dagger}Ap_{k+1}$
  \State $k \gets k+1$
  \Until{Error is sufficiently small, or max iteration count exceeded} 
\end{algorithmic}
\end{algorithm}
Several things to note:
\begin{itemize}
\item Each step requires on application of the normal operator $A^{\dagger}A$. Except for the first step, only the action of the normal  operator is required.
\item The initial step requires an application of $A^{\dagger}$ to the
  data. A common choice is $x_0 = 0$.
\item In the form presented here, the algorithm requires explicitly only the model space inner product (of course, the data space inner
  product is implicit in the definition \ref{eqn:wadj} of
  $A^{\dagger}$). 
\end{itemize}

The algorithm described above is equivalent to the well-known
Preconditioned Conjugate Gradient algorithm \cite[]{gvl}. Define
$A^*$ to be the adjoint of $\doF$ with respect to the (unaltered)
$L^2$ norm in model space and the weighted norm defined above in data
space. That is,
\begin{eqnarray}
A^* &=& A^T W_d \nonumber\\
A^{\dagger} &=& W_m^{-1}\doF^*
\label{eqn:dwadj}
\end{eqnarray}
 Also, using the notation of the
algorithm above, set 
\begin{eqnarray}
r_k &=& A^*(b - Ax_k) = W_m g_k\nonumber \\
q_k &=& W_m w_k
\label{eqn:pcgvars}
\end{eqnarray}
Then for example
\[
\langle g_k, g_k\rangle_m = \langle g_k, W_m g_k \rangle = \langle
g_k,r_k \rangle
\]
and so on, so the Conjugate Gradient Algorithm is equivalent to
\begin{algorithm}[H]
\caption{Preconditioned Conjugate Gradient Algorithm}
\begin{algorithmic}[1]
\State Choose $x_0$ somehow
  \State $r_0 \gets A^*(b-Ax_0)$
  \State $p_0 \gets W_m^{-1}r_0$
  \State $g_0 \gets p_0$
  \State $q_0 \gets A^*Ap_0$
  \State $k \gets 0$
  \Repeat
  \State $\alpha_k \gets \frac{\langle g_k,r_k \rangle}{\langle p_k,q_k\rangle}$
  \State $x_{k+1} \gets x_k + \alpha_k p_k$
  \State $r_{k+1} \gets r_k - \alpha_kq_k$
  \State $g_{k+1} \gets W_m^{-1}r_{k+1}$
  \State $\beta_{k+1} \gets \frac{\langle g_{k+1},r_{k+1}\rangle_m}{\langle g_k,r_k\rangle_m}$
  \State $p_{k+1}\gets g_{k+1}+\beta_{k+1}p_k$
  \State $q_{k+1} \gets A^*Ap_{k+1}$
  \State $k \gets k+1$
  \Until{Error is sufficiently small, or max iteration count exceeded} 
\end{algorithmic}
\end{algorithm}
Note that the Preconditioned Conjugate Gradient algorithm involves
only one application of {\em model-unweighted} normal operator
$A^*A$ and one applicatino of the model weight operator $W_m^{-1}$
per step, and only the unweighted (Euclidean) inner product in the
model space - that is, (weighted) the model space inner product has
disappeared as well. This algorithm is considerably simpler to
implement than the Conjugate Gradient Algorithm in weighted spaces, so
we will use it in all experiments reported below.

\section{Practical Computation of Adjoint and Normal Operators}
For absorbing boundary (deghosted) data, 
\[
\iddoF = -32 \kappa^3 \beta K_z (I_t\doF)^T  K_t^{-3} D_{z_s}D_{z_r} = -32 \kappa^3 \beta K_z (D_{z_s}D_{z_r}I_t\doF)^T  K_t^{-3} 
\]
\begin{equation}
\label{eqn:abadj}
\approx -32 \kappa^3 \beta K_z \frac{1}{4z_sz_r}(I_t\doF_0)^T Z^T K_t^{-3}
\end{equation}
using the dipole relation \ref{eqn:opdipolebis} and the symmetry of
the ``ghosting'' operator.

As noted earlier, the $\Psi$DOs $K_z$ and $K_t$ can be computed via 1D
Fourier transforms using tapered high- and low-pass cutoffs to
regulate frequency content to avoid aliasing, and $(I_t\doF_0)^T $ is
computed via the adjoint state method.  The normal operator is simply
the composition of the operator expressed in \ref{eqn:abadj} with
$I_t\doF_0$.

The upshot: equation \ref{eqn:abadj} provides a computation of the
approximate inverse of $I_t \doF$ that does not require explicit
computation of the ``ghosting'' operator $D_{z_s}D_{z_r}$.

Similarly, \ref{eqn:appinvfs} states that
\[
\iddoFf = -32 \kappa^3 \beta K_z (ZI_t \doF_0)^T K_t^{-3}
(4z_sz_r)^{-2} (D_{z_s}D_{z_r})^{-1}Z
\]
 \begin{eqnarray}
\label{eqn:fsadj}
\approx -32 \kappa^3 \beta K_z \frac{1}{4z_sz_r}(I_t\doF)^T K_t^{-3} Z
\end{eqnarray}
which also provides a route to computation of $\iddoFf$ via the adjoint
state method and simple filters, without requiring explicit
computation of the ``de-ghosting'' operator $(D_{z_s}D_{z_r})^{-1}$,
as expression \ref{eqn:appinvfs} appeared to require.

Note that the expressions \ref{eqn:abadj} and \ref{eqn:fsadj} depend
on the shallow tow depth assumption,  appropriate for marine towed
streamer acquisition only, and then only for a limited range of
frequencies (up to the first ghost notch).
For some other configurations, similar constructions should be
possible. For example, OBS node arrays record 4C data, including
vertical velocity data. Vertical velocity is precisely $I_tD_{z_r}$
applied to pressure data. With shallow tow depth sources as typically
used in node surveys, the $I_tD_{z_s}$ factor can be handled by free
surface modeling and the method of images as is done here. For
streamers at deeper tow depth, or to go beyond the notch, some
additional data is necessary. If source tow depth is shallow enough
that the method of images is adequate to approximate $I_t D_{z_s}$,
then the receiver side vertical velocity data could be supplied by
Geostreamer$^{TM}$ for instance without regard to tow depth . These
possibilities remain to be investigated.

It is important to note that the adjoint computations are
inexact, and that the weights as presented above are not quite SPD
operators, so need modification if they are to precisely define inner
products. Besides the microlocal ellipticity of the weights, the model
weight $W_m^{-1}$ is clearly not symmetric, but can easily be 
symmetrized:
\[
W_m^{-1} \leftarrow 32\kappa^{3/2}\beta^{1/2} K_z \beta^{1/2}\kappa^{3/2}
\]

In the special case of a constant-property layer (ideal ocean) near
$z=0$, the data weights $W_d$ and $W_d^0$ are prima facie symmetric, as their
factors are all convolutions hence commute. Thus the inner products
are actually inner products, except for being semidefinite, and act as
inner products for high-frequency microlocalized data.

The adjoint formulae \ref{eqn:abadj} and \ref{eqn:fsadj} are only as
good as the ``shallow tow depth'' assumption. That is, there is an
error of size $O(k_z^2z_rz_s)^2$ between computed and actual adjoint,
in which $k_z$ is the largest frequency carrying significant
energy. Since the error likely contributes to a nonsymmetric error in
the normal operator, CG iteration is likely to stall eventually,
requiring restart. Note that direct assessment of the error in the
adjoint is difficult, as the data side inner product is not easily
computable. However the asymmetry of the computed normal operator can
be assessed by random vector computations.

A final note on implementation: the (microlocal) definite property of
the ghosting operator is dependent on the DSR assumption, as explained
by \cite{HouSymes:15}: the source wavefields must be downgoing,
receiver wavefields upcoming, at least insofar as these carry
significan energy in the computations explained above. If these
requirements are violated, then the partial weighted adjoint $A^*A
(A=I_t \doF)$ may acquire negative eigenvalues, causing the conjugate
gradient iteration to diverge.

The downgoing/upcoming property
could be enforced by dip filtering of source and receiver
data. However for the simplified marine configuarions used in the
experiments reported below, it was simpler to separate the scattering
region from the source-receiver manifold by a horizontally homogeneous
zone (water layer).  The separation is accomplished by a depth
tapering operator, which tapers the extended bulk modulus perturbation
$\delta \bar{\kappa}$ to zero near the surface. A simple ray analysis
show that the source and receiver wavefields then must have the
required properties, and consequently we observe convergence in the
conjugate gradient iteration. We emphasize that without this
precaution, horizontally propagating energy inevitably develops during
the iteration and leads to divergence.

%%%%%%%%%%%%%%%% numerics %%%%%%%%%%%%%%

\inputdir{project}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Example 0: Three Reflectors, Homogeneous
  Background}
This example is adapted from the first example in
\cite[]{HouSymes:15}, and reproduces it almost exactly.

The background parameters are $\kappa$=2.5 GPa, $\rho$ = 1.0 g/cc. The
density perturbation vanishes, the bulk modulus perturbation is
plotted in \ref{fig:boxdbulk}.

\plot{boxdbulk}{width=\textwidth}{Example 0: Bulk modulus perturbation for three
  layer example. Max excursion = 0.4 GPa.}

\plot{boxborn37}{width=\textwidth}{Example 0: Born simulated free surface traces,
  center midpoing 37 of 76, 
  (integrated) [2.5,5.0, 35, 40] Hz bandpass filter source.}

\plot{boxfreeinvp1it0import}{width=\textwidth}{Example 0: Asymptotic extended inversion of
  free surface data.}

\plot{boxfreerebornp1it0shot37}{width=\textwidth}{Example 0: Resimulation of data
  in Figure \ref{fig:boxborn37}.}

\plot{boxfreerebornp1it0diff37}{width=\textwidth}{Example 0: Data residual for shot
  depicted in Figure \ref{fig:boxborn37}.}

\plot{boxfreeinvp1it0stack}{width=\textwidth}{Example 0: Asymptotic extended
  inversion of free surface data, stacked over offset - best estimate
  of bulk modulus perturbation.}

\plot{boxcomp150}{width=\textwidth}{Example 0: comparison of central
  depth trace, blue (target) vs. red (asymptotic inversion).}

\newpage

\section{Example 1: Marmousi LSM, Coarse Grid}

The aim in this example and in examples 2 and 4 is to use the
zero-offset restriction of the approximate inverse \ref{eqn:fsadj} to
accelerate (ordinary, non-extended) least squares migration (LSM). The
example is based on the Marmousi model \cite[]{BoLaVe:91} treated as a
constant density acoustic medium, $\rho = 1$. The bulk modulus is shown
in Figure \ref{fig:coarse1bulk}.

\plot{coarse1bulk}{width=\textwidth}{Example 1: Bulk modulus derived from 
  Marmousi model, with $rho = 1$ - that is, velocity 
  squared. Subsampled by factor of 6 in both directions, so that $dx =
  dz = 24$ m.}

The scale division used in this experiment is displayed in Figures
\ref{fig:coarse1bulkbig} and \ref{fig:coarse1dbulk}. The smoothed bulk
modulus displayed in Figure \ref{fig:coarse1bulkbig} is the result of
4 repetitions of 6 $\times $ 6 box averaging of the subsampled
Marmousi model (Figure \ref{fig:coarse1bulk}) on the 24 m grid. The
  bulk modulus perturbation (Figure \ref{fig:coarse1dbulk}) is simply
  the difference.

\plot{coarse1bulkbig}{width=\textwidth}{Example 1: Smooth background bulk modulus
  resulting from 6 $\times$ 6 box averaging of the field in Figure
  \ref{fig:coarse1bulk}, repeated 4 times.}

\plot{coarse1dbulk}{width=\textwidth}{Example 1: Difference between unsmoothed
  and smoothed bulk moduli, used as bulk modulus perturbation to
  generate Born data for this experiment series.}

This model is used to simulate 90 shots spaced 96 m apart, with 190
receivers in a fixed (shot-independent) array. Shot and receiver
depths are 24 m. The acquisition geometry is summarized in the output
of {\tt surange}:
\begin{verbatim}
17100 traces:
tracl    1 17100 (1 - 17100)
tracr    1 17100 (1 - 17100)
fldr     0 89 (0 - 89)
tracf    1 190 (1 - 190)
offset   -8760 8856 (-216 - 312)
gelev    -24
selev    -24
sx       240 8784 (240 - 8784)
gx       24 9096 (24 - 9096)
ns       501
dt       8000

Shot coordinate limits:
        North(240,0) South(240,0) East(8784,0) West(240,0)

Receiver coordinate limits:
        North(24,0) South(24,0) East(9096,0) West(24,0)

Midpoint coordinate limits:
        North(132,0) South(132,0) East(8940,0) West(132,0)
\end{verbatim}

Sources and receivers are point and isotropic. The source wavelet is
is a once-integrated bandpass filter with corner frequencies 2.5, 5,
12.5, and 15 Hz. The pressure (source) field is a
bandpass-filtered Green's function.

The source wavelet is scaled by the value of the bulk modulus near the
source positions, that is, in the water column, to compensate for the
lack of such scaling in the IWAVE acoustic staggered grid
implementation: otherwise, the computed fields would not correspond to
solutions of equation \ref{eqn:awe}.

The surface $z=0$ is pressure-free, that is, the boundary condition is
$p=0$ on this surface.

We use a very conventional staggered-grid finite difference method of
second order in time and 8th order in space to simulate all wavefields
contributing the results reported here \cite[]{Vir:84,Vir:86}.

Shot gather 45, at source position 4560 m, is displayed in Figure
\ref{fig:coarse1bornr45}.

\plot{coarse1bornr45}{width=\textwidth}{Example 1: Shot gather 45, $x_s$ = 4560
  m. }

Migration (Reverse Time Migration = RTM) of this data with yields the
image displayed in Figure \ref{fig:coarse1mbulk}. Our implementation
of RTM has a couple of features worthy of note:
\begin{itemize}
\item it realizes the {\em adjoint} or {\em transpose} of Born
  modeling. This choice differs from some more conventional choices of
  RTM in filtering and scaling. We verify the adjointness relation
  between RTM and Born modeling operators by the well-known {\em dot
    product test} \cite[]{claer:PVI} which passes at a low multiple of
  machine (single) precision.
\item For reconstruction of the source wavefield, we use the optimal
  checkpointing method
  \cite[]{Griewank:92,Griewank:00,Griewank:book,Symes:06a-pub}, which
  has the virtue of achieving the adjoint relation to machine
  precision for virtually any time-domain simulation method, but the
  drawback of somewhat more computational complexity than would be
  required for some specialized techniques suited for special systems
  such as our acoustic problem. Generally the cost of RTM via optimal
  checkpointing is roughly 4 -5 times the cost of forward Born
  modeling for problems with $O(10000)$ time steps, such as those
  solved here.
\end{itemize}

Several aspects are immediately apparent from comparison with the of
the RTM image \ref{fig:coarse1mbulk} with the target image, that is,
the actual bulk modulus perturbation used to generate the data, see
Figure \ref{fig:coarse1dbulk}:
\begin{itemize}
\item locations and orientation of events in the image
  \ref{fig:coarse1mbulk} are correct, up to limits imposed by the bandlimited
  nature of the image, which follows from the bandlimited nature of
  the data.
\item the overal amplitude is grossly incorrect, and the relation
  between reflector strengths in various parts of the image is also
  much different: in particular, the deeper part of the image is much
  fainter than the shallow part. This feature follows from the uneven
  illumination of the subsurface: both the fall-off of source
  wavefield and receiver wavefield amplitude with distance from the
  surface, and the focusing and defocusing of these wavefields as they
  traverse the heterogeneous refractive velocity model, contribute to
  this effect.
\item some evidence of transmitted wavefield is present: the
  low-frequency swings visible in the image are due to refracted
  (diving) wave components in source and receiver field contributing
  to image amplitude along the entire diving wave ray path, not just
  at a reflector location. These signal components are actually quite
  important, and are the principal contribution to FWI. However in
  reflector imaging they are often regarded as noise.
\end{itemize}

\plot{coarse1mbulk}{width=\textwidth}{Example 1: Reverse time migration of Born
  data. Notice the uneven illumination, and the faint but visible
  low spatial frequency diving wave artifacts.}

LSM is the name acquired in recent years by the data fitting problem:
given data (regarded as) resulting from a Born or perturbational
model, estimate the corresponding model perturbation by minimization
the mean-square misfit between predicted and given Born data. This
task defines a linear inverse problem, and an equally good (and older)
synonym is ``linearized inversion''. It has long been recognized that
reducing or minimizing data misfit tends to resolve the illumination
and amplitude defects of migrated images, noted above
\cite[]{BourJiaLai:89}.  Data misfit minimization is commonly
implemented via conjugate gradient iteration or an equivalent
\cite[]{Nemeth:99}. Each iteration requires one Born forward modeling
operation and one (variant of) Reverse Time Migration.

The 2017 SEG Annual International Meeting and Exposition featured
several sessions on LSM, including elastic LSM for estimation of shear
properties, acoustic LSM as is demonstrated here, viscoelastic LSM for
correct estimation of amplitudes in the presence of attenuation and/or
attentuation parameters, and a variety of case studies.

We use Born data (one shot of which was just displayed) as target data
in our LSM exercise. Twenty iterations of conjugate gradient iteration
produces the estimate of the bulk modulus perturbation displayed in
FIgure \ref{fig:coarse1bulkinvp0it20}.

\plot{coarse1bulkinvp0it20}{width=\textwidth}{Example 1: Estimate of bulk modulus
  perturbation resulting from least squres migration, 20 conjugate
  gradient iterations. Plotted on same color scale as target bulk
  modulus perturbation (Figure \ref{fig:coarse1dbulk}).}

LSM has indeed balanced the amplitudes throughout the image: the
relative amplitudes between various parts are much closer to those of
the target image \ref{fig:coarse1dbulk}. The overall amplitude is also
much closer to correct, so one to some extent regard the image point
values as local and band-limited estimates of physical
parameter (perturbation in bulk modulus). The ampltudes are generally
low, partly because of the spatial averaging effect of bandlimitation,
partly because further iterations would decrease data residual and
recover model parameters more accurately.

LSM achieves its aim by presuming that achieving data fit will result
in a more accurate image of the subsurface. Figure
\ref{fig:coarse1rebornp0it20r45} shows the predicted data for
shot 45 based on the bulk modulus perturbation in Figure
\ref{fig:coarse1bulkinvp0it20}. These traces indeed closely resemble
the target data \ref{fig:coarse1bornr45}; the difference (residual
shot record) appears as \ref{fig:coarse1residrebornp0it20r45}. The RMS
fit level achieved is part of the information displayed in the
following table:

\begin{verbatim}
========================== BEGIN CGNE =========================
Iteration   |  Residual Norm |  Gradient Norm
         1    2.50883907e-01   2.06906028e-04
         2    2.11527959e-01   1.23586302e-04
         3    1.93715423e-01   1.00169396e-04
         4    1.79163083e-01   9.69018802e-05
         5    1.64154008e-01   9.32183611e-05
         6    1.47441313e-01   8.68227726e-05
         7    1.32597432e-01   7.44872232e-05
         8    1.20940886e-01   6.83080798e-05
         9    1.11524783e-01   5.57000349e-05
        10    1.03391834e-01   5.03123520e-05
        11    9.70416740e-02   4.44237339e-05
        12    9.14672911e-02   4.11237088e-05
        13    8.63379091e-02   3.77963879e-05
        14    8.15358013e-02   3.53451178e-05
        15    7.70760477e-02   3.34034703e-05
        16    7.26942420e-02   3.14817880e-05
        17    6.85755685e-02   3.00511110e-05
        18    6.44731000e-02   2.88680549e-05
        19    6.06474578e-02   2.66172028e-05
        20    5.71314394e-02   2.49373679e-05
-----------------------------------------------
        21    5.40441014e-02   2.32873826e-05
=========================== END CGNE ==========================

 ****************** CGNE summary *****************  
initial residual norm      = 2.50883907e-01
residual norm              = 5.40441014e-02
residual redn              = 2.15414777e-01
initial gradient norm      = 2.06906028e-04
gradient norm              = 2.32873826e-05
gradient redn              = 1.12550527e-01
\end{verbatim}
Evidently convergence is slow but sure. The left-hand column lists the
data residual, which is (for mathematically consistent implementations
of the conjugate gradient algorithm) guaranteed to decrease at each
step. The right-hand column is the gradient length, or normal
residual, whcih also converges to zero but not (necessarily)
monotnically.

The reduction in data misfit achieved over 20 iterations is
approximately 21.5\%.

\plot{coarse1rebornp0it20r45}{width=\textwidth}{Example 1: Predicted data from 
  model of Figure \ref{fig:coarse1bulkinvp0it20}.}

\plot{coarse1residrebornp0it20r45}{width=\textwidth}{Example 1: Data residual from 
  model of Figure \ref{fig:coarse1bulkinvp0it20}. RMS error is
  approximately 21.5\%.}

We can reach two tentative conclusions from this example:
\begin{itemize}
\item data fitting, that is, LSM, does indeed tend to extract image 
  amplitudes so that they resemble those in the ``earth'', hence
  incidentally correct uneven illumination;
\item convergence can be slow, and the premium in computational cost
  over RTM is a large factor (20X, in this example).
\end{itemize}

The cost of LSM is a major impediment to its use. To speed it up,
either each iteration can be made less expensive, or the number of
necessary iterations can be reduced. We use the approximate
inverse constructed in the preceeding sections to accelerate the
convergence of the LSM loop. Note that in this section we are dealing
with a physical or non-extended model. The second form of the
approximate inverse  
involves a model weight operator that acts diagonally on the various
offset sections of the extended model, and in particular on the
zero-offset section, which is the result of conventional RTM such as
the image displayed above. There is no reason to think particularly
that the zero-offset section of the approximate inverse output is an
approximate inverse of the zero-offset (physical) problem, however we
observe that indeed the amplitudes are considerably
improved over basic RTM, so we may attempt to use it as a
preconditioner.

We explain the link between the zero-offset weighted adjoint and a
{\em bona fide} approximate inverse in the discussion of Example 3.

The preconditioned conjugate gradient iteration described earlier is
applied to estimate $\delta \kappa$. The algorithm approximates a
minimizer of the least-squares objective
\begin{equation}
\label{eqn:elsmex1}
\|I_t\dF \delta \kappa -I_t\delta d\|^2_d = \langle (I_t \doF \delta \kappa \delta(h) -
I_d \delta d) W_d^0 (I_t \doF \delta \kappa \delta(h)- I_d \delta d)\rangle). 
\end{equation}
The expression on the RHS uses the extension relation $\dF \delta
\kappa = \doF \delta \kappa \delta(h)$. The code implements $\dF$
through this relation, with care to avoid inefficiency. The data
weight operator $W_d^0$ is defined in \ref{eqn:wddeffs}. As pointed
out above, the PCG algorithm does not require explicit computation of
$W_d^0$, a good thing as it would be quite expensive. THe expression
\ref{eqn:elsmex1} emphasizes that the data residual is minimized in
the sense of the weighted norm, not in the sense of the unweighted
($L^2$) norm. Since we do not compute the weighted norm explicitly,
the implemented PCG algorithm displays records the $L^2$ norm, rather
than the weighted norm, and the former is not guaranteed to decrease
monotonically.

Application of the asymptotic inverse \ref{eqn:appinvfs}, in the
convenient form \ref{eqn:fsadj} provided by the shallow tow depth hypothesis,  yields the image displayed in Figure
\ref{fig:coarse1bulkinvp1it0}. This image has the same overall
amplitude balance as the 20-iteration LSM image
\ref{fig:coarse1bulkinvp0it20}, but results from a single migration
pass, is less noisy, and has even more
correct relative amplitudes. On the other hand the overall amplitude
is roughly 4X too small, and the data residual is still large - around 80\% in
fact. However if you scale the image by a factor of 4, the difference
drops to less than 50\%, indicating that the approximate inverse
zero section \ref{fig:coarse1bulkinvp1it0} may be a good search
direction. We will explain why this is so in the discussion of Example 3.

\plot{coarse1bulkinvp1it0}{width=\textwidth}{Example 1: Approximate inversion or
  true amplitude image = zero offset section of subsurface offset
  approximate inversion. Plotted on same color scale as target bulk
  modulus perturbation in Figure \ref{fig:coarse1dbulk}.}

Ten iterations of PCG produces the approximate inversion in Figure
\ref{fig:coarse1bulkinvp1it10}. The amplitudes are roughly correct,
allowing for the spatial averaging influence of band limitation, and the
resolution improved somewhat over that displayed in Figure
\ref{fig:coarse1bulkinvp1it0}. More than that, this is a good
inversion: the RMS error is less than 10\%. The output of the RVL CGNE
tool describes the convergence:

\begin{verbatim}
========================== BEGIN PCGNE =========================
Iteration   |  Residual Norm |  Gradient Norm
         1    2.50883907e-01   6.89441109e+00
         2    1.18942171e-01   2.66392207e+00
         3    7.03184232e-02   1.10846794e+00
         4    5.15148342e-02   6.83876038e-01
         5    4.13431711e-02   4.61543351e-01
         6    3.50607708e-02   3.48029107e-01
         7    3.10531463e-02   2.71396458e-01
         8    2.83174179e-02   2.21747950e-01
         9    2.63601393e-02   1.83933064e-01
        10    2.49932427e-02   1.60866663e-01
-----------------------------------------------
        11    2.39612553e-02   1.46033272e-01
=========================== END PCGNE ==========================

 ****************** CGNE summary *****************  
initial residual norm      = 2.50883907e-01
residual norm              = 2.39612553e-02
residual redn              = 9.55073461e-02
initial gradient norm      = 6.89441109e+00
gradient norm              = 1.46033272e-01
gradient redn              = 2.11813990e-02
\end{verbatim}

The predicted data shot 45 is displayed in Figure
\ref{fig:coarse1rebornp1it10r45}, plotted on the same scale as Figure
\ref{fig:coarse1bornr45} - it is hard to see a difference between the
two. The residual practically disappears (Figure \ref{fig:coarse1residrebornp1it10r45}).

\plot{coarse1bulkinvp1it10}{width=\textwidth}{Example 1: Estimated bulk modulus
  perturbation from ten precondiitioned CG iterations. Compare to
  Figure \ref{fig:coarse1dbulk}.}

\plot{coarse1rebornp1it10r45}{width=\textwidth}{Example 1: Predicted shot record 45 from
  inversion displayed in Figure \ref{fig:coarse1bulkinvp1it10} -
  compare with Figure \ref{fig:coarse1bornr45}, which shows target
  data plotted on same grey scale, or with next figure, which
  replicates Figure \ref{fig:coarse1bornr45} for the reader's convenience.}

\plot{coarse1bornr45bis}{width=\textwidth}{Example 1: Replication of Figure \ref{fig:coarse1bornr45} for
convenient comparison with previous figure. Note that the only readily
visible difference is the appearance of water column energy in the
resimulated data (Figure \ref{fig:coarse1rebornp1it10r45}), absent in
the target data (Figure \ref{fig:coarse1bornr45}).}

\plot{coarse1residrebornp1it10r45}{width=\textwidth}{Example 1: Residual =
  difference of data displayed in previous two figures, plotted on
  same grey scale.}

These results suggest the following conclusions: 
\begin{itemize}
\item The zero-offset restriction of the asymptotic 
  inversion operator \ref{eqn:appinvfs} (implemented via \ref{eqn:fsadj}) is an effective preconditioner for LSM: it gives a initial
  search direction and continues to accelerate CG in later
  iterations. Its deficient overall amplitude is implicitly corrected
  by the CG step length calculation.
Convergence is much faster with than without preconditioning: the
preconditioned algorithm takes 4 iterations to reach (or better) the
residual obtained by the unpreconditioned algorithm in 20 iterations.
\item Good RMS convergence does not guarantee convergence of the
  estimated bulk modulus (or other parameter) to the target model {\em
    pointwise} -
  only to within the resolution of the data bandwidth, which affects
  estimated amplitudes via spatial averaging. 
\end{itemize}

\newpage
\section{Example 2: Marmousi LSM, Fine Grid}

\inputdir{project}

This section replicates the results of the preceding section with
higher frequency data and finer spatial grid. The Marmousi bulk
modulus on this 12 m $\times$ 12 m grid is shown 
in Figure \ref{fig:fine1bulk}.

\plot{fine1bulk}{width=\textwidth}{Example 2: Bulk modulus derived from 
  Marmousi model, with $rho = 1$ - that is, velocity 
  squared. Subsampled by factor of 6 in both directions, so that $dx =
  dz = 12$ m.}

The scale division used in this experiment is displayed in Figures
\ref{fig:fine1bulkbig} and \ref{fig:fine1dbulk}. The smoothed bulk
modulus displayed in Figure \ref{fig:fine1bulkbig} is the result of
4 repetitions of 6 $\times $ 6 box moving average of the subsampled
Marmousi model (Figure \ref{fig:fine1bulk}) on the 12 m grid. The
  bulk modulus perturbation (Figure \ref{fig:fine1dbulk}) is the
  difference between bulk moduli corresponding to the Marmousi model
  and the result of 2 repititions of 3 $\times$ 3 box moving average.

\plot{fine1bulkbig}{width=\textwidth}{Example 2: Smooth background bulk modulus
  resulting from 6 $\times$ 6 box averaging of the velocity
  corresponding to the field in Figure
  \ref{fig:fine1bulk}, repeated 4 times.}

\plot{fine1dbulk}{width=\textwidth}{Example 2: Difference between base bulk modulus (Fig
  \ref{fig:fine1bulk} and the bulk modulus obtained by a milder
  smothing (3 $\times$ 3 moving average repeated twice) of the
  Marmousi velocity field. Used as bulk modulus perturbation to
  generate Born data for this experiment series.}

This model is used to simulate 180 shots spaced 48 m apart, with 380
receivers in a fixed (shot-independent) array. Shot and receiver
depths are 12 m. The acquisition geometry is summarized in the output
of {\tt surange}:
\begin{verbatim}
68760 traces:
tracl    1 68760 (1 - 68760)
tracr    1 68760 (1 - 68760)
fldr     0 179 (0 - 179)
tracf    1 382 (1 - 382)
offset   -8820 8916 (-228 - 324)
gelev    -12
selev    -12
sx       240 8832 (240 - 8832)
gx       12 9156 (12 - 9156)
ns       1001
dt       4000

Shot coordinate limits:
        North(240,0) South(240,0) East(8832,0) West(240,0)

Receiver coordinate limits:
        North(12,0) South(12,0) East(9156,0) West(12,0)

Midpoint coordinate limits:
        North(126,0) South(126,0) East(8994,0) West(126,0)
\end{verbatim}

Sources and receivers are point and isotropic. The source wavelet is
is a once-integrated bandpass filter with corner frequencies 2.5, 5,
25, and 30 Hz. The pressure (source) field is a
bandpass-filtered Green's function.

The surface $z=0$ is pressure-free, that is, the boundary condition is
$p=0$ on this surface.

Shot gather 90, at source position 4560 m, is displayed in Figure
\ref{fig:fine1bornr90}.

\plot{fine1bornr90}{width=\textwidth}{Example 2: Shot gather 90, $x_s$ = 4560
  m. }

RTM once again produces an image with major reflectors correctly
positioned, but with badly unbalanced amplitudes and low-frequency
diving wave artifacts (``rabbit ears'' - in FWI applications not
useful, but in fact completely useless here, as the velocity model is
already precisely correct).

\plot{fine1mbulk}{width=\textwidth}{Example 2: Reverse time migration of Born
  data. Notice the uneven illumination, and the clearly visible
  low spatial frequency diving wave artifacts.}

The additional bandwidth appears to make this problem more difficult
(more ill-conditioned) than the problem discussed in the last section.
LSM via twenty iterations of conjugate gradient iteration
produces the estimate of the bulk modulus perturbation displayed in
FIgure \ref{fig:fine1bulkinvp0it20}.

\plot{fine1bulkinvp0it20}{width=\textwidth}{Example 2: Estimate of bulk modulus
  perturbation resulting from least squres migration, 20 conjugate
  gradient iterations. Plotted on same color scale as target bulk
  modulus perturbation (Figure \ref{fig:fine1dbulk}).}

LSM has partly balanced the amplitudes throughout the image: the
relative amplitudes between various parts are much closer to those of
the target image \ref{fig:fine1dbulk}, but not as close as the
analogous result from the narrow-band low frequency roblem, Figure
\ref{fig:coarse1bulkinvp0it20}. Figure
\ref{fig:fine1rebornp0it20r90} shows the predicted data for
shot 90 based on the bulk modulus perturbation in Figure
\ref{fig:fine1bulkinvp0it20}. These traces resemble
the target data \ref{fig:fine1bornr90}, but not as closely as the
analogous comparison of the lower-frequency data. The difference (residual
shot record) appears as \ref{fig:fine1residrebornp0it20r90}. The RMS
fit level achieved is part of the information displayed in the
following table:

\begin{verbatim}
========================== BEGIN CGNE =========================
Iteration   |  Residual Norm |  Gradient Norm
         1    1.32567376e-01   9.79450924e-05
         2    1.20665237e-01   1.74745495e-04
         3    1.10067330e-01   1.08682645e-04
         4    1.05383545e-01   7.54519788e-05
         5    1.00764051e-01   6.28481284e-05
         6    9.69513506e-02   6.03487533e-05
         7    9.27119926e-02   4.58884570e-05
         8    8.90481845e-02   4.39115865e-05
         9    8.59621167e-02   4.09871645e-05
        10    8.23227838e-02   4.35815928e-05
        11    7.92605504e-02   4.27014675e-05
        12    7.76133314e-02   5.97550788e-05
        13    7.50623047e-02   4.16568146e-05
        14    7.14619383e-02   3.48134454e-05
        15    6.85999393e-02   3.68092878e-05
        16    6.53088316e-02   3.76127427e-05
        17    6.37752339e-02   3.79335834e-05
        18    6.12660609e-02   3.35733057e-05
        19    5.85174002e-02   2.77393119e-05
        20    5.61746396e-02   2.67642426e-05
-----------------------------------------------
        21    5.39411120e-02   2.79296019e-05
=========================== END CGNE ==========================

 ****************** CGNE summary *****************  
initial residual norm      = 1.32567376e-01
residual norm              = 5.39411120e-02
residual redn              = 4.06895816e-01
initial gradient norm      = 9.79450924e-05
gradient norm              = 2.79296019e-05
gradient redn              = 2.85155714e-01

\end{verbatim}
Evidently convergence is somewhat slower than before: the final data
misfit achieved after 20 iterations is approximately 40\%.

\plot{fine1rebornp0it20r90}{width=\textwidth}{Example 2: Predicted data from 
  model of Figure \ref{fig:fine1bulkinvp0it20}.}

\plot{fine1residrebornp0it20r90}{width=\textwidth}{Example 2: Data residual from 
  model of Figure \ref{fig:fine1bulkinvp0it20}. RMS error is
  approximately 40.7\%.}

\plot{fine1comp}{width=\textwidth}{Example 2: Relative mean square 
  error vs. iteration for CG (red curve) and PCG (blue curve).}

\plot{fine1compnt}{width=\textwidth}{Example 2: Relative mean square 
  error vs. iteration for CG (red curve) and PCG (blue curve), without
  depth tapering in water layer.}

Application of the asymptotic inverse (zero offset section of
\ref{eqn:appinvfs}) yields the image displayed in Figure
\ref{fig:fine1bulkinvp1it0}. This image better relative amplitudes than the 20-iteration LSM image
\ref{fig:fine1bulkinvp0it20}, and is less noisy, but the overall
amplitude is quite small compared to the target bulk modulus
perturbation (Figure \ref{fig:fine1dbulk}). As was true for the lower
frequency example, the good relative amplitudes and reflector
waveforms  indicates that the approximate inverse
zero section \ref{fig:fine1bulkinvp1it0} may be a good search
direction.

\plot{fine1bulkinvp1it0}{width=\textwidth}{Example 2: Approximate inversion or
  true amplitude image = zero offset section of subsurface offset
  approximate inversion. Plotted on same color scale as target bulk
  modulus perturbation in Figure \ref{fig:fine1dbulk}.}

Twenty iterations of preconditioned
CG produces the approximate inversion in Figure
\ref{fig:fine1bulkinvp1it20}. Comparison with Figure
\ref{fig:fine1dbulk} suggests that the amplitudes and waveforms are
roughly correct,
allowing for the spatial averaging influence of band limitation, and the
resolution improved somewhat over that displayed in Figure
\ref{fig:fine1bulkinvp1it0}. More than that, this is a good
inversion: the RMS error is less than 10\%. The output of the RVL CGNE
tool describes the convergence:

\begin{verbatim}
========================== BEGIN PCGNE =========================
Iteration   |  Residual Norm |  Gradient Norm
         1    1.32567376e-01   2.19551253e+00
         2    8.64119902e-02   8.03773761e-01
         3    6.29559085e-02   4.04524982e-01
         4    5.03380559e-02   2.41569892e-01
         5    4.18944918e-02   1.65786415e-01
         6    3.59481946e-02   1.23230457e-01
         7    3.13160010e-02   9.72592682e-02
         8    2.77100820e-02   7.86418393e-02
         9    2.49173287e-02   6.50660247e-02
        10    2.26768292e-02   5.65260053e-02
        11    2.08535902e-02   4.91942577e-02
        12    1.93884857e-02   4.33817171e-02
        13    1.81311555e-02   3.99240106e-02
        14    1.70452520e-02   3.66419256e-02
        15    1.61075015e-02   3.42206433e-02
        16    1.53041314e-02   3.18240337e-02
        17    1.46190701e-02   3.01125068e-02
        18    1.40266838e-02   2.82831118e-02
        19    1.35276383e-02   2.65362691e-02
        20    1.30786821e-02   2.57860348e-02
-----------------------------------------------
        21    1.26670329e-02   2.50269361e-02
=========================== END PCGNE ==========================

 ****************** CGNE summary *****************  
initial residual norm      = 1.32567376e-01
residual norm              = 1.26670329e-02
residual redn              = 9.55516621e-02
initial gradient norm      = 2.19551253e+00
gradient norm              = 2.50269361e-02
gradient redn              = 1.13991313e-02

\end{verbatim}

The predicted data shot 90 is displayed in Figure
\ref{fig:fine1rebornp1it20r90}, plotted on the same scale as Figure
\ref{fig:fine1bornr90} - it is hard to see a difference between the
two. The residual practically disappears (Figure \ref{fig:fine1residrebornp1it20r90}).

\plot{fine1bulkinvp1it20}{width=\textwidth}{Example 2: Estimated bulk modulus
  perturbation from 20 precondiitioned CG iterations. Compare to
  Figure \ref{fig:fine1dbulk}.}

\plot{fine1rebornp1it20r90}{width=\textwidth}{Example 2: Predicted shot record 90 from
  inversion displayed in Figure \ref{fig:fine1bulkinvp1it20} -
  compare with Figure \ref{fig:fine1bornr90}, which shows target
  data plotted on same grey scale, or with next figure, which
  replicates Figure \ref{fig:fine1bornr90} for the reader's convenience.}

\plot{fine1bornr90bis}{width=\textwidth}{Example 2: Replication of Figure \ref{fig:fine1bornr90} for
convenient comparison with previous figure. Note that the only readily
visible difference is the appearance of water column energy in the
resimulated data (Figure \ref{fig:fine1rebornp1it20r90}), absent in
the target data (Figure \ref{fig:fine1bornr90}).}

\plot{fine1residrebornp1it20r90}{width=\textwidth}{Example 2: Residual =
  difference of data displayed in previous two figures, plotted on
  same grey scale.}

These results support the same conclusions as in the preceding
section. The approximate inverse provides an effective
preconditioner, in fact the data residual of the preconditioned loop is
smaller at 4 iterations than the data residual of the
un-preconditioned loop after 20.

\newpage
\section{Example 3: Marmousi Extended LSM}
The setting and parameters for this example are exactly as in the
last,  except that the bulk modulus perturbation is extended in a
subsurface offset interval of [-125 m, 125 m], or 10 grid points on
either side of zero offset. The free surface data is precisely the
same as in the previous example (shot 90 shown in Figure \ref{fig:fine1bornr90bis}).

The asymptotic inverse operator $\iddoFf$ produces the $\delta
\bar{\kappa}$ shown in Figure \ref{fig:fine10bulkinvp1it0}. Recall
from Part 1, Theorem 3, that the a left inverse for the physical
modeling operator may be extracted from $\delta \bar{\kappa} $ by
stacking over subsurface offset. The stack in Figure
\ref{fig:fine10bulkinvp1it0stack} is plotted on the same color scale
as the bulk modulus (Figure \ref{fig:fine1dbulk}.

\plot{fine10bulkinvp1it0}{width=\textwidth}{Example 3: Extended bulk
  modulus estimate, application of asymptotic inverse operator to data
  of Example 2.}

\plot{fine10bulkinvp1it0stack}{width=\textwidth}{Example 3: Stack of extended
  bulk modulus inversion - approximate left inverse of physical
  forward modeling operator.}

\plot{fine10bulkinvp1it0stack0}{width=\textwidth}{Example 3: Zero
  offset section of the extended asymptotic inversion. Note that color
scale differs by factor of 4 from that in Figure
\ref{fig:fine10bulkinvp1it0stack}.}

\plot{fine10bulkinvp1it0g6000}{width=\textwidth}{Example 3: Subsurface
  offset gather at position $x_s=$ 6000 m indicated in Figure
  \ref{fig:fine10bulkinvp1it0}, plotted on same grey scale. Note apparent concentration of energy
within 50 m of $h=0$.}

\plot{fine10bulkinvp1it0g6000rat}{width=\textwidth}{Example 3: Subsurface
  offset gather at position $x_s=$ 6000 m, depth range 2000 m $<z<$
  3200 m, plotted to 1:1 aspect ratio
  and padded with zeroes to left and right to obtain square
  region. Plotted on same grey scale as previous figure. Note that extend of energy spread is approximately one
  apparent wavelength.}

Note that the similarity of FIgures \ref{fig:fine10bulkinvp1it0stack}
and \ref{fig:fine1bulkinvp1it0} - the two differ mostly by scale, with
the latter being an order of magnitude smaller. In fact, the
zero-offset asymptotic inverse shown in Figure \ref{fig:fine1bulkinvp1it0} is
exactly the same as the zero offset section of the extended inversion
(Figure \ref{fig:fine10bulkinvp1it0stack0}. An explanation for the deficient
amplitude of the zero-offset asymptotic inversion is now evident: the
stack has the proper amplitude, but it is clear from Figure
\ref{fig:fine10bulkinvp1it0} that the energy in the image is spread
over a number of offsets, exactly how many depending on depth.

The extent of the spread depends on the ray geometry (and on the surface
offset range, as will become evident by comparison with Example
5). Figure \ref{fig:fine10bulkinvp1it0g6000} shows the subsurface
offset image gather at $x=$ 6000 m: the image energy appears to
concentrate within a few grid cells of $h=$0. To visualize the extent
of this concentraation more clearly, Figure
\ref{fig:fine10bulkinvp1it0g6000rat} displays the part of the same
data between $z=$ 2000 and 3200 m, with enough zero padding on the
sides that the plot has 1:1 aspect ratio. Evidently the energy in this
gather is maximally focused, in view of Heisenberg's principle: the
width of the corridor containing most of the energy is about a
wavelength. It has been speculated that image gather energy would
focus in a Fresnel zone, but that is clearly not correct, at least for
a full aperture example such as this one.

In any case the amplitude of each offset section near zero offset is
more or less the same as any other as one sees from Figure
\ref{fig:fine10bulkinvp1it0g6000}, so the stack multiplies the
amplitude, and by the theory developed in part 1 produces the
``proper'' amplitude. Therefore the amplitudes of the indivitual
offset sections, in particular zero offset, must be a fraction of this
``proper'' amplitude. In the middle part of the gather, the two grid
cells on either side of zero offset contain the bulk of the energy, so
the multiplier would be roughly 5 (in view of the maximal focusing,
this is the same as the number of spatial grid points per wavelength
at peak frequency, which indeed is roughly 5 in this example). The
extent of the energy spread is not constant, and therefore the
amplitude of the zero-offset inversion is not precisely a constant
multiple of the target amplitude, but it varies slowly with depth. So
the asymptotic zero-offset inverse acts as a slowly varying scaling of
the target bulk modulus perturbation hence is usable as as
preconditioner, even though it is not itself an accurate approximate
inversion - as we have seen.

On the other hand, we have shown that \ref{eqn:appinvfs} is an
asymptotic inverse for the extended forward map $\iddoFf$, provided
that most of the energy in the image volume is focused near $h=0$, as
it appears to be in this example. Therefore PCG should converge
rapidly, to a data-fitting extended model. Figure
\ref{fig:fine10bulkinvp1it20} displays the result of 20 PCG
iterations. As was the case for the zero-offset verson of this
example, the data residual is roughly 7.5\%, and the stack (Figure
\ref{fig:fine10bulkinvp1it20stack} closely approximates the target
bulk modulus perturbation (Figure \ref{fig:fine1dbulk}), allowing for
resolution limitation due to limited bandwidth. In fact, it differs
little from the zero-offset PCG result \ref{fig:fine1bulkinvp1it20}
and therefore also generates a data misfit via physical modeling of
less than 10%.

\plot{fine10bulkinvp1it20}{width=\textwidth}{Example 3: Extended bulk
  modulus estimate, estimated from 20 PCG iteratoins from data
  of Example 2.}

\plot{fine10bulkinvp1it20stack}{width=\textwidth}{Example 3: Stack of extended
  bulk modulus inversion from 20 PCG iterations - near-exact left inverse of physical
  forward modeling operator.}

Inversion, rather than asymptotic inversion, also refines the focusing
of energy in the image volume. Figure
\ref{fig:fine10bulkinvp1it20g6000} shows the gather at $x=$ 6000 m
from the volume depicted in Figure \ref{fig:fine10bulkinvp1it20}. The
additional focusing is even more evident in the 1:1 aspect ratio plot
(Figure \ref{fig:fine10bulkinvp1it20g6000rat}), constructed in the same
way as Figure \ref{fig:fine10bulkinvp1it20g6000rat}.

\plot{fine10bulkinvp1it20g6000}{width=\textwidth}{Example 3: Subsurface
  offset gather of 20 PCG iteration inversion at position $x_s=$ 6000 m indicated in Figure
  \ref{fig:fine10bulkinvp1it20}, plotted on same grey scale. Note apparent concentration of energy
within 50 m of $h=0$.}

\plot{fine10bulkinvp1it20g6000rat}{width=\textwidth}{Example 3:
  Subsurface offset gather of 20 PCG iteration inversion at position
  $x_s=$ 6000 m, depth range 2000 m $<z<$ 3200 m, plotted to 1:1
  aspect ratio and padded with zeroes to left and right to obtain
  square region. Plotted on same grey scale as previous figure. Note
  that extent of energy spread is very close to one apparent
  wavelength.}

\newpage
\section{Example 4: Marmousi LSM, Towed Streamer Geometry}
This example differs from the preceding three in that it uses towed
streamer, rather than full aperture, acquisition geometry. The
parameters are as
specified by IFP in the original 1989 data distribution
\cite[]{BoLaVe:91}, with only one change: the shot points are subsampled by a factor of 4, leaving 60
shots with 100 m interval. The output of surange gives the details:
\begin{verbatim}
5760 traces:
tracl    1 22752 (1 - 22752)
tracr    658 23409 (658 - 23409)
fldr     3000 8900 (3000 - 8900)
tracf    1 96 (1 - 96)
cdp      1 568 (1 - 568)
cdpt     1 48 (1 - 1)
trid     1
duse     1
offset   -2575 -200 (-2575 - -200)
gelev    -8
selev    -12
sx       3000 8900 (3000 - 8900)
gx       425 8700 (425 - 8700)
ns       751
dt       4000

Shot coordinate limits:
        North(3000,0) South(3000,0) East(8900,0) West(3000,0)
Receiver coordinate limits:
        North(425,0) South(425,0) East(8700,0) West(425,0)
Midpoint coordinate limits:
        North(1712.5,0) South(1712.5,0) East(8800,0) West(1712.5,0)
\end{verbatim}

The spatial grid is subsampled by a factor of 5 from IFP
model, so that $dz = dx = $ 20 m. 
The density in this example is once again set = 1.0 g/cc. The velocity
is smoothed by by 4 passes of a 6-point rectangular average and
squared to create the background velocity model 
(Figure \ref{fig:marm1bulkbig}). The bulk modulus perturbation is
computed by creating a slightly less smoothed bulk modulus (velocity
smoothed by 2 passes of a 3 point rectangular average, squared) then
subtracting that from the unsmoothed bulk modulus. The perturbation is
displayed in Figure \ref{fig:marm1dbulk}.

\plot{marm1bulkbig}{width=\textwidth}{Example 4: background bulk
  modulus, 20 m grid.}

\plot{marm1dbulk}{width=\textwidth}{Example 4: bulk modulus perturbation.}

The source pulse ($w(t)$ in the
system \ref{eqn:awe}) is a [2.5, 5.0, 15, 20] Hz bandpass filter. The
time integral in $I_t \doF$ is accommodated by trapezoidal rule
integration of this pulse.

Figure \ref{fig:marm1r6000} shows the synthetic at shot position 6000
m.

\plot{marm1r6000}{width=\textwidth}{Example 4: linearized shot gather at $x_s$ =
  6000 m.}

The behaviour of the asymptotic inverse for this non-extended
(zero-offset) example is no different than in previous similar
experiments: it gives a good relative amplitude map, but with small
overall amplitude. It is not in itself a usable inverse, but functions
as a quite successful preconditioner. Ten PCG iterations generate a quite accurate
inversion displayed in Figure \ref{fig:marm1invp1it10}.

The iteration record illustrates the successful convergence
acceleration - we do not show the record for unpreconditioned CG, but
it is considerably slower, as in previous examples.

\begin{verbatim}
========================== BEGIN PCGNE =========================
Iteration   |  Residual Norm |  Gradient Norm
         1    4.29975577e-02   2.53198719e+00
         2    2.50279326e-02   1.15877616e+00
         3    1.54467793e-02   6.48228705e-01
         4    1.08529218e-02   3.65152717e-01
         5    8.42937827e-03   2.28351489e-01
         6    6.79261191e-03   1.60824731e-01
         7    5.62188262e-03   1.16707794e-01
         8    4.78196330e-03   9.19547901e-02
         9    4.19104146e-03   7.66246468e-02
        10    3.71039868e-03   6.24982305e-02
-----------------------------------------------
        11    3.33797187e-03   5.21011576e-02
=========================== END PCGNE ==========================

 ****************** CGNE summary *****************  
initial residual norm      = 4.29975577e-02
residual norm              = 3.33797187e-03
residual redn              = 7.76316598e-02
initial gradient norm      = 2.53198719e+00
gradient norm              = 5.21011576e-02
gradient redn              = 2.05771811e-02

\end{verbatim}

The resimulation at shot location 6000
m is plotted in Figure \ref{fig:marm1rebornp1it10r6000}, and the
residual on the same grey scale in Figure \ref{fig:marm1residp1it10r6000}.

\plot{marm1invp1it10}{width=\textwidth}{Example 4: inverted bulk
  modulus perturbation, 10 PCG iterations.}
\plot{marm1rebornp1it10r6000}{width=\textwidth}{Example 4: resimulation of shot gather at $x_s$ =
  6000 m from inverted bulk modulus perturbation (Figure
  \ref{fig:marm1invp1it10}).}
\plot{marm1residp1it10r6000}{width=\textwidth}{Example 4: data
  residual at $x_s$= 6000 m.}

Evidently the original rather short cable (at least, rather short by
modern standards) produces data more than sufficient for an effective
inversion, provided that the background model is kinematically
consistent with the data.

\newpage

\section{Example 5: Marmousi Extended LSM, Towed Streamer Geometry}
This example bears the same relation to Example 4 that Example 3 did
to Example 2. All parameters are the same, except that the bulk
modulus perturbation is extended, -200 m $< h < $ 200 m, so a total of
21 grid points in $h$ (for computational efficiency, $\Delta h$ =
$\Delta x$).

For this example, we show only the result of PCG inversion with 10
iterations. Figure \ref{fig:marm10freeinvp1it10} shows the extended
inversion for $\delta \bar{\kappa}$, and Figure
\ref{fig:marm10freeinvp1it10stack} the stack (optimal estimate of $\delta
\kappa$). Comparison with Figure \ref{fig:marm1invp1it10} reveals that
the dipping layers on the left of the model are truncated several
hundred meters further to the right in Figure
\ref{fig:marm10freeinvp1it10stack}. This truncation is due to the lack of
energy in image gathers for $h<0$ in the left hand part of the model,
which in turn results from the off-end cable geometry and the slope of
the layer package. Thus image energy available at offset zero in Example
4 is diluted by missing energy in $h<0$ in Example 5, in the left
side of the model.

\plot{marm10freeinvp1it10}{width=\textwidth}{Example 5: PCG extended inversion, 10
  iterations, Marmousi with original streamer geometry, 20 m spatial
  grid, 2.5-5-15-20 Hz bandpass source, free surface, subsurface
  offset axis $-200 < h < 200$ m.}

\plot{marm10freeinvp1it10stack}{width=\textwidth}{Example 5: Stack of data in
  Figure \ref{fig:marm10freeinvp1it10}.}

Figure
\ref{fig:marm10freeinvp1it10g6000} shows the gather at $x=$ 6000 m
from the volume depicted in Figure \ref{fig:marm10freeinvp1it10}. The
 1:1 aspect ratio plot
(Figure \ref{fig:marm10freeinvp1it10g6000rat}), constructed in the same
way as Figure \ref{fig:fine10bulkinvp1it0g6000rat}, indicates that
once again the image energy is mostly confined to a corridor one
wavelength wide, although the wavelength is longer in this example,
corresponding to the lower source freequency band ([2.5,5,15,20] Hz,
as opposed to [2.5,5,25,30] Hz for Example 3).

\plot{marm10freeinvp1it10g6000}{width=\textwidth}{Example 5: Subsurface
  offset gather of 10 PCG iteration inversion at position $x_s=$ 6000 m indicated in Figure
  \ref{fig:marm10freeinvp1it10}, plotted on same grey scale. Note apparent concentration of energy
within 100 m of $h=0$.}

\plot{marm10freeinvp1it10g6000rat}{width=\textwidth}{Example 5: Subsurface
  offset gather of 10 PCG iteration inversion at position $x_s=$ 6000 m, depth range 2000 m $<z<$
  3200 m, plotted to 1:1 aspect ratio
  and padded with zeroes to left and right to obtain square
  region. Plotted on same grey scale as Figure \ref{fig:marm10freeinvp1it10g6000}. Note that extend of energy spread is very close to one
  apparent wavelength.}

%\newpage
%\plot{vg1invp0it0agc}{width=\textwidth}{Example6: RTM image after AGC, 
%Viking Graben data.}

%\plot{vg1invp1it0}{width=\textwidth}{Example6: Asymptotic inverse, 
%Viking Graben data.}

\section{Plots for DL Lecture}

\plot{coarse1unbornr45}{width=\textwidth}{Simulation of shot 45
  (nonlinear)}
\plot{coarse1bornr45c}{width=\textwidth}{Born simulation of shot 45}
%\plot{fine1diffr90}{width=\textwidth}{Difference between target,
%  background simulations of shot 90 (nonlinear)}
%\plot{fine1linerrr90}{width=\textwidth}{Linearization error}

\bibliographystyle{seg}
\bibliography{../../bib/masterref}