### 2019-10-23 repro/papers/trans/paper.tex
update

\title{Waveform Inversion via Source Extension: Overview as of Q1 2019}
\author{William. W. Symes \thanks{The Rice Inversion Project,
Department of Computational and Applied Mathematics, Rice University,
Houston TX 77251-1892 USA, email {\tt symes@caam.rice.edu}.}}

\extrafloats{100}

\lefthead{Symes}

\righthead{Source Extension}

\maketitle
\begin{abstract}
FWI via source extension is a workaround for cycle skipping. For pure transmission (including diving wave) data, a close link to traveltime tomography explains the good results observed in experiments. No such link is known to exist in the case of pure reflection data, and numerical results are mixed. This proposal outlines several numerical experiments that should complement theoretical investigations and decide the effectiveness of these algorithms.
\end{abstract}

\section{Overview}
FWI can be described in terms of a wave operator $L[c]$, depending on an array of space-dependent coefficients $c$, a trace sampling operator $P$, a wavefield $u$, and a source function (of space and time) $f$. The basic FWI problem is: given $d$, find $c$ so that 
\begin{equation}
\label{eqn:fwi}
Pu \approx d \mbox{ and } L[c]u = f.
\end{equation}
Sometimes $f$ is regarded as given, sometimes only some aspects are given (for example, localization at a source point) and others are to be determined as part of the solution. A typical formulation is via nonlinear least squares:
\[
\mbox{choose } c \mbox{ to minimize } \|PL[c]^{-1}f -d \|^2 
\]
usually with some sort of regularization thrown in.

Note that $f$ and $d$ are typically vectors of trace gathers, one for each source position in the survey (or possibly some other parameter, produced by processing - will stick with source position $\bx_s$).

As is well-known, local optimization methods are the only feasible approach given the dimensions of a typical instance of \ref{eqn:fwi}, and those have a tendency to stall due to ``cycle-skipping''. Source extension is one approach to avoiding this problem. It consists in imposing the wave equation as a soft as opposed to hard constraint, by allowing thw source to have more degrees of freedom than is permitted by the basic model. For sake of argument (and because almost all work so far has assumed it), adopt the isotropic point source as the model: for the data gather at source positon $\bx_s$, the source takes the form
\begin{equation}
\label{eqn:ptsrc}
f(\bx,t;\bx_s) = w(t)\delta(\bx-\bx_s).
\end{equation}
Every method explained below can be altered to accommodate more complex source models.

Source extension replaces $f(\bx,t;\bx_s)$ with an artificial source $g(\bx,t; \bx_s)$ with more degrees of freedom than merely a single common source wavelet $w(t)$ as in definition \ref{eqn:ptsrc}. Assuming that \ref{eqn:ptsrc} is actually correct physics, these extra degrees of freedom have to be suppressed in the optimal solution, via an {\em annihilator} $A$, an operator acting on extended sources $g$, whose null space is precisely the physical sources (of the form \ref{eqn:ptsrc}). Extended source inversion replaces the FWI objective with a penalty function
\begin{equation}
\label{eqn:esi}
\mbox{choose } c, g \mbox{ to minimize } \|PL[c]^{-1}g -d \|^2 + \alpha \|A[g]\|^2 
\end{equation}

The way in which degrees of freedom are introduced, and the choice of annihilator, distinguish the various source extension methods.

\subsection{Wavefield Reconstruction Inversion}
Introduced by \cite{LeeuwenHerrmannWRI:13} and significantly modified by \cite{WangYingst:SEG16}. Extended source space: arbitrary functions of space-time $g(\bx,t;\bx_s)$, one for each source position. Annihilator: 
\begin{equation}
\label{eqn:wri}
A[g] = g-f
\end{equation}
With these choices, WRI takes the form introduced by \cite{WangYingst:SEG16}. The original approach is equivalent: \cite{LeeuwenHerrmannWRI:13} use the dyamic wavefield $u$ as the additional degrees of freedom, and minimize
\[
\|Pu-d\|^2 + \alpha \|L[c]u-f\|^2
\]
Since the wave equation (any reasonable version!) has a unique solution for each right-hand side, replacing $L[c]u$ turns this original form of WRI into the form \ref{eqn:esi}.

Note that in princple at least, the actual source $f$ (positions and wavelet) must be known {\em a priori}.

The Wang-Yingst variant leads to a huge reduction in the number of degrees of freedom: either the original or the modified versions, as originally presented, appear to require storage of an entire space-time source, which is prohibitive for 3D and/or time domain formulation. I will explain in an appendix how the Wang-Yingst approach drastically shrinks the storage requirements of this method, and makes it feasible for time-domain 3D application, which Wang and Yingst and their co-authors have demonstrated.

\subsection{Space-Time Extension}
\cite{HuangSymes:SEG16b}. Differs from WRI by choice of annihilator:
\begin{equation}
\label{eqn:msste}
A[g](\bx,t;\bx_s) = |\bx-\bx_s|g(\bx,t;\bx_s).
\end{equation}
This annihilator merely enforces the position constraint on the source, without relying on {\em a priori} knowledge of the source wavelet. In fact, the source wavelet is a by-product of the minimization \ref{eqn:esi}.

Appears to require storage of full space-time source volume - the trick explained in the appendix for WRI does not work, due the form of the annihilator.

\subsection{Volume Extension}
\cite{HuangSymes:SEG16a}. Assumes that source signature deconvolution has been performed, so that the actual wavelet is (a bandlimited version of) $\delta(t)$. The extended source is an {\em exploding reflector}:
\begin{equation}
\label{eqn:msvol}
g(\bx,t;\bx_s) = h(\bx;\bx_s)\delta(t).
\end{equation}
Uses the same annihilator as the Space-Time Extension, but loses the advantage of solving for the source as by-product - the source must be known, so that decon can be carried out prior to inversion.

However, only requires a spatial storage volume per source position, essentially equivalent to a second data volume, therefore practical for 3D (though I don't believe it has been tried except in 2D).

The variable projection method applied to this extension has essentially exploding reflector inversion (as opposed to exploding reflector migration) as its inner problem, i.e. estimation of $h$ in \ref{eqn:msvol}. The same problem arises in so-called photoacoustic tomography, and has received a lot of attention in the mathematical inverse problems and biomedical imaging literature. In fact there is a quite accurate approximate inverse available for the inner problem, consisting of converting pressure traces to normal velocity traces and back-projecting (this is known as ``time reversal'') in the IP literature).

A disadvantage of this approach, if you want to handle reflections, is that unlike post-stack migration the actual physical wave velocity is used in modeling, so that the the region in which $h$ must be defined may considerably larger than the ordinary modeling domain - reflection two-way time is effectively modeled by one-way time.

The volume extension concept is similar to an idea discussed by \cite{ZhangGao:08}.

\subsection{Surface Extension}
No publication yet - subject of my workshop talk at the 2016 SEG, and GH has described it in talks at Total. Similar to the volume extension, but the source is spread over the source space-time surface, rather than over space at $t=0$. In the simplest case, where the sources all lie on $z=z_s$, the extended source space consists of functions (distributions) of the form ($\bx=(x,y,z)$)
\begin{equation}
\label{eqn:mssur}
g(\bx,t;\bx_s) = h(x,y,t)\delta(z-z_s).
\end{equation}
As for the previous two cases, the annihilator is given by the localization penalty \ref{eqn:msste}. Therefore the source wavelet is a by-product of the inversion, and does not need to be specified {\em a priori}. Like the volume extension, the additional data volume required is on the same order of size as the data.

\subsection{Adaptive Waveform Inversion = Source-Receiver Extension}
 Popularized by \cite{Warner:14}, \cite{Warner:16}, has many antecedents, listed in \cite[]{HuangSymes:17}. Each trace regarded as separate gather, with separate point source waveform. In effect,
\begin{equation}
\label{eqn:awi}
g(\bx,t;\bx_s) = w(t;\bx_s)\delta(\bx-\bx_s) 
\end{equation}
with repeated occurances of $\bx_s$ regarded as distinct. Usually written by making the wavelet a function of the receiver position also, to distinguish redundant source positions:
\begin{equation}
\label{eqn:awi-alt}
g(\bx,t;\bx_r,\bx_s) = w(t;\bx_r,\bx_s)\delta(\bx-\bx_s) 
\end{equation}
hence the alternate name.

Various annihilators have been suggested - see references. Warner proposes a variant (``AWI'')  that (in principle) requires the source wavelet to be known, and deconvolved from the data traces. Then uses multiplication by $t$, which annihilates $\delta(t)$:
\begin{equation}
\label{eqn:awi-ann}
A[g](\bx,t;\bx_r,\bx_s) = t g (\bx,t;\bx_r,\bx_s)
\end{equation}
AWI distinguished from earlier variants by normalizing the objective with the estimated wavelet $L^2$ norm:
\begin{equation}
\label{eqn:awi-obj}
\mbox{choose } c, g \mbox{ to minimize } \sum_{\bx_r,\bx_s}\frac{\int dt\,|t|^2 |w(t;\bx_r,\bx_s)|^2}{\int dt\,|w(t;\bx_r,\bx_s)|^2}
\end{equation}
Note the order of summation and normalization: each trace is normalized independently (this is not clear from the notation in \cite[]{Warner:16}).

The justification for normalization given by \cite{Warner:16} is that it avoids a familar source/reflector ambiguity. Actually it is more important than that, and plays an important role when no reflectors are present. It is possible to recast the objective \ref{eqn:awi-obj} as equivalent to \ref{eqn:esi} with a weighted norm used in the latter. [I am writing this up - take my word for it!] The weighted norm makes the linear map from source to data {\em unitary} (at least to leading order in frequency), and that is an essential property in relating AWI to traveltime tomography.

\section{What is known}
For several of the source extension algorithms, the objective is known to be closely linked to traveltime tomography in one form or another, and to inherit its convergence properties. These connections are known {\em only} for pure transmission, and not always then. Nothing theoretical is known about any of these methdods for reflection data.

The connection to traveltime tomography was first (so far as I know) established for the source-receiver extension in several TRIP technical reports related to Hua Song's PhD thesis \cite[]{Song:94c,SoSy:92,SongSymes:94a,SongSymes:94b,Symes:94c}. This work established a form of differential semblance velocity estimation for crosswell tomography, and the related tomographic objective was a form of slope, or stereo-, tomography. The time moment objective \ref{eqn:awi-obj} (without normalization) first appears in \cite[]{Plessix:00a,Plessix:00}. A similar idea appears in \cite[]{LuoSava:11}. Neither of these refferences made an explicit connection to tomography - that was done in \cite[]{HuangSymes2015SEG,HuangSymes:17}.

The defect, in the relation between AWI/source-receiver extention and traveltime tomography, is that it collapses when the ray field (must be able to speak of rays, to invoke traveltime tomography) develops caustics. This was already demonstrated both theoretically and by numerical example in \cite[]{Symes:94c}, though the fact does not seem to have percolated very far. \cite{HuangSymes:17} give an explicit illustration of this failure in a diving wave setting. As soon as the diving wave field is triplicated, AWI is just as bad as FWI.

On the other end of the list, almost nothing is known about why WRI converges, or when it does. It may be possible to repeat a version of the analysis mentioned next for the volume and surface extensions, based on geometric optics and the Wang-Yingst variant as described in the appendix, though that has not been tried yet.

The volume and surface extensions have a well-defined relation to tomographic principles (slope or traveltime, depending on type of annihilator). This relation holds regardless of ray geometry (with or without triplication) in transmission configuration.

Again, nothing is known theoretically about the behaviour of any of these methods for reflection data, and numerical evidence is ambiguous.

%\input{prop.tex}

\bibliographystyle{seg}
\bibliography{../../bib/masterref}

\input{append1.tex}
%\input{append2.tex}
%\input{append3.tex}
%\input{talkfigs.tex}
\end{document}