### 2019-12-12 repro/papers/sserefl/paper.tex
many edits

\title{Surface Source Extension, Reflection Case}
\author{William. W. Symes \thanks{The Rice Inversion Project,
Department of Computational and Applied Mathematics, Rice University,
Houston TX 77251-1892 USA, email {\tt symes@caam.rice.edu}.}}

\lefthead{Symes}

\righthead{Reflection SSE}

\maketitle
\begin{abstract}
FWI via surface source extension (SSE) applies equally well to reflection or transmission data. In the transmission case, it is asymptotically equivalent to a form of travel time tomography, fitting event slowness rather than arrival time. Use of this method for reflection data simply involves permitting reflectors in the velocity and other subsurface mechanical parameters. It is possible to analyze a simple layered medium special case in the plane wave domain. The analysis suggests that in fact reflection SSE cannot constructively update velocity in absence of correctly positioned reflectors, so more closely resembles so-called Reflection FWI than transmission SSE in its velocity updating capabilities. This theoretical result is supported by numerical experiments with 2D reflection data.
\end{abstract}

\section{Introduction}
FWI based on source extension has been around since the early 1990's,
and some versions are provably robust - converge from substantially
incorrect initial velocity models. However the theoretical robustness
results all pertain to purely transmission configurations, such as
crosswell or diving wave geometry. It has been an open question
whether source extended FWI remains robust if all kinematic
information in the data is in the form of reflection
moveout. Some numerical evidence for such robustness has appeared, but
in each case it is not obvious that post-critical events did not play
a role in the inversion.

The goal of this paper is to produce a proof that in one precisely
defined case and in a precisely defined sense, one particular variant
of source-extended FWI is {\em not} robust for inversion of reflection
data. This counterexample negates any general claim of robustness for
such methods of reflection inversion, of course.

Surface source extension, defined below, is the method investigated
here. The objective of the reasoning laid out in the following sections is
to explicitly calculate the gradient of a particular surface source
extended FWI objective. A context that permits such explicit
calculation is plane-wave propagation in a layered medium. An explicit
computation of the gradient at a homogeneous background medium with
data containing a single reflector shows that the gradient has
vanishing low-frequency components as the central data frequency goes
to infinity: in fact, the mean square in any fixed frequency band goes
to zero. Therefore a step in the negative gradient direction cannot
affect traveltimes significantly and therefore this technique cannot
produce a productive velocity update in this case.

It is important to note the contrast with other extension methods. The
surface source extended FWI objective in the closest analogue for
transmission, namely the estimation of a homogeneous velocity from one
or more transmission traces, has been shown to be locally convex
independent of data frequency content. Also, the subsurface offset
extended FWI objective for surface reflection data has been shown to
have similar properties.

The first of the following sections sets out the framework of this
study. It does {\em not} include mathematically complete definitions
or fully qualified statements of results - these may be found in their
entirety in later sections. Instead, it is intended to introduce the
concepts that figure in the main results, which are stated succinctly
in the third section. The  I will review plane-wave modeling, and construct
an asymptotic approximation to single-reflector plane wave data. In
the final section I will present the surface source extended FWI
objective and compute its gradient at a homogeneous background with
single reflector data. The main result will follow immediately from
this calculation.

I include several appendices, on 1D wave propagation and plane wave
reflection modeling, as this material is difficult to find in
sufficiently precise form.

Whenever convenient, I abbreviate ``surface source extended FWI'' as SSE-FWI.

\section{Preliminaries}

This paper will use linear acoustics in velocity-pressure form as its
model for wave motion. In this model, the space-time fields $p$ is excess
pressure, $\bv$ is particle velocity about equilibrium, $f$
represents an energy source as a defect in the acoustic constitutive
law. The bulk modulus $\kappa$ and density $\rho$ are positive functions of
spatial position. The acoustic wave system couples these quantities:
\begin{eqnarray}
\label{eqn:awe}
\frac{\partial p}{\partial t} + \kappa \nabla \cdot {\bf v} &=&
                                                                f \nonumber\\
\rho \frac{\partial {\bf v}}{\partial t} + \nabla p = 0 \nonumber\\
p,{\bf v}&=&0, t \ll 0
\end{eqnarray}
In terms of the solution of the system \ref{eqn:awe}, the modeling
operator is defined by
\begin{equation}
\label{eqn:fwd}
S[\kappa]f = p|_{z=z_r}
\end{equation}
Two special types of right-hand side are important in the discussion
to follow: isotropic point sources, of the form
\begin{equation}
  \label{eqn:ipr}
  f(\bx,t;\bx_s) = w(t)\delta(\bx-\bx_s).  
\end{equation}
and surface sources, of the form
\begin{equation}
  \label{eqn:ss}
  f(\bx,t;\bx_s) = u(x,y,t;x_s,y_s)\delta(z-z_s). 
\end{equation}
Point sources are a special case of surface sources. Introduce appropriately
parametrized versions of the modeling operator for these two cases:
\begin{eqnarray}
 f(\bx,t;\bx_s) = w(t)\delta(\bx-\bx_s) & \Rightarrow & S[\kappa]f = S_p
  [\kappa]w\\ \nonumber
  \label{eqn:specsrc}
f(\bx,t;\bx_s) = u(x,y,t;x_s,y_s)\delta(z-z_s) & \Rightarrow & S[\kappa]f =
  S_s[\kappa]u
\end{eqnarray}

The standing assumption in this work is that observed data to be fit
is to good approximation the result of a point source. That is, there
are bulk modulus $\kappa_*(\bx)$ and source wavelet or pulse $w_*(t)$
so that
\begin{equation}
  \label{eqn:datadef}
  d = S_p[\kappa_*]w_*
\end{equation}
In equation \ref{eqn:datadef}, $d$ is a function of receiver position
$\bx_r$, source position $\bx_s$, and time $t$.

Full waveform inversion (FWI), as construed in this paper, means:
\begin{equation}
  \label{eqn:deffwi}
  \mbox{Given } d, \mbox{ find }\kappa \mbox{ and }w \mbox{ so that }d
  \approx S_p[\kappa]w.
\end{equation}
In the following paragraphs, I will discuss four approaches to FWI:

\subsection{``Straight'' FWI}

The standard approach to FWI is to formulate the task \ref{eqn:deffwi}
as a nonlinear least squares problem:
\begin{equation}
  \label{eqn:defols}
  \mbox{Given } d, \mbox{ find }\kappa \mbox{ and }w \mbox{ to
    minimize }
  \| d -  S_p[\kappa]w\|^2.
\end{equation}
Here the symbol $\|\cdot\|^2$ signifies {\em mean-square}, that is,
sum the squares of every value of the field between the vertical bars.

This concept was introduced in the 1980's
(\cite{BamChavLai:79,Tara:84a,KolbColLai:86,Crasetal:90},
and many other papers since then). Usually some form of regularization
is applied, to compensate for poorly determined aspects of $\kappa$
and/or $w$, as is explained in Tarantola's influential book
\cite[]{Tarantola:05}. The approach was of only academic interest
until roughly fifteen years ago, when 3D modeling and related computations
first became feasible at industrial scale, and evidence began to emerge 
that FWI offered potentially decisive improvements in imaging accuracy
over other inversion (imaging) methods, especially for construction of
large-scale earth model features
\cite[]{Plessix:10,Vigh:10}. In the last ten years, FWI has become
both a production service offered by the geophysical industry, and a
leading topic of both industry-oriented and academic research.

As was understood very early in the evolution of FWI, the basic
least-squares formulation \ref{eqn:defols} suffers from a major
defect: the parameter spaces tend to be so large that only iterative
optimization (that is, a variant
of Newton's method) may feasibly be applied, due to computational cost
- this remains true today and for the foreseeable future - but these
method tend not to produce usable results unless the initial iterate
is in some sense ``good'' (predicts travel times to within a
half-wavelength) \cite[]{GauTarVir:86,Plessix:10}.``Good'' initial models are not always
available. Accordingly, a number of alternative approaches have been
proposed, many of them variations on the least squares theme, amongst
which are the three approaches described below.

\subsection{Source Extensions}

Careful analysis of the defect mentioned in the last paragraph shows
that the failure of iterative optimization applied to \ref{eqn:defols}
to converge (or at least to converge usefully quickly) appears to
arise from failure to fit the data: some parts of the oscillatory time
series $d(\cdot,t)$ overlap oscillations in $S_p[\kappa]w$, while
others do not. It seems reasonable to guess that this ``cycle skipping'' phenomenon 
could be avoided if the data could be fit throughout the inversion
process. That requirement in turn implies a modeling operator with
more degrees of freedom, that is, a bigger domain, than the basic
modeling operator $S_p$, which should produce the same results as
$S_p$ when applied to pairs of bulk moduli and isotropic point sources \ref{eqn:ipr}.

There is
well-established mathematical terminology for this relationship:
operator A is an {\em extension} of operator B if (i) the domain of $A$
contains the domain of $B$, and whenever $v$ is in the domain of $B$,
$Av = Bv$. Two extensions of $S_p$ already appear
above: the isotropic point radiator \ref{eqn:ipr} is a special case of
both the surface source \ref{eqn:ss}, which is in turn a special case
of the generic right-hand side in the system \ref{eqn:awe}. So the
domain of $S_s$ contains the domain of $S_p$, as does the domain of
$S$. From the relations \ref{eqn:specsrc}, $S$ and $S_s$ produce the
same result as $S_p$ when applied to sources in their common domain,
so these operators are both extensions of $S_p$.

Since these extensions of $S_p$ involve additional degrees of freedom
for the energy source in the system \ref{eqn:awe}, they have come to
be called {\em source extensions}. These are not the only source
extensions - see \cite{HuangNammourSymesDollizal:SEG19} for an
overview of source extensions and their properties. Extensions may
also be constructed by adding degrees of freedom to the coefficients
in the wave system (bulk modulus, in the present case), 
rather than the right hand side - see \cite{geoprosp:2008} for a
description of some of these {\em medium extensions}.

The purpose of any extension is to add options for data fitting,
thereby reducing the chances of cycle-skipping. As will be seen below,
both extension $S$ and $S_s$ permit arbitrarily accurate fitting of
essentially any data with essentially any bulk modulus field $\kappa$,
with proper choice of $f$ or $w$ respectively. However, the standing
assumption on the data $d$ is that it derives from a point source, per
\ref{eqn:datadef}. So the added degrees of freedom represented by the
extensions $S$ and $S_s$ must be suppressed somehow. As explained by
\cite{geoprosp:2008}, a convenient mechanism for suppression of the
extension degrees of freedom is a penalty term added to the objective
of the extended
version of the least squares problem \ref{eqn:defols}. Choice of penalty
term, along with choice of extension, characterizes each of the extended
inversion methods detailed in the next few paragraphs.

\subsubsection{Wavefield Reconstruction Inversion}
This concept was introduced by \cite{LeeuwenHerrmannWRI:13}, and further
developed by \cite{LeeuwenHerrmann:16,WangYingst:SEG16} and other
authors. It is based on the extension $S$ (equation \ref{eqn:fwd}), with full
space-time volume source $f$, and presumes that the correct point
source wavelet $w_*$ is known. 
\begin{equation}
  \label{eqn:defwri}
  \mbox{Given } d, \mbox{ find }\kappa \mbox{ and }f \mbox{ to
    minimize }
  \| d -  S[\kappa]f\|^2+\lambda^2\|f-w_*\delta(\cdot-\bx_s)\|^2.
\end{equation}
Evidently the difference $f-w_*\delta(\cdot-\bx_s)$ should be such
that the scaled sum-of-squares makes sense - this is always the case
for discretizations, but for the continuum limit entails an actual
assumption.

\cite{LeeuwenHerrmannWRI:13} posed this problem slightly
differently, in terms of the second order wave equation for the
pressure wavefield $p$. From this viewpoint, the 
second terms becomes a scaled mean square of the residual in the
wave equation. The formulation presented here is equivalent, and was
introduced by \cite{WangYingst:SEG16}. \cite{LeeuwenHerrmann:16} used
the variable projection method \cite[]{GolubPereyra:03}, eliminating the source $f$ in the
inner step and updating the bulk modulus (or an equivalent quantity)
in the outer step. In any penalty method, control of the penalty
parameter has a large influence on the speed of
convergence. \cite{Aghamiry:19} use an augmented Lagrangian algorithm
to minimize the influence of the penalty weight choice.

The Achilles' heel of Wavefield Reconstruction Inversion (WRI), in any of
the formulations described so far, is the necessity of storing a full
space time field, namely $f$ (or $p$). Especially for time-domain
formulation and high frequency waves, this storage requirement can be
formidable. In fact, ``most'' source fields $f$ are non-radiating,
that is, $S[\kappa]f=0$. The orthocomplement of this subspace of
non-radiating sources is the range of the adjoint operator
$S[\kappa]^T$. Therefore the sources may be re-parametrized as images of
$S[\kappa]^T$. Write $f = g + w_*\delta(\cdot-\bx_s)$: then the second term in
the objective displayed in \ref{eqn:defwri} is simply $\|g\|^2$.
Decompose $g = S[\kappa]^Te + n$, in
which $e$ is the same type of object as $d$ and $S[\kappa]n=0$ (that
is, $n$ is a non-radiating
source), and note that the decomposition is orthogonal. Then the WRI problem \ref{eqn:defwri} may be reformulated as
\begin{equation}
  \label{eqn:defwrireform}
  \mbox{Given } d, \mbox{ find }\kappa \mbox{ and }e \mbox{ to
    minimize }
  \| d -  S[\kappa]w_*\delta(\cdot-\bx_s) - S[\kappa]S[\kappa]^Te\|^2+\lambda^2\|S[\kappa]^Te\|^2.
\end{equation}
Note that $n$ has disappeared: {\em a priori} it does not contribute to the first
term, and $n \ne 0$ makes the second term larger. So minimization
necessarily entails the choice $n=0$.

This reformulation does {\em not} require storage of $d+1$-dimensional
fields, as does the original WRI formulation \ref{eqn:defwri}. The
field $e$ has taken the place of the space-time source $f$, and its
size is precisely the same as that of the data field $d$. Therefore
the extra memory cost over modeling is the same as another copy of
data. A close examination reveals that in computing the composition
$SS^T$ via time-stepping methods, the same simultaneity issue arises
as in all applications of the adjoint state method. However good
solutions are known \cite[]{Griewank:book}, and since precisely the
same issue arises in time-domain FWI, the cost is the same. Therefore,
it is this reformulation of WRI that will be analyzed below.

\subsubsection{Space-Time Extension}
\cite{HuangSymes:Geo18a} introduced a variant of WRI, also using the space-time extension, but not
requiring knowledge of the target point source wavelet $w_*$: in fact,
an estimate of this wavelet is a by-product of the algorithm. This
approach relies on an {\em annihilator}, a linear operator whose null
space is precisely the desired class of source models, in this case
isotropic point radiators. Such an operator can be used to generate a
quadratic form whose zero set is also the desired class, hence can be
used as a penalty in a nonlinear least squares objective, designed to
drive the source towards to the desired type. See \cite{geoprosp:2008}
for a discussion of annihilators in extended inversion.

The annihilator proposed by \cite{HuangSymes:Geo18a} is multiplication
by $|\bx-\bx_s|$, that is,
\begin{equation}
  \label{eqn:defann}
  Af(\bx,t;\bx_s) = |\bx-\bx_s|f(\bx,t;\bx_s).
\end{equation}
The space-time extension approach consists of solving the optimization
problem:
\begin{equation}
  \label{eqn:defste}
  \mbox{Given } d, \mbox{ find }\kappa \mbox{ and }f \mbox{ to
    minimize }
  \| d -  S[\kappa]f\|^2+\lambda^2\|Af\|^2.
\end{equation}
This problem can be reformulated in the same way as the WRI problem. Write
$f = w\delta(\cdot-\bx_s) + S[\kappa]^Te + n$, where $n$ is
non-radiating as before: then the problem \ref{eqn:defste} is
equivalent to
\[
   \mbox{Given } d, \mbox{ find }\kappa, w, e \mbox{ and }n \mbox{ to
     minimize }
 \]
\begin{equation}
  \label{eqn:defstereform}
  \| d -  S[\kappa]w_*\delta(\cdot-\bx_s) - S[\kappa]S[\kappa]^Te\|^2+\lambda^2\|A(S[\kappa]^Te+n)\|^2.
\end{equation}
Note that the non-radiating component $n$ must remain in the second
term of the objective, because there is no reason for $n$ to be
$A^TA$-orthogonal to the range of $S[\kappa]^T$. Also, the isotropic
point radiator component $w\delta(\cdot-\bx_s)$ must also remain as
part of the unknowns - simply because it is unknown! It does not
appear in the second term because $A$ annihilates it.

\subsubsection{Surface Source Extension}
Beside the reformulation of WRI described above, it is also possible
to avoid the storage problem of WRI by limiting the support of the
extended source $f$ to lower dimensional sets. Of several
possibilities discussed in the literature, I will discuss the 
  surface source extension approach introduced by
\cite{HuangNammourSymesDollizal:SEG19}. The extension in question is
defined in equation \ref{eqn:defss}, and denoted $S_s$. The same
annihilator \ref{eqn:defann} is used in this approach, this time
interpreted as acting in the plane $z=z_s$, that is,
\[
  Au(x,y,t;x_s,y_s) = \sqrt{(x-x_s)^2+(y-y_s)^2}u(x,y,t;x_s,y_s).
\]
The problem formulation is
\begin{equation}
  \label{eqn:defsse}
  \mbox{Given } d, \mbox{ find }\kappa \mbox{ and }u \mbox{ to
    minimize }
  \| d -  S_s[\kappa]u\|^2+\lambda^2\|Au\|^2.
\end{equation}
Evidently the extended source has as many parameters as the data, so
the memory complexity of this problem is comparable to that of FWI or
reformulated WRI. The computational complexity compares favorably to
that of reformulated WRI.

\subsection{Layered Media, Plane Waves}
As explained in the introduction, this study explores the properties
of the various FWI variants explained above for the special case of
layered media: $\kappa = \kappa(z)$,
$\rho=\rho(z)$. Then the modeled shots differ from
each other by a horizontal translation, so that there is in effect
only one shot, for which I take the location to be
$\bx_s=(0,0,0)$ and suppress $\bx_s$ from the notation.

If $\log \kappa$ and $\log \rho$ are globally bounded and $f$ has
compact support (and some other properties that will be made precise
later), there is $\xi_{\rm max} > 0$ so that for
$\|\xi\|<\xi_{\rm max}$, the integrals defining slant stack fields
$\tilde{p},\tilde{\bv}$, namely
\begin{eqnarray}
\label{eqn:slant}
\tp(\xi,z,t)& = &\int dx\, dy\, p(x,y,z,t+\xi x),\nonumber\\
\tv(\xi,z,t)&=&\int dx\, dy\, v_z(x,y,z,t+\xi x),\nonumber\\
\tf(\xi,z,t)&=&\int dx\,dy\, f(x,y,z,t+\xi x),
\end{eqnarray}
are absolutely convergent, and $\tp, \tv$ satisfy the plane wave
version of the acoustic system \ref{eqn:awe}:
\begin{eqnarray}
\label{eqn:awepw}
(1-c^2\xi^2)\frac{\partial \tp}{\partial t} + \kappa\frac{\partial
  \tv}{\partial z} &=& \tf \nonumber\\
\rho \frac{\partial \tv}{\partial t} + \frac{\partial \tp}{\partial
  z}&=&0\nonumber\\
\tp,\tv&=&0, t \ll 0.
\end{eqnarray}
Here $c=\sqrt{\kappa/\rho}$ is the sound velocity. I use $\xi$ instead
of the more usual $p$ for plane wave slowness since $p$ already has
been assigned to mean pressure. Note that for a isotropic point
source, respectively a surface source,
\begin{eqnarray}
  f(\bx,t;\bx_s)=w(t)\delta(\bx) & \Rightarrow &
                                               \tf(\xi,z,t)=w(t)\delta(z)\\ \nonumber 
  f(\bx,t;\bx_s) = u(x,y,t)\delta(z) & \Rightarrow &
                                                     \tf(\xi,z,t) =
                                                     \tu(\xi,t)\delta(z)\\
  \nonumber
  \tu(\xi,t) &=& \int dx\,dy\, u(x,y,t+\xi x)
\end{eqnarray}
The plane wave domain modeling
operator for space-time extended sources samples $\tp$ on $z=z_r$:
\begin{equation}
\label{eqn:modpw}
\tS[\kappa]\tf = \tp|_{z=z_r}
\end{equation}
Similarly, the plane wave domain versions of $S_p$ and $S_s$ are
\begin{eqnarray}
 \tf(\xi,z,t) = w(t)\delta(z) & \Rightarrow & \tS[\kappa]\tf = \tS_p
  [\kappa]w\\ \nonumber
  \label{eqn:specsrc}
\tf(\xi,z,t) = \tu(\xi,t)\delta(z-z_s) & \Rightarrow & \tS[\kappa]\tf =
  \tS_s[\kappa]\tu
\end{eqnarray}

For homogeneous $\kappa=\kappa_0, \rho=\rho_0, c=c_0$, a computation
recapitulated in the Appendix C, relying on computations for the 1D
acoustics system detailed in Appendices A and B, yields
\begin{equation}
\label{eqn:modhom}
\tS_s[\kappa_0]\tu(\xi,t) = \frac{1}{2(1-c_0^2\xi^2)}\tu
\left(\xi,t-z_r\sqrt{\frac{1}{c_0^2}-\xi^2}\right).
\end{equation}
That is, at a homogeneous model, the plane wave modeling operator is
simply scaling by a $\xi$-dependent factor.

\section{Main Results}

The space dimension is not terribly important in the results to
follow. Generally I will assume it to take the physical value $d=3$,
though the numerical examples will use $d=2$ for reasons of
computational cost. For spatial position, write $\bx = (x,y,z)$ (or $\bx = (x,z)$ for
2D).

\cite{BlazekStolkSymes:13} give conditions under which the system
\ref{eqn:awe} has a strong solution, that is, in which all of the
terms in equations above are interpretable as functions of space-time. Assume that

\indent Assumption A: $\log \kappa, \log \rho$ are bounded and
measureable. \\

\indent Assumption B: $f$ is square-integrable, has a square-integrable time
  derivative, and \\has compact support.\\

Then Theorem 1 in \cite{BlazekStolkSymes:13} implies that there exists
a unique causal (vanishing for $t \ll 0$) strong solution
$(p,\bv)$ of the system \ref{eqn:awe}, that is, $p$ and $\bv$ are
$C^1$ functions of time with square-integrable (in space) values,
$\nabla p$ and $\nabla \cdot \bv$ are $C^0$ functions of time with
square-integrable values,  and the equations \ref{eqn:awe} are
satisfied pointwise a. e.

The central object in this study is the relation between the
coefficients $\kappa$ and $\rho$, on the one hand, and the fields $p$
and $\bv$ on the other. The influence of the wave velocity $c$ is the
most important, so from here on $\rho$ will be fixed and constant:
$\rho(\bx) = \rho_0 \in \bR^+$. Thus view $p$ and $\bv$ as functions of
$\kappa$ and $f$.

In seismic and ultrasonic experiments, the acoustic
fields are recorded at or near a hypersurface (the surface of the
earth, or the skin, or...). Theorem 2 in \cite[]{BlazekStolkSymes:13}
assures that the trace of $p$ on a cylindrical hypersurface $\Gamma
\times [0,T]$ is well-defined, as is the trace of the normal component
of $\bv$. These traces will serve as idealized data in the discussion
to follow, ignoring issues of sampling and other details of actual
measurement processes.

This discussion further idealizes acquisition by setting $\Gamma =
\{\bx: z=z_r\}$. This is not a bad approximation for some field
settings, and is convenient for the eventual use of the Radon
transform.

Active (artificial) sources (right-hand sides in the wave equations)
are frequently localized in space, relative to other active length scales. The
simplest such localized source is the {\em isotropic radiator},
\begin{equation}
  \label{eqn:ipr}
  f(\bx,t;\bx_s) = w(t)\delta(\bx-\bx_s).  
\end{equation} 
in which $\bx_s$ is the location of the radiator. The radiator is a
special case of the {\em surface source}, in which the support of the
right-hand side in \ref{eqn:awe} is confined to a hypersurface. For
present purposes, it suffices to consider sources localized on the
hyperplane $z=z_s$:
\begin{equation}
  \label{eqn:ss}
  f(\bx,t;\bx_s) = w(x,y,t)\delta(z-z_s). 
\end{equation}
Neither distribution defined by \ref{eqn:ipr}, \ref{eqn:ss} is 
square-integrable, so the theory in \cite[]{BlazekStolkSymes:13} does
not apply directly to the system \ref{eqn:awe} with $f$ defined by
equation \ref{eqn:ipr}. However a simple trick produces a unique
(necessarily weak) solution in both cases, provided an additional
constraint is placed on $\kappa$:

\indent Assumption C: There exists $\epsilon >0$ so that $\kappa =
\kappa_0 \in \bR^+$ in $\{\bx: |z-z_s|<\epsilon\}$.

\noindent That is, $\kappa$ is constant near the plane $z=z_s$. See
Appendix \appSingular for a proof of this statement.

Define the coefficient-to-data map (``forward map'', ``modeling
operator'') $S$ by
\begin{equation}
\label{eqn:fwd}
S[\kappa]f = p|_{z=z_r}
\end{equation}
The notation reminds the reader that $S$ is linear in $f$ but
nonlinear in $\kappa$.

Theorem 2 in \cite[]{BlazekStolkSymes:13} assures that $S[\kappa]$ is
well-defined as a map from right-hand sides $f$ having
square-integrable time derivatives to square-integrable traces. With
more regularity in time, the map is also differentiable in its
dependence on $\kappa$. In Appendix \ref{appSingular}, this result is
extended to the singular sources \ref{eqn:ipr}, \ref{eqn:ss}.

In fact, stronger results hold under Assumption C. The special cases
considered below have stronger continuity properties.

After Radon transform, this source takes the form
$\tw(\xi,t)\delta(z)$, where $\xi$ is the plane wave slowness.

The underlying hypothesis is that the data comes from a point
source. Extension methods encode such information in an annihilator,
an operator that vanishes for models that satisfy the hypothesis. For
SSE-FWI, a natural annilator is a variant on the operator introduced
for this purpose by \cite{HuangSymes:Geo18a,HuangSymes:Geo18b} for two
similar source extension methods, namely multiplication by offset,
also reminiscent of a now-standard annihilator for subsurface offset
introduced by \cite{stolk2001c}. In the plane wave domain,
multiplication by offset becomes 
\[
\tA = D_{\xi}I_t,
\]
that is, differentiation in $\xi$ and integration in $t$, in either
order.

The forward map $\tS$ for plane-wave transformed inverse problem maps
extended plane wave source wavelets $\tw(\xi,t)$ to plane wave data
traces (sampling of the plane-wave field at $z=0$, assuming this to be the
depth of sources and receivers, as I do here). I will regard this operator as a
function of the bulk modulus $\kappa(z)$, by assuming the density
$\rho$ to be fixed and homogeneous.

$\tS[\kappa]$ is 
actually invertible, at least for $\kappa$ close to homogeneous. This conclusion follows for instance
from explicit expressions in two special cases: for homogeneous models
$\kappa=\kappa_0$, 
\[
\tS[\kappa_0]\tw(\xi,t) = a_0(\xi)\tw(\xi,t),
\]
and for a smooth model $\kappa_1$ with an embedded reflector at $z=z_1$,
\[
\tS[\kappa_1](\xi,t) = a_0(\xi)\wl(t) +
R(\xi)\wl\left(t-2\tau(\xi,z_1)\right) + O(\lambda)
\]
Here the source is non-extended: $\tw(\xi,t)=\wl(t)$, and 
\[
\wl(t)=\frac{1}{\sqrt{\lambda}}w_1\left(\frac{t}{\lambda}\right)
\]
is a family of wavelets depending on central wavelength $\lambda$. The
vertical travel time $\tau(\xi,z_1)$ is given by
\[
\tau(\xi,z_1) = \int_0^{z_1}\sqrt{\frac{1}{c^2}-\xi^2},\,\,c=\sqrt{\kappa/rho}.
\]

Since $\tS$ is invertible, it is possible to use the
simplified objective presented by
\cite{Symes:EAGE15}:
\begin{equation}
\label{eqn:dsobj}
J_0[\kappa]=\frac{1}{2}\|\tA \tS[\kappa]^{-1}\td\|^2
\end{equation}
A derivation of this limit case from the standard penalty formulation
of extended FWI 
appears in the last section. Inversion based on this type of objective
has been called ``inversion velocity analysis'' in the subsurface
offset context \cite[]{Herve2017},\cite[]{HouSymes:Geo18}, and is also the
basis of Adaptive Waveform Inversion
\cite{Warner:14,Warner:16}.

The main result pertains to the gradient at homogeneous
$\kappa=\kappa_0$ for data resulting from a single flat reflector
embedded in a smooth background:
\[
\tdl = \tS[\kappa_1]\wl
\]
I calculate the gradient of $J_0$ in this configuration.
The result is given in equation \ref{eqn:jgrad} in the last
section. Converted to spatial frequency via the Fourier transform, it
becomes
\begin{equation}
\label{eqn:jgradft-intro}
\widehat{\nabla J[\kappa_0]}(k) = -\frac{c_0}{16}\int\, d\xi\,a_1(\xi) 
\frac{1}{ik}\hat{u}_{\lambda}(\xi,k)D_{\xi}^2\overline{\hat{u}_{\lambda}(\xi,k)}
\end{equation}
where
\[
\hat{u}_{\lambda}(\xi,k) = \left(1 +
  \frac{R(\xi)}{a_0(\xi)}e^{4ik\tau(\xi,z_1)/\tc(\xi)}\right)\hat{\wl}\left(\frac{2k}{\tc(\xi)}\right) 
\]
and 
\[
\tc(\xi) = \left(\frac{1}{c_0}^2-\xi^2\right)^{-1/2}
\]
is the vertical plane wave velocity for homogenous wave velocity
$c_0 = \sqrt{\kappa_0/rho}$.

As is reasonable, even mandatory, for simulation of seismic data,
assume that the ``mother wavelet'' $w_1$ has at least two vanishing
moments. Then for any $k_{\rm max} >0$, $\xi_{\rm max} > 0$,
\[
\max \{\left|\hat{w}_{\lambda}\left(2k/\tc{\xi}\right)\right|:
|k|\le k_{\rm max}, |\xi|\le \xi_{\rm max} \} \rightarrow 0\mbox{ as }
\lambda \rightarrow 0.
\]
It follows immediately from equation \ref{eqn:jgradft-intro} that the
same is true of the gradient:
\begin{equation}
\label{eqn:bigdeal}
\max \{\left|\widehat{\nabla J[\kappa_0]}(k)\right|:
|k|\le k_{\rm max}, |\xi|\le \xi_{\rm max} \} \rightarrow 0\mbox{ as }
\lambda \rightarrow 0.
\end{equation}

Therefore the rate at which traveltime changes for any plane wave
component in the specified range, as the bulk modulus is changed along
the direction of the negative gradient, is vanishingly small as
$\lambda \rightarrow 0$.

Otherwise put, for high-frequency reflection data, this method cannot
update the velocity away from a homogenous initial guess in a way that
is kinematically constructive.

It may even be possible to show that the update may be destructive, by
more careful examination of the gradient approximation. However, who
cares, really: it is established that SSE-FWI has similar behaviour to
FWI for at least one simple instance of reflection data.

\section{Plane Wave Modeling}

Assume that $f$ is square-integrable, $\kappa = \kappa(z)$,
$\rho=\rho(z)$ (layered medium). Then the modeled shots differ from
each other by a horizontal translation, so that there is in effect
only one shot, for which I take the location to be
$\bx_s=(0,0,z_s)$ and suppress $\bx_s$ from the notation.

To make things even simpler, set $z_s=z_r=0$. Permitting arbitrary
values just makes the arithmetic more complicated, without affecting
the essential results.

 If $\log \kappa$ and $\log \rho$ are globally
bounded, there is $\xi_{\rm max} > 0$ so that for $\|\xi\|<\xi_{\rm
  max}$, the integrals defining slant stack fields
$\tilde{p},\tilde{\bv}$, namely
\begin{eqnarray}
\label{eqn:slant}
\tp(\xi,z,t)& = &\int dx p(x,z,t+\xi x),\nonumber\\
\tv(\xi,z,t)&=&\int dx \bv(x,z,t+\xi x),\nonumber\\
\tf(\xi,z,t)&=&\int dx f(x,z,t+\xi x),
\end{eqnarray}
are absolutely convergent, and $\tp, \tv$ satisfy 
\begin{eqnarray}
\label{eqn:awepw}
(1-c^2\xi^2)\frac{\partial \tp}{\partial t} + \kappa\frac{\partial
  \tv}{\partial z} &=& \tf \nonumber\\
\rho \frac{\partial \tv}{\partial t} + \frac{\partial \tp}{\partial
  z}&=&0\nonumber\\
\tp,\tv&=&0, t \ll 0.
\end{eqnarray}
Here $c=\sqrt{\kappa/\rho}$ is the sound velocity. I use $\xi$ instead
of the more usual $p$ for plane wave slowness since $p$ already has
been assigned to mean pressure. Note that if
$f(\bx,t;\bx_s)=w(t)\delta(\bx)$, then
$\tf(\xi,z,t)=w(t)\delta(z)$. Likewise for a surface extended source,
$f(\bx,t;\bx_s) = w(x,t)\delta(z)$, $\tf(\xi,z,t) = \tw(\xi,t)\delta(z)$.

From now on assume a surface extended source, so in plane wave domain
$\tf(\xi,z,t) = \tw(\xi,t)\delta(z)$. The plane wave domain modeling
operator samples $\tp$ on $z=z_r=0$:
\begin{equation}
\label{eqn:modpw}
\tS[\kappa]\tw = \tp|_{z=0}
\end{equation}

For homogeneous $\kappa=\kappa_0, \rho=\rho_0, c=c_0$, a computation
recapitulated in the Appendix C, relying on computations for the 1D
acoustics system detailed in Appendices A and B, yields
\begin{equation}
\label{eqn:modhom}
\tS[\kappa_0]\tw(\xi,t) = \frac{1}{2(1-c_0^2\xi^2)}\tw(\xi,t).
\end{equation}
That is, at a homogeneous model, the plane wave modeling operator is
simply scaling by a $\xi$-dependent factor.

The derivative of the plane wave modeling operator $\tS$ with respect
to $\kappa$ is the $z=0$ trace of the pressure perturbation field
$\delta \tp$, a solution component of the perturbational plane wave
system
\begin{eqnarray}
\label{eqn:dawepw}
(1-c^2\xi^2)\frac{\partial \delta \tp}{\partial t} -
  \delta \kappa \frac{\xi^2}{\rho}\frac{\partial \tp}{\partial t}
  \nonumber\\
+\kappa\frac{\partial
  \delta \tv}{\partial z} + \delta \kappa\frac{\partial
  \tv}{\partial z} &=& 0 \nonumber\\
\rho \frac{\partial \delta \tv}{\partial t} + \frac{\partial \delta \tp}{\partial
  z}&=&0\nonumber\\
\delta \tp,\delta \tv&=&0, t \ll 0.
\end{eqnarray}
The derivative of the modeling operator $\tS$ is simply the trace at
$z=0$ of the plane wave pressure field $\tp$:
\begin{equation}
\label{eqn:dmodpw}
D(\tS[\kappa]\tw)\delta \kappa = \delta \tp|_{z=0}
\end{equation}
For homogeneous background model $\kappa=\kappa_0$, the appendices
explain how to develop an explicit expression for this operator: one
obtains
\begin{equation}
\label{eqn:dmodhom}
D(\tS[\kappa]\tw)_{\kappa=\kappa_0}\delta \kappa(\xi,t) =
\frac{c_0}{16\kappa_0 (1-c_0^2\xi^2)^{5/2}}\int \,ds\,\tw(\xi,t-s)(\delta \kappa)'  \left(\frac{c_0
      s}{2\sqrt{1-c_0^2\xi^2}}\right)
\end{equation}
To compute the gradient of the SSE objective function, the transpose
of this operator (as a linear map on $\delta \kappa$) is required:
that is
\begin{equation}
\label{eqn:dmodhomadj}
D(\tS[\kappa]\tw)_{\kappa=\kappa_0}^T\td(z) =
- 
\frac{1}{8} \int \,d\xi\,\frac{1}{\kappa_0 (1-c_0^2\xi^2)^2}\frac{\partial}{\partial z} \int \,dt\,
  w\left(\xi,t-\frac{2z\sqrt{1-c_0^2\xi^2}}{c_0}\right)d(t).
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple Reflection Data}

Inclusion of reflections simply means permitting $\kappa$ to be
non-smooth. The simplest example is the layer-over-half-space. The
data to be described in this section belongs to a model having the
same properties at the measurement surface $z=0$ as the homogeneous
reference model: that is, $\kappa(0)=\kappa_0$. So to introduce a
reflector with different moveout than that implied by the reference
model, it is necessary to allow $\kappa$ to be non-constant between
the surface an the reflector depth, say $z_1 > 0$.

Suppose that $\kappa_1$ is smooth in the intervals $[0,z_1)$ and
$(z_1,\infty)$, but undergoes a nonzero jump at $z=z_1$, 
\[
[\kappa_1]_{z=z_1} \ne 0.
\]
The proper notion of solution for wave equations such as equation
\ref{eqn:awe} or \ref{eqn:awepw} with discontinuous coefficients such
as $\kappa_1$ is that of {\em weak solution}, see
\cite{BlazekStolkSymes:13}. For the point source case, $\tw(\xi,t) =
w(t)$, with smooth $w$, a weak solution can be constructed that is
smooth in $z \ne z_1$: this construction is the familiar combination
of reflected and transmitted waves. Since causal weak solutions are
unique, this is {\em the} solution. The smooth fields on either side
of the reflector at $z=z_1$ can approximated via geometric optics,
leading to an asymptotic
approximation to the solution. To express the asymptotic nature of
this approximation, consider a family of source wavelets, rather than
a single wavelet
\[
\wl = \frac{1}{\sqrt{\lambda}}w_1\left(\frac{t}{\lambda}\right)
\]
where $w_1\in C_0^{\infty}(\bR)$ has zero mean. Then [REFERENCE - WHERE?]

The pressure field at the receiver 
location $z_r=0$ takes the familiar form  
\begin{equation}
\label{eqn:modlohs}
\tS[\kappa_1]\wl(\xi,t) = \frac{1}{2c_0\sqrt{1-c_0^2\xi^2}}\wl(t) +
R(\xi)\wl\left(t-2\tau(\xi,z_1)\right) + O(\lambda)
\end{equation}
where $\tau$ is the vertical plane wave travel time:
\begin{equation}
\label{eqn:pwtt}
\tau(\xi,z) = \int_0^{z} \sqrt{\frac{1}{c(z)^2}-\xi^2}.
\end{equation}
The reflection coefficient $R$ depends on $\xi$ and on
the values of $\kappa_1$ and $\rho$ near $z=z_1$: the precise form is not
important for the argument to follow.

The remainder in \ref{eqn:modlohs} is $O(\lambda)$ in the sense of
$L^2$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SSE Objective and Gradient: vanishing low frequencies}
In this section I will formulate the SSE objective, and calculate its
gradient at homogeneous background for single reflector data as
described in the last section. I will show that this gradient has
low-frequency content as the data wavelength $\lambda$, introduced in
the last section and in the introduction, tends to zero.

FWI via SSE requires specification of an {\em annihilator}, the null
space of which consists of non-extended models, as explained in the
introduction.  Construct a suitable plane-wave domain annihilator
$\tA$ for
SSE by intertwining the Radon transform $R$ with the space-domain
annihilator $A = $ multiplication by offset. In order to do this
properly, it is necessary to introduce a cutoff in offset: not a
hardship, since infinite offsets are unavailable. Denote by $\chi$ a
offset mute. Define
\[
Rw(\chi,t) = \int \,dx\,\chi(x) w(x,t+\xi x)
\]
\[
Aw(x,t)=x w(x,t)
\]
\[
\tilde{A} Rw (\xi,t) = R Aw (\xi,t)
\]
\[
=\int dx x \chi(x) w(x,t+\xi x) = \int dx \chi(x)\frac{\partial}{\partial \xi}
I_tw(x,t+\xi x)
\]
\[
= D_{\xi} I_t R  w \equiv \tilde{A} Rw (\xi,t) 
\]

Since $\tS[\kappa]$ is continuous in $\kappa$
\cite[]{BlazekStolkSymes:13}, and $\tS[\kappa_0]$ is a multiple of the
identity, $\tS[\kappa]$ is invertible in sup-norm neighborhood of
$\kappa$. Introduce the the scaled reduced penalty function objective
\[
\kappa \mapsto \min_{\tw}(\frac{1}{\alpha^2}\|\tS[\kappa]\tw-\td\|^2 + \|\tA
\tw\|^2)
\]
\[
=\frac{1}{\alpha^2}\|\tS[\kappa]\tw_{\alpha}[\kappa]-\td\|^2 + \|\tA
\tw_{\alpha}[\kappa]\|^2
\]
where $\tw_{\alpha}[\kappa]$ solves the normal equation:
\[
\tS[\kappa]^T\tS[\kappa]\tw  + \alpha^2\tA^T\tA\tw = \tS[\kappa]^T\td.
\]
Since $\tS[\kappa]$ is invertible, $\tw_{\alpha}[\kappa]$ is uniformly
bounded as $\alpha \rightarrow 0$. Let $\tw_0[\kappa] =
\tS[\kappa]^{-1}\td$. Then
\[
\tS[\kappa]^T\tS[\kappa](\tw_{\alpha}[\kappa]-\tw_0[\kappa])+\alpha^2\tA^T\tA\tw_{\alpha}[\kappa]
= 0,
\]
so $\tw_{\alpha}[\kappa] \rightarrow \tw_0[\kappa]$ as $\alpha
\rightarrow 0$. Consequently the scaled objective above becomes
\[
\frac{1}{\alpha^2}\|\tS[\kappa]\tw_{\alpha}[\kappa]-\td\|^2 + \|\tA 
\tw_{\alpha}[\kappa]\|^2
=\frac{1}{\alpha^2}\|\tS[\kappa](\tw_{\alpha}[\kappa]-\tw_0[\kappa])\|^2 + \|\tA 
\tw_{\alpha}[\kappa]\|^2
\]
\[
=\frac{1}{\alpha^2}\langle (\tw_{\alpha}[\kappa]-\tw_0[\kappa]),
\tS[\kappa]^T\tS[\kappa](\tw_{\alpha}[\kappa]-\tw_0[\kappa])\rangle+ \|\tA 
\tw_{\alpha}[\kappa]\|^2
\]
\[
=\frac{1}{\alpha^2}\langle (\tw_{\alpha}[\kappa]-\tw_0[\kappa]),
-\alpha^2\tA^T\tA \tw_{\alpha}[\kappa] \rangle + \|\tA\tw_{\alpha}[\kappa]\|^2
=\langle (\tw_{\alpha}[\kappa]-\tw_0[\kappa]),
-\tA^T\tA \tw_{\alpha}[\kappa] \rangle + \|\tA\tw_{\alpha}[\kappa]\|^2
\]
\begin{equation}
\label{eqn:j0def}
\rightarrow \|\tA\tw_0[\kappa]\|^2 \equiv 2 J_0[\kappa,\td].
\end{equation}
This $\alpha \rightarrow 0$ limit is the simplified objective used by
\cite{Symes:EAGE15} and mentioned in the introduction.

%\[
%DJ[\kappa_0]\delta \kappa = -\langle D(\tS[\kappa_0]\tw)\delta \kappa,
%\tA^T\tA S[\kappa_0]^{-1}\td \rangle , \,\tw=\tS[\kappa_0]^{-1}\td
%\]
Now assume that $\td$ is data computed from a simple reflector as
described in the previous section, depending on the wavelength
parameter $\lambda$:
\[
\td_{\lambda}(\xi,t) = a_0(\xi)\wl(t) +
R(\xi)\wl\left(t-2\tau(\xi,z_1)\right) + O(\lambda)
\]
Then
\begin{equation}
\label{eqn:wtildedef}
\tS[\kappa_0]^{-1}\td_{\lambda}(\xi,t) = \wl(t) +
\frac{R(\xi)}{a_0(\xi)}\wl\left(t-2\tau(\xi,z_1)\right) + O(\lambda)
\equiv \tw(\xi,t) + O(\lambda)
\end{equation}
Write 
\[
a_1(\xi)=\frac{1}{\kappa_0 (1-c_0^2\xi^2)^2}, \tc_0(\xi) =
\frac{c_0}{\sqrt{1-c_0^2\xi^2}}
\]
%Then
%\[
%DJ[\kappa_0]\delta \kappa =
%\left\langle \frac{c_0}{16\kappa_0 (1-c_0^2\xi^2)^{5/2}}\int
%\,ds\,\tw(\xi,t-s)(\delta \kappa)'
%\left(\frac{c_0s}{2\sqrt{1-c_0^2\xi^2}}\right), D_{\xi}^2I_t^2
%\tw(\xi,t) \right\rangle
%\]
%Per the derivation of simple reflection, presume that (1) $w_1(t) = 0$
%for $t>1$, and $w_1$ (hence $\wl$) has two vanishing moments, and (2) $\delta \kappa = 0 $ in an interval of positive
%length about $z=0$, and (3) $\lambda$ is small enough that $\wl(t)=0$
%if $\delta(\kappa)(t\tc/2) \ne 0$ for all $\xi$ for which $\chi(\xi)
%\ne 0$ (I.e. the range of the $\xi$ integral).

%Then the convolution of $\delta \kappa(t\tc/2)$ and $\wl$ vanishes for
%small $t$, hence the product with $\wl(t)$

%and

From the definition \ref{eqn:j0def} of $J_0$ and a simple calculation,
\[
\nabla J[\kappa_0,\td] = (D\tS[\kappa_0]\tw)^T(\tS[\kappa_0]^T)^{-1}\tA^T\tA
\tS[\kappa_0]^{-1}\td,  \,\tw=\tS[\kappa_0]^{-1}\td
\]
Since $D_{\xi}$ and $I_t$ are both skew-symmetric and commute, $\tA$
is symmetric. $\tS$ is $\xi-$ dependent scaling. For simplicity
abbreviated the relation \ref{eqn:modhom} as
\[
\tS[\kappa_0]\tw(\xi,t) = a_0(\xi)\tw(\xi,t)
\]
Then
\[
\nabla J[\kappa_0](z) =
- 
\frac{1}{8} \int \,d\xi\,a_1(\xi)\frac{\partial}{\partial z} \int \,dt\, 
  \tw\left(\xi,t-\frac{2z}{\tc_0}\right)\left(D_{\xi}^2I_t^2 \tw(\xi,t)\right)
\]

\begin{equation}
\label{eqn:jgrad}
=-\frac{c_0}{16} \int \,d\xi\,a_1(\xi)\frac{\partial}{\partial z} \int \,dz'\, 
  \tw\left(\xi,\frac{2(z'-z)}{\tc_0}\right)\left(D_{\xi}^2I_z^2
    \tw\left(\xi,\frac{2z'}{\tc_0}\right)\right)
\end{equation} 
Set $u(\xi,z) = \tw(\xi,2z/\tc)$. Denote by $\hat{f}$ the Fourier
transform of $f$.  
\begin{equation}
\label{eqn:jgradft}
\widehat{\nabla J[\kappa_0]}(k) = -\frac{c_0}{16}\int\, d\xi\,a_1(\xi) 
\frac{1}{ik}\hat{u}(\xi,k)D_{\xi}^2\overline{\hat{u}(\xi,k)}
\end{equation}
From the definition \ref{eqn:wtildedef} of $\tw$,
\[
\hat{\tw}(\xi,\omega) = \left(1 + \frac{R(\xi)}{a_0(\xi)}e^{2i\omega\tau(\xi,z_1)}\right)\hat{\wl}(\omega)
\]
whence
\[
\hat{u}(\xi,k) = \left(1 +
  \frac{R(\xi)}{a_0(\xi)}e^{4ik\tau(\xi,z_1)/\tc(\xi)}\right)\hat{\wl}\left(\frac{2k}{\tc(\xi)}\right) 
\]
Together with the expression \ref{eqn:jgradft} for the Fourier
transform of the gradient, this observation establishes the assertion
stated in the introduction.

\append{Singular RIght-hand Sides and Weak Solutions}
Theorem 1 in \cite[]{BlazekStolkSymes:13} provides a strong solution of the system \ref{eqn:awe}: $p$, respectively $\bv$, is continuous
in time with values in the Sobolev space $H^1(\bR^d)$, respectively
$H^1_{\rm div}(\bR^d)$ and
also continuously differentiable with values in $L^2(\bR^d)$,
respectively $(L^2(\bR^d))^d$. Therefore all of the terms in
\ref{eqn:awe} are square-integrable functions.

Theorem 4 in \cite[]{BlazekStolkSymes:13} implies that for each $t$, $p(\cdot,t)$ and $\bv(\cdot,t)$
have compact support, and that the support expands in $t$ at the rate
bounded by the sup of the velocity $c=\sqrt{\kappa/\rho}$, as one
would expect. Therefore, for any fixed max time $T$, it is possible to
treat the problem for $0\le t \le T$ as a Dirichlet problem in a
domain $\Omega$ of sufficiently large diameter. Without loss of
generality, assume that $\mbox{supp }f \subset \Omega \times [0,T]$.

The isotropic point radiator \ref{eqn:ipr} is not
square-integrable, so the theory in \cite[]{BlazekStolkSymes:13} does
not apply directly to the system
\begin{eqnarray}
\label{eqn:aweptsrc}
\frac{\partial p}{\partial t} + \kappa \nabla \cdot {\bf v} &=&
                                                                w(t)\delta(\bx-\bx_s) \nonumber\\
\rho \frac{\partial {\bf v}}{\partial t} + \nabla p = 0 \nonumber\\
p,{\bf v}&=&0, t \ll 0
\end{eqnarray}
However there is a simple trick that makes this
theory applicable, provided one assumes that $\kappa$ is constant
($=\kappa_0 \in \bR^+$) near
$\bx=\bx_s$ (Assumption C). With this constraint, the problem
\ref{eqn:aweptsrc} may be shown to have a unique causal weak solution.,

Note that the existence of weak solutions for problems of this type
has been 
well-known for decades, with the additional assumption that the coefficients are
smooth, see for example \cite{Lax:PDENotes}. Using the progressing
wave expansion of the fundamental solution for smooth coefficient
problems (for example, \cite{Friedlander:75}) in place of the analytic
expression for the constant-coefficient point radiator solution given
below, it is possible to obtain a similar result under the weaker
\begin{itemize}
\item[Assumption C']: $\kappa$ is smooth near $\bx_s$.
\end{itemize}

Denote by $(p_0,\bv_0)$ the (weak) solution of the
system \ref{eqn:awe} with $\kappa$ replaced by $\kappa_0$. For $d=3$,
\[
 p_0(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-\frac{ |\bx-\bx_s|}{c_0}\right)
\]
in which $c_0=\sqrt{\kappa_0/\rho_0}$. There is a similar expression
for $\bv_0$, and for other space dimensions $d$. Because $w$ is smooth, both $p$ and $\bv$ are smooth
except at $\bx=\bx_s$. Pick $\phi \in C_0^{\infty}(\bR^d)$ so that
$\phi = 1$ in a neighborhood of $\bx_s$ and $\kappa=\kappa_0$ on
$\mbox{supp }\phi$. Then
\begin{eqnarray}
\label{eqn:awecut}
\frac{\partial (\phi p_0)}{\partial t} + \kappa \nabla \cdot (\phi {\bf v}_0) &=&
                           w(t)\delta(\bx-\bx_s) + f_0  \nonumber\\
\rho \frac{\partial (\phi {\bf v}_0)}{\partial t} + \nabla (\phi p_0) = {\bf g}_0 \nonumber\\
\end{eqnarray}
in which $f_0$ and ${\bf g}_0$ are smooth of compact support. Denote by
$(p_1,\bv_1)$ the strong solution of
\begin{eqnarray}
\label{eqn:awesupp}
\frac{\partial p_1}{\partial t} + \kappa \nabla \cdot {\bf v}_1 &=&
                                                                -f_0 \nonumber\\
\rho \frac{\partial {\bf v}_1}{\partial t} + \nabla p_1 = -{\bf g}_0 \nonumber\\
p_1,{\bf v}_1&=&0, t \ll 0
\end{eqnarray}
as provided by Theorem 1 of \cite{BlazekStolkSymes:13}. Then
\begin{eqnarray*}
  p & = & \phi p_0 + p_1\\
  \bv & = & \phi \bv_0 + \bv_1
\end{eqnarray*}
defines the unique weak solution of the system \ref{eqn:aweptsrc}.

\cite{SantosaSymes:00} use an analogous construction for the
second-order wave equation to give $L^2$ bounds on the point radiator
solution.

As mentioned in the Preliminaries section, a similar argument
establishes the existence of weak solutions for the surface source problem
\begin{eqnarray}
\label{eqn:awess}
\frac{\partial p}{\partial t} + \kappa \nabla \cdot {\bf v} &=&
                                                                w(x,y,t)\delta(z-z_s) \nonumber\\
\rho \frac{\partial {\bf v}}{\partial t} + \nabla p = 0 \nonumber\\
p,{\bf v}&=&0, t \ll 0
\end{eqnarray}
under Assumption D, that $\kappa=\kappa_0$ near $z=z_s$. An analytic
solution for $\kappa \equiv \kappa_0$ follows from that for the point
radiator problem, by superposition. Then a similar cutoff construction
and appeal to Theorem 1 of \cite{BlazekStolkSymes:13} proveds a weak
solution to \ref{eqn:awess}.

This construction also works via recourse to the progressing wave
expansion for the weaker
\begin{itemize}
\item[Assumption D']: $\epsilon > 0$ exists for which $\kappa$ is
smooth in the collar neigborhood $\{\bx: |z-z_s|<\epsilon\}$.
\end{itemize}

\append{1D Radiation Problem}
Begin with the 1D acoustics point source system. 
\begin{eqnarray}
\label{eqn:awe1d}
\frac{\partial p}{\partial t} +\kappa\frac{\partial 
  v}{\partial z} &=& w(t)\delta(z-z_s) \nonumber\\
\rho \frac{\partial v}{\partial t} + \frac{\partial p}{\partial 
  z}&=&0\nonumber\\
 p,v&=&0, t \ll 0. 
\end{eqnarray}
Since the right hand side is singular, so is the solution, so it must
be a solution in the weak sense; since it is causal, it is uniquely
determined by the system above. It follows from the weak solution
conditions that the pressure is continuous at $p=0$, whence $v$ must
have a discontinuity.

Now assume that $\kappa, \rho$ are constant. 
In $z \ne z_s$, the right hand side 
vanishes, so the solution must be locally a combination of plane
waves; causality implies that
\[
p(z,t)=a\left(t \mp \frac{z-z_s}{c}\right), \, v(z,t)=\pm b\left(t \mp
  \frac{z_s}{c}\right)
\]
From the second dynamical equation (Newton's law) it follows that $b =
a/(\rho c)$. The singularity on the LHS of the first dynamical
equation (constitutive law) is
\[
\kappa [v]_{z=z_s}\delta(z-z_s) =
2\frac{\kappa}{c}b\delta(z-z_s) = 2a\delta(z-z_s).
\] 
This must in turn equal the RHS of the constitutive law, whence
$a=w/2$. Thus
\begin{eqnarray}
\label{eqn:sol1d}
p(z,t) &=& \frac{1}{2}w\left(t \mp \frac{z-z_s}{c}\right) \nonumber \\
v(z,t) &=& \pm\frac{1}{2\rho c}w\left(t \mp \frac{z-z_s}{c}\right)
           \nonumber \\
\pm z-z_s &>& 0.
\end{eqnarray}

\append{1D Born Approximation}

The perturbational system for \ref{eqn:awe1d} is:
\begin{eqnarray}
\label{eqn:dawe1d}
\frac{\partial \delta p}{\partial t} +\kappa\frac{\partial
  \delta v}{\partial z} + \delta \kappa\frac{\partial
  v}{\partial z} &=& 0 \nonumber\\
\rho \frac{\partial \delta v}{\partial t} + \frac{\partial \delta p}{\partial
  z}&=&0\nonumber\\
\delta p,\delta v&=&0, t \ll 0.
\end{eqnarray}

If $\kappa, \rho $ are constant, it is possible to develop explicit
expressions for the solution of this system. Only $\kappa$ will be
subject to perturbation in any of the problems considered in this
report, so assume that that $\delta\rho=0$.

The key is to understand the perturbational impulse response. That is,
set $w(t)=H(t)$. Then the identity \ref{eqn:sol1d} implies that the
inhomogeneous term in the first equation of the system \ref{eqn:dawe1d}
is 
\begin{equation}
  \label{eqn:pertrhs}
-\delta \kappa(z) \frac{\partial v}{\partial z}(z,t) = \frac{\delta
  \kappa(z_s)}{\rho c}H(t)\delta(z-z_s) +
\frac{\delta\kappa(z)}{2\kappa}\delta\left(t - \frac{|z-z_s|}{c}\right)
\end{equation}
Correspondingly, the solution $(\delta p, \delta v)$ is the sum of
three components:
\begin{eqnarray}
  \label{eqn:pertrhsterms}
  \delta p & = & \delta p_s + \delta p_d \nonumber\\
  \delta v & = & \delta v_s + \delta v_d 
\end{eqnarray}
The first component satisfies
\begin{eqnarray}
  \label{eqn:dawe1ds}
  \frac{\partial \delta p_s}{\partial t} +\kappa\frac{\partial
  \delta v_s}{\partial z}   &=&  \frac{\delta
  \kappa(z_s)}{\rho c}H(t)\delta(z-z_s)\nonumber\\
  \rho \frac{\partial \delta v_s}{\partial t} + \frac{\partial \delta p_s}{\partial
  z}&=&0\nonumber\\
  \delta p_s,\delta v_s&=&0, t < 0.
\end{eqnarray}
The initial conditions are free to be chosen, and the choice above is convenient.

Note that \ref{eqn:dawe1ds} is precisely a system of the form
\ref{eqn:awe1d}, so from the formulae \ref{eqn:sol1d} with $w(t)=\frac{\delta
  \kappa(z_s)}{\rho c}H(t)$
\begin{eqnarray}
\label{eqn:dsol1ds}
\delta p_s(z,t) &=& \frac{\delta
  \kappa(z_s)}{2\rho c}H\left(t - \frac{|z-z_s|}{c}\right) \nonumber \\
\delta v_s(z,t) &=& \pm\frac{\delta \kappa(z_s)}{2\rho^2 c^2}H\left(t - \frac{|z-z_s|}{c}\right)
\end{eqnarray}

Define $P_d(z,t;\zeta), V_d(z,t;\zeta)$ to be the solution of the system
\begin{eqnarray}
  \label{eqn:dawe1dd}
  \frac{\partial P_d}{\partial t} +\kappa\frac{\partial
  V_d}{\partial z}   &=&   \frac{\delta\kappa(\zeta)}{2\kappa}\delta(z-\zeta)\delta\left(t - \frac{|\zeta-z_s|}{c}\right)\nonumber\\
  \rho \frac{\partial V_d}{\partial t} + \frac{\partial P_d}{\partial
  z}&=&0\nonumber\\
  P_d,V_d&=&0, t < 0.
\end{eqnarray}
Then
\begin{eqnarray}
  \label{eqn:intrep1}
  \delta p_d(z,t) & = & \int d\zeta\, P_d(z,t;\zeta)\\
  \delta v_d(z,t) & = & \int d\zeta\, V_d(z,t;\zeta)
\end{eqnarray}

The system \ref{eqn:dawe1dd} is precisely of the form
\ref{eqn:awe1d} with $z_s$ replaced by $\zeta$ and
\[
  w(t)= \frac{\delta\kappa(\zeta)}{2\kappa}\delta\left(t - \frac{|\zeta-z_s|}{c}\right)
\]
Thus from \ref{eqn:sol1d}
\begin{eqnarray*}
  P_d(z,t;\zeta) &=& \frac{\delta\kappa(\zeta)}{4\kappa} \delta\left(t - \frac{|z-\zeta|+|\zeta-z_s|}{c}\right) \nonumber \\
  V_d(z,t;\zeta) &=& \pm\frac{\delta\kappa(\zeta)}{4\kappa \rho c} \delta\left(t - \frac{|z-\zeta|+|\zeta-z_s|}{c}\right) 
\end{eqnarray*}

\begin{equation}
\label{eqn:imp1d}
D(S[\kappa]H)\delta \kappa (t) = \delta p(0,t)=  \frac{\delta 
  \kappa\left(\frac{ct}{2}\right)}{8\kappa} 
\end{equation}
whence for arbitrary causal wavelet $w$,
\begin{equation}
\label{eqn:fwd1d}
D(S[\kappa]w)\delta \kappa (t) = \int_0^t\,ds\,w'(t-s)\frac{\delta 
  \kappa\left(\frac{cs}{2}\right)}{8\kappa} 
\end{equation}

\append{Plane Wave Modeling of Reflection Traces}
Re-write the plane wave system \ref{eqn:awepw} as
\begin{eqnarray}
\label{eqn:awepw1}
\frac{\partial \tp}{\partial t} + \tk\frac{\partial
  \tv}{\partial z} &=& (1-c^2\xi^2)^{-1}\tf \nonumber\\
\rho \frac{\partial \tv}{\partial t} + \frac{\partial \tp}{\partial
  z}&=&0\nonumber\\
\tp,\tv&=&0, t \ll 0,
\end{eqnarray}
in which $\tk(\xi,z)=(1-c(z)^2\xi^2)^{-1}\kappa(z)$ is a scaled version of
bulk modulus. Similarly define the vertical plane wave velocity as
\[
\tc(\xi,z)=\frac{c(z)}{1-c(z)^2\xi^2} =
\left(\frac{1}{c(z)^2}-\xi^2\right)^{-1/2}.
\]
If $\tf(\xi,z,t)=\tw(\xi,t)\delta(z)$ is a surface
extended source, plane wave version, this is equivalent to replacing
$\tw(\xi,t)$ by $(1-c(0)^2\xi^2)^{-1}\tw(\xi,t)$.

Since the system \ref{eqn:awepw1} has the form of the 1D system
\ref{eqn:awe1d} for each $\xi $, both $\tS[\kappa_0] $ and
$D\tS[\kappa_0]$ can be computed from the corresponding 1D
results. From \ref{eqn:sol1d},
\[
\tS[\kappa_0]\tw(\xi,t)=\frac{1}{2}(1-c_0^2\xi^2)^{-1}\tw(\xi,t)
\]
which is equation \ref{eqn:modhom}.

From \ref{eqn:fwd1d},
\[
D(\tS[\kappa]\tw)_{\kappa=\kappa_0}\delta \kappa = \int_0^t\,ds\,(1-c_0^2\xi^2)\tw'(\xi,t-s)\frac{\delta 
  \tk \left(\frac{\tc s}{2}\right)}{8\tk} 
\]
Note that 
\[
\delta \tk = \delta \left(\frac{\kappa}{1-\kappa \xi^2/\rho}\right) = 
\]
\[
\delta \kappa (1-c^2\xi^2)^{-2}
\]
Consequently 
\[
D(\tS[\kappa]\tw)_{\kappa=\kappa_0}\delta \kappa(\xi,t) =
\int_0^t\,ds\,\tw'(\xi,t-s)\frac{\delta \kappa  \left(\frac{\tc 
      s}{2}\right)}{8\kappa_0 (1-c_0^2\xi^2)^2} 
\]
\[
= \frac{c_0}{2}\int_0^t\,ds\,\tw(\xi,t-s)\frac{(\delta \kappa)'  \left(\frac{\tc 
      s}{2}\right)}{8\kappa_0 (1-c_0^2\xi^2)^{5/2}} 
\]
which is equation \ref{eqn:dmodhom}.

\bibliographystyle{seg}
\bibliography{../../bib/masterref}