### 2020-02-02 repro/papers/ssevelan/talkfigs.tex

\section{Figures for talk}
\plot{pulse00}{width=\textwidth}{Bandpass filter pulse}
\plot{fwitrbulkmodinvlh0}{width=\textwidth}{Inversion for bulk modulus of data displayed in Figure \ref{fig:ptpwindl0}. Initial model: constant bulk modulus = 4.0 GPa. Algorithm: 20 iterations of trust region Krylov-Gauss-Newton algorithm with preconditioning by box smoother of radius 5 points, repeated 10 times.  Each KGN step computed by up to 20 Conjugate Gradient iterations.} 
\plot{fwitrdatapreslh0}{width=\textwidth}{Data residual produced by 20 itersions of KGN FWI for bulk modulus, starting at constant bulk modulus. Residual RMS approximately 26\% of data RMS.}
\plot{cgwhestsourceplh0}{width=\textwidth}{Extended inversion for source on 1500 m $\le x \le$ 5500 m, $z$=3000 m, data from Figure \ref{fig:ptpwindl0}, homogeneous bulk modulus = 4.0 GPA.}
\plot{cgwhresdataplh0}{width=\textwidth}{Residual data from extended source inversion shown in Figure \ref{fig:cgwhestsourceplh0}. RMS is 3\% of data RMS.}
\plot{bmh0s}{width=\textwidth}{Iteration 1}
\plot{cgw1bulkupdlh0}{width=\textwidth}{Iteration 1 SSE bulkmod estimate}
\plot{cgw2bulkupdlh0}{width=\textwidth}{Iteration 2 SSE bulkmod estimate}
\plot{cgw3bulkupdlh0}{width=\textwidth}{Iteration 3 SSE bulkmod estimate}
\plot{cgw4bulkupdlh0}{width=\textwidth}{Iteration 4 SSE bulkmod estimate}
\plot{cgw5bulkupdlh0}{width=\textwidth}{Iteration 5 SSE bulkmod estimate}
\plot{cgw6bulkupdlh0}{width=\textwidth}{Iteration 6 SSE bulkmod estimate}
\plot{cgw7bulkupdlh0}{width=\textwidth}{Iteration 7 SSE bulkmod estimate}
\plot{cgw8bulkupdlh0}{width=\textwidth}{Iteration 8 SSE bulkmod estimate}
\plot{cgw9bulkupdlh0}{width=\textwidth}{Iteration 9 SSE bulkmod estimate}
\plot{cgw10bulkupdlh0}{width=\textwidth}{Iteration 10 SSE bulkmod estimate}
\plot{cgw11bulkupdlh0}{width=\textwidth}{Iteration 11 SSE bulkmod estimate}
\plot{cgw12bulkupdlh0}{width=\textwidth}{Iteration 12 SSE bulkmod estimate}
\plot{cgw13bulkupdlh0}{width=\textwidth}{Iteration 13 SSE bulkmod estimate}
\plot{cgw14bulkupdlh0}{width=\textwidth}{Iteration 14 SSE bulkmod estimate}
\plot{cgw15bulkupdlh0}{width=\textwidth}{Iteration 15 SSE bulkmod estimate}
\plot{cgw22bulkupdlh0}{width=\textwidth}{Iteration 22 SSE bulkmod estimate}
\plot{cgw30bulkupdlh0}{width=\textwidth}{Iteration 30 SSE bulkmod estimate}
\plot{fwitrsse1bulkmodinvlh0}{width=\textwidth}{Inversion for bulk modulus of data displayed in Figure \ref{fig:ptpwindl0}. Initial model: result of 15 steepest descent steps of SSE, Figure \ref{fig:cgw15bulkupdlh0}. Algorithm: 20 iterations of trust region Krylov-Gauss-Newton algorithm with preconditioning by box smoother of radius 5 points, repeated 10 times.  Each KGN step computed by up to 20 Conjugate Gradient iterations.} 
\plot{fwitrsse1datapreslh0}{width=\textwidth}{Data residual produced by 20 itersions of KGN FWI for bulk modulus, starting at output of 15 steepest descent steps of SSE, Figure \ref{fig:cgw15bulkupdlh0}. Residual RMS approximately 1.2\% of data RMS.}
\plot{cgw0estsourceplh0}{width=\textwidth}{Iteration 1 SSE source estimate alpha=4.36 $\times 10^{-6}$}
\plot{cgw1estsourceplh0}{width=\textwidth}{Iteration 2 SSE source estimate alpha=5.93 $\times 10^{-6}$}
\plot{cgw2estsourceplh0}{width=\textwidth}{Iteration 3 SSE source estimate alpha=6.82 $\times 10^{-6}$}
\plot{cgw3estsourceplh0}{width=\textwidth}{Iteration 4 SSE source estimat alpha=7.93 $\times 10^{-6}$e}
\plot{cgw4estsourceplh0}{width=\textwidth}{Iteration 5 SSE source estimate alpha=9.29 $\times 10^{-6}$}
\plot{cgw5estsourceplh0}{width=\textwidth}{Iteration 6 SSE source estimate alpha=1.1 $\times 10^{-5}$}
\plot{cgw6estsourceplh0}{width=\textwidth}{Iteration 7 SSE source estimate alpha=1.32 $\times 10^{-5}$}
\plot{cgw7estsourceplh0}{width=\textwidth}{Iteration 8 SSE source estimate alpha=1.32 $\times 10^{-5}$}
\plot{cgw8estsourceplh0}{width=\textwidth}{Iteration 9 SSE source estimate alpha=1.72 $\times 10^{-5}$}
\plot{cgw9estsourceplh0}{width=\textwidth}{Iteration 10 SSE source estimate alpha=2.2 $\times 10^{-5}$}
\plot{cgw10estsourceplh0}{width=\textwidth}{Iteration 11 SSE source estimate alpha=2.79 $\times 10^{-5}$}
\plot{cgw11estsourceplh0}{width=\textwidth}{Iteration 12 SSE source estimate alpha=2.79 $\times 10^{-5}$}
\plot{cgw12estsourceplh0}{width=\textwidth}{Iteration 13 SSE source estimate alpha=2.79 $\times 10^{-5}$}
\plot{cgw13estsourceplh0}{width=\textwidth}{Iteration 14 SSE source estimate alpha=2.79 $\times 10^{-5}$}
\plot{cgw14estsourceplh0}{width=\textwidth}{Iteration 15 SSE source estimate alpha=2.79 $\times 10^{-5}$}
\plot{cgw21estsourceplh0}{width=\textwidth}{Iteration 22 SSE source estimate alpha=5.73 $\times 10^{-5}$}
\plot{cgw29estsourceplh0}{width=\textwidth}{Iteration 30 SSE source estimate alpha=9.1 $\times 10^{-5}$}
\plot{cgw0resdataplh0}{width=\textwidth}{Iteration 1 SSE data residual}
\plot{cgw1resdataplh0}{width=\textwidth}{Iteration 2 SSE data residual}
\plot{cgw2resdataplh0}{width=\textwidth}{Iteration 3 SSE data residual}
\plot{cgw3resdataplh0}{width=\textwidth}{Iteration 4 SSE data residual}
\plot{cgw4resdataplh0}{width=\textwidth}{Iteration 5 SSE data residual}
\plot{cgw5resdataplh0}{width=\textwidth}{Iteration 6 SSE data residual}
\plot{cgw6resdataplh0}{width=\textwidth}{Iteration 7 SSE data residual}
\plot{cgw7resdataplh0}{width=\textwidth}{Iteration 8 SSE data residual}
\plot{cgw8resdataplh0}{width=\textwidth}{Iteration 9 SSE data residual}
\plot{cgw9resdataplh0}{width=\textwidth}{Iteration 10 SSE data residual}
\plot{cgw10resdataplh0}{width=\textwidth}{Iteration 11 SSE data residual}
\plot{cgw11resdataplh0}{width=\textwidth}{Iteration 12 SSE data residual}
\plot{cgw12resdataplh0}{width=\textwidth}{Iteration 13 SSE data residual}
\plot{cgw13resdataplh0}{width=\textwidth}{Iteration 14 SSE data residual}
\plot{cgw14resdataplh0}{width=\textwidth}{Iteration 15 SSE data residual}