### 2020-03-04 repro/papers/dso1d/paper.tex
more

\title{Full Waveform Inversion by Source Extension: Why it works}
\author{William. W. Symes \thanks{The Rice Inversion Project,
Department of Computational and Applied Mathematics, Rice University,
Houston TX 77251-1892 USA, email {\tt symes@caam.rice.edu}.}}

\lefthead{Symes}

\righthead{Why it works}

\maketitle
\begin{abstract}
An extremely simple example shows how extended modeling plus variable projection plus a suitable annihilator lead to constructive velocity updates.
\end{abstract}

\section{Introduction}
Full Waveform Inversion (FWI) is now well-established as
a useful tool for probing the earth's subsurface
\cite[]{VirieuxOperto:09,Fichtner:10}, but its use is still
impeded by so-called ``cycle-skipping''. Because the computational
size of field inversion tasks is very large, only iterative local (descent)
minimization of the data misfit function is computationally
feasible. However such local descent methods tend to stagnate as
suboptimal and geologically uninformative earth models, unless initial
models are already quite close to optimal, in the sense of predicting
the arrival times of seismic events to within a small multiple of a
dominant wavelenght \cite[]{GauTarVir:86,Plessix:10}.

This paper concerns one of the many ideas that have been advanced to
overcome cycle-skipping, namely so-called extended inversion
\cite[]{geoprosp:2008}. ``Extended'' signifies that addional degrees
of freedom are provided to the modeling process, in the hope of
opening up more effective routes to geologically informative models
with acceptable data fit. Since these extended degrees of freedom are
not part of the basic physics chosen to model the data acquisition
process, they should be suppressed in the eventual solution. Extended
inversion methods differ by the choice of additional degrees of
freedom, and by choice of penalty applied to eliminate them in the final
result.

Many of these extension concepts sound
plausible, and appear to work at least to some extent as one might
hope from their heuristic justifications. However very few of these
approaches have been underwritten by mathematical argument: in
essence, they are mostly justified only ``in the rear-view mirror'',
with no assurance that failure is not just around the corner, at the
next example. On top of that, some of these approaches, for example
those based on the computationally attractive Variable Projection
Method (``VPM'') of \cite{GolubPereyra:03}, are cast in such form that the reasons for success are
not readily apparent.

This note shows exactly how VPM leads to successful velocity updates
for a particular extended inversion approach to a very simple inverse
problem, which asks that a homogeneous velocity field be deduced from
one trace at known offset. I put forward this inverse problem and
extension-based solution not because there are not simpler
ways of answering the question it poses - there certainly are - but
because the formal ingredients of waveform-based velocity estimation
in this very simple setting are common to many similar extended
inversion algorithms, and because in this case every computation can
be done analytically, nearly to completion. In particular, it becomes
clear why the VPM gradient formula produces a constructive update,
with no possibility of stagnation at local minimizers.

The extended inversion approach
developed here uses a {\em source extension}, in which source
parameters form the additional degrees of freedom. This type of
extension presumes that the actual or target source is constrained in
some way; the extended source is allowed to violate the
constraint. For the problem considered here, the source model reduces
to a wavelet, and the target wavelet is assumed to be non-zero only in
a short time
interval (an approximate impulse). The extension consists in permitting energy to spread
in time at intermediate iterations of the inversion. A simple penalty
for energy spread (second moment of square amplitude) drives the
extended source towards a focused source satisfying the assumed
constraint. Not all penalties are created equal: the penalty used here
has a critical {\em (pseudo-)differential} attribute necessary for avoidance of cycle-skipping.

%despite not incorporating explicitly the extension measure
%(``annihilator'') central to the approach.

I begin with a quick sketch of constant density acoustics,
and describe the single-trace transmission inverse problem. To make
the role of data frequency content clear, I introduce a family of
noise-free data parametrized by wavelength. For completeness, I show how
the standard FWI approach to this problem generates multiple local
minima that will be found by any descent method unless the initial
estimate predicts travel time from source to receiver with an error
on the order of a wavelength. The next section describes the source
extension objective, and the reduced objective
produced by VPM. As VPM eliminates the extended source, this
function depends only on the velocity, just as does the FWI objective.  A
nearly-explicit calculation of the VPM gradient 
shows that the only stationary points are ``within a wavelength'' of
the correct velocity, used to build the data: that is, cycle-skipping
cannot occur. The paper ends with a discussion of the parallels
between the calculations presented here and the structure of other
extended inversion methods applicable to field-scale velocity
estimation, and the critical role that the differential nature of the
extension penalty plays in the success of this and other extension methods.

\section{Preliminaries}
Assume small amplitude (linearized) acoustic propagation, constant
density, and isotropic point source and receiver. Denote by $m(\bx)$
the slowness (reciprocal velocity) at spatial position $\bx$, $f(t)$
the time dependence of the point source (``wavelet'') at location
$\bx=\bx_s$. Then the (excess) pressure field $p(\bx,t)$ obeys a
scalar wave equation:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m(\bx)^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         f(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0
\end{eqnarray}
Suppose that a single trace is recorded, at distance $r>0$ from the
source position $\bx_s$. The dominant information in a single
trace is the transient signal time of arrival, constraining only the mean slowness in the
region
between source and receiver, so assume that the
slowness is constant, that is, independent of position $\bx$. The pressure field is simply the  the source
wavelet $f(t)$ convolved with the
acoustic Green's function, for which an analytic expression is
available in the constant $m$ case \cite[]{CourHil:62}:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}f\left(t-m|\bx-\bx_s|\right).
\end{equation}

The receiver location $\bx_r$ lies at distance $r$ from the source
location $\bx_s$, that is, $|\bx_r-\bx_s|=r$. The predicted signal at
$p(\bx_r,t)$ depends nonlinearly on the slowness $m$ and linearly on the
source wavelet $f$. Therefore it is naturally represented as the
action of a $m$-dependent linear operator $S[m]$ on $f$:
\begin{equation}
\label{eqn:mod}
S[m]f (t)= p(\bx_r,t) = \frac{1}{4\pi r}f\left(t-mr\right).
\end{equation}

Ignoring amplitude, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many descriptions of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above, which
amounts essentially to a time shift. To make the link with wavelet
frequency content manifest, I introduce a family $\{\fl\}$ of
wavelets parametrized by a nondimensional wavelength $\lambda$,
\begin{equation}
  \label{eqn:deffl}
  \fl(t) = \frac{1}{\lambda}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
The ``mother wavelet'' $f_1$ vanishes for $t>1$, and has positive
mean-square. No other constraints need be placed on $f_1$. Note that
the scaling is such that the mean-square
\[
  \|\fl\|^2 = \int\,dt\,|\fl(t)|^2
\]
is independent of $\lambda$.

Note that in relations such as ``$t>1$'' a time unit is implicit in
the right-hand side.

To this family of wavelets and a choice of target slowness $m_*$  corresponds a family of noise-free data
\begin{equation}
  \label{eqn:defdata}
  \dl=S[m_*]\fl.
\end{equation}
This family of data in turn defines a family of inverse problems, to
which I now turn.

\section{Full Waveform Inversion}
The preceding section provided all of the raw ingredients to define
full waveform inversion for estimation of $m$ from a single trace.
It is only $m$ that is to be determined: the $\lambda-$dependent family of wavelets
$\{\fl\}$ is regarded as known, along with the data family $\{\dl\}$. The aim
is to chose $m$ to minimize 
\begin{equation}
  \label{eqn:FWIfix}
  J_{\rm fix}[m] = \frac{1}{2}\|S[m]\fl-\dl\|^2.
\end{equation}
for all values of $\lambda > 0$.

Written out in detail, this objective function is
\[
 J_{\rm fix}[m] =  \frac{1}{32\pi^2
    r^2}\int\,dt\,\left|\fl\left(t-mr\right)-\fl\left(t-m_*r\right)\right|^2
\]
Since $f_1$ vanishes for $|t|>1$, $\fl$ vanishes for $|t|>\lambda$,
and $S[m]\fl$ vanishes if $|t-mr|>\lambda$. So if $|mr-m_*r|
= |m-m_*|r > 2\lambda$, then $|t-mr|+|t-m_*r| \ge |mr-m_*r| >
2\lambda$ so either $|t-mr|>\lambda$ or $|t-m_*r|>\lambda$, that is,
either $S[m]\fl(t)=0$ or $S[m_*]\fl(t)=0$. Therefore $S[m]\fl$ and
$S[m_*]\fl$ are orthogonal in the sense of the $L^2$ inner product:
\begin{equation}
  \label{eqn:ortho}
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, \langle S[m]\fl,
  S[m_*]\fl\rangle = \int\,dt\,S[m]\fl(t)S[m_*]\fl(t) = 0
\end{equation}
But $\dl = S[m_*]\fl$, so this is the same as saying that $\dl$ is
orthogonal to $S[m]\fl$. So conclude that
\begin{equation}
  \label{eqn:iso}
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, J_{\rm fix}[m]=\frac{1}{16\pi^2
    r^2}\|f_1\|^2.
\end{equation}
using the previously observed independence of $\|\fl\|$ from
$\lambda$.

That is, for slowness $m$ in error by more than $\lambda/r$ from the
target slowness $m_*$, the FWI objective \ref{eqn:FWIfix} is perfectly
flat: all nearby values of $m$ are local minima. Therefore the local
exploration of the objective
gives no useful information whatever about constructive search
directions towards the global minimizer $m=m_*$, for which the
objective value is of course = 0, if the initial estimate of $m$ in
error by an amount greater than a fixed multiple of $\lambda$. This is
precisely the behaviour of FWI noted many times in the literature: the
model must be known to ``within a wavelength''.

\section{Source Extension Inversion}
An obvious extended model for this example is $\{m,f\}$, that is, pairs of velocity and source wavelet, with the extended modeling operator already defined by equation \ref{eqn:mod}. With an annihilator $A$, to be chosen below, the penalty form of extended inversion is: minimize over $\{v,f\}$
\begin{equation}
\label{eqn:obj}
J[v,f] = \frac{1}{2}(\|S[m]f-d\|^2 + \alpha^2 \|Af\|^2).
\end{equation}
VPM \cite[]{GolubPereyra:03} takes advantage of $J$ being quadratic in $f$ to solve for $f$ given $v$, thus producing a reducted objective of $v$ alone:
\begin{equation}
\label{eqn:red}
J_{\rm red}[v] = \min_f J[v,f] = J[v,f[v]],
\end{equation}
where $f[v]$ is the minimizer of $J$ over $f$ for given $v$.
For this example, $J_{\rm red}$ is simple to compute. First observe that apart from amplitude, $S[m]$ is unitary:
\begin{equation}
\label{eqn:tran}
S[m]^T g (t) = \frac{1}{4\pi r}g\left(t+\frac{r}{v}\right)
\end{equation}
so
\begin{equation}
\label{eqn:unit}
S[m]^TS[m] f (t) = \frac{1}{(4\pi r)^2} f(t).
\end{equation}
Therefore the normal equation for the minimizer on the RHS of equation \ref{eqn:red} is
\begin{equation}
\label{eqn:norm1}
\left(\frac{1}{(4\pi r)^2} I + \alpha^2 A^TA\right)f = S[m]^Td.
\end{equation}

At this point I have to come clean about the actual choice of $A$. As usual with source extensions, the proper annihilator depends on modeling assumptions, not fundamental physics. Assume that signature decon has been applied and the target source $f_*$ is a zero-phase bandpass filter, reasonably well focused at time $t=0$. Since there is no spatial structure to exploit in this hyper-simple problem, we construct $A$ to penalize energy away from $t=0$:
\begin{equation}
\label{eqn:ann}
Af(t) = tf(t).
\end{equation}
Then the normal operator on the LHS of \ref{eqn:norm1} is simply multiplication by a positive function of time, and can be inverted by inspection. Using the identity \ref{eqn:tran} for the adjoint operator, obtain
\begin{equation}
\label{eqn:soln} 
f[v](t) = \left(\frac{1}{(4\pi r)^2} + \alpha^2 t^2\right)^{-1}\frac{1}{4\pi r}d\left(t+\frac{r}{v}\right)  
\end{equation} 
Suppose that the data is noise-free: 
\begin{equation}
\label{eqn:nonoise}
d(t) = S[v_*]f_*(t) = \frac{1}{4\pi r}f_*\left(t-\frac{r}{v_*}\right)
\end{equation}
Then
\begin{equation}
\label{eqn:solnnon}
f[v](t) = \left(\frac{1}{(4\pi r)^2} + \alpha^2 t^2\right)^{-1}\frac{1}{(4\pi r)^2}f_*\left(t-\frac{r}{v_*}+\frac{r}{v}\right) 
\end{equation}
so the residual is, after a little algebra,
\begin{equation}
\label{eqn:resid}
(S[m]f[v]-d)(t) =\frac{1}{4\pi r}[ (1+(4\pi r)^2 \alpha^2 (t-r/v)^2)^{-1} - 1]f_*(t-r/v_*)
\end{equation}
The gradient of $J_{\rm red}$ is given by the famous VPM formula (well-known to those who know it well):
\begin{equation}
\label{eqn:grad}
\nabla J[v] = (DS[m]f[v])^*(S[m]f[v]-d)
\end{equation}.
In this formula, $DS[m]$ is the derivative with respect to $v$ - since we can make explicit use of the Green's function, easy to see that
\begin{equation}
\label{eqn:deriv}
(DS[m]\delta v)f (t) = \frac{\delta v}{4\pi v^2}\frac{df}{dt}(t-r/v) = S[m](Q[v]\delta v)f (t),
\end{equation}
where 
\begin{equation}
\label{eqn:defq}
(Q[v]\delta v)f = \frac{r \delta v}{v^2} \frac{df}{dt}.
\end{equation}
That is, $Q[v]\delta v$ is a skew-adjoint operator depending linearly on $\delta v$ - more on this below.
The adjoint $(DS[m]f)^*$ is the adjoint of the map from model space to data space:
\[
(DS[m]f): \delta v \mapsto (DS[m]\delta v)f
\]
and is NOT the same as the adjoint denoted by $S[m]^T$, which is the data-space-to-data-space adjoint (got that?).
That is, this adjoint make this relation true, for all $\delta v,f,$ and $d$:
\begin{equation}
\label{eqn:vadj}
\delta v \cdot (DS[m]f)^*d = \int \,dt\,[(DS[m]\delta v)f] (t) d(t) 
\end{equation}
On the left side is the dot product in model space - since model space is just 1D in this example, that's just the numerical product. On the right is the dot product in data space, in the idealized continuum limit. Using equations \ref{eqn:deriv}, \ref{eqn:defq} and \ref{eqn:vadj},
\[
\delta v \cdot \nabla J[v] =  \int \,dt\,[(DS[m]\delta v)f[v]] (t) (S[m]f[v]-d)(t)
\]
\[
= \int \,dt\, [S[m](Q[v]\delta v)f[v](t)] (S[m]f[v]-d)(t) = \int
\,dt\,(Q[v]\delta v)f[v](t) S[m]^T(S[m]  f[v]-d)(t)
\]
Now invoke the normal equation \ref{eqn:norm1}: and replace the last factor:
\begin{equation}
\label{eqn:comm1}
= -\alpha^2\int \,dt\, (Q[v]\delta v)f[v](t)A^TAf[v](t)
\end{equation}
Since $Q[v]\delta v$ is skew-symmetric, shift it onto the other factor in this $L^2$ inner product (why not):
\[
= \alpha^2\int \,dt\, f[v](t)(Q[v]\delta v)A^TAf[v](t) 
\]
\begin{equation}
\label{eqn:comm2}
= \alpha^2\int \,dt\, (A^TAf[v](t)(Q[v]\delta v)f[v](t) + f[v](t)[(Q[v]\delta v),A^TA]f[v](t))
\end{equation}
where I swapped $Q$ and $A^TA$ at the cost of introducing a term involving the commutator $[Q,A^TA] = QA^TA-A^TAQ$, and rearranged the first term using the symmetry of $A^TA$. Now notice that this first term is exactly the same as the RHS of equation \ref{eqn:comm1}, except for the minus sign - so subtracting the RHS of \ref{eqn:comm2}
from \ref{eqn:comm1} and rearranging, get
\[
-\alpha^2\int \,dt\, (Q[v]\delta v)f[v](t)A^TAf[v](t) = \frac{1}{2}\alpha^2\int \,dt\,f[v](t)[(Q[v]\delta v),A^TA]f[v](t)
\]
hence 
\begin{equation}
\label{eqn:gradcomm}
\delta v \cdot \nabla J[v] = \frac{1}{2}\alpha^2\int \,dt\,f[v](t)[(Q[v]\delta v),A^TA]f[v](t)
\end{equation}
Remember that $A^TA$ amounts to multiplying by $t^2$, and $Q$ is the scaled time derivative (equation \ref{eqn:defq}), so
\begin{equation}
\label{eqn:comm3}
[(Q[v]\delta v),A^TA]=\frac{2r\delta v}{v^2}t
\end{equation}
Insert this identity into equation \ref{eqn:gradcomm} to obtain
\begin{equation}
\label{eqn:comm4}
\delta v \cdot \nabla J_{\rm red}[v] =  \delta v\frac{r \alpha^2}{v^2}\int \,dt\,tf[v]^2(t)
\end{equation}
Combine this identity with the formula \ref{eqn:solnnon} for the solution of the inner problem, and divide out the common factor $\delta v$, to obtain
\[
\nabla J_{\rm red}[v] = \frac{r\alpha^2}{v^2}\int \,dt\,t\left(\frac{1}{(4\pi r)^2} + \alpha^2 t^2\right)^{-1}\frac{1}{(4\pi r)^2}f_*\left(t-\frac{r}{v_*}+\frac{r}{v}\right)^2
\]
\begin{equation}
\label{eqn:gradfinal}
= \frac{r\alpha^2}{v^2}\int \,dt\,\left(t+\frac{r}{v_*}-\frac{r}{v}\right)\left(1 + (4\pi r)^2\alpha^2 \left(t+\frac{r}{v_*}-\frac{r}{v}\right)^2\right)^{-1}f_*(t)^2
\end{equation}

Recall that the definition of $A$ (multiplication by $t$) was justified by a modeling assumption, that the actual source is (possibly after designature) highly localized at $t=0$ - ideally a Dirac delta, but because of bandlimitation not quite. Still, a high degree of localization, as implicit in this choice of annihilator, means that the integral above is approximately proportional to the integrand at $t=0$, that is, 
\[
\approx \frac{r\alpha^2}{v^2}\left(\frac{r}{v_*}-\frac{r}{v}\right)\left(1 + (4\pi r)^2\alpha^2 \left(\frac{r}{v_*}-\frac{r}{v}\right)^2\right)^{-1}
\]
Note that to the extent that this approximation is accurate,
\begin{itemize}
\item if $v > v_*$, then $\nabla J_{\rm red}[v] > 0$, and
\item if $v < v_*$, then $\nabla J_{\rm red}[v] < 0$, and
\end{itemize}
That is, within the domain of validity of this approximation (setting $t=0$ in equation \ref{eqn:gradfinal}), {\em there are no local mins:} the gradient has the correct sign and velocity updates computed from it will be constructive.

On the other hand, careful examination of equation \ref{eqn:gradfinal} (which involves no approximations) shows that if $\alpha$ is too large, then the finite-frequency spread of $f_*$ is guaranteed to result in non-convex behaviour, since the side-lobes will dominate. This message is consistent with the import of many other similar computations.

\section{Discussion}
There remain two important points to be made.

First, the formal computations centering on the operator $Q$ (equations \ref{eqn:deriv} through \ref{eqn:gradcomm}) depend only on the relation \ref{eqn:deriv} and the skew-symmetry of $Q$, which hold with minor modifications for other more complex waveform inversion problems. These other more complex problems do not submit to such a simple treatment as is shown in the equations following \ref{eqn:gradcomm}, but for example it is possible in some cases to use the relation \ref{eqn:deriv} to extract a relation between extended waveform inversion and traveltime tomography, via analysis of the Hessian at a zero-residual global minimizer. See for instance \cite{tenKroode:IPTA14,Symes:IPTA14,Symes:Madrid}. Up to that point the reasoning is quite general, and central to the understanding gained so far of extended inversion methods.

The cited papers are set in the context of subsurface-offset extended Born inversion, and demand a single scattering approximation and high-frequency asymptotics - that is, microlocal anallysis. Somewhat surprisingly, it has recently proven possible to establish the existence of a variant of the operator $Q$ for acoustic problems with transmission configuration but without the assumed smoothness of the material parameter fields essential for microlocal analysis. That is, at least in some circumstances $DS=SQ$ with approximately skew-symmetric $Q$, even when reflections (even multiple) occur in the material model. It remains to be seen whether enough can be established about these more general versions of $Q$ (which are for example in general not pseudodifferential0 to gain any hold over the corresponding inverse problems.

The second point to be made has to do with the necessity of nested optimization, of which VPM is a variant. The aim is to find a minimizer of the bivariate objective $J[v,f]$ (and ultimately of data fitting, that is, FWI), but the bivariate problem is computationally intractable. The framework developed here can be used to explain why this is so: optimization in $v$ and $f$ simultaneously is very ill-conditioned (in the continuum limit, infinitely so), whereas reduced objetives (of which VPM produces one possibility) are much easier to optimize. Yin Huang's thesis \cite[]{YinHuang:16} shows this contrast between global and nested optimization very clearly, using a different extension.

\section{Conclusion}
Desipite its simplicity, the single-trace inversion transmission problem proves typical of many more complex waveform inversion problems. The structure of the derivative is similar in many of these problems, and for the particularly simple one explained here, can be analysed on paper to the point of showing explicitly why this approach to waveform inversino works.

\bibliographystyle{seg}
\bibliography{../../bib/masterref}