### 2020-06-15 repro/papers/discrep/vpmloss.tex

\section{Variable projection method with loss of derivative}

While minimization of $J_{\alpha}$ might be tackled directly - by
alternately minimizations between $m$ and $w$, or by computing updates
for $m$ and $w$ simultaneously - such joint mimization performs
poorly, as \cite{YinHuang:16} has shown. The reason for this poor
performance is that $J_{\alpha}$ has dramatically different
sensitivity to $m$ versus $w$,
as will be explained below.
The separable nature of the least-squares inverse problems defined in \ref{eqn:obj} invites use of the variable projection method, a nested optimization approach. First, in the inner loop, the objective function is optimized over linear parameter $w$ with the nonlinear parameter $m$ fixed. 
The gradient of the objective function $\Ja[m, w]$ with respect to $w$ is
\begin{equation}
\nabla _{w} \Ja[m, w;d] =F[m]^T (F[m] w - d) 
+ \alpha A^T A  w 
\label{eqn:grad1}
\end{equation}
where $T$ denotes transpose.  A stationary point of equation \ref{eqn:obj} satisfies the {\em normal equation}
\begin{equation}
(F[m]^T F[m] + \alpha A^T A ) w = F[m]^T d.
\label{eqn:condition}
\end{equation}
In view of Assumption A above, the Lax-Milgram Theorem (\cite{Yosida}, p. 92)
applies to show that the system \ref{eqn:condition} has a
unique solution. Since $m$, $d$ and $\alpha$ determine the operator on the LHS
of equation \ref{eqn:condition} and its RHS, its solution becomes a
function $\aw[m;d]$ of these quantities. Continuity in $d$ is a
consequence of the Lax-Milgram result.

As observed by \cite{GolubPereyra:73,GolubPereyra:03}, $\Ja$ may be minimized over $m,w$ by minimizing the reduced objective
\begin{equation}
  \label{eqn:red}
  \tJa[m; d] = \Ja[m,\aw[m;d];d]
\end{equation}
over $m$. It is natural, and the goal of this paper, to use local,
Newton-like methods to minimize. However such
methods require at least a gradient, and therefore derivatives of $F$,
and so far I have not specified any regularity properties of $F$.

In fact, if the theory is to encompass inverse problems in wave
propagation in which wave velocities are amongst the model parameters
to be estimated, then $F$ cannot be presumed to be differentiable, or
even locally uniformly continuous, as a ${\cal B}(W,D)$-valued
function on $M$. Instead, it is {\em differentiable with loss of
  regularity}, as I will now explain.

The mapping properties of typical modeling operators $F$ may be
described with the help of a {\em scales of Hilbert spaces}
$\{W^s, D^s\}_{s=0}^{\infty}$, with the following properties:

\noindent {\bf Assumption C:}
\begin{itemize}
\item $\{W^s\}_{s=0}^{\infty}$ and $\{D^s\}_{s=0}^{\infty}$ are {\em decreasing}, that is,
  $W^{s+1} \subset W^s$ for $s \ge 0$, and similarly for $\{D^s\}_{s=0}^{\infty}$;
\item denoting by $\|\cdot\|_s$ the norm in $W^s$, $\|w\|_t \ge
  \|w\|_s$ if $t \ge s$ and $w \in W^t$, with a similar property for $\{D^s\}_{s=0}^{\infty}$;
\item $W^0=W$, $D^0=D$;
\item 
  \begin{equation}
    \label{eqn:reg0}
    F[m] \in {\cal B}(W^s,D^s),\,F[m]^T \in {\cal B}(D^s,W^s) \mbox{
      for any } s\ge 0, m\in M
  \end{equation}
  (here $F[m]^T$ denotes the adjoint of $F[m] \in {\cal B}(W^0,D^0)$,
  that is, $\langle F[m]w,d \rangle_{D^0} = \langle w,F[m]^T
  \rangle_{W^0}$ for $w \in W^0=W, d \in D^0=D$);
\item $A \in {\cal B}(W^s,W^s)$ for any $s \ge 0$;
\item For each $\alpha > 0$, 
  \begin{equation}
    \label{eqn:normsm}
    F^TF+\alpha^2A^TA \in C^{\infty}(M, {\cal I}(W^s,W^s)).
  \end{equation}
  There exists $C_{s,\alpha}>0$ so  that for all
    $m\in M$, 
    \begin{equation}
      \label{eqn:nopco}
      \|(F[m]^TF[m]+\alpha^2A^TA)^{-1}\|_{{\cal B}(W^s,W^s)} \le
      C_{s,\alpha}.
    \end{equation}
\item 
  \begin{eqnarray}
    F &\in& C^k(M, {\cal B}(W^s,D^t)) \mbox{ provided that }s > t+k, t, 
    k \ge 0;\\
    \label{eqn:reg1}
    F^T &\in& C^k(M, {\cal B}(D^s,W^t)) \mbox{ provided that }s > t+k, t, 
    k \ge 0;
    \label{eqn:reg2}
  \end{eqnarray}
\end{itemize}
Note that the inequality of indices in the condition \ref{eqn:reg1} is {\em
  strict}. In particular, $F$ is not assumed to be continuous as a map
from $M$ to ${\cal B}(W^s,D^s)$ for any $s$. This constraint is
necessary, as even the simplest examples show. For instance, the
example considered later in this paper has this property, as I will
show explicitly. On the other hand,
\cite{BlazekStolkSymes:13} show that versions of conditions
\ref{eqn:reg0}, \ref{eqn:reg1} hold in
general for mappings defined by symmetric hyperbolic
integro-differential systems, encompassing most common physical models
of wave propagation.

Note also that, despite the relatively weak differentiability properties of $F$, as
indicated in condition \ref{eqn:reg1}, its normal operator is actually
smooth as a ${\cal I}(W^s,W^s)$-valued function for all $s\ge 0$
(condition \ref{eqn:nopco}). This condition also abstracts properties
common to many (though not all) extended inversion frameworks. It 
certainly holds for the example described in this paper.

The scales appearing in the examples that motivated Assumption C
calibrate regularity: that is, members of $W^t$ are more regular as
functions on space-time than members of $W^s$ if $t>s$. Thus the
condition \ref{eqn:reg1} expresses differentiability (of $F$ as a
function on $M$) with loss of regularity (of the arguments as
functions on space-time).

A first consequence of Assumption C is that the solution $\aw[m;d]$ of
the normal equation \ref{eqn:condition} is a continuously differentiable $W^s$-valued
function on $M \times D^{s+2}$ for any $s \ge 0, \alpha>0$. Indeed, condition
\ref{eqn:nopco} implies that the inverse of the normal operator is a
smooth ${\cal B}(W^s,W^s)$-valued function on $M$, and condition
\ref{eqn:reg2} that the RHS of the normal equation \ref{eqn:condition}
depends differentiably on $(m,d) \in M \times D^t$ so long as $t>s+1$.

From its definition
the definition of $\Ja$ (\ref{eqn:obj}), and the conclusion of the
preceding paragraph, conclude that $\tJa[m;d] \in C^1(M \times D^4)$.
Indeed, $\aw \in C^1(M \times D^4,W^2)$ from the preceding paragraph,
and the data residual $(m,w,d) \mapsto F[m]w-d$ is in $C^1(M \times
W^2 \times D^0)$ according to condition \ref{eqn:reg1}.

The main result of \cite[]{GolubPereyra:73} is a formula for the
gradient of the reduced objective \ref{eqn:red}.  Justification of this
formula in the present context rests on the mapping properties of the
derivative $DF$, which are naturally posed in
terms of the tangent space of the domain of $F$.  Recall that $M$ is
an open subset of a Hilbert space, which has so far gone
un-named. This ambient Hilbert space can be identified with the
tangent space $T_mM$ of $M$ at any point $m \in M$, and these tangent
spaces are all isomophic to the ambient space. Taking advantage of
this fact, I will drop the subscript and refer to $TM (=T_mM$ for any
$m \in M)$ as the tangent space of $M$. Thus assumption \ref{eqn:reg1}
implies that for $m \in M$, $DF[m] \in {\cal B}(TM,{\cal B}(W,D))$.

A particular instance of \ref{eqn:reg1} is the assertion that $F \in
C^1(M,{\cal B}(W^2,D^0))$. Consequently for $m \in M$, $(\delta m,w)
\rightarrow (DF[m]\delta m)w$ is continuous and bilinear $:TM \times
W^2 \rightarrow D^0$. The derivative of $\tJa$ is
\[
    D\tJa[m;d]\delta m = \langle 
  (DF[m]\delta m)\aw[m,d],F[m]\aw[m;d]-d\rangle_{D^0}
\]
\begin{equation}
  \label{eqn:grad1}
 +\langle D\aw[m;d]\delta m,F[m]^T(F[m]\aw[m;d]-d)+\alpha^2A^TA\aw[m;d]\rangle_{W^0}
\end{equation}
for any $\delta m \in TM$, provided that the derivatives appearing on
the RHS make sense, which they do if $d \in D^4$, as noted above. Note
that the second term on the RHS of equation \ref{eqn:grad1} vanishes
thanks to the normal equation \ref{eqn:condition}.

Given $w \in
W^2$, the adjoint of the continuous linear map $T_mM \rightarrow D^0$
given by $\delta m \rightarrow
(DF[m]\delta m)w$ will be denoted $DF[m]^*(w,\cdot)$: that is,
\begin{equation}
  \label{eqn:madj}
  \langle (DF[m]\delta m)w,d\rangle_{D^0} = \langle \delta m,
  DF[m]^*(w,d)\rangle_{TM}.
\end{equation}

With this convention, the result \ref{eqn:grad1} can be re-written
\begin{equation}
  \label{eqn:deriv}
  D\tJa[m;d]\delta m = \langle \delta m, DF[m]^*(\aw[m;d], F[m]\aw[m;d]-d)\rangle_{TM}
\end{equation}
Therefore

\begin{theorem}\label{eqn:vpmgrad}
  Under Assumptions A, B, and C,  $\tJa[\cdot;d] \in C^k(M)$ provided
  that $k \ge 1$ and $d \in D^{3+k}$. Its gradient is given by
 \begin{equation}
   \nabla _{m} \tJa[m; d] = D F[m]^*(\aw[m;d],F[m]\aw[m;d]-d))
   \label{eqn:grad2}
 \end{equation}
 \end{theorem}
The identity \ref{eqn:grad2} has precisely the form of the formula derived
\cite{GolubPereyra:03,GolubPereyra:73} for differentiable separable
least squares problems on $\bR^n$. The existence of higher derivatives
than the first under the circumstances indicated follows from very
similar reasoning to that used to establish the assertion for $k=1$.