### 2020-08-01 repro/papers/wri1d/rev1/madfigs.tex

\inputdir{project}

\plot{wavelet}{width=\textwidth}{Ricker Wavelet, zero phase, 20 Hz
  center frequency.}

\plot{data}{width=\textwidth}{Predicted data for 1D transmission
  problem, source wavelet = 20 Hz Ricker wavelet of Figure
  \ref{fig:wavelet}, source-receiver offset = 1 km, velocity = 2.5
  km/s.}

\plot{j0}{width=\textwidth}{Blue curve = FWI objective, red curve =
  VPM-WRI objective, plotted as functions of slowness, $\alpha$= 0.1
  s/$\sqrt{\rm km}$.}

\plot{j3}{width=\textwidth}{Blue curve = FWI objective, red curve =
  VPM-WRI objective, plotted as functions of slowness, $\alpha$= 1.0
  s/$\sqrt{\rm km}$.}