### 2020-08-19 repro/papers/discrep/paper.tex
interim

\title{Solution of an Acoustic Transmission Inverse Problem by Extended Inversion}
\author{William W. Symes, Orcas Island, Huiyi Chen* and Susan E. Minkoff, The University of Texas at Dallas}

\lefthead{Symes}

\righthead{Discrepancy}

\maketitle
\begin{abstract}
  Study of an extremely simple single-trace transmission example shows
  how an extended source formulation of full waveform inversion can
  produce an optimization problem without spurious local minima
  (``cycle skipping''), hence efficiently solvable via Newton-like
  local optimization methods. The data consist of a single trace
  recorded at a given distance from a point source. The velocity or
  slowness is presumed homogeneous, and the target source wavelet is
  presumed quasi-impulsive or localized near zero time lag. The source
  is extended by permitting energy to spread in time, and the spread
  is controlled by adding a weighted mean square of the extended
  source wavelet to the data misfit, to produce the extended inversion
  objective. The objective function and its gradient can be computed
  explicitly, and it is easily seen that the error in any local
  minimizer is bounded by a multiple of the permitted energy spread,
  which may be interpreted as a wavelength. The derivation shows several
  important features of all similar extended source algorithms. For
  example, nested optimization, with the source estimation in the
  inner optimization (variable projection method), is essential. The
  choice of the penalty operator, controlling the extended source
  degrees of freedom, is critical: in order to produce an objective
  immune from cycle-skipping, the penalty operator must be
  (pseudo-)differential. The penalty weight is chosen dynamically,
  according to a version of the Discrepancy Principle, in order to
  accelerate convergence to a near-global minimum. Like many other
  extended formulations, the example problem considered here permits
  the discrepancy criterion to be satisfied with zero penalty
  weight. This unusual feature removes a common obstacle to the use of
  the discrepancy principle, namely the determination of a
  satisfactory initial estimate of the penalty weight: the initial weight may
  be set to zero, and the dynamic algorithm updates the weight until
  convergence is achieved.

\end{abstract}

\section{Introduction}
Model-based parameter estimation by least squares data fitting is a
widely used and successful technique for data analysis in science and
engineering \cite[]{Bertero:98,Vogel:02}, particularly in the
geosciences \cite[]{Parker:94,Tarantola:05}. In its application to
seismology, least-squares data fitting has come to be known as Full
Waveform Inversion (FWI), and is now well-established as a useful tool
for probing the earth's subsurface
\cite[]{VirieuxOperto:09,Fichtner:10,Schuster:17}. However, in its
application to seismology, FWI encounters a serious
practical obstacle, so-called ``cycle-skipping''. This term signifies
the tendency of the mean-square error function to exhibit many
stationary points, and local minima far from the global minimizer (we
show an explicit example of this phenomenon later in this paper). Because of the
typical dimensions of field FWI problems, only local gradient-based
minimization algorithms are computationally practical, and these find
local minimizers, so may stagnate at suboptimal and geologically
uninformative earth models. Local descent methods avoid suboptimal
stagnation only if initial models are already quite close to optimal,
in the sense of predicting the arrival times of seismic events to
within a small multiple of the dominant data wavelength
\cite[]{GauTarVir:86,VirieuxOperto:09,Plessix:10}.

This paper examines one of a number of proposed alternatives to least squares
FWI, based on an extension of the model-to-data mapping.  We show
exactly how an example of this {\em extended inversion} approach leads
to successful solution of a very simple acoustic transmission inverse
problem. Despite its simplicity, this problem exhibits several
important characteristics of all inverse problems in wave propagation
in which wave velocities are amongst the unknowns. It has the
advantage of permitting explicit calculations that reveal clearly the
failure mechanism for FWI, that is, straightforward nonlinear least
squares, and also the reasons for success of an extended inversion
approach.

In the remainder of this introduction, we will briefly describe the conceptual
framework, the inverse problem, and sketch our main results. We leave out
some technical details that are supplied in the next section,
containing precise statements of the results.

The idealized physical setting is the propagation of wave through a
homogeneous acoustic fluid occuyping $\bR^3$ \cite[]{Frie:58}. The
fluid's salient characteristic for the present discussion is the
slowness, or reciprocal wave velocity, $m > 0$. A source of acoustic
energy localized at a point in $\bR^3$ radiates uniformly in all
directions with time-dependent intensity $w(t)$. Both the acoustic
pressure field and the intensity (``wavelet'') $w$ vanish at large
negative time. A receiver records the pressure values at a distance
$r>0$ from the source position, between the times $t_{\rm min}$ and
$t_{\rm max}$. As explained in the next section, the
recorded presssure is predicted to be
\begin{equation}
\label{eqn:mod}
\frac{1}{4\pi r}w\left(t-mr\right)  = F[m]w(t)
\end{equation}
Ignoring the amplitude factor $1/(4\pi r)$, this map implements a $m$-dependent time
shift. This time shift operator is the basis of many descriptions of
the cycle-skipping phenomenon (for example, \cite{VirieuxOperto:09},
Figure 7), so it is unsurprising that an analysis of cycle-skipping
can be based on the simple modeling operator described above.

Evidently if the wavelet $w$ is square-integrable, so is the predicted
recorded pressure - its $L^2$ norm is the cumulative power trasferred
from the fluid to the receiver \cite[]{SantosaSymes:00}. Thus this relation defines a map
$F: \bR^+ \times L^2(\bR) \rightarrow
L^2([t_{\rm min},t_{\rm max}])$, linear in its second argument.

A naive version of the inverse problem investigated in this paper
would be: given recorded data
$d \in L^2([t_{\rm min},t_{\rm max}])$, find $m \in \bR$ and $w \in L^2(\bR)$ so that
$F[m]w \approx d$. Clearly this problem statement is not interesting as it
stands, since it is always possible to solve it with no error and nonuniquely,
and it does not constrain the slowness $m$ at all. To be useful, the statement must
be augmented with a constraint on the model $(m,w)$. We choose 
$\lambda > 0$ and search for $w$ with $\mbox{supp }w \subset
[-\lambda,\lambda]$. The quantity $\lambda$ is to be considered
``small'': as we shall show, it is related to a mean wavelength of
$w$.

Abusing notation slightly, we identify $L^2([-\lambda,\lambda])$ with
a subspace of $L^2(\bR)$ via extension by $0$. Denote by $\lF$
restriction of $F$ to $\bR^+ \times L^2([-\lambda,\lambda])$ defined by equation
\ref{eqn:mod}. Then the FWI problem with support constraint,
described in the last paragraph, can be crudely phrased: given $d \in
L^2([t_{\rm min},t_{\rm max}])$, find $(m,w)$ in the domain of $\lF$ so that $\lF[m,w]
\approx d$.

Despite its simplicity, this problem exhibits the
cycle-skipping pathology characteristic of field-scale FWI. Specifically, we show that the reduced mean square
error function, the objective function of FWI for this problem, 
\begin{equation}
  \label{eqn:redms}
  (m,w) \mapsto \lerr[m,w,d]=\frac{1}{2}\|\lF[m]w-d\|^2,
\end{equation}
has local minimizers far from any global minimizer, even for noise-free
data (Theorem \ref{thm:fwi}).

This fact implies that local optimization must fail to solve a precise
version of the FWI problem: given $\epsilon, \lambda > 0$, $(m,w)$ is
a solution if it lies in the domain of $\lF$ and
$\lerr[m,w,d] \le 1/2 \epsilon^2\|d\|^2$ (Problem Statement
\ref{eqn:probstat0} below).  In fact, we show
that``large residual'' local minima of $\lerr[\cdot,\cdot,d]$ exist,
at which its value is much greater than its global minimum
value. Therefore if $\epsilon$ is near the global minimum value, then
local minimization of $\lerr[\cdot,\cdot,d]$ can fail to find a
solution of the inverse problem. This phenomenon is what is meant by
``cycle-skipping''.

Extended inversion involves an extension of the model-to-data map, in
this case $\lF$.  The extension used here consists in dropping the
support constraint on $w$, so the extended map is simply $F$ as
defined above ($\lF$ is a restriction of $F$, so $F$ is an extension
of $\lF$). To compensate for the resulting indeterminacy in the
estimation of $m$, we add a penalty term to the objective: for $m \in
\bR^+,w\in L^2(\bR), d \in L^2([t_{\rm min},t_{\rm max}])$, define
\begin{eqnarray}
  \label{eqn:pen}
  e[m,w;d] & = & \frac{1}{2}\|F[m]w-d\|^2,\\
  g[w] & = & \frac{1}{2}\|A\|^2, \\
  \Ja[m, w;d] &=& e[m,w;d] + \alpha^2 g[w] \nonumber\\
  &=& \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2). 
\label{eqn:pen}
\end{eqnarray}
The only obvious requirement on the penalty operator $A$ is that it be
nonzero for large $|t|$, so that minimizers of $\Ja$ should be
concentrated near $t=0$. Thus the penalty is a soft stand-in for the
support constraint. We
show that a good choice for $A$ is given by $(Aw)(t)=tw(t)$, and the
remainder of the discussion in this section makes this assumption. The
penalty weight $\alpha$ must be chosen somehow; we review a simple
algorithm for controlling $\alpha$ later in the paper.

Assume that $\mu>0$ and $d = F[m_*]w_* + n$, with $\mbox{supp }w_*
\subset [-\mu,\mu]$, and denote by $\eta = \|n\|/\|F[m_*w_*]$
the noise-to-signal ratio. We show that any local minimizer $(m,w)$
of $\Ja$ satisfies 
\begin{itemize}
\item If $\eta \le 0.6$,
  \begin{equation}
    \label{eqn:mbnd}
    |m-m_*| \le (1+f(\eta))\frac{\lambda}{r},
  \end{equation}
  where $f(\eta)=2\eta+O(\eta^2)$ (Theorem \ref{thm:mnoiseres});
\item 
  $(m,{\bf 1}_{[-\lambda,\lambda]}w)$ is a solution of the FWI problem
  as stated above, that is,
  \begin{equation}
    \label{eqn:success}
    \lerr[m, {\bf 1}_{[-\lambda,\lambda]}w,d] \le \frac{1}{2}\epsilon^2,
  \end{equation}
  so long as
  \begin{eqnarray}
    \lambda & \ge & (2+f(\eta))\mu,\nonumber\\
    \epsilon &\ge & (4\pi r \alpha \lambda)^2 +\eta.
    \label{eqn:solcond}
  \end{eqnarray}
  (Theorem \ref{thm:ipnoisesuf})
\end{itemize}.
That is, the minimization of the extended objective $\Ja$ provides a
solution of the FWI problem, with constraints on the achievable
performance and bounds on the errors. All of these
bounds are shown to be sharp, and the constraints cannot be
significantly weakened. For example, the support radius $\lambda$ in
the projection onto $L^2([-\lambda,\lambda])$ must in general be at least twice the
support radius of the ``noise free'' wavelet $w_*$, plus an additional
allowance for noise.

The remainder of the paper begins with a complete description of the acoustic transmission problem and precise
statements of our main results, filling in the
few details not discussed in the preceding sketch. Then we provide an
discussion of the potential theoretical and practical import of these
results, including relations with prior work, some numerical
illustrations, and directions for further research. Following this
discussion section, we give complete proofs of all of our results.
An appendix describes the computational algorithm used in producing
the numerical examples, including a method for controlling the penalty
weight $\alpha$ based on the Discrepancy Principle
\cite[]{EnglHankeNeubauer,FuSymes2017discrepancy}. A second appendix
explains the relation between the support constraint and well-known
criteria for success of FWI based on data frequency content. The basis
for this relation is the Heisenberg inequality \cite[]{Folland:07}.

\section{An Acoustic Transmission Inverse Problem}
In this section, we describe solution of the radiation
problem for acoustic wave
propagation in a homogeneous fluid, and various operators based on
this solution. Using these operators, we formulate the single-trace
acoustic transmission inverse problem. Finally, we state the main
results of our paper.

We shall use the abbreviations ${\cal B}(X,Y)$ and ${\cal I}(X,Y)$ for 
the algebra of bounded linear operators from the Hilbert space $X$ to 
the Hilbert space $Y$, and its subalgebra of invertible operators.

\subsection{Acoustic transmission}
Assume small amplitude (linearized) acoustic propagation, constant
density, and isotropic point source and receiver. Denote by $m(\bx)$
the slowness (reciprocal velocity) at spatial position $\bx$, $w(t)$
the time dependence of the point source (``wavelet'') at location
$\bx=\bx_s$. Then the (excess) pressure field $p(\bx,t)$ obeys a
scalar wave equation:
\begin{eqnarray}
  \label{eqn:awe}
  \left(m(\bx)^2\frac{\partial^2 p}{\partial t^2} - \nabla^2\right) p(\bx,t) &=&
                                                                         w(t)\delta(\bx-\bx_s) \nonumber\\
  p(\bx,t)&=&0, t\ll 0
\end{eqnarray}
Suppose that a single trace is recorded, at distance $r>0$ from the
source position $\bx_s$. The dominant information in a single
trace is the transient signal time of arrival, constraining only the mean slowness in the
region
between source and receiver, so assume that the
slowness is constant, that is, independent of position $\bx$. The pressure field is simply the  the source
wavelet $w(t)$ convolved with the
acoustic Green's function, for which an analytic expression is
available in the constant $m$ case \cite[]{CourHil:62}:
\begin{equation}
  \label{eqn:homsol}
  p(\bx,t) = \frac{1}{4\pi |\bx-\bx_s|}w\left(t-m|\bx-\bx_s|\right).
\end{equation}

The receiver location $\bx_r$ lies at distance $r$ from the source
location $\bx_s$, that is, $|\bx_r-\bx_s|=r$. The predicted signal at
$p(\bx_r,t)$ depends nonlinearly on the slowness $m$ and linearly on the
source wavelet $w$. Therefore it is naturally represented as the
action of a $m$-dependent linear operator $F[m]$ on $w$, as described
in \ref{eqn:mod}.
%\begin{equation}
%\label{eqn:mod}
%F[m]w(t)= p(\bx_r,t) = \frac{1}{4\pi r}w\left(t-mr\right).
%\end{equation}

The slowness $m$ must be positive, as follows from basic acoustics,
and in fact reside in a range $M=(m_{\rm min}, m_{\rm max})$ characteristic of the
material model: for crustal rock, a reasonable choice would be
$m_{\rm min}=0.125, m_{\rm max}=0.6$ s/km.

Field recording takes place over finite time intervals (and with
finite sample rate, a detail we shall ignore in this paper). Denote by
$[t_{\rm min},t_{\rm max}]$ the recording interval for the
hypothetical single trace data. Then natural choices for domain and
range of $F[m]$ are $W = L^2(\bR)$ and $D=L^2([t_{\rm min},t_{\rm
  max}])$ respectively (that is, interpret the definition \ref{eqn:mod}
as holding for $t \in [t_{\rm min},t_{\rm  max}]$). So
\begin{equation}
  \label{eqn:mapprop}
  \mbox{for }m \in M, F[m] \in {\cal B}(W,D),\mbox{ and }\|F[m]\| =
  \frac{1}{4\pi r}.
\end{equation}
Note also that $F[m]$ is surjective for every $m \in M$.

In computational practice, $W$ will have to be replaced by a
finite-dimensional subspace of $L^2(\bR)$. Many such choices will
implicitly limit the support of $w \in W$ to a bounded interval, say
$[T_{\rm min},T_{\rm max}]$. To maintain the surjective property,
these bounds should be chosen so that $[t_{\rm min}, t_{\rm max}]
\subset [T_{\rm min}+mr,T_{\rm max}+mr]$ for all $m \in M$, that is,
\begin{eqnarray}
  \label{eqn:dombounds}
  t_{\rm min} &\ge & T_{\rm min}+m_{\rm max}r,\nonumber\\
  t_{\rm max} &\le & T_{\rm max}+m_{\rm min}r.
\end{eqnarray}
We will ignore these computational necessities in this work,
maintaining the definition $W=L^2(\bR)$.

\subsection{An inverse problem}

Part of the definition of an inverse problem based on the model
definition \ref{eqn:mod} is that observed data $d \in D$ be fit: that
is, $m, w$ must be found so that $F[m]w \approx d$. However this
formulation does not provide enough information to determine the model
$(m,w) \in M \times W$: indeed, the definition
\ref{eqn:mod} shows that $F[m]$ is surjective for any $m \in M$, so
the data can always be fit precisely by appropriate choice of $w$, and
$m$ is completely
unconstrained.  In order that data fit constrain $m$, it is necessary
to further constrain $w$ {\em a priori} - that is, to add information
beyond the trace data $d$.

One natural constraint is to assume that the support of $w$ is
limited. An assumption of small support may be justified as
follows. In practical seismic data processing, actual sources may act
over considerable time intervals - for example, the most commonly used
marine source (airgun) produces an oscillating pulse that dies away
slowly. It is commonplace to estimate this pulse then deconvolve it
from the data by safeguarded Fourier division or other means. This
so-called signature deconvolution results in modified data
corresponding to a source equal to signature deconvolution of the
pulse estimate itself. The deconvolved source pulse approximates an
impulse ($\delta(t)$): it cannot have point support, due to the finite
frequency nature of seismic data, hence the bandlimited nature of the
pulse estimate, but its support is ultimately much smaller than that
of the original pulse estimate.

Therefore, supplement the observed data $d \in D$ with a presumed 
source support radius $\lambda$. Define 
$\lW=\{w \in W: \mbox{ supp }w \subset [-\lambda,\lambda]\}$, and 
denote by $\lF$ the restriction of $F$, defined in equation 
\ref{eqn:mod}, to $M \times \lW$.

Note that for $w \in \lW$, $m \in M$,
$\mbox{supp }F[m]w \subset [ m_{\rm min}r-\lambda,
m_{\rm max}r+\lambda] \cap
[t_{\rm min},t_{\rm max}]$. Also, $\lF[m]$ is coercive for all $m \in
M$ only if
$[ m_{\rm min}r-\lambda,m_{\rm max}r+\lambda] \subset
[t_{\rm min},t_{\rm max}]$.

We will assume that $\lambda_{\rm max}
>0$ is chosen so that $M, r,\lambda_{\rm max}$ and $[t_{\rm min},t_{\rm max}]$ satisfy
\begin{equation}
  \label{eqn:fullrec}
  [ m_{\rm min}r-\lambda_{\rm max}, m_{\rm max}r+\lambda_{\rm max}]
  \subset [t_{\rm min},t_{\rm max}].
\end{equation}
Then for 
$m \in M$, $0 <\lambda \le \lambda_{\rm max}$, $\lF[m] \in {\cal B}(\lW,D)$ is coercive.

Note that condition \ref{eqn:fullrec} implies
\begin{eqnarray}
  \lambda_{\rm max} & \le & t_{\rm max}-m_{\rm max}r,\nonumber \\
  t_{\rm min}-m_{\rm min}r & \le & -\lambda_{\rm max} \label{eqn:lminc}
\end{eqnarray}
hence for any $m \in M$,
\begin{equation}
  [-\lambda_{\rm max}, -\lambda_{\rm max}] \subset[ t_{\rm min}-mr ,
  t_{\rm max}-mr].
  \label{eqn:zeroinc}
\end{equation}

In terms of this infrastructure, the inverse problem studied in 
this paper is:

\noindent {\bf Problem Statement:} given data $d \in D$, relative 
error level $\epsilon \in [0,1)$, and support radius $\lambda \in (0, 
\lambda_{\rm max}]$, find 
$(m,w) \in M \times \lW$ for which 
\begin{equation}
  \label{eqn:probstat0}
  \|\lF[m]w-d\| \le \epsilon\|d\|. 
\end{equation}

\noindent{\bf Remark:} The constraint $\epsilon < 1$ imposed on the
target noise level is eliminates the obvious choice $(m,0)$, which
satisfies the data misfit constraint for any $m \in M$.

\noindent{\bf Remark:} The support constraint is closely linked to the
folk theorem about FWI noted many times in the literature: convergence
of a descent method requires that the initial slowness must be known
to ``within a (fraction of a) wavelength''. The relation is a
consequence of Heisenberg's inequality, and will be reviewed in
Appendix B.

The best case for data fitting
is clearly the one in which the data can be fit precisely: that is,
there exists $(m_*,w_*) \in M\times \lW$ so that
\begin{equation}
  \label{eqn:defdatanonoise}
  d=\lF[m_*]w_*.
\end{equation}
Such data $d$ is {\em noise-free}, in the range of the map $\lF$. For
such data a solution of the Problem Statement
exists with arbitrarily small $\epsilon>0$.

\subsection{Soft constraints and penalty formulation}

As mentioned in the introduction, a straightforward nonlinear least
squares formulation of the problem stated in \ref{eqn:probstat0}
exhibits local minimizers with large fit error, and these will be
found by any local optimization algorithm unless it is initialized at
a quite accurate initial model. The first result stated in the next
section establishes this assertion for the single trace transmission
problem, however it appears to be valid for field scale problems, for
essentially the same reasons, as widely reported in the literature on
seismic inversion.

One (of many) alternative optimization formulation of this and similar
problems involves introducing an extension of the model-to-data
operator, and compensating for the physically spurious degrees of
freedom so introduced by adding a penalty term controlling them to the
objective.  The application of this idea presented here extends $\lF$
by dropping the support constraint on its second argument, that is,
using the expression \ref{eqn:mod} to define a map $F: M \times W
\rightarrow D$. The objective begins with an extended version of the
mean-square error,
\begin{equation}
  \label{eqn:mserr}
  e[m,w;d]=\frac{1}{2}\|F[m]w-d\|^2.
\end{equation}
We will choose a penalty to control the extra degrees of freedom, that
is, $w$ outside of $[-\lambda,\lambda]$. This  penalty is quadratic, constrains only $w\in W$ (as it is a soft
version of the condition $w \in \lW$), and is
$m-$independent. Therefore the penalty function takes the form
\begin{equation}
\label{eqn:pen}
p[w] = \frac{1}{2}\|Aw\|_Y^2 
\end{equation}
where $A \in {\cal B}(W,Y)$, $Y$ being another suitable Hilbert
space.

The penalty approach asks that a linear combination of $e,p$,
depending on a relative weight $\alpha$:
\begin{equation}
\Ja[m, w;d] = e[m,w;d] + \alpha^2 p[w] = \frac{1}{2}(\|F[m]w-d\|^2 + \alpha^2 \|Aw\|^2)
\label{eqn:pen}
\end{equation}
be minimized over $(m,w)\in M \times W$.

The choice of penalty weight $\alpha$ has a profound influence on the
character of this optimization problem. For the moment, we will mandate
only that $\alpha \ge 0$. A complete description of an algorithm using
the penalty function \ref{eqn:pen} to solve the inverse problem
\ref{eqn:probstat0} must include a selection principle for $\alpha$. We
describe one such principle in Appendix A.

The main theoretical device used in the proofs of our main results on
extended inversion is the reduction of $\Ja$ to a function $\tJa$ of
$m$ alone, by minimization over $w$. This reduction is the foundation
of the {\em Variable Projection Method} (VPM)
\cite[]{GolubPereyra:03} for solution of partly linear (``separable'')
least squares problems. We use the VPM in our numerical illustrations
of the theory, as described in Discussion section and Appendix A.

\begin{theorem}
\label{thm:red}
Assume that the Hessian operator of
$\Ja[m,\cdot;d]$ is invertible:
\begin{equation}
\label{eqn:norminv}
F[m]^TF[m] +\alpha^2 A^TA \in {\cal I}(W,W).
\end{equation}
(The superscript $T$ denotes the adjoint with respect to the inner products 
in $W=L^2(\bR)$ and $D=L^2([t_{\rm min},t_{\rm max}])$.)
Then the unique stationary point $\aw[m;d] \in W$ of
$\Ja[m,\cdot;d]$ is the solution of the {\em normal equation}:
\begin{equation}
  \label{eqn:norm}
  (F[m]^TF[m]+\alpha^2A^TA)w= F[m]^Td,
\end{equation}
Define the reduced or variable projection objective by
\begin{equation}
  \label{eqn:redexp}
  \tJa[m;d] =\Ja[m,w_{\alpha}[m;d];d] = \min_w \Ja[m,w;d]
\end{equation}
Then $(m,w)$ is a local minimizer of $\Ja[\cdot,\cdot;d]$ if and only
if $m$ is a local minimizer of $\tJa[\cdot;d]$ and $w=\aw[m;d]$.
\end{theorem}
\noindent{\bf Remark:} If $\Ja[\cdot,\cdot;d]$ and $\tJa[\cdot;d]$ are differentiable, then
  ``local minimizer'' in the conclusion of the preceding theorem can
  be replaced by ``stationary point''. However, for the problem
  addressed in this paper, $\Ja[\cdot,\cdot;d]$ {\em is not}
  differentiable without added smoothness constraints on $w$, whereas
  $\tJa[\cdot;d]$ {\em is} differentiable for proper choice of penalty
  operator $A$, as will be shown
  below. This phenomenon is replicated for a wide variety of extended
  approaches for inverse problems in wave propagation. We will have
  more to say about it in the Discussion section.

\subsection{Main Results}
The first main result establishes the existence of large (100 \%)
residual local minimizers for the basic FWI objective $\lerr$, even
for noise-free data.
\begin{theorem}
  \label{thm:fwi}
  Suppose that $0 <\lambda\le \lambda_{\rm max}$,  $m_* \in M, w_*
  \in \lW, d=\lF[m_*]w_*$ is noise-free data per definition \ref{eqn:defdatanonoise},
  Under assumption \ref{eqn:fullrec}, for any $m \in M$ with $|m-m_*|r>2\lambda$,
\begin{equation}
  \label{eqn:isovpm}
 \min_w \lerr[m,w;d]=\lerr[m,0,d]=\frac{1}{2}\|d\|^2,
\end{equation}
and any such $(m,0)$ is a local minimizer of $e$.
\end{theorem}

Turning to extended inversion, we examine two choices of penalty
operators. For each choice, we ask first whether local minimizers 
of the resulting VPM
objective $\tJa[\cdot;d]$ occur far from a slowness $m_*$. Both choices for $A$ considered here are scalar 
multiplication operators defined by a choice of positive multiplier $a \in L^{\infty}(\bR)$:
\begin{equation}
  \label{eqn:annmult}
  A w(t)= a(t)w(t), \, t\in \bR: \,a(t) > 0 \mbox{ for }t \ne 0.
\end{equation}

A penalty operator $A$ of which $\lW$ is the null space would be a
natural choice. Such operators have come to be called
``annihilators'', since they map all members of the constraint
subspace $\lW$ to zero.  A simple example is
\begin{eqnarray}
  A = E^c_{\lambda}&=&I - E_{\lambda},\mbox{ where } \nonumber \\
  E_{\lambda}w(t) &=&{\bf 1}_{[-\lambda,\lambda]}(t)w(t). 
                      \label{eqn:ann0}
\end{eqnarray}
That is, $E_{\lambda}$ is the orthogonal projector onto $\lW$,
and $E_{\lambda}^c$ is the orthogonal projector onto its
orthocomplement, an operator of the form \ref{eqn:annmult} with $a
= 1 - {\bf 1}_{[-\lambda,\lambda]}$. %Note that this choice of $a$
%satisfies condition \ref{eqn:abnd} with $C \equiv 1$.

\begin{theorem}
  \label{thm:boxcarbad}
  Suppose that
  \begin{itemize}
  \item[1. ] $\lambda \in (0,\lambda_{\rm max}]$;
  \item[2. ] $m_*\in M, w_*\in \lW$, and $d=F[m_*]w_*$ (noise-free
    data);
  \item[3. ] $A=E^c_{\lambda}$, that is, $a=1-{\bf 1}_{-\lambda,\lambda]}$ in
    the definition \ref{eqn:annmult}.
  \end{itemize}
  Then if $|m-m_*| >  2\lambda/r$,
  \[
    \tJa[m;d] = \frac{\alpha^2}{2(1+(4 \pi r)^2 \alpha^2)}\|w_*\|^2.
  \]
\end{theorem}

\noindent {\bf Remark.}
One might have thought that $A=E^c_{\lambda}$ would be a better choice
of annihilator, as for noise-free
data, the solution set defined by the problem statement
\ref{eqn:probstat0} is the same as the set of global minimizers of
$\tJa$ in this case. However, for this choice of annihilator,
$\tJa$ exhibits the same feature as the mean square error $e$, namely
a continuum of stationary points at any distance from the global
minimizer $m_*$ greater than a multiple of $\lambda$. Therefore the
extended inversion with this choice of annihilator is no more amenable
to local optimization than is FWI.

A second possible 
penalty operator penalizes energy away from 
$t=0$: choose $\tau > 0$ and set 
\begin{equation}
  \label{eqn:ann}
  a(t) = \min(|t|, \tau). 
\end{equation}
The cutoff $\tau$ will be chosen large enough to be effectively inactive: 
specifically, hindsight suggests 
\begin{equation}
  \label{eqn:taudef}
  \tau = \max\{|t_{\rm min}-m_{\rm min}r|,|t_{\rm min}-m_{\rm max}r|, |t_{\rm max}-m_{\rm min}r|, |t_{\rm max}-m_{\rm max}r|\}. 
\end{equation}
This 
particular annihilator has been employed in earlier papers on extended 
source inversion 
\cite[]{Plessix:00a,LuoSava:11,Warner:14,HuangSymes:SEG15a,Warner:16,HuangSymes:GEO17}.

\begin{theorem}
  \label{thm:mnoiseres}
  Asumme that
  \begin{itemize}
  \item[1. ] $\alpha, \mu> 0$,
  \item[2. ] $m_* \in M, w_* \in W_{\mu}$,
  \item[3. ] $d_* = F[m_*]w_*$,
  \item[4. ] $n \in D$ and $d = d_* + n$.
  \end{itemize}
  Set $\eta = \|n\|/\|d_*\|$, and assume further that
  \begin{itemize}
  \item[5. ]
    \begin{equation}
      \label{eqn:mnoisecond}
      \eta < \frac{\sqrt{5}-1}{2}, and
    \end{equation}
  \item[6. ] $m$ is a stationary point of $\tJa[\cdot;d]$.
  \end{itemize}
  Then
  \begin{equation}
    \label{eqn:mnoisesuff}
    |m-m_*| \le \left(1+\frac{2\eta(1+\eta)}{1-\eta(1+\eta)}\right)\frac{\mu}{r}.
  \end{equation}
\end{theorem}

\noindent {\bf Remark.} That is, {\em with the choice of penalty
  multiplier $a$ given in equation \ref{eqn:ann}, and support radius
  $\mu$ of the ``noise-free'' wavelet $w_*$, $\tJa$ has no local
  minima further than $(1+O(\eta))\mu/r)$ from the slowness used to
  generate the data}.

\noindent {\bf Remark.} The estimate $|m-m_*|r<\mu(1+O(\eta))$ for stationary
points $m$ of $\tJa$ is sharp: it is possible to choose $w_* \in W_{\mu}$
so that $\mu - |m-m_*|r$ is as small as you like. In particular,
the ``exact'' or ``true'' slowness $m_*$ is not necessarily the only 
stationary point, or even a stationary point at all, and in particular
is not (necessarily) a global minimizer of $\tJa$.

\noindent {\bf Remark.} No similar bound could hold for much larger
noise levels than specified in condition \ref{eqn:mnoisecond}, the
right-hand side of which is a bit larger than 0.6. For example, if the noise is
the predicted data for the same wavelet $w_*$ with a substantially different
slowness $m_{\flat}$, that is, $n=F[m_{\flat}]w_*$, then by symmetry
the average slowness $0.5(m_*+m_{\flat})$ is a stationary point of
$\tJa[\cdot,d_*+n]$. The difference with $m_*$ is not constrained at
all by the support of $w_*$. So for this example with 100\% noise, no
bound of the type given by conclusion 2 could possibly hold. We will
illustrate this phenomenon numerically later in the paper.

\noindent{\bf Remark.} Note that $\alpha$ plays no role in the
conclusions of this theorem. It is only required that $\alpha >0$.

Unless the data is noise-free, there is no reason to suppose that the
estimated wavelet $\aw[m;d]$ (Theorem \ref{thm:red}) will lie in
$\lW$, unless the support of the noise $n$ is not restricted. In order
to construct a solution of the inverse problem \ref{eqn:probstat0}, we
project $\aw[m;d]$ onto $\lW$. For sufficiently large $\lambda,
\epsilon$, the result is a solution of the inverse problem:

\begin{theorem}
  \label{thm:ipnoisesuf}
  Assume the hypotheses of Theorem \ref{thm:mnoiseres}, and that
  inequality \ref{eqn:mnoisecond} holds, and $\mu \in
  (0,\lambda_{\rm max}]$. Then the pair
  \[
    (m,{\bf 1}_{[-\lambda,\lambda]}\aw[m,d])
  \]
  solves the inverse problem as stated in \ref{eqn:probstat0} if
  \begin{equation}
    \label{eqn:ipnoiselam}
    \left( 2+\frac{2\eta(1+\eta)}{1-\eta(1+\eta)}\right)\mu \le \lambda
    \le \lambda_{\rm max}, 
  \end{equation}
  and
  \begin{equation}
    \label{eqn:ipnoiseeps}
    \epsilon \ge (4 \pi r \alpha\lambda)^2(1+\eta^2)^{1/2} 
  \end{equation}    
\end{theorem}

\noindent {\bf Remark:} Note that the sufficient condition \ref{eqn:ipnoiselam} for
$\lambda$ is independent of $\alpha$. It follows that for any choice
of $\lambda$ consistent with this bound, $(m,{\bf
  1}_{[-\lambda,\lambda]}\aw[m,d])$ is a solution of the inverse
problem for any $\epsilon > 0$ provided that $\alpha$ is chosen
sufficiently small ($O(\sqrt{epsilon})$).

\section{Discussion}

Use of model extension to reveal the influence of earth mechanical
parameters on data is an old idea, and underlies several classes of
seismic data processing methods in widespread use for decades
\cite[]{geoprosp:2008}. Use of model extension to create optimization
principles for inversion is a more recent development
\cite[]{SymesCar:91,Plessix:00a,Symes:09,LuoSava:11,BiondiAlmomin:SEG12,LeeuwenHerrmannWRI:13,LiuSymesLi:14,Warner:14,ChaurisGP:14,Warner:16,LeeuwenHerrmann:16,Herve2017,HouSymes:Geo18}. Extended
inversion methods differ by the choice of additional degrees of
freedom, and by choice of penalty applied to eliminate them in the
final result. The approach studied in this paper belongs to the {\em
  source extension} variety: the extra parameters are provided by
permitting the energy source component of the experimental model to
have more, or less constrained, parameters than the experimental
design suggests. \cite{HuangNammourSymesDollizal:SEG19} give an
overview and taxonomy of source extension methods, with extensive
references.

Extended inversion is not the only alternative to straightforward
least-squares data fitting that may overcome cycle-skipping. For
example, evidence has recently emerged that the Wasserstein metric
from arising in the theory of optimal transport may provide a measure
of error between model-predicted and observed seismic data less
oscillatory than the mean-square
\cite[]{EngquistYang:GEO18,Metivier:GEO18}.

Many inversion concepts sound plausible, and appear
to work at least to some extent as one might hope from their heuristic
justifications. However very few of these approaches have been
underwritten by mathematical argument: in essence, they are mostly
justified only by numerical example, ``in the rear-view mirror'', with no assurance that
failure is not just around the corner, at the next example. On top of
that, some of these approaches, for example those based on the
computationally attractive Variable Projection Method (``VPM'') of
\cite{GolubPereyra:03}, are cast in such form that the reasons for
success are not readily apparent.

In many cases, the constraint posed by membership of the second component $w$ in the subspace $\lW$ is an approximation to an idealized limit $\lambda \rightarrow 0$.
It may be that $W_0 = \cap_{\lambda > 0}\lW$ is non-trivial, in which
case $F$ may be viewed as an extension of $F_0 = F|_{ M \times
  W_0}$. In that case, $F_0$ and $F$ realize the idealized extension
structure discussed in many prior works on extended inversion (for
example, \cite[]{geoprosp:2008}): $F_0$ is the physical modeling
operator, and $W_0$ consists of second components of physical or
feasible models. However, in many of these examples, $W_0$ is not
computationally accessible, or - even if it is - gives results that
closely approximate those obtained with $\lW$ for small $\lambda$. For
other examples, for instance that  discussed later in this paper (and
implicitly in \cite[]{Warner:16}, for instance), $W_0=\{0\}$ and the
family $\{\lW: \lambda > 0\}$ actually embodies all of the necessary
physics.

While minimization of $\Ja$ might be tackled directly - by
alternating minimizations over $m$ and $w$, or by computing updates
for $m$ and $w$ simultaneously - such joint minimization performs
poorly, as \cite{YinHuang:16} has shown. The reason for this poor
performance is that $\Ja$ has dramatically different
sensitivity to $m$ versus $w$, as observed in the cited reference and elsewhere.
%especially for high frequency $f$, as
%the reader will see below.
%as will be established below.
Instead, a nested approach, in which $w$ is
eliminated in an inner optimization,
generally gives far better numerical performance.  This takes advantage of
$\Ja$ being quadratic in
$w$ to eliminate $w$ from the problem statement, as described in the
following theorem.
%\section{Theory}

%\cite{FuSymes2017discrepancy} introduce a version of the discrepancy
%principle and its application to separable nonlinear least squares
%problems in penalty form \ref{eqn:basic}. The algorithm developed
%there is incomplete, however: it lacks  termination criteria. In this
%section, I review the discrepancy-based algorithm and introduce an
%appropriate stopping rule, thus completing the development of the
%algorithm presented in \cite{FuSymes2017discrepancy}.

\section{Full Waveform Inversion: hard constraint}
A natural formulation of the Problem Statement \ref{eqn:probstat0} is
via least squares: that is, given $d \in D$, $\lambda >0$, find
$(m,w) \in M \times \lW$ to minimize
$\lerr: M \times W \rightarrow \bR$, defined in display
\ref{eqn:redms}.  A pair $(m,w) \in M\times \lW$ satisfies the
condition \ref{eqn:probstat0} iff $\lerr[m,w;d] \le \epsilon^2/2$, so a
solution exists iff a global minimizer of $\lerr$ satisfies this
inequality.

In this discussion, I will presume that $d$ is 
noise-free, as defined at the end of the last section (equation
\ref{eqn:defdatanonoise}). Thus {\em a priori} a global minimizer of
$e$ is a solution of the Problem Statement for any
$\epsilon > 0$. The only
question is how to find such a global minimizer. As mentioned in the
introduction, the simple model problem studied here is a proxy for
large-scale data fitting problems for which the only computationally
feastible approach to minimization of analogs of $e$
is some form of descent-based iterative local optimization. Any such
algorithm requires the specification of an initial iterate
$(m_0,w_0)$. Therefore the key question is: what
constraints must be imposed on $(m_0,w_0)$ in order that the resulting
sequence of iterates convertes to a global minimizer?

To answer this question, note that 
\[
 \lerr[m,w;d] =  \frac{1}{32\pi^2
    r^2}\int_0^T\,dt\,\left|w\left(t-mr\right)-w_*\left(t-m_*r\right)\right|^2
\]
Since $w_*, w$ vanish for $|t|>\lambda$,
$\lF[m_*]w_*(t)$ vanishes if $|t-m_*r|>\lambda$ and $\lF[m]w$ vanishes if $|t-mr|>\lambda$. So if $|mr-m_*r|
= |m-m_*|r > 2\lambda$, then $|t-mr|+|t-m_*r| \ge |mr-m_*r| >
2\lambda$ so either $|t-mr|>\lambda$ or $|t-m_*r|>\lambda$, that is,
either $\lF[m]w(t)=0$ or $\lF[m_*]w_*=0$. Therefore $\lF[m]w$ and
$\lF[m_*]w_*$ are orthogonal in the sense of the $L^2$ inner product
$\langle \cdot,\cdot \rangle_D$ on $D$:
\begin{equation}
  \label{eqn:ortho}
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, \langle F[m]w,
  F[m_*]w_*\rangle_D = 0
\end{equation}
But $d = \lF[m_*]w_*$, so this is the same as saying that $d$ is
orthogonal to $F[m]w$. So conclude after a minor manipulation that
\begin{equation}
  \label{eqn:iso}
  |m- m_*|r > 2\lambda \,\, \Rightarrow \,\, e[m,w;d]=\frac{1}{32\pi^2
    r^2}(\|w\|^2 + \|w_*\|^2).
\end{equation}
That is, for slowness $m$ in error by more than $2\lambda/r$ from the 
target slowness $m_*$, the means square error (FWI objective) $e$ is independent of
$m$, and its minimum over $w$ is attained for $w=0$:
\begin{theorem}
  \label{thm:fwi}
  Suppose that $0 <\lambda\le \lambda_{\rm max}$,  $m_* \in M, w_*
  \in \lW, d=\lF[m_*]w_*$ is noise-free data per definition \ref{eqn:defdatanonoise},
  Under assumption \ref{eqn:fullrec}, for any $m \in M$ with $r|m-m_*|>2\lambda$,
\begin{equation}
  \label{eqn:isovpm}
 \min_w e[m,w;d]=e[m,0,d]=\frac{1}{32\pi^2 r^2}\|w_*\|^2,
\end{equation}
and any such $(m,0)$ is a stationary point of $e$.
\end{theorem}

Therefore local minimizers of $e$ abound, as far as you like from the
global minimizer $(m_*,w_*)$. Local exploration of the FWI objective
$e$ gives no useful information whatever about constructive search
directions, and descent-based optimization tends to fail if the
initial estimate $m_0$ is in error by more than $2\lambda/r$
(``further than a multiple of a wavelength'', per discussion
above). In fact the actual behaviour of FWI itererations is worse
(failure if $m_0$ is in error by ``half a wavelength''), as follows
from a more refined analysis of the ``cycle-skipping'' local behaviour of $e$ near its
global minimizer.

\section{All stationary points are (near) global minimizers -
  sometimes}
It is not immediately apparent why use of a soft constraint should
produce an optimization problem more amenable to descent methods than
does the hard-constrained inverse problem (FWI). In this section, I
will show that the penalty formulation can yield a reduced problem
with for which any stationary point is close to a global minimizer,
{\em for some choices of penalty operator, but not for other choices
  that might appear equally or even more suitable}.

For the single-trace transmission inverse problem, $\tilde{J}_{\alpha}$ is explicitly
computable. First observe that apart from amplitude, $F[m]$ is
unitary: for $g \in D$,
\begin{equation}
\label{eqn:tran}
F[m]^T g (t) =
\left\{
  \begin{array}{c}
    \frac{1}{4\pi r}g\left(t+mr\right), \, t \in [t_{\rm min}-mr,
    t_{\rm max}-mr],\\
    0, \mbox{ else.}
  \end{array}
\right.
\end{equation}
so
\begin{equation}
  \label{eqn:unit}
  F[m]^TF[m] = \frac{1}{(4\pi r)^2}{\bf 1}_{[t_{\rm min}-mr,  
    t_{\rm max}-mr]}
\end{equation}
in which ${\bf 1}_{S}$ denotes
multiplication by the characteristic function of a measurable 
$S \subset \bR$.

Therefore the normal equation for the minimizer on the RHS of equation \ref{eqn:redexp} is
\begin{equation}
  \label{eqn:norm1}
  \left(\frac{1}{(4\pi r)^2} {\bf 1}_{[t_{\rm min}-mr,  
      t_{\rm max}-mr]} + \alpha^2 A^TA\right)w= F[m]^Td.
\end{equation}

The role of $A$ is to penalize distance from the constrained
subspace $\lW$, in some sense.

With these choices, the normal equation \ref{eqn:norm1} becomes
\begin{equation}
\label{eqn:norm2}
\left(\frac{1}{(4\pi r)^2}  {\bf 1}_{[t_{\rm min}-mr,  
      t_{\rm max}-mr]} + \alpha^2a^2\right)w= F[m]^Td.
\end{equation}

\begin{theorem}
  \label{thm:norminv}
  Assume the conditions \ref{eqn:mod}, \ref{eqn:fullrec},
  \ref{eqn:annmult}. Also assume that $\lambda \in (0,\lambda_{\rm
    max}], \alpha > 0,$ and that $a \in L^{\infty}(\bR)$
  mentioned in condition \ref{eqn:annmult} satisfies
  \begin{equation}
    \label{eqn:abnd}
    |t| > \lambda \Rightarrow a(t) \ge C,
  \end{equation}
  in which $C>0$ may depend on $\lambda$.
  Then
  \begin{itemize}
  \item[1. ]the normal operator $F[m]^TF[m] + \alpha^2A^TA$ is
    invertible for any $m \in M$, $\alpha > 0$;
  \item[2. ]for any $d \in D$, the normal equation \ref{eqn:norm} has
    a unique solution $\aw[m;d]\in W$, and 
    \begin{equation}
      \label{eqn:normsol}
      \aw[m;d](t) = \left\{
        \begin{array}{c}
          \left(\frac{1}{(4\pi r)^2} + \alpha^2
          a^2(t)\right)^{-1}\frac{1}{4 \pi r}d(t+mr), t \in [t_{\rm
          min}-mr, t_{\rm max}-mr];\\
          0, \mbox{ else;}
        \end{array}
      \right.
    \end{equation}
  \item[3. ]if in addition $d=F[m_*]w_*, w_* \in \lW$ is noise-free, as in equation
    \ref{eqn:defdatanonoise},
    \begin{equation}
      \label{eqn:solnonoise}
      \aw[m,d](t)= \left(1+ (4\pi r)^2\alpha^2 a(t)^2\right)^{-1}w_*\left(t+(m-m_*)r\right).
    \end{equation}
  \end{itemize}
\end{theorem}

\begin{proof}
  \begin{itemize}
  \item[1. ]Note that thanks to \ref{eqn:zeroinc}, if $|t|\le
    \lambda \le \lambda_{\rm max}$, then ${\bf 1}_{[t_{\rm min}-mr,  
      t_{\rm max}-mr]}(t) = 1$, whereas if $|t|>\lambda$,
    then $a(t) \ge C$, whence
    \[
      \frac{1}{(4\pi r)^2}  {\bf 1}_{[t_{\rm min}-mr,  
        t_{\rm max}-mr]} + \alpha^2a^2  \ge \min\{(4\pi r)^2,
      \alpha^2\min\{1/(4\pi r)^2,C^2\} \}> 0.
    \]
    Therefore the normal operator is invertible under the stated
    conditions.

  \item[2. ]From the identity \ref{eqn:tran}.
    \[
      \mbox{supp }F[m]^Td \subset [t_{\rm min}-mr,t_{\rm max}-mr].
    \]
    Define $w_{\rm tmp}$ to be the right-hand side of equation \ref{eqn:normsol}. Then
    from the previous observation and identity \ref{eqn:tran},
    \[
      \mbox{supp }w_{\rm tmp} \subset [t_{\rm min}-mr,t_{\rm max}-mr].
    \]
    From the identity \ref{eqn:unit}, for any $w \in W$,
    \[
      t \in [t_{\rm min}-mr,t_{\rm max}-mr] \Rightarrow F[m]^TF[m]w(t)
      = \frac{1}{(4 \pi r)^2}w(t).
    \]
    It follows from this and the previous two observations that
    $w_{\rm tmp}$ solves the normal equation \ref{eqn:norm}, and
    therefore that $\aw[m;d]=w_{\rm tmp}$.

  \item[3. ]Follows by inserting the definition
    \ref{eqn:defdatanonoise} of $d$ in \ref{eqn:normsol} and
    rearranging.
  \end{itemize}
\end{proof}

Since the normal operator is invertible, the reduced
objective $\tJa$ is given by the expression \ref{eqn:objexp}.
The results of the previous theorem imply those of

\begin{theorem}
  \label{thm:epjgen}
  Assume the hypotheses of Theorem \ref{thm:norminv}. Then
  \begin{equation}
  \label{eqn:residnormgen}
  e[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,(4\pi r \alpha a(t-mr))^4(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)^2
\end{equation}
\begin{equation}
  \label{eqn:anninormgen}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,(4\pi r a(t-mr))^2(1 +
  (4\pi r \alpha a(t-mr))^2)^{-2}d(t)
\end{equation}

\begin{equation}
  \label{eqn:expjgen}
\tJa[m;d] = \frac{1}{2}\int\,dt\,(4\pi r \alpha a(t-mr))^2(1+(4\pi r \alpha 
a(t-mr))^2)^{-1}d(t)^2. 
\end{equation}
Finally, if $a \in W^{1,1}_{\rm loc}(\bR)$ (distributions with integrable 
derivatives), then $\tJa[\cdot;d]$ is differentiable, and 
\begin{equation}
  \label{eqn:dexpjgen}
  \frac{d}{dm}\tJa[m;d] = -(4 \pi r \alpha)^2 \int \,dt \, 
  \left(a\frac{da}{dt}\right)(t-mr)(1+(4\pi r \alpha 
  a(t-mr))^2)^{-2}d(t)^2. 
\end{equation}
\end{theorem}

\begin{proof}
  From equation \ref{eqn:normsol},
  \[
    F[m]\aw[m;d](t) = 
    \frac{1}{4 \pi r}\left(\frac{1}{(4\pi r)^2} + \alpha^2
      a^2(t-mr)\right)^{-1}\frac{1}{4 \pi r}d(t),
  \]
  so
  \[
    (F[m]\aw[m;d]-d)(t) = (1 + (4 \pi r\alpha
    a(t-mr))^2)^{-1}-1)d(t)
  \]
  \[
    =(4 \pi r\alpha a(t-mr))^2(1 + (4 \pi r\alpha
    a(t-mr))^2)^{-1}d(t).
  \]
  Half the integral of the square of this data residual is
  $e[m,\aw[m;d],d]$, which proves identity \ref{eqn:residnormgen}.

  To compute $p[m,\aw[m;d],d]$, note that
  \[
    A\aw[m;d](t)=a(t) \left(\frac{1}{(4\pi r)^2} + \alpha^2
      a^2(t)\right)^{-1}\frac{1}{4 \pi r}d(t+mr)
  \]
  \[
    = 4\pi r a(t) (1 + (4\pi r \alpha a(t))^2)^{-1}d(t+mr)
  \]
  for $ t \in [t_{\rm min}-mr, t_{\rm max}-mr]$, so squaring,
  integrating, and changing integration variables $t \mapsto t-mr$
  gives the result \ref{eqn:anninormgen}

  That the VPM objective $\tJa$ is given by \ref{eqn:expjgen} follows from equations \ref{eqn:pen},
  \ref{eqn:redexp}, \ref{eqn:residnormgen}, and
  \ref{eqn:anninormgen}. Differentiation under the integral sign
  yields the expression \ref{eqn:dexpjgen} for its derivative with
  respect to $m$, for smooth $a$, and then by a limiting argument for
  $a \in W^{1,1}_{\rm loc}(\bR)$.
\end{proof}

\begin{corollary}
  \label{thm:epjnonoise}
  Assume the hypotheses of Theorem \ref{thm:norminv}, item 3. Then 
\begin{equation}
  \label{eqn:residnorm}
  e[m,\aw[m,d];d] 
= 8 \pi^2 r^2 \alpha^4\int\,dt\,a(t-(m-m_*)r)^4(1+(4\pi r)^2 \alpha^2 
    a(t-(m-m^*)r)^2)^{-2}w_*(t)^2.
\end{equation}
\begin{equation}
  \label{eqn:anninorm}
  p[m,\aw[m,d];d] = \frac{1}{2}\int \,dt\,  a(t-(m-m_*)r)^2 \left(1+ (4\pi r)^2\alpha^2
    a(t-(m-m_*)r)^2\right)^{-2}w_*(t)^2.
\end{equation}
so 
\begin{equation}
\label{eqn:expjnonoise}
\tJa[m;d] = \frac{\alpha^2}{2}\int\,dt\,a(t-(m-m_*)r)^2(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{-1}w_*(t)^2. 
\end{equation}
Finally, if $a \in W^{1,1}_{\rm loc}(\bR)$, then $\tJa[\cdot;d]$ is differentiable, and 
\begin{equation}
  \label{eqn:dexpjnonoise}
  \frac{d}{dm}\tJa[m;d] = -r \alpha^2 \int \,dt \, 
  \left(a\frac{da}{dt}\right)(t-(m-m_*)r)(1+(4\pi r)^2 \alpha^2 
  a(t-(m-m_*)r)^2)^{-2}w_*(t)^2. 
\end{equation}
\end{corollary}

\begin{proof} of Theorem \ref{thm:boxcarbad}

 This is clear from the definition $a = 1-{\bf 1}_{-\lambda,\lambda]}$ and equation \ref{eqn:expjnonoise}, as the supports of $w_*$ and ${\bf
    1}_{[-\lambda,\lambda]}(t-(m-m_*))$ are disjoint for the range of
  $m$ identified in the theorem.
\end{proof}

\begin{theorem}
  \label{thm:rampgood}
  Suppose that
  \begin{enumerate}
  \item $m_* \in M$;
  \item $0 < \mu \le \lambda$, and $w_* \in W_{\mu}$;
  \item $d_* = F[m_*]w_*$;
  \item $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
    with $\tau$ given by equation \ref{eqn:taudef}; and
  \item $\alpha > 0$.
  \end{enumerate}
  Then for any $m \in M$, 
  \begin{equation}
    | (m - m_*)r| > \lambda  \Rightarrow  \left|\frac{d}{dm}\tJa[m;d_*]\right| > \alpha^2 
    \frac{\lambda-\mu}{(1+(4\pi r)^2\alpha^2 
      (\lambda+\mu)^2)^{2}} \|w_*\|^2 
    \label{eqn:gradbndnonoise}
  \end{equation}
\end{theorem}
\begin{proof}
  As observed before, $\mbox{supp }\aw[m;d_*] \subset [t_{\rm
    min}-mr,t_{\rm max}-mr]\subset [-\tau,\tau]$, with $\tau$ defined
  in \ref{eqn:taudef}. Therefore, $a(t) = |t|$ $a a'(t) = t$ in the
  support of the integrand on the RHS of equation
  \ref{eqn:dexpjnonoise}, which therefore 
  becomes (after change of integration variable)
  \begin{equation}
    \label{eqn:gradfinal}
    \frac{d}{dm}\tJa[m;d_*] = -r \alpha^2 \int \,dt \, 
  t(1+(4\pi r)^2 \alpha^2 
  t^2)^{-2}w_*(t+(m-m_*))^2.
  \end{equation}
  Recall that $w_*(t+(m-m_*)r)$
  vanishes if $|t+(m-m_*)r| > \lambda$. Therefore the integral on the
  RHS of equation \ref{eqn:gradfinal} can be re-written
  \[
    = -r\alpha^2\int_{-(m-m_*)r-\lambda}^{-(m-m_*)r+\lambda}
    \,dt\, t(1+(4\pi r)^2\alpha^2 t^2)^{-2}w_*\left(t+(m-m_*)r\right)^2
  \]
  Suppose that $\mu \le \lambda$ and $w_* \in W_{\mu}$. 
  If $m > m_*+\lambda/r$, then $t+(m-m_*)r \in \mbox{supp }w_*$
  implies $-\mu - \lambda < t < \mu-\lambda<0$, so 
  \[
    t(1+(4\pi r)^2\alpha^2 t^2)^{-2} < (\mu-\lambda)(1+(4\pi r)^2\alpha^2 (\mu+\lambda)^2)^{-2}<0
  \]
  in the support of the integrand in equation
  \ref{eqn:gradfinal}. Arguing similarly for $m<m_*-\lambda/r$, obtain
  a similar inequality, implying the conclusion \ref{eqn:gradbndnonoise}.
\end{proof}

\begin{corollary}
  \label{thm:rampreallygood}
  Suppose that
  \begin{enumerate}
  \item $m_* \in M$;
  \item $0 <  \lambda$, and $w_* \in \lW$;
  \item $d_* = F[m_*]w_*$;
  \item $a(t)=\min\{|t|,\tau\}$ in the definition \ref{eqn:annmult},
    with $\tau$ given by equation \ref{eqn:taudef}; 
  \item $\alpha > 0$; and
  \item$m \in M$ is a stationary point of $\tJa[\cdot;d_*]$.
  \end{enumerate}
  Then $|m-m_*| < \lambda /r$.
\end{corollary}

\begin{proof} Follows directly from Theorem \ref{thm:rampgood} by
  taking $\mu=\lambda$.
\end{proof}

HERE INSERT RESULT: TJA DIFFBL RIGHTARROW A DIFFBLE - relate in
Discussion to Stolk 03

\section{Extended Inversion}
The preceding theorem established that a proper choice of annihilator
leads to a reduced penalty objective all of whose stationary points
are within $O(\lambda)$ of the target slowness $m_*$, provided that
the data are noise-free in the sense of equation
\ref{eqn:defdatanonoise}. This result leaves open two questions:
\begin{itemize}
\item how does one use this reduced penalty minimization to produce
  a solution of the inverse problem as in problem statement
  \ref{eqn:probstat0}? 
\item how does one answer the same question for noisy data?
\end{itemize}

\subsection{Noise-free Data}
\begin{theorem}
  \label{thm:ipnonoisesuf}
  Suppose that $a$ is given by definition \ref{eqn:ann}, $\alpha$,
  $\mu \in (0,\lambda_{\rm max}]$,
  $d$ is given by
  \ref{eqn:defdatanonoise} with $w_* \in W_{\mu}$, and  $m_{\infty}$ is a stationary
  point of $\tJa[\cdot;d]$. Then $(m_{\infty},\aw[m_{\infty};d])$ is a
  solution of the inverse problem \ref{eqn:probstat0} for any $\lambda
  \ge 2\mu$ and $\epsilon \ge (4\pi r \lambda \alpha)^2$.
\end{theorem}

\begin{proof}
  From the assumption $w_* \in W_{\mu}$ and Corollary
  \ref{thm:rampreallygood}, $|(m_{\infty}-m_*)r|\le \mu$. From the
  identity \ref{eqn:solnonoise},
  $\mbox{supp }\aw[m_{\infty};d] \subset
  [(m_{\infty}-m_*)r-\mu,(m_{\infty}-m_*)r+\mu] \subset
  [-2\mu,2\mu]$. Because of the support limitation, $a(t)=|t|$ in the
  interval of integration appearing in \ref{eqn:residnorm}, so
\[
  e[m_{\infty},\aw[m_{\infty},d];d] 
= 8 \pi^2 r^2 \alpha^4\int^{\mu}_{-\mu}\,dt\,|t-(m_{\infty}-m_*)r|^4(1+(4\pi r)^2 \alpha^2 
|t-(m_{\infty}-m^*)r|^2)^{-2}w_*(t)^2.
\]
and therefore
\begin{equation}
  \label{eqn:estresidnorm}
e[m_{\infty},\aw[m_{\infty};d];d] \le 8 \pi^2 r^2 (2\mu\alpha)^4 \|w_*\|^2 =
\frac{1}{2}(8 \pi r \mu \alpha)^4 \|d\|^2
\end{equation}
\end{proof}

The inequality \ref{eqn:estresidnorm} can be interpreted as a bound 
on $\alpha$, given $\epsilon$ and $\lambda$, for a
stationary point of $\tJa$ to yield a solution of the inverse
problem: one obtains a solution, provided that $\alpha$ is
sufficiently small. On the other hand, it is clear that $\alpha$
cannot be too large if stationary points of $\tJa$ are to yield
solutions: the integrand in \ref{eqn:residnorm} is increasing in
$\alpha$ for every $t$ and $m$, and the multiplier
\[
t \mapsto (4\pi r \alpha(t-(m_{\infty}-m_*)r))^4(1+(4\pi r)^2 \alpha^2 
|t-(m_{\infty}-m^*)r|^2)^{-2}
\]
tends monotonically to $1$ as $\alpha \rightarrow \infty$, uniformly
on the complement of any open interval containing
$t=(m_{\infty}-m_*)r$. Therefore
\begin{equation}
  \label{eqn:elimit}
  \lim_{\alpha \rightarrow \infty}e[m_{\infty},\aw[m_{\infty};d];d] =
  \frac{1}{2}\frac{1}{(4 \pi r)^2}\|w_*\|^2 = \frac{1}{2}\|d\|^2.
\end{equation}
Consequently, there exists $\alpha_{\rm max}(\epsilon,\lambda,d)$ so
that
\[
  e[m_{\infty},\aw[m_{\infty};d];d]  \le \frac{1}{2}\epsilon^2\|d\|^2
  \Rightarrow \alpha \le \alpha_{\rm max}(\epsilon,\lambda,d).
\]
The existence of this limiting penalty weight has been inferred
indirectly; the next section is devoted to a constructive algorithm
for its approximation.

\subsection{The Effect of Noise}
Suppose that the data trace $d$ takes the form
\begin{equation}
  \label{eqn:defdatanoisy}
  d = F[m_*]w_* + n = d_*+n,
\end{equation}
with $m_* \in M, w_* \in W_{\mu}$, $0<\mu<\lambda$, and noise trace $n \in
D$. Since no support assumptions can be made about $n$, equation
\ref{eqn:normsol} implies that $\aw[m;d] \notin \lW$ for any values of
$\alpha$ and $\lambda$.  Therefore minimization of $\tJa$ cannot by itself yield a
solution of the inverse problem as defined in the problem statement
\ref{eqn:probstat0}. In this section, we explain how a solution may
nontheless be constructed from a stationary point of $\tJa$.

First we examine the effect of additive noise on the estimation of the
slowness $m$. In expressing the result, we use the dimensionless
relative data error
\begin{equation}
  \label{eqn:defeta}
  \eta = \frac{\|n\|}{\|d_*\|}. 
\end{equation}

\begin{theorem}
  \label{thm:mnoise}
  Assume the hypotheses of Theorem \ref{thm:rampgood}, and that $d$ is
  given by definition \ref{eqn:defdatanoisy}. Suppose that $m \in M$
  is a stationary point of $\tJa[\cdot;d]$, and that
  \begin{equation}
    \label{eqn:mnoisebnd}
    \eta(1+\eta) \le \frac{16}{3\sqrt{3}}\frac{4\pi r \alpha
      (\lambda-\mu)}{(1+(4\pi r\alpha(\lambda+\mu))^2)^2}
  \end{equation}
  Then $|m-m_*| \le \lambda /r$.
\end{theorem}

\begin{proof}
  From equation \ref{eqn:dexpjgen}, $d \tJa / dm$ is the
value of a quadratic form in $d$ with (indefinite) symmetric operator
$B = $ multiplication by
\[
  b(t;m,\alpha)  = -\frac{(4 \pi r \alpha)^2 (t-mr)}{(1+(4\pi r \alpha (t-mr))^2)^{2}}
\]
Therefore
\begin{equation}
  \label{eqn:gradlip}
  \left|\frac{d}{dm}\tJa[m;d]-\frac{d}{dm}\tJa[m;d_*]\right| =
  |\langle (d+d_*),B(d-d_*)\rangle| \le \max_{t \in
    \bR}|b(t;m,\alpha)|\eta(1+\eta)\|d_*\|^2
\end{equation}
A straightforward calculation shows that
\[
  \max_{t \in \bR} b(t;m,\alpha) = \frac{3\sqrt{3}}{16} 4\pi r\alpha.
\]
For a stationary point $m$ of
$\tJa[\cdot;d]$, the inequality \ref{eqn:gradlip} implies
\[
  \left|\frac{d}{dm}\tJa[m;d_*]\right| \le \frac{3\sqrt{3}}{16} 4\pi
  r\alpha \eta(1+\eta)\|d_*\|^2
\]
On the other hand, the conclusion \ref{eqn:gradbndnonoise} of Theorem
\ref{thm:rampgood} implies that if also
\[
  \frac{3\sqrt{3}}{16} 4\pi r\alpha \eta(1+\eta)\|d_*\|^2 \le (4 \pi r
  \alpha)^2 \frac{\lambda-\mu}{(1+(4\pi r)^2\alpha^2
    (\lambda+\mu)^2)^{2}} \|d_*\|^2
\]
then $|m-m_*|\le \lambda/r$. Rearranging, obtain the conclusion.
\end{proof}

\begin{corollary}
  \label{thm:mnoisecor}
  Asumme the hypotheses of Theorem \ref{thm:mnoise}, in particular
  that $m$ is a stationary point of $\tJa[\cdot;d]$, $d=d_*+n$. Then
  \begin{itemize}
    \item[1. ] the bound \ref{eqn:mnoisebnd} holds only if 
      noise level $\eta=\|n\|/\|d_*\|$ satisfies
      \begin{equation}
        \label{eqn:mnoisecond}
        \eta < \frac{1+\sqrt{5}}{2} < 1;
      \end{equation}
    \item[2. ] if condition \ref{eqn:mnoisecond} holds, then
%      $|m-m_*|<\lambda/r$ provided that
%      \begin{equation}
%        \label{eqn:mnoisesuff}
%        \frac{\lambda}{\mu} \ge 1+\frac{2\eta(1+\eta)}{1-\eta(1+\eta)}. 
%      \end{equation}
      \begin{equation}
        \label{eqn:mnoisesuff}
        |m-m_*| \le \left(1+\frac{2\eta(1+\eta)}{1-\eta(1+\eta)}\right)\frac{\mu}{r}. 
      \end{equation}
    \end{itemize}
\end{corollary}

\begin{proof}
  Write $\lambda = (1+\delta)\mu$, and $x=4 \pi r \alpha \mu$. Then
  the right-hand side of equation \ref{eqn:mnoisebnd} may be written as
  \begin{equation}
    \label{eqn:mnoisebndrev}
    \frac{16}{3\sqrt{3}}\frac{4\pi r \alpha
      (\lambda-\mu)}{(1+(4\pi r\alpha(\lambda+\mu))^2)^2} = D
    \frac{x}{(1+C^2 x^2)^2},
  \end{equation}
  where
  \[
    D=\frac{16}{3\sqrt{3}}\delta,\,C=2+\delta.
  \]
  The positive stationary point of the quantity on the right-hand side
  of \ref{eqn:mnoisebndrev} is a maximum, and occurs at
  $x=1/(\sqrt{3}C)$, that is
  \[
    4 \pi r \alpha \mu = \frac{1}{\sqrt{3}(2+\delta)}.
  \]
  Thus
  \[
    1+C^2x^2 = \frac{4}{3}
  \]
  hence the maximum value is
  \[
    \frac{D3\sqrt{3}}{16C} = \frac{\delta}{2+\delta}.
  \]
  This maximum value must be larger than the left hand side of inequality
  \ref{eqn:mnoisebnd} in order that there be any solutions at all, but
  is less than $1$. This observation establishes conclusion 1 of the
  theorem. Also, existence of $\lambda$ satisfying \ref{eqn:mnoisebnd}
  implies that
  \[
    \eta(1+\eta) \le \frac{\delta}{2+\delta}.
  \]
  Solving this inequality for $\delta$ and unwinding the definitions
  leads to conclusion 2.
\end{proof}

\noindent {\bf Remark:} We emphasize that Theorem \ref{thm:mnoise} and
Corollary \ref{thm:mnoisecor} state {\em sufficient} conditions for a bound on
the slowness error $|m-m_*|$ in terms of the relative data noise level $\eta$,
giving an additional ``fudge factor'' beyond the support size $\mu$
of the noise-free wavelet $w_*$ for an interval within which the slowness error is
guaranteed to lie.

Conclusion 1 in Corollary \ref{thm:mnoisecor} constrains the range of
noise level to which these results apply to a bit more than 60\%. That
is, the bound given by conclusion 2 is useful only for small noise. In
the limit as $\eta \rightarrow 0$, conclusion 2 becomes
$\lambda/\mu \gtrsim 1 + 2\eta$, that is, the ``fudge factor'' beyond
the noise-free bound is approximatly twice the noise level.

On the other hand, stronger bounds than given by Corollary
\ref{thm:mnoisecor} are possible, given additional constraints on the
noise $n$. A natural example is uniformly distributed random noise,
filtered to have the same spectrum as the source. The expression
\ref{eqn:dexpjgen} implies that the interaction of noise $n$ and
signal $d_*$ in the derivative of $\tJa$ is local, so that the
coefficient of $\eta$ on the left-hand side of inequality
\ref{eqn:mnoisebnd} is effectively much less that 1, resulting in a
larger range of allowable $\eta$. While we will not formulate such a
result, one of the numerical examples below suggests its feasibility.

    % Recall that $\lE$ is the projection of $W$ onto $\lW$
%(i.e. multiplication by ${\bf 1}_{[-\lambda,\lambda]}$.

\begin{proof} of Theorem \ref{thm:ipnoisesuf}:
From Theorem \ref{thm:norminv}, 
\[
  \aw[m;d](t) = (1+ (4 \pi r \alpha t)^2)^{-1}(w_*(t+(m-m_*)) + 4\pi r 
  n(t+mr)) 
\]
\[
  = \aw[m;d_*] + (1+ (4 \pi r \alpha t)^2)^{-1}4\pi r n(t+mr)
\]

From Corollary \ref{thm:mnoisecor},
\[
  |m-m_*| \le \left(1+\frac{2\eta(1+\eta)}{1-\eta(1+\eta)}\right)\frac{\mu}{r} 
\]
\[
  =\left(2+\frac{2\eta(1+\eta)}{1-\eta(1+\eta)}\right)\frac{\mu}{r}
  -\frac{\mu}{r} \le \left(\frac{\lambda}{\mu}-1\right)\frac{\mu}{r}
\]
That is,
\[
  |m-m_*|r \le \lambda-\mu.
\]
From Theorem \ref{thm:norminv},
\[
  \mbox{supp }\aw[m;d_*] \subset [-\mu-(m-m_*)r, \mu-(m-m_*)r]
  \subset [-\lambda,\lambda]
\]
so
\[
  E_{\lambda}\aw[m;d](t) = \aw[m;d_*](t) + E_{\lambda}(1+ (4 \pi r \alpha
  t)^2)^{-1}4\pi r n(t+mr)
\]
From the definition of $F[m]$, for any $t_1<t_2$, $w \in W$,  
\[
  F[m]{\bf 1}_{[t_1.t_2]}w = {\bf 1}_{[t_1+mr,t_2+mr]}F[m]w  
\]
Thus the data residual after projection is
\[
  F[m]E_{\lambda}\aw[m,d](t) -d(t) = F[m]\aw[m,d_*](t) -d_*(t)  
\]
\[
  + {\bf 1}_{[-\lambda+mr,\lambda+mr]}(4 \pi r \alpha (t-mr))^2 (1+ (4 \pi
  r \alpha (t-mr))^2)^{-1} n(t)
\]
\[
  -(1- {\bf 1}_{[-\lambda+mr,\lambda+mr]})n(t)
\]
From \ref{eqn:residnorm} and the bound on $m-m_*$,
\[
  \| F[m]E_{\lambda}\aw[m,d_*] -d*\|^2 = (4 \pi r \alpha)^4
  \int_{-\lambda+(m-m_*)r}^{\lambda+(m-m_*)r}\,dt\, (t-(m-m_*)r)^4
\]
\[
  \times (1+(4\pi r \alpha)^2 
  (t-(m-m^*)r)^2)^{-2}d_*(t)^2
\]
\[
  \le (4 \pi r \alpha \lambda)^4\|d_*\|^2
\]
Similarly, the norm squared of the sum of the last two terms is 
\[
  \le (4 \pi r \alpha \lambda)^2 \|{\bf 1}_{[-\lambda+mr,\lambda+mr]}
  n\|^2 + \|(1 - {\bf 1}_{[-\lambda+mr,\lambda+mr]}n\|^2
\]
Without additional hypotheses to outlaw the accumulation of $n$ near
$t=mr$, all that can be said is that this is
\[
  \le \max \{(4 \pi r \alpha \lambda)^2, 1\} \|n\|^2
\]
Putting this all together,
\[
  \|F[m]E_{\lambda}\aw[m,d]-d\| \le 4 \pi r \alpha
  \lambda(\max\{1,(4\pi r \alpha \lambda)^{-2}\}\eta^2 + (4
  \pi r \alpha \lambda)^2)^{1/2}\|d_*\|
\]
as asserted.
\end{proof}

\bibliographystyle{seg}
\bibliography{../../bib/masterref}

\append{Numerics and penalty selection via the Discrepancy Principle}
As described in the introduction, the discrepancy principle (in one of
its guises) involves setting an acceptable range of data misfit
$[e_-,e_+]$, and adjusting the penalty weight $\alpha$ in the reduced
objective definition \ref{eqn:red} so that the data error term at the
inner solution, $e[m,\aw[m;d],d]$, lies in this range. Since updating
$m$ changes the inner problem, the appropriate condition for such
separable problems is that $e$ {\em stays} in the range $[e_-,e_+]$ as
$m$ is updated. In this section, we examine the dependence of $e$
on $\alpha$ with a view to understanding how to reset $\alpha$ when
$m$ changes.

Accordingly, regard $m$ as fixed and suppress it from the notation for the remainder of ;dthis section, and introduce the abbreviations
\begin{eqnarray}
\label{eqn:erralph}a
e(\alpha) &=& e[m,\aw[m;d],d]\\
\label{eqn:penalph}
p(\alpha) &=& p[m,\aw[m;d],d]
\end{eqnarray}
Since the normal matrix $F^T F + \alpha^2 A^T A$ has been assumed
invertible (condition \ref{eqn:nopco}), the solution of $\aw$ of the
normal equation \ref{eqn:condition} is smooth in $\alpha^2$.
Differentiating equation \ref{eqn:condition} with respect to  $\alpha^2$ leads to the relation
\begin{equation}
(F^T F + \alpha^2 A^T A ) \frac{dw}{d\alpha^2} = -A^T A w
\label{eqn:dnorm}
\end{equation}
whence
\begin{align}
\frac{de}{d\alpha^2} 
&=\left\langle\frac{dw}{d\alpha^2},F^T(Fw-d) \right\rangle \nonumber \\
&=-\alpha^2\left\langle\frac{dw}{d\alpha^2},A^TAw\right\rangle \nonumber \\ 
&=\alpha^2 \langle A^TAw,(F^TF + \alpha^2 A^TA)^{-1}A^TAw\rangle \nonumber \\
&\ge 0
\label{eqn:de}
\end{align}
Note that the inequality in equation \ref{eqn:de} is {\em strict}  if $p > 0$ hence $A^TAw \ne 0$, since the normal operator is assumed to be positive definite.
The derivative of $p(\alpha^2)$ with respect to $\alpha^2$ is
\begin{align}
\frac{dp}{d\alpha^2} &=  -\langle A^T Aw,(F^TF + \alpha^2 A^TA)^{-1}A^TAw \rangle \nonumber \\
&\leq 0
\label{eqn:dp}
\end{align}
similarly a strict inequality if $p > 0$.

Equations \ref{eqn:de} and \ref{eqn:dp} show that increasing $\alpha^2$ implies increasing $e$ while decreasing $p$, and
\begin{align}
&\langle A^TA w,(F^TF + \alpha^2 A^TA)^{-1} A^TA w \rangle \nonumber \\ 
=& \langle (A^TA)^{1/2}w,[(A^TA)^{-1/2}F^TF(A^TA)^{-1/2} + \alpha^2 I]^{-1}(A^TA)^{1/2}w \rangle \nonumber \\ 
\le& \frac{1}{\alpha^2} \langle A^TA w, w\rangle = \frac{2}{\alpha^2}p
\end{align}
In view of equation \ref{eqn:de},
\begin{equation}
\label{eqn:lep}
\frac{de}{d\alpha^2}  \le 2p.
\end{equation}
with this inequality also being strict if $p > 0$.

Suppose the current weight is $\alpha^2_c$ and denote a candidate for an updated weight by $\alpha^2_+$. Then from inequality \ref{eqn:lep},
\begin{equation}
e(\alpha^2_+)-e(\alpha^2_c) \le \int_{\alpha^2_c}^{\alpha^2_+} 2p d\alpha^2
\end{equation}
If $\alpha^2_+ \ge \alpha^2_c$, then in view of inequality \ref{eqn:dp}, the above is
\begin{equation}
\label{eqn:basic}
\le 2 p(\alpha^2_c) (\alpha^2_+-\alpha^2_c)  
\end{equation}

Let us suppose that $e(\alpha^2_c) < e_{+}$. Then setting 
\begin{equation}
\label{eqn:alphasecant}
\alpha^2_+ = \alpha^2_c + \frac{e_{+}-e(\alpha^2_c)}{2p(\alpha^2_c)} 
\end{equation}
implies via inequality \ref{eqn:basic} that 
\begin{equation}
e(\alpha^2_+)-e(\alpha^2_c) \le e_{+}-e(\alpha^2_c)
\end{equation}

Assuming that $p(\alpha^2_+) > 0$, hence $p(\alpha^2)>0$ for $\alpha^2_c \le \alpha^2 \le \alpha^2_+$,  we conclude that if $\alpha^2_+$ is given by the rule \ref{eqn:alphasecant},
then
\begin{equation}
\label{eqn:assert}
e(\alpha^2_c) < e(\alpha^2_+) \le e_{+}
\end{equation}
That is, unless $p(\alpha^2_+) = 0$ (in which case a physical solution of the inverse problem has been reached),  $e(\alpha^2_+)$ is larger than $e(\alpha^2_c)$ but in any case does not exceed $e_+$. The rule \ref{eqn:alphasecant} therefore provides a feasible updated $\alpha^2$ consistent with the upper bound in the discrepancy principle.

\append{Heisenberg, support, and wavelength}

[THE NEXT BIT MAY OR MAY NOT BE WORTH KEEPING - IT WAS INTERESTING TO
WORK OUT]

Without further contraints on data or solution, nothing more can be
said about bounds on $\alpha$. If the data $d$ (and the target wavelet
$w_*$) are assumed to have a square integrable derivative, then a
necessary condition follows from the Heisenberg
inequality (see for example \cite{Folland:07}, p. 255). To formulate
this result in its most general form, introduce the Hilbert subspaces
$V^0 \subset L^2(\bR), V^1 \subset H^1(\bR)$:
\begin{eqnarray}
  V^0 & = & \{ f \in L^2(\bR): Af \in L^2(\bR)\}, \nonumber\\
  V^1 & = & V^0 \cap H^1(\bR).
            \label{eqn:vdef}
\end{eqnarray}
$V^0$ is the domain of $A$, and equipped with the graph norm of $A$. A
natural norm in $V^1$ is
\[
  \|f\|^2_{V^1} = \|f\|_{V^0}^2 + \|f\|_{H^1}^2.
\]
$V^j$ is the completion of $C_0^{\infty}(\bR)$ in the corresponding
norm, j=0,1.

\begin{proposition}
  \label{thm:heis}
For $w \in V^1$,
  \begin{equation}
    \label{eqn:heis}
    \|Aw\|\|w'\| \ge \frac{1}{2}\|w\|^2
  \end{equation}
\end{proposition}

\begin{proof}
  For $w \in C_0^{\infty}(\bR)$,
  \[
    \int w^2 = \left|\int\,dt\, t (w(t)^2)' \right|= \left|2\int\,dt\,tw(t)w'(t)\right| \le
    2\|Aw\| \|w'\|
  \]
  by the Cauchy-Schwarz inequality. Since $C_0^{\infty}(\bR)$ is dense
  in $V^1$, the conclusion follows by continuity.
\end{proof}

In the conventional formulation of the Heisenberg inequality, the $L^2$ norm of
$w'$ is replaced by the its equivalent in terms of the Fourier
transform $\hat{w}$. Adopting temporarily the use of dummy variables
in the expression of functions, the identity \ref{eqn:heis} turns into
the usual form of the Heisenberg inequality: for $w \in V^1$,
\begin{equation}
\label{eqn:fheis}
\|tw(t)\|\|k\hat{w}(k)\| \ge \frac{1}{4\pi}\|w\|^2.
\end{equation}

Define $\krms[w]$, the root mean square estimator of frequency of $w
\in V^1$, by
\begin{equation}
  \label{eqn:krms}
  \krms[w]=\frac{1}{2\pi}\frac{\|w'\|}{\|w\|} = \left(\int
    \,dk\,\frac{|\hat{w}(k)|^2}{\|\hat{w}\|^2} k^2\right)^{1/2}.
\end{equation}
Then the inequalities \ref{eqn:heis}, \ref{eqn:fheis} can be rewritten as
\begin{equation}
  \label{eqn:frmsheis}
  \|Aw\| \ge \frac{\|w\|}{4\pi \krms[w]}.
\end{equation}

For $\lambda >0$, define
\begin{equation}
  \label{eqn:wlam1}
  \lW^1 = \lW \cap H^1(\bR).
\end{equation}
Note that $\lW^1 \subset V^1$ is a closed subspace for any
$\lambda>0$.

\begin{proposition}
  \label{thm:klam}
  For $w \in \lW^1$,
  \[
    \krms[w] \ge \frac{1}{4 \pi \lambda}
  \]
\end{proposition}

\begin{proof}
  Follows directly from inequality \ref{eqn:frmsheis} and the obvious bound
  $\|A|_{\lW}\|\le \lambda$.
\end{proof}

\noindent{\bf Remark:} This result is the link mentioned earlier between the support
constraint and the well-known frequency-based criteria for success of
FWI. The result of Theorem \ref{thm:fwi} can be rephrased as showing the
existence of many stationary points of the mean-square error function
for which the travel time is in error by more than
$\lambda \ge 1/{4 \pi \krms[w]|}$. In fact, the usual error criterion mentioned in the
literature is that the initial estimate of travel time must be in error by at most ``half a
wavelength'' if FWI is to converge reliably. This is correct in some
circumstances, depending on features of the target wavelet $w_*$ of
which the arguments in this paper do not take account.

\begin{proposition}
  \label{thm:heis2}
  For $\lambda>0, w \in \lW^1$,
  \begin{equation}
    \label{eqn:heis2}
    \|A^2w\| \ge \frac{\|w\|}{24 \lambda (2\pi\krms[w])^3}.
  \end{equation}
\end{proposition}

\begin{proof}
  Since $A$  preserves $\lW^1$, inequality \ref{eqn:heis} implies
  \begin{equation}
    \label{eqn:heissq}
    \|A^2w\| \ge \frac{\|Aw\|^2}{2\|(Aw)'\|}.
  \end{equation}
  From the definition of $A$, $(Aw)'=w +Aw'$, so for $w \in \lW^1$,
  \[
    \|(Aw)'\| \le \|w\|+ \lambda\|w'\|.
  \]
  Applying \ref{eqn:heis} again,
  \[
    \|w\|^2 \le 2\|w'\|\|Aw\|\le 2\lambda \|w'\|\|w\|,
  \]
  so obtain the 1D Poincar\'{e} inequality: for $w \in \lW^1$,
  \[
    \|w\| \le 2\lambda\|w'\|.
  \]
  Thus
  \[
    \|(Aw)'\| \le 3\lambda\|w'\|.
  \]
  Apply this estimate together with the basic Heisenberg estimate
  \ref{eqn:heis} to the inequality \ref{eqn:heissq} to obtain
  \[
    \|A^2w\| \ge \frac{\|w\|^4}{24\lambda \|w'\|^3}
  \]
  Rearranging and using the definition \ref{eqn:krms} of $\krms[w]$,
  arrive at inequality \ref{eqn:heis2}.
\end{proof}

\begin{theorem}
  \label{thm:ipnonoisenec}
  Suppose that $a$ is given by definition \ref{eqn:ann}, $\alpha$,
  $\mu \in (0,\lambda_{\rm max}/2]$, $d$ is given by
  \ref{eqn:defdatanonoise} with $w_* \in W_{\mu}^1$, and that a
  stationary point $m_{\infty}$ of $\tJa[\cdot;d]$ yields a solution
  of the inverse problem \ref{eqn:probstat0} for $\lambda \ge 2\mu$,
  $\epsilon>0$. Then
\begin{equation}
  \label{eqn:epsalpha}
  3 (4\pi \lambda \krms[w_*])^3 \epsilon \ge \frac{(4\pi  r\alpha\lambda)^2}
  {(1+(4\pi r \alpha \lambda)^2)}
\end{equation}
\end{theorem}

\begin{proof}
  Rearranging the identity \ref{eqn:residnorm}, and observing as in
  the proof of Theorem \ref{thm:ipnonoisesuf} that the support of the
  integrand is contained in $[-2\mu, 2\mu] \subset [\lambda,\lambda]$,
  \[
  e[m_{\infty},\aw[m_{\infty},d];d] 
= 8 \pi^2 r^2 \alpha^4\int\,dt\,t^4(1+(4\pi r)^2 \alpha^2 t^2)^{-2}w_*(t+(m_{\infty}-m_*))^2
\]
\[
  \ge \frac{8\pi^2r^2\alpha^4}{(1+(4\pi r)^2 \alpha^2
    \lambda^2)^{2}}\int\,dt\,t^4w_*(t+(m_{\infty}-m_*))^2
\]
\begin{equation}
  \label{eqn:residnormbis}
  \ge \frac{8\pi^2r^2\alpha^4}{(1+(4\pi r)^2 \alpha^2
    \lambda^2)^{2}}\|A^2w_*(\cdot + (m_{\infty}-m_*)r)\|^2.
 \end{equation}
Since $\krms[w]$ is invariant under translation of $w \in V^1$,
Proposition \ref{thm:heis2} implies that this is
\[
  \ge \frac{8\pi^2r^2\alpha^4}{(1+(4\pi r)^2 \alpha^2
    \lambda^2)^{2}}\frac{\|w_*\|^2}{(24)^2 \lambda^2 (2\pi\krms[w_*])^6}
\]
\[
  = \frac{1}{2}\frac{(4\pi r\alpha\lambda)^4}{(1+(4\pi r \alpha
    \lambda)^2)^{2}}\frac{\|d\|^2}{(24)^2 (2\pi
    \lambda \krms[w_*])^6}
\]
The pair $(m_{\infty}, \aw[m_{\infty},d])$ is presumed to solve the inverse
problem as stated in \ref{eqn:probstat0}, in particular
\[
  \epsilon \ge (2 e[m_{\infty}, \aw[m_{\infty},d],d])^{1/2}/\|d\|
\]
the inequality \ref{eqn:epsalpha} follows.

\end{proof}

Inequality \ref{eqn:epsalpha} couples the dimensionless quantites
$\epsilon$,  $4\pi r \lambda \alpha$, and $4 \pi \lambda
\krms[w_*]$. Proposition \ref{thm:klam}, implies that the left
hand side is $\ge 3\epsilon$. Since the right hand side is $\le 1$,
the inequality implies no limitation on $\alpha$ if the left hand side
is $\ge 1$. For small $\epsilon$, the the largest permissible $\alpha$
is $O(\sqrt{\epsilon})$. The permissible range of $\alpha$ increases with
nondimensionalized RMS frequency $4\pi \lambda \krms[w_*]$. Since
$\krms$ is invariant under translation and scaling of its argument,
$\krms[w_*]=\krms[d]$, that is, the nondimensionalized RMS frequency
is an observable property of the data, in the noise-free case at
least.