### 2020-09-27 repro/papers/discrep/alg.tex
added alg, figs, makefigs, modified fig captions

\title{Discrepancy Algorithm for Single-Trace Transmission Inverse Problem}
\author{William W. Symes}

\lefthead{Symes}

\righthead{Algorithm}

\maketitle
\begin{abstract}
  Concise statement of the discrepancy algorithm for solving the single-trace transmission inverse problem.
\end{abstract}

\section{Definitions}
``The paper'' throughout refers to ``Solution of an Acoustic Transmission Inverse Problem by Extended Inversion'', current draft.

Note that I have changed the notation for the penalty term to $g$, rather than $p$ as has been the case before, in order to avoid conflict with the conventional notation for the acoustic pressure field.

The mean square error $e$, penalty term $g$, and extended objective $\Ja $ are defined in equations 4, 5, and 6. These quantities are normalized via division by the norm-squared of the data. In recent drafts I also introduced a notation for the relative $L^2$ error = noise-to-signal ratio:
\begin{equation}
  \label{eqn:rmsgen}
  r[m,w;d] = \|F[m]w-d\|/\|d\| = \sqrt{2 e[m,w;d]}
\end{equation}
The algorithm presented here assumes that the multiplier $a$ appearing in equation 39 (and in the definition of the annihilator $A$) is given by $a(t)=|t|$. The paper explains how this choice can be regarded as defining a bounded $A$, due to the bounded range of time allowed in the data.

The minimizer of $\Ja $ over the wavelet $w$ is given by solution of the normal equation. Under the conditions stated in Proposition 2, this minimizer is given in equation 39 of the paper:
\begin{equation}
  \label{eqn:normsol}
  \aw[m;d](t) = \left\{
    \begin{array}{c}
      \left(\frac{1}{(4\pi r)^2} + \alpha^2
      t^2\right)^{-1}\frac{1}{4 \pi r}d(t+mr), t \in [t_{\rm
      min}-mr, t_{\rm max}-mr];\\
      0, \mbox{ else;}
    \end{array}
  \right.
\end{equation}
The reduced quantities for the Variable Projection Method are obtained by substituting $\aw[m;d]$ for $w$ in equations 4, 5, and 6 to obtain equations 41, 42, and 43, then using $a(t)=|t|$:
\begin{equation}
  \label{eqn:residnormgen}
  e[m,\aw[m,d];d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}} \,dt\,(4\pi r \alpha (t-mr))^4(1 +
  (4\pi r \alpha (t-mr))^2)^{-2}d(t)^2
\end{equation}
\begin{equation}
  \label{eqn:anninormgen}
  g[m,\aw[m,d];d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}} \,dt\,(4\pi r (t-mr))^2(1 +
  (4\pi r \alpha (t-mr))^2)^{-2}d(t)
\end{equation}

\begin{equation}
  \label{eqn:expjgen}
\tJa[m;d] = \frac{1}{2\|d\|^2}\int_{t_{\rm min}}^{t_{\rm max}}\,dt\,(4\pi r \alpha (t-mr))^2(1+(4\pi r \alpha 
(t-mr))^2)^{-1}d(t)^2. 
\end{equation}
The derivative with respect to $m$ of $\tJa $ is
\begin{equation}
  \label{eqn:dexpjgen}
  \frac{d}{dm}\tJa[m;d] = -\frac{(4 \pi r \alpha)^2}{\|d\|^2} \int_{t_{\rm min}}^{t_{\rm max}} \,dt \,  (t-mr)(1+(4\pi r \alpha (t-mr))^2)^{-2}d(t)^2. 
\end{equation}

The data $d$ is sampled on a uniform grid of cell size $dt$ (in matlab notation, $t_{\rm min}:dt:t_{\rm max}$). The integrals in the preceding equations, and the integral in the defintion of $\|d\|^2$, should be approximated by a quadrature rule. The trapezoidal rule should be accurate enough, provided that $d$ is bandlimited. In our numerical experiments, both noise-free data and noise are obtained by convolving numerical delta functions (``spikes'') with sampled Ricker wavelets. If the center frequency of the Ricker wavelet is $f$, then the usual rule of thumb is that Fourier components above frequencies of $2.5f$ are negligible. Sampling at 10 gridpoints per wavelength is a roughly adequate choice for the trapezoidal rule.

For example, if the center frequency is 20 Hz, Fourier components at frequencies above 50 Hz are negligible. 50 Hz corresponds to a wavelength of 0.02 s, so a suitable sample interval for our calculations is $dt = 0.002$ s.

\section{Algorithm}
The ideal goal of the discrepancy algorithm is:

\begin{quote}
  Given data $d$, a target noise-to-signal ratio $r_{\rm tgt}$ (or equivalently a target signal-to-noise ratio of $1/r_{\rm tgt}$), and a fudge factor $C<1$, find a stationary point $m_{\infty}$ of the VPM objective $\tJa[\cdot;d]$ so that
\begin{equation}
  \label{eqn:discrep}
  r_-=Cr_{\rm tgt} \le r[m,\aw[m;d];d] \le r_+= C^{-1}r_{\rm tgt}, 
\end{equation}
or equivalently
\begin{equation}
  \label{eqn:discrepsq}
  e_-=C^2e_{\rm tgt} \le e[m,\aw[m;d];d] \le e_+=C^{-2}e_{\rm tgt}.
\end{equation}
\end{quote}
Note: in the Fu paper, $e_{\pm}$ were denoted $X_{\pm}$, and $e_{\rm tgt}$ was denoted $X$.

This statement is ideal because an actual stationary point cannot be found exactly. Instead, accept an approximate stationary point $m_{\rm approx}$ defined by a choice of tolerance $\delta$ for the norm of the gradient:
\begin{equation}
  \label{eqn:approx}
  \left|\frac{d \tJa}{dm}[m_{\rm approx};d]\right| \le \delta.
\end{equation}

Schematic statement of the discrepancy algorithm:
\begin{algorithm}[H]
\caption{Discrepancy algorithm for updating $m, \alpha$}
\begin{algorithmic}[1]
  \State print parameters: $e_{\rm tgt},e_-,e_+,\delta$, max number of iterations allowed for outer loop, $\alpha$ update loop, $m$ update loop.
  \State print target slowness $m_*$, plot target wavelet $w_*$, noise-free data $d_*=F[m_*]w_*$, noise $n$, data $d=d_*+n$
  \State Choose $m\in M$, $\alpha=0$;
  \Repeat
  \Repeat
  \State print iteration number, $m$, $\alpha$, $r$, $e$, $g$, $\tJa$ before update
  \State update $\alpha$ for fixed $m$
  \State print $m$, $\alpha$, $r$, $e$, $g$, $\tJa$ after update  
  \Until inequalities \ref{eqn:discrepsq} are satisfied
  \Repeat
  \State print iteration number, $m$, $\alpha$, $r$, $e$, $g$, $\tJa$ before update
  \State update $m$ for fixed $\alpha$
  \State print $m$, $\alpha$, $r$, $e$, $g$, $\tJa$ after update  
  \Until{inequalities \ref{eqn:discrepsq} are not satisfied, or $\|\nabla \tJa[m;d]\|$
    sufficiently small}
  \Until{inequalities \ref{eqn:discrepsq} are satisfied}
  \State print final values of $m$, $\alpha$, $r$, $e$, $g$, $\tJa$
  \State plot estimated wavelet $\aw[m;d]$, estimated data $F[m]\aw[m;d]$, and residual $F[m]\aw[m;d]-d$
\end{algorithmic}
\end{algorithm}

Note: Each loop (outer, $m$, $\alpha$ needs to be supplied with a limit on the number of iterations. If a loop comes to the limit without satisfying the convergence criterion for that loop (``until''), then the algorithm should exit with an appropriate failure message.

Note: plots of data ($d_*$, $n$, $d$, $F[m]\aw[m;d]$, $F[m]\aw[m;d]-d$) should have same vertical scale. Same for plots of wavelets ($w_*$, $\aw[m;d]$).

\section{Update $\alpha $}
The Fu paper (repeated in Appendix A in the current draft) sketches a proof that if $e_{\rm curr} = e[m,w_{\alpha_{\rm curr}}[m;d];d]< e_+$, then  $\alpha_{\rm next}$ given by  
\begin{equation}
\label{eqn:alphasecant}
\alpha^2_{\rm next} = \alpha^2_{\rm curr} + \frac{e_{+}-e_{\rm curr}}{2p[m,w_{\alpha_{\rm curr}}[m;d];d]} 
\end{equation}
satisfies
\[
  e_{\rm curr}  < e_{\rm next} =  e[m,w_{\alpha_{\rm next}}[m;d];d]  < e_+.
\]
That is, the update rule \ref{eqn:alphasecant} produces an increasing sequence of values of $e$ in the interval $[0,e_+)$.

We are currently lacking a proof that in fact the sequence of $e$ values so generated converges to $e_+$ for fixed $m$, but numerical evidence is very strong that this is so. Therefore we expect the stopping criterion in line 9 of the algorithm to be satisfied after some number of iterations, so this is a suitable update rule for $\alpha$.

This iteration does appear to be fairly slow, compared to some alternatives. However it is reliable. Since the current computations are very inexpensive, use it.

\section{Update $m$}

Any local optimization algorithm may be used, provided that is has the property called ``global convergence'' in the literature of optimization. This property does {\em not} mean ``converges to the global optimum''. In fact it means ``converges globally (starting from any initial guess) to a local minimum''.

An example of such an algorithm is steepest descent with backtracking line search. The line search should include a ``sufficient decrease'' test. Huiyi's recent algorithm is a good choice. Also, good implementations of Brent's method for one-variable minimization are also globally convergent.

The intent of the discrepancy algorithm design is that eventually the gradient decrease criterion will be satisfied, that is, that the iteration will converge before the bounds \ref{eqn:discrepsq} are violated. That is why we use membership in an interval containing $e_{\rm tgt}$, rather than equality with $e_{\rm tgt}$, as the termination criterion. When the $m$ update converges (small gradient) while the bounds \ref{eqn:discrepsq} are satisfied, the outer iteration (over both $m$ and $\alpha$) will terminate, with a satisfactory solution (line 15).

Again, there is convincing numerical evidence that this will generally happen, but a proof is currently lacking.