### 2025-05-29 repro/papers/ssevelan/slides/slidesSEG19.tex

\documentclass[xcolor=dvipsnames,12pt,aspectratio=169]{beamer}
\usecolortheme[named=violet]{structure}

%\usetheme{umbc2} 
\usepackage{pgf}
%\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage{amssymb,amsmath}
\usepackage[english]{babel}
\mode<presentation>

 \usepackage{multimedia}
\usefonttheme{structurebold}

%\useinnertheme{umbcboxes}

%\setbeamercolor{umbcboxes}{bg=violet!15,fg=black}

\setbeamercovered{transparent}
\setbeamertemplate{navigation symbols}{}
\input{./math-macros}
\setlength{\parskip}{2ex}   % place a blank line between paragraphs
\def\Logo{\@ifnextchar(\beamer@Logo\logo}
\def\beamer@Logo(#1,#2){\logo}

\AtBeginSection[]{
{\setbeamertemplate{logo}{}
 \begin{frame}<beamer>
    \frametitle{Agenda}
    \tableofcontents[currentsection]
  \end{frame}
}
}
% Images
% Macros
\newcommand{\peq}{\,+\hspace{-0.15cm}=}
\newcommand{\bm}{\bar{m}}
\newcommand{\dm}{\delta m}
\newcommand{\dom}{\delta \bar{m}}
\newcommand{\doma}{\delta \bar{m}_{\alpha}}
\newcommand{\om}{\bar{m}}
\newcommand{\bH}{{\bf H}}
\newcommand{\bR}{{\bf R}}
\newcommand{\bcF}{{\bar{\cal F}}}
\newcommand{\cF}{{\cal F}}
\newcommand{\oF}{\bar{F}}
\newcommand{\of}{\bar{f}}
\newcommand{\ou}{\bar{u}}
\newcommand{\odF}{\bar{F}^{\dagger}}
\newcommand{\oS}{\bar{S}}
\newcommand{\oM}{\bar{M}}
\newcommand{\oJa}{\bar{J}_{\alpha}}
\newcommand{\Na}{N_{\alpha}}
\newcommand{\dd}{\delta d}
\newcommand{\ddl}{\delta d_{\alpha}}
\title[]{Waveform Inversion via Source Extension}
\author[]{Guanghui Huang \inst{1} \and Rami Nammour \inst{2} \and \\William W. Symes \inst{3} \and Mohamed Dolliazal \inst{2}}
\institute[]{\inst{1} Michigan State University \and \inst{2} Total $E\&P$ $R\&T$ USA \and \inst{3} Rice University}
\date{September 2019}

\begin{document}
\frame{
\titlepage
}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

%\begin{frame}\frametitle{The Plan}
%\begin{itemize}
%\item What is ``extended modeling", and why bother
%\item FWI via surface source extension - examples
%\item Theory: why \& when it works
%\item Zoology of source extensions
%\end{itemize}
%\end{frame}

\section{What is ``extended modeling'', and why bother}

\begin{frame}%\frametitle{Conventional FWI}
FWI: adjust model to minimize data misfit (observed - modeled)

Wave equation:
\[
L[m]u = f 
\]
$m$ = model parameters, $L$ = wave operator, $u$ = modeled wavefield(s), $f$ = energy source model

FWI objective: 
\[
J_{\rm FWI}[m] = \sum (R u - d)^2
\]
$R$ = sampling operator, $Ru$ = modeled data, $d$ = observed data, $\sum$ = sum over all components

Note: Wave equation $\Rightarrow$ $u = u[m,f]$
\end{frame}

\begin{frame}
A simple example: acoustic transmission inversion w/ isotropic point source @ $\bx_s$: $f = w(t)\delta(\bx-\bx_s)$
%\[
%\frac{1}{\kappa}\frac{\partial p}{\partial t} = -\nabla \cdot {\bf v} + w(t)\delta(\bx-\bx_s), \,\rho \frac{\partial {\bf v}}{\partial t} = - \nabla p; \,p,{\bf v}=0, t \ll 0
%\]
\vspace{-1.25in}
%\vspace{-0.5in}
\begin{center}
\hspace{-0.5in}\includegraphics[height=4.0in]{Fig/bml0c.pdf}\\
%\includegraphics[height=2.0in}{Fig/bml0.pdf}\\
\vspace{-1in}
target bulk modulus $\kappa$, source at $x_s=3500$ m, $z_s=3000$ m, receivers at $x_r=2000-6000$ m, $z_r=1000$ m
\end{center}
\end{frame}

\begin{frame}
%Modeling operator $(S[\kappa]f)(t,x_r) = \phi(x_r) p(t,x_r,z_r)$ - linear in $f$, nonlinear in $\kappa$
Model $m=[\kappa]$, wavefield $u=[p,\bv]$, $L[m]u=$ acoustic wave system, $R$ = sample $p$ at receivers, 
\begin{center}
\includegraphics[height=1.5in]{Fig/pulse00.pdf} \includegraphics[height=1.5in]{Fig/ptpwindl0.pdf}\\
\vspace{0.25in}
Left: [1.0, 2.0, 7.5, 12.5] Hz trapzoidal bandpass filter pulse $w$; Right: pressure traces $Ru$ at $z_r=1000 m$ for source at $x_s=3500 m, z_s=3000 m$. 
\end{center}
\end{frame}

\begin{frame}
Minimizing $J_{\rm FWI}[m] = \sum (R u - d)^2$:

\begin{itemize}
\item Modeling (computing $u$ etc.): time-domain Finite Difference
\vspace{0,5cm}
\item Efficient comp'n of $\nabla J_{\rm FWI}$ = ``adjoint state method'' (Chavent-Lemmonier 74,...) - $\sim$ RTM
\vspace{0.5cm}
\item Iteratively update $m$ to reduce $J_{\rm FWI}$ - Newton's or related gradient/descent method, plus smoothing regularization
\end{itemize}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=1.25in]{Fig/bml0.pdf}\includegraphics[height=1.25in]{Fig/fwitrbulkmodinvlh0.pdf}\\
\vspace{0.5cm}
Left: target model; 
Right: FWI estimate of $m$ = bulk modulus: initial model $\kappa_0 \equiv 4.0$ GPa, 20 iterations trust region Krylov-Gauss-Newton with smoothing preconditioner. Each KGN step =  up to ``inner'' 20 CG steps $\approx$ 400 fwd models
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=1.2in]{Fig/ptpwindl0.pdf}\includegraphics[height=1.2in]{Fig/fwitrreslh0wind.pdf}\\
\vspace{0.5cm}
Left: target data; 
Right: data residual after 20 KGN steps - final RMS residual reduction = 0.26
%\|d\|=9.72 \|S-d\|=2.51
\end{center}
\end{frame}

\begin{frame}
Big problem: "cycle skipping"
\begin{itemize}
\item $J[m]$ appears to have many (local) minima
\item to avoid getting stuck at wrong estimates, initial $m$ (and $f$) must predict event times in $u$ ``correct within a wavelength''
\end{itemize}

OK with sufficiently low frequency data with high S/N, but not always available
\end{frame}

\begin{frame}\frametitle{An Analogy}
%\vspace{0.5cm}
Task: Given latitude of Vancouver ($49^{\circ} 15' N$), find its longitude by traveling there from somewhere in Canadian Rockies ($\approx 121^{\circ} W$)

Algorithm: start at correct latitude (eg. Sunshine Valley, $49^{\circ} 16' N$, $121^{\circ} 11' W$) and go {\color{blue} downhill and due west} until altitude minimized 
\vspace{-3cm}
\begin{center}
\includegraphics[width=0.6\textwidth,angle=270]{Fig/bc.pdf}
\end{center}
\vspace{-3cm}
\footnotetext{\tiny{https://www.google.com/maps/@49.2332511,-122.8181263,9z, 2019.08.18}}
\end{frame}

\begin{frame}
You, looking west. Now, {\color{blue} go downhill and due west}... oops.
\vspace{-1.5cm}
\begin{center}
\includegraphics[width=0.6\textwidth,angle=270]{Fig/mtns.pdf}
\end{center}
\vspace{-1cm}
%\tiny{{https://www.dreamingandwandering.com/2016/01/10/25-photos-of-the-canadian-rockies-that-will-make-you-pack-your-bags-and-go/}, 2019.08.21}
\footnotetext{\tiny{\noindent https://www.dreamingandwandering.com/2016/01/10/25-photos-of-the-canadian-rockies-that-will-make-you-pack-your-bags-and-go/}, 2019.08.21}
\end{frame}

\begin{frame}
Analogy w/ FWI: altitude $\sim$ $J_{\rm FWI}[m] =$ mean square error

Can't reduce to desired min = 0 while going downhill ($\sim$ using descent method) {\color{blue} and} maintaining latitude of Vancouver ($\sim$ known physical properties of solution)

Descent methods - mandatory because of computational size, can't drop

Alternative: drop ``physical constraint'' (latitude), search in larger ``unphysical'' space 
\end{frame}

\begin{frame}
With full 2D search available, can go to local creek, then float down to local river, join Fraser River near Hope ($49^{\circ} 23' N$) ... float your boat to Vancouver, downhill all the way!
\vspace{-2.8cm}
\begin{center}
\includegraphics[width=0.5\textwidth]{Fig/fraser.pdf}
\end{center}
\vspace{-3cm}
Moral: dropping some known properties (physical constraints) of solution {\color{blue} during search} may make descent search feasible

\footnotetext{\tiny{https://images.search.yahoo.com/yhs/search?p=fraser+river}, 2019.08.21}

\end{frame}

\section{FWI via source extension}
\begin{frame}

{\color{blue} Extended modeling:} allow interim models to violate known ``physical'' constraints on true solution

Active source data (usually) involves {\em localized} source model with known location, eg. ($\bx=(x,y,z)$)
\[
f(\bx,t;x_s,y_s) = w(t) \delta(\bx-\bx_s) = \mbox{ isotropic point source at }\bx=\bx_s
\]
{\color{blue} Source Extension}: replace $f$ with {\em non-localized} source model $\of$

Extended wavefield $\ou$  solves $L[m]\ou = \of$, data prediction = $R\ou$ - computational complexity similar to FWI

\end{frame}

\begin{frame}
Source extension methods:
\begin{itemize}
\item
Wavefield Reconstruction Inversion (WRI):
van Leeuwen and Herrmann GJI 2013, Inv. Probs. 2016, Wang \& Yingst SEG 2016, Aghamiry et al GEO 2019 
\item Adaptive Waveform Inversion (AWI): Warner \& Guatsch, GEO 2014, 2016
\item ... (see abstract)
\item {\color{blue} surface source extension (SSE):} $\of(\bx,t;x_s,y_s) = w(x,y,t; x_s,y_s)\delta(z-z_s)$
\end{itemize}

Common feature: additional source degrees of freedom permit good fit to data {\em for large set of $m$}
\end{frame}

\begin{frame}
Detailed Formulation:

At solution, must recover known constraints on model (``get closer to Vancouver's latitude as you get close to the sea''):

$\Rightarrow$ introduce {\color{blue} annihilator} $A$, vanishing on models obeying these constraints - minimize output jointly with data residual

%Many possibilities - MS difference w/ known source (WRI), mean time lag (AWI)

Many choices. For SSE: scale by offset 
\[
A\of(\bx,t;x_s,y_s) = \sqrt{(x-x_s)^2+(y-y_s)^2}w(x,y,t;x_s,y_s)
\]

{\color {blue} Extension Principle:} $m,\of$ so  that $R\bar{u} \approx d$, $A\of \approx 0$ $\Rightarrow$ approx FWI solution
\end{frame}

%\begin{frame}
%Extended objective function: data misfit + annihilator output,

%\[
%J_{\rm SSE}[m] = \min_{\of} \,\frac{1}{2} \sum (R\ou - d)^2 + \alpha^2 \sum (A\of)^2
%\]

%Reduced: ``pre-minimized'' over source

%FWI via surface source extension: given $d$ and initial $m_0$, use numerical optimization method to find minimizer $m$ of
%$J_{\rm SSE}[m]$.
%\end{frame}

%\begin{frame}
%Examples:

%(1) diving wave data (sampling op may include window in defn)

%(2) reflection data

%In both cases, method is [??? VPM, CG for inner problem, steepest descent for outer???]
%\end{frame}

%\begin{frame}
%Mohamed's examples 
%\begin{itemize}
%\item include straight FWI for both
%\item also some info about objective values, residuals, convergence, etc.
%\item plots of models, but also sources, residuals - especially want to show extended source from early iteration
%\end{itemize}
%\end{frame}

\begin{frame}
FWI via SSE: given $d$ and initial $m_0$, use numerical optimization method to find minimizer $m$ of
\[
J_{\rm SSE}[m] = \mbox{ min}_{\of}\frac{1}{2} \sum (R\ou - d)^2 + \alpha^2 \sum (A\of)^2
\]
over $m$

\begin{itemize}
\item time-domain FD, gradient via adjoint state
\item update $\of$ using Conjugate Gradient Iteration (inner loop)
\item update $m$ using steepest descent, line search (outer loop) plus smoothing regularization
\item update $\alpha$ using Discrepancy Principle (keep data residual $\in (e_-, e_+)$) - {\color{blue} can always choose $\of$ to fit data} $\Rightarrow$ can start with $\alpha=0$ (Fu \& S. GEO 2017)
\end{itemize}
\end{frame}

\begin{frame}
\begin{center}
Example 1: Lens model
\end{center}
\begin{center}
\includegraphics[height=1.25in]{Fig/cgw1bulkupdlh0.pdf}\includegraphics[height=1.25in]{Fig/cgw30bulkupdlh0.pdf}\\
\vspace{0.25in}
Start at $\kappa_0=4.0$ GPa, use $e_-=0.1\|d\|, e_+=0.2\|d\|$. \\
Left: Iteration 1 ($\alpha=4.83 \times 10^{-6}$)\\
Right: Iteration 30 ($\alpha = 9.1\times 10^{-5}$)
\end{center}
\end{frame}

\begin{frame}
$m=[\kappa]$ converges $\leftrightarrow$ $\alpha$ increases, extended source $\of$ (initially spread over $z=z_s=1000$ m) {\color{blue} focuses} at $x=x_s=3500$ m.
\begin{center}
\hspace{-1cm}\includegraphics[height=1.25in]{Fig/cgw1estsourceplh0.pdf}\includegraphics[height=1.25in]{Fig/cgw7estsourceplh0.pdf}\includegraphics[height=1.25in]{Fig/cgw29estsourceplh0.pdf}\\
\vspace{0.25in}
Left to right: $\of$ at Iterations 1 ($\alpha=4.83 \times 10^{-6}, r=2000$ m), 8 ($\alpha= 1.32 \times 10^{-5}, r=1000$ m) and 30 ($\alpha=9.1 \times 10^{-5}, r=250$ m).
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=1.25in]{Fig/fwitrsse1bulkmodinvlh0.pdf}\includegraphics[height=1.25in]{Fig/bml0.pdf}\\
\vspace{0.25in}
Left: ``Jump to $\alpha=\infty$": bulk modulus estimated by FWI, starting with iteration 30 of SSE. 18 steps KGN, each $\le$ inner 20 CG steps, final RMS residual reduction = 0.012;
Right: Target bulk modulus field
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{-0.7in}
\includegraphics[height=4.5in]{HMC1.pdf}
\end{center}
\end{frame}

\begin{frame}
HMC Salt/Carbonate Acoustic Model

Streamer acquisition $z_s = 7$ m, $\Delta x_s=37.5$ m, 315 shots, $z_r=15$ m, $\Delta x_r=12.5$ m, 475 recvs, max offset = $6$ km

Pulse for data generation = [2, 3, 6, 10] Hz trapezoidal bandpass filter

Free surface @ $z=0$, other boundaries absorbing, 2-8 staggered FD scheme

Regularization: penalize MS velocity gradient, weight (also $\alpha$) via several trials

Data contains diving waves, pre- \& post-critical reflections, multiple reflections, refracted multiple reflections, head waves, free surface multiples, ....

\end{frame}

\begin{frame}
\begin{center}
\vspace{-0.7in}
\includegraphics[height=4.5in]{HMC2.pdf}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{-0.7in}
\includegraphics[height=4.5in]{HMC3.pdf}
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\vspace{-0.7in}
\includegraphics[height=4.5in]{HMC4.pdf}
\end{center}
\end{frame}

\section{Theory: why \& when it works}

\begin{frame}
FWI vs. SSE for transmission (diving wave, crosswell) data:

$J_{\rm FWI}[m]$ and $J_{\rm SSE}[m]$ = regular functions of $m$, locally close to quadratic approximations.

$d$ consistent: $d=Ru[m,f]$, $\of = f = w(t)\delta(\bx-\bx_s)$, $A\of=0$ $\Rightarrow$

\begin{itemize}
\item  $J_{\rm FWI}[m]=J_{\rm SSE}[m]=0$
\vspace{0.5cm}
\item $J_{\rm FWI}[m + \delta m] \approx \mbox{const. }\omega^2 \|\delta m\|^2$ for some $\delta m$ (kinematic perturbations), $\omega$ = center frequency in source pulse $w(t)$
\vspace{0.5cm}
\item $J_{\rm SSE}[m+\delta m] \le \mbox{const. }\|\delta m\|^2$ independent of $\omega, \delta m$
\end{itemize}
\end{frame}

\begin{frame}
$\Rightarrow$ radius of convexity region:
\begin{itemize}
\item $J_{\rm FWI}: O(1/\omega)$, must initialize ``within a wavelength''
\item $J_{\rm SSE}: O(1)$, converges even with (rel.) poor initial guess, HF data
\end{itemize}

Method of proof: ray theory (``mocrolocal analysis'')

More detailed analysis: minimizing $J_{\rm SSE}$ equivalent to a form of transmission travel time tomography: {\color{blue} tomography without picking}

NB: for reflection data... ???
\end{frame}

\begin{frame}\frametitle{Late Breaking News}
Theory for special reflection case: layered ($\kappa(z)$) medium, dip filter to eliminate postcritical/diving wave energy

Explicit computation of $\nabla J_{\rm SSE}$ for 
\begin{itemize}
\item homogenous background 
\item data from single reflector in otherwise slowly varying $\kappa$, wavelet with central frequency $\omega$
\end{itemize}
$\Rightarrow$ $\nabla J_{\rm SSE} \rightarrow 0$ as $\omega \rightarrow \infty$

Same behavior as $\nabla J_{\rm FWI}$!!!

Upshot: for pure reflection data, SSE may not provide constructive velocity updates
\end{frame}

\begin{frame}
{\color{blue} {\bf Summary:}} Surface Source Extension

\begin{itemize}
\item overcomes cycle-skipping for transmission data, solid theoretical foundation
\item same computational complexity as FWI
\item open question: theory for reflections {\color{blue} (initial result not positive)}
\end{itemize}
%\end{frame}

%\begin{frame}
Other source extension methods (WRI, AWI, ...)

\begin{itemize}
\item Many promising examples 
\item Little known theoretically - connection to travel time tomography???
\end{itemize}
\end{frame}

\begin{frame}\frametitle{Model Extension}
Add degrees of freedom to $v$, $\kappa$, instead of source

Annihilator: flatten/focus gathers

{\color{blue}Extracts velocity updates from reflection and refraction events} (S. 86, Kern \& S. 94, S. GP 08, Shen \& S. GEO 08, Shen EAGE 12, Biondi \& Almomin 12, 14, Chauris \& Cocher GEO 17, Hou \& S. GEO 18, Barnier \& Biondi SEG 19)

Chief challenge: expense(model extension) $\gg$ expense(FWI) (Note: expense(source extension) $\approx$ expense(FWI))
\end{frame}

\begin{frame}\frametitle{Acknowledgements}
\begin{itemize}
\item Total E\&P USA
\item Sponsors of The Rice Inversion Project
\item Organizers of SS 1 (Recent Advances and Road Ahead) at SEG 2019
\end{itemize}
\end{frame}

\end{document}

\begin{frame}
Matched Source Waveform Inversion (Hua \& Symes 1993, 4, Huang \& Symes 2017)

\begin{itemize}
\item similar to AWI (source wavelet for each trace), w/o normalization
\item if
\begin{itemize}
\item transmission data (crosswell, diving wave)
\item unique ray paths
\end{itemize}
then mean square moment $\approx$ mean square travel time misfit: {\color{blue} tomography without picking}

\item relation with tomography fails ($\approx$ FWI) if multiple rays connect sources and receivers - contrast SSE
\end{itemize}
\end{frame}

\begin{frame}
\begin{center}
Example 2: Deep Water Marmousi
\end{center}
\begin{center}
\includegraphics[height=2in,width=4in]{Fig/marm2km/Marm2kmVt.pdf}\\
Modified Marmousi Model: 2 km water layer added to minimize non-reflected energy, also box-snaped target
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=2in]{Fig/marm2km/Marm2kmShot1}\includegraphics[height=2in]{Fig/marm2km/Marm2kmShot81}\\
Data: isotropic point source with 8 Hz Ricker wavelet, $z_r=0.04$ km, $x_r=$ 0.05 km to 9.15 km, $\Delta x_r=$ 0.05 km, $z_s=0.08$ km, $x_s=0.6$ km to $x_s=8.6$, $\Delta x_s$ = 0.1 km. Shots at $x_s=0.6$ km and $x_s=8.6$ km displayed.
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=2in,width=4in]{Fig/marm2km/Marm2kmV0.pdf}\\
Initial $v(z)$ model for both FWI and SSE
\end{center}
\end{frame}

\begin{frame}
\vspace{-0.1cm}
\begin{center}
\includegraphics[height=2in,width=4in]{Fig/marm2km/Marm2kmFWI.pdf}\\
FWI estimate of $v$: 200 iterations LBFGS, frequency domain FD 5-8 Hz
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=2in,width=4in]{Fig/marm2km/Marm2kmSurf.pdf}\\
SSE estimate of $v$: 150 iterations LBFGS, frequency domain FD 5-8 Hz, fixed $\alpha$ set by trial
\end{center}
\end{frame}

\begin{frame}
%\vspace{0.07cm}
\begin{center}
\includegraphics[height=2in,width=4in]{Fig/marm2km/Marm2kmVt.pdf}\\
Modified Marmousi Model: 2 km water layer added to minimize non-reflected energy, also box-snaped target
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=2in]{Fig/marm2km/Marm2kmShot1Inv}\includegraphics[height=2in]{Fig/marm2km/Marm2kmShot81Inv}\\
Data predicted from SSE inverted model: shots at $x_s=0.6$ km and $x_s=8.6$ km.
\end{center}
\end{frame}

\begin{frame}
\begin{center}
\includegraphics[height=2in]{Fig/marm2km/Marm2kmShot1}\includegraphics[height=2in]{Fig/marm2km/Marm2kmShot81}\\
Target data:\\ shots at $x_s=0.6$ km and $x_s=8.6$ km.
\end{center}
\end{frame}

\begin{frame}
Wavefield Reconstruction Inversion: $\of(\bx,t;\bx_s)$ distributed throughout space-time

van Leeuwen and Herrmann 2013, 2016, Wang \& Yingst 2016

similar: space-time extension Huang \& S. 2018

\begin{itemize}
\item[+ ]many good examples, convincing numerical evidence
\item[+ ]3D field data - Wang \& Yingst 2016
\item[- ]basic form: store complete space-time (or space-frequency) fields (6D)
\end{itemize}
\end{frame}

\begin{frame}
Adaptive Waveform Inversion (Warner \& Guatsch, 2014, 2016): $\of(\bx,t;\bx_r,\bx_s) = w(t;\bx_r)\delta(\bx-\bx_s)$ localized but depends on receiver coordinate $\bx_r$ - one wavelet per data trace

similar: source-receiver extension Hua \& S. 1994, Huang \& S. 2017

\begin{itemize}
\item[+ ]additional storage: $\approx$ data volume (5D)
\item[+ ]minor mod of standard FWI workflow
\item[- ]source-receiver: loses advantage over FWI when sources, receivers connected by multiple ray paths (fails for lens model, above)
\end{itemize}
\end{frame}

\begin{frame}
Intermediate: source distributed over space-time surface
\begin{itemize}
\item const. t surface = volume source extension - Huang \& S. 2018
\item const z surface = {\color{blue} surface source extension (SSE):} 
\[
\of(\bx,t;x_s,y_s) = w(x,y; x_s,y_s)\delta(z-z_s)
\]
\end{itemize}
\begin{itemize}
\item[+ ]additional storage = O(data volume) (5D)
\item[+ ]no problem with multiple ray paths
\item[+ ]theory for transmission data (eg. diving wave): {\color{blue} tomography w/o picking}
\end{itemize} 
\end{frame}